### 2020-06-29 paper.tex
added notes for paper

Notes for paper:
- list requirements and assumptions (e.g. Files are RO)
- make sure it's clearly explained why Sea over writing your own script to manage filesystems
    - Sea will manage disk and memory for you (abstraction of the script to be written)
    - Difficulties of intercepting libc calls. Describe and explain custom code that has been added
- Spark vs Burst Buffer vs Lustre caching vs Sea. Need to benchmark Sea vs burst buffer to measure
  applicability when Burst Buffer options are available
- Sea vs Neuroimaging workflow engines
- Splitting data into chunks to maximize storage space or defined constant with max file size
- Intercepting system calls vs libc calls.
- Making Sea flushing pipeline-aware through scheduler hooks
- Flushing: Pros and cons of single flusher/node vs multiple flushers/node (maximize network bandwidth usage)
- Distributed Sea? Sea will function in a distributed manner with Sub. Not for this paper

### 2020-07-09 experiments/incrementation/run_bench.sh
code fixes and some results

@@ -3,33 +3,54 @@
 #set -e
 set -u 
 
+typeset -i i END
+
+let END=$1
+
 export SEA_HOME=/home/vhs/Sea
 experiments=("${SEA_HOME}/sea_launch.sh sea_parallel.sh")
 experiments+=("bash ./lustre_parallel.sh")
-for e in {0..2}
+results_file=./results/result.csv
+curr_exp=""
+
+order=( 0 1 2 3 )
+
+echo "experiment,repetition,runtime,max_mem" > ${results_file}
+
+for ((i=0;i<END;i++))
 do
-    for i in {1..3}
+    order=( $(shuf -e "${order[@]}") )
+    for e in "${order[@]}"
     do
         echo $e $i
         rm /mnt/lustre/vhs/output/* sea.log ; rm -rf /disk0/vhs/seatmp /dev/shm/seatmp
         echo 3 | sudo tee /proc/sys/vm/drop_caches
         echo running experiment ${experiments[0]}
+
+        exec 3>&1 4>&2
+
         if [[ $e == 0 ]]
         then
+            curr_exp="mem_all"
+            cp .sea_flushlist_all ${SEA_HOME}/.sea_flushlist
             echo "mem ${experiments[0]}" 
             cp mem_sea.ini ${SEA_HOME}/sea.ini
-            time ${experiments[0]}
-        elif [[ $e == 1 ]]
+            res=$({ command time -f "%e,%M" ${experiments[0]} 1>&3; } 2>&1 | grep "[0-9]*\.[0-9]*,[0-9]*" )
+        elif [[ $e == 2 ]]
         then
-            echo "disk ${experiments[0]}"
-            cp disk_sea.ini ${SEA_HOME}/sea.ini
-            time ${experiments[0]}
+            curr_exp="mem_final"
+            cp .sea_flushlist_mem ${SEA_HOME}/.sea_flushlist
+            echo "mem_final ${experiments[0]}" 
+            cp mem_sea.ini ${SEA_HOME}/sea.ini
+            res=$({ command time -f "%e,%M" ${experiments[0]} 1>&3; } 2>&1 | grep "[0-9]*\.[0-9]*,[0-9]*" )
         else
+            curr_exp="lustre"
             echo "lustre ${experiments[1]}"
-            time ${experiments[1]}
-        fi 
-        echo 'done'
+            res=$({ command time -f "%e,%M" ${experiments[1]} 1>&-; } 2>&1 )
+        fi
[…]

### 2020-07-09 sea_flusher.sh

@@ -13,7 +13,7 @@ flush_file=${SEA_HOME}/.sea_flushlist
 
 flush () {
         re_flush=""
-        for s in $sources_arr
+        for s in "${sources_arr[@]}"
         do
             while IFS="" read -r re || [ -n "$re" ]
             do
@@ -22,14 +22,14 @@ flush () {
                     re_flush+=" "
                 fi
 
-                re_flush+=$sources_arr/$re
+                re_flush+=$s/$re
 
             done < $flush_file
         done
 
         if [[ $re_flush != "" ]]
         then
-            found_files=$(find ${re_flush} -type f -follow -print || echo "")
+            found_files=$(find ${re_flush} -type f -follow || echo "")
             
             if [[ $found_files != "" ]]
             then
@@ -42,7 +42,7 @@ flush () {
                         if [[ "$f" == "${s}/"* ]]
                         then
                             file_out=$(echo "${f/$s/$base_source}")
-                            echo "flushing and evicting $f to $file_out"
+                            #echo "flushing and evicting $f to $file_out"
                             rsync -a --no-owner --no-group --remove-source-files "$f" "$file_out" || ( sleep 5 && rsync -a --no-owner --no-group --remove-source-files "$f" "$file_out" )
                             break
                         fi
@@ -56,7 +56,7 @@ flush () {
 flush_process () {
     while :
     do
-        flush 
+        flush
         sleep 5 &
         wait $!
     done

### 2020-07-09 sea.cpp
code fixes and some results

@@ -98,7 +98,7 @@ int sea_getpath(const char* oldpath, char passpath[PATH_MAX], int masked_path, i
     struct config sea_config = get_sea_config();
 
     int match = 0;
-    char tmp_passpath[PATH_MAX] = { '\0' };
+    int exists = 0;
 
     if (sea_lvl != -1) {
         match = pass_getpath(oldpath, passpath, masked_path, sea_lvl);
@@ -110,21 +110,11 @@ int sea_getpath(const char* oldpath, char passpath[PATH_MAX], int masked_path, i
             //printf("passpath %s\n", passpath);
 
             if ( masked_path == 0 && match == 1) {
-                int exists = sea_checkpath(passpath);
+                exists = sea_checkpath(passpath);
                 //printf("exists %d %s\n", exists, passpath);
 
                 if (exists)
                     return match; 
-                else if (tmp_passpath[0] == '\0') {
-                    set_internal();
-                    struct statvfs buf;
-                    int ret;
-                    if((ret = statvfs(sea_config.source_mounts[i], &buf) == 0) && (buf.f_bavail * buf.f_bsize > sea_config.max_fs * sea_config.n_threads)) {
-                        // if doesn't exist, create at top of hierarchy
-                        strcpy(tmp_passpath, passpath);
-                    }
-                    unset_internal();
-                }
             }
             else if ( masked_path == 1 ) {
                 return match;
@@ -132,9 +122,17 @@ int sea_getpath(const char* oldpath, char passpath[PATH_MAX], int masked_path, i
         }
     }
 
-    if (match == 1 && tmp_passpath[0] != '\0') {
-        // printf("tmppath %s\n", tmp_passpath);
-        strcpy(passpath, tmp_passpath);
+    if (match == 1 && exists == 0) {
+        for ( int i=0 ; i < sea_config.n_sources; ++i ) {
+            set_internal();
+            struct statvfs buf;
+            int ret;
+            if((ret = statvfs(sea_config.source_mounts[i], &buf) == 0) && (buf.f_bavail * buf.f_bsize > sea_config.max_fs * sea_config.n_threads)) {
+                match = pass_getpath(oldpath, passpath, masked_path, i);
+                unset_internal();
+                return match;
+            }
+        }
     }
 
     return match;

### 2020-07-09 experiments/incrementation/lustre_parallel4.sh
added some additional results

@@ -1,0 +1,2 @@
+#!/bin/bash
+parallel --jobs 4 < lustre_jobs.txt

### 2020-07-09 experiments/incrementation/sea_parallel4.sh

@@ -1,0 +1,1 @@
+parallel --jobs 4 < sea_jobs.txt

### 2020-07-09 paper/paper.tex
rough paper outline

\documentclass[10pt,journal,compsoc]{IEEEtran}

\begin{document}
\title{Sea: A hierarchical filesystem for processing Big Neuroimaging Data on HPC}

\author{Val\'erie Hayot-Sasson and Tristan Glatard}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
\IEEEtitleabstractindextext{%
\begin{abstract}
\end{abstract}
}

% make the title area
\maketitle

\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{N}{euroimaging} datasets have significantly grown in size in the past decade,
resulting in Petabytes of data. As a result, processing pipelines have now shifted from compute to 
data intensive. Some efforts have been made to reduce the data transfer costs in neuroimaging by
introducing new methods to generate pipelines, such as through the adoption of Big Data frameworks
like Apache Spark \cite{Spark} for the pipeline definition \cite{thunder} \cite{other papers}. 
However, such pipelines would have to be entirely rewritten for the framework of choice and Big Data
frameworks like Spark and Dask were found to be difficult to use in describing neuroimaging pipelines \cite{mehta}.
Furthermore, the neuroimaging community relies heavily on the use of well-established command-line tools \cite{fsl, afni, freesurfer}, resulting in
data transfers being inevitable, even with a Big Data framework.

High Performance Computing (HPC) clusters are often the infrastructure of choice for researchers looking to processing their data.
While there are many benefits to using such clusters, such as large amounts of resources available at little to no cost to
the researchers, the resources are shared amongst users concurrently.

Goal - leverage local storage to speed up computation/Sea

\subsection{Lustre}
\subsubsection{Burst Buffers}
\subsubsection{Other Luster optimizations}
\subsection{Big Data Frameworks and File systems}
HDFS Alluxio Hadoop Spark Dask
\subsubsection{What are they/how do they work}
\subsubsection{Why are the suboptimal for neuroimaging/on HPC}
\subsection{The linux page cache}
\subsubsection{What is it/how does it work}
\subsubsection{Why do we still need Sea}
\subsection{File System implementations}
\subsubsection{Kernel-space file systems}
\subsubsection{FUSE}
\subsubsection{System call interception with ptrace}
\subsubsection{LD\_PRELOAD}

\section{Materials and Methods}

\subsection{Sea design and implementation}
\subsubsection{Requirements and Assumptions}
\subsubsection{Benefits}
User does not need to create logic to determine what file systems to write to / implement flushing

\subsubsection{libc interception}
\subsubsection{Flushing}
\subsubsection{Configuration File}
\subsubsection{Program execution}
\subsubsection{Limitations}

\subsection{Experiments}
\subsubsection{Infrastructure}
\subsubsection{Pipelines}
Big Brain incrementation
fmri processing?

\section{Results}

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% biography section

\begin{IEEEbiography}{Val\'erie Hayot-Sasson}
\end{IEEEbiography}
\begin{IEEEbiography}{Tristan Glatard}
\end{IEEEbiography}
\end{document}

### 2020-07-10 sea_launch.sh
some folder cleanup and addition of figures

@@ -1,28 +1,0 @@
-#!/bin/bash
-
-set -e
-set -u
-
-trap "exit" INT TERM
-#trap "kill 0" EXIT
-
-conf_file=${SEA_HOME}/sea.ini
-
-n_sources="$(($(awk -F "=" '/n_sources/ {print $2}' ${conf_file} | tr -d ' ;') - 1))"
-
-for ((i=0;i<${n_sources};i++))
-do
-    source=($(cat ${conf_file} | grep "source_$i" | cut -d "=" -f 2 | tr -d ' ;'))
-    mkdir -p ${source} 
-done
-
-bash ${SEA_HOME}/sea_flusher.sh &
-pid=$!
-echo "flusher pid" ${pid}
-
-trap 'trap - SIGTERM && kill $pid' SIGINT SIGTERM EXIT
-
-echo "Launching script $1"
-bash $1
-echo "killing flush" && kill $pid
-

### 2020-07-11 src/config.cpp
some bug fixes

@@ -49,6 +49,7 @@ void create_config_file(char * config_file)
             "log_level = 1 ; # DEBUG=4, INFO=3, WARNING=2, ERROR=1, NONE=0 (use <= 3 tests)\n"
             "n_sources = 1 ;\n"
             "source_0 = %s/source ; # use absolute paths\n"
+            "max_fs = 100"
             "\n", sea_home, sea_home, sea_home);
     fclose(ini);
 }
@@ -99,7 +100,7 @@ void parse_config()
         printf("Missing log_file in config file %s\n", config_file);
         exit(1);
     }
-    if((sea_config.max_fs = (unsigned long) iniparser_getlongint(config_dict, "sea:max_fs", 0))==0)
+    if((sea_config.max_fs = atol((char *) iniparser_getstring(config_dict, "sea:max_fs", NULL)))==0)
     {
         printf("Missing max_fs in config file %s\n", config_file);
         exit(1);

### 2020-07-11 src/sea.cpp

@@ -102,6 +102,7 @@ int sea_getpath(const char* oldpath, char passpath[PATH_MAX], int masked_path, i
 
     if (sea_lvl != -1) {
         match = pass_getpath(oldpath, passpath, masked_path, sea_lvl);
+        return match;
         //printf("match %d %s\n", match, passpath);
     }
     else {

### 2020-07-11 tests/test_custom.cpp

@@ -1,7 +1,7 @@
 #include <gtest/gtest.h>
-#include "../passthrough.h"
-#include "../sea.h"
-#include "../config.h"
+#include "../src/passthrough.h"
+#include "../src/sea.h"
+#include "../src/config.h"
 #include <stdlib.h>
 
 // need to add symlink tests here

### 2020-07-11 src/config.cpp
added the ability to specific multiple sources of the same level

@@ -3,6 +3,11 @@ extern "C"{
 }
 #include <linux/limits.h>
 #include <sys/stat.h>
+#include <vector>
+#include <sstream>
+#include <algorithm>
+#include <ctime>
+#include <cstdlib>
 #include "config.h"
 #include "logger.h"
 #include "passthrough.h"
@@ -47,12 +52,15 @@ void create_config_file(char * config_file)
             "mount_dir = %s/mount_dir ; # use absolute paths\n"
             "log_file = %s/sea.log ;\n"
             "log_level = 1 ; # DEBUG=4, INFO=3, WARNING=2, ERROR=1, NONE=0 (use <= 3 tests)\n"
-            "n_sources = 1 ;\n"
+            "n_levels = 1 ;\n"
             "source_0 = %s/source ; # use absolute paths\n"
             "max_fs = 100"
             "\n", sea_home, sea_home, sea_home);
     fclose(ini);
 }
+
+// random generator function:
+int random_gen (int i) { return std::rand()%i;}
 
 void parse_config()
 {
@@ -70,25 +78,49 @@ void parse_config()
         printf("Cannot load config file %s\n", config_file);
         exit(1);
     }
-    if((sea_config.n_sources = iniparser_getint(config_dict, "sea:n_sources", 0))==0)
+    if((sea_config.n_levels = iniparser_getint(config_dict, "sea:n_levels", 0))==0)
     {
          printf("Missing n_sources in config file %s\n", config_file);
          exit(1);
     }
-    sea_config.source_mounts = new char* [sea_config.n_sources];
-    for(int i = 0 ; i <  sea_config.n_sources ; i++)
+    std::vector<std::string> all_sources;
+
+    // for random number generation
+    std::srand ( unsigned ( std::time(0) ) );
+
+    for(int i = 0 ; i <  sea_config.n_levels ; i++)
     {
+        std::vector<std::string> current_lvl;
         char source_name[15];
         sprintf(source_name, "sea:source_%d", i);
+        char * lvl_str;
         //sea_config.source_mounts[i] = new char[PATH_MAX];
-        if((sea_config.source_mounts[i] = (char *) iniparser_getstring(config_dict, source_name, NULL))==0)
+        if((lvl_str = (char *) iniparser_getstring(config_dict, source_name, NULL))==0)
         {
             printf("Missing %s in config file %s\n", source_name, config_file);
             exit(1);
         }
+        std::stringstream s_stream(lvl_str);
[…]

### 2020-07-11 src/sea.cpp
added the ability to specific multiple sources of the same level

@@ -154,7 +154,6 @@ int sea_getpath(const char* oldpath, char passpath[PATH_MAX], int masked_path, i
  */
 void populateFileSet(char *basePath, int sea_lvl, struct config sea_config, std::set<std::string> &sea_paths)
 {
-    //printf("base %s \n", basePath);
     char path[PATH_MAX];
     struct dirent *dp;
     DIR *dir = ((funcptr_opendir)libc_opendir)(basePath);
@@ -216,11 +215,14 @@ void initialize_sea() {
     sea_files.clear();
     sea_internal = 0;
     struct config sea_config = get_sea_config();
-    char ** source_mounts = sea_config.source_mounts;
 
     for (int i=0; i < sea_config.n_sources; i++){
-        populateFileSet(source_mounts[i], i, sea_config, sea_files);
+        populateFileSet(sea_config.source_mounts[i], i, sea_config, sea_files);
+        //printf("sources %s\n", sea_config.source_mounts[i]);
     }
+    //for (auto it=sea_files.begin(); it != sea_files.end(); ++it) 
+    //    printf("seafiles %s\n", it->c_str());
+    //printf("done\n");
 }
 
 static pthread_once_t sea_initialized = PTHREAD_ONCE_INIT;

### 2020-07-11 bin/sea_flusher.sh
accidentally removed the launch and the flusher. added them back

@@ -1,0 +1,85 @@
+#!/bin/bash
+
+
+set -e
+set -u
+
+base_source=""
+sources_arr=()
+n_sources=0
+conf_file=${SEA_HOME}/sea.ini
+flush_file=${SEA_HOME}/.sea_flushlist
+
+
+flush () {
+        re_flush=""
+        for s in "${sources_arr[@]}"
+        do
+            while IFS="" read -r re || [ -n "$re" ]
+            do
+                if [[ $re_flush != "" ]]
+                then
+                    re_flush+=" "
+                fi
+
+                re_flush+=$s/$re
+
+            done < $flush_file
+        done
+
+        if [[ $re_flush != "" ]]
+        then
+            found_files=$(find ${re_flush} -type f -follow || echo "")
+            
+            if [[ $found_files != "" ]]
+            then
+                tobe_flushed=$(ls -tur ${found_files})
+
+                for f in $tobe_flushed
+                do
+                    for s in $sources_arr
+                    do
+                        if [[ "$f" == "${s}/"* ]]
+                        then
+                            file_out=$(echo "${f/$s/$base_source}")
+                            #echo "flushing and evicting $f to $file_out"
+                            rsync -a --no-owner --no-group --remove-source-files "$f" "$file_out" || ( sleep 5 && rsync -a --no-owner --no-group --remove-source-files "$f" "$file_out" )
+                            break
+                        fi
+                    done
+                done
+            fi
+        fi
+}
+
+
+flush_process () {
+    while :
+    do
+        flush
+        sleep 5 &
+        wait $!
+    done
+}
+
+
+get_sources () {
+
+    # refers to the last index and not the total number of sources
+    n_sources="$(($(awk -F "=" '/n_sources/ {print $2}' ${SEA_HOME}/sea.ini | tr -d ' ;') - 1))"
+
+    for ((i=0;i<${n_sources};i++))
+    do
+        sources_arr+=($(cat ${conf_file} | grep "source_$i" | cut -d "=" -f 2 | tr -d ' ;'))
+        echo source_$i ${sources_arr[$i]}
+    done
+
+    base_source="$(cat ${conf_file} | grep "source_$i" | cut -d "=" -f 2 | tr -d ' ;')"
+    echo base dir ${base_source}
+
+}
+
+trap 'flush' SIGINT SIGTERM EXIT
+get_sources
+flush_process
+

### 2020-07-11 bin/sea_launch.sh
accidentally removed the launch and the flusher. added them back

@@ -1,0 +1,28 @@
+#!/bin/bash
+
+set -e
+set -u
+
+trap "exit" INT TERM
+#trap "kill 0" EXIT
+
+conf_file=${SEA_HOME}/sea.ini
+
+n_sources="$(($(awk -F "=" '/n_sources/ {print $2}' ${conf_file} | tr -d ' ;') - 1))"
+
+for ((i=0;i<${n_sources};i++))
+do
+    source=($(cat ${conf_file} | grep "source_$i" | cut -d "=" -f 2 | tr -d ' ;'))
+    mkdir -p ${source} 
+done
+
+bash ${SEA_HOME}/bin/sea_flusher.sh &
+pid=$!
+echo "flusher pid" ${pid}
+
+trap 'trap - SIGTERM && kill $pid' SIGINT SIGTERM EXIT
+
+echo "Launching script $1"
+bash $1
+echo "killing flush" && kill $pid
+

### 2020-07-21 bin/sea_flusher.sh
fixed sea flusher bugs

@@ -1,23 +1,27 @@
 #!/bin/bash
 
-
-set -e
+#set -e
 set -u
+#set -x
 
 base_source=""
 sources_arr=()
 n_sources=0
 conf_file=${SEA_HOME}/sea.ini
 flush_file=${SEA_HOME}/.sea_flushlist
+tmpfile=$(mktemp /tmp/.sea_flush.inprog.XXXXXX)
 
-get_rp_mv () {
-    tobe_flushed=$1
-    OLDTIME=$2
-    task=$3
+mv_old () {
+
+    found_files=$1 # files that currently exist in source mounts
+    OLDTIME=$2 # user specified expiration time
+    task=$3 # the name of the calling task (either "process" or "cleanup")
+
     flushlist=""
-    for f in $tobe_flushed
+    for f in $found_files
     do
-
+        
+        # check if file exists otherwise continue (in case it has been flushed by another process)
         if [ ! -f $f ]
         then
             continue
@@ -25,20 +29,21 @@ get_rp_mv () {
         # code adapted from here https://stackoverflow.com/a/28341234
         # Get current and file times
         CURTIME=$(date +%s)
-        FILETIME=$(stat $f -c %Y || true)
+        FILETIME=$(stat $f -c %Y || true && continue)
 
-        if [[ $FILETIME == "" ]]
+        if [ $FILETIME == "" ]
         then
             continue
         fi
 
+        # calculate how old the file is (based on access time)
         TIMEDIFF=$(($CURTIME - $FILETIME))
 
         # Check if file older
         if [ $TIMEDIFF -ge $OLDTIME ]; then
             echo "flushing and evicting $f to $base_source"
-            #[ ! -f $f ] ||  &
 
+            # some string formatting
             if [[ "$flushlist" != "" ]]
             then
                 flushlist+=" "
@@ -48,21 +53,34 @@ get_rp_mv () {
            
         fi
     done
+
+    # if we have some files that older that OLDTIME, move them
     if [[ ${flushlist} != "" ]]
     then
-        (mv $flushlist ${base_source} || true) &
-        pid=$!
+
+        # mv each image individually (as opposed to all at once)
+        for f in $flushlist
+        do
+            ([ ! -f $f ] || mv $f ${base_source} || true) &
+        done
+
+        # wait for all images to be moved
         wait
-        #echo "flush and evict ${flushlist}"
[…]

### 2020-07-21 experiments/incrementation/increment.py

@@ -24,6 +24,7 @@ for i in range(it):
     benchmark("read_start", in_file)
     im = nib.load(in_file)
     data = np.asanyarray(im.dataobj)
+    data += 0 # make sure it's in memory
     benchmark("read_end", in_file)
     benchmark("inc_start", in_file)
     inc = data + 1

### 2020-07-22 bin/sea_launch.sh
added eviction to in-memory

@@ -16,13 +16,15 @@ do
     mkdir -p ${source} 
 done
 
-${SEA_HOME}/bin/sea_flusher.sh &
+${SEA_HOME}/bin/sea_flusher.sh 1 &
 pid=$!
 echo "flusher pid" ${pid}
 
-trap 'trap - SIGTERM && kill $pid' SIGINT SIGTERM EXIT
+#trap 'trap - SIGTERM && kill $pid' SIGINT SIGTERM EXIT
 
 echo "Launching script $1"
 bash $1
 echo "killing flush" && kill $pid
+${SEA_HOME}/bin/sea_flusher.sh 2
+
 

### 2020-07-22 experiments/incrementation/run_bench.sh

@@ -52,7 +52,7 @@ launch_exp () {
 
     if [[ curr_exp != "lustre" ]]
     then
-        grep -i "flushing*" ${all_file} > ${flush_file}
+        grep -Ei "mv|cp|rm" ${all_file} > ${flush_file}
         flushtime=$(grep 'real' ${all_file} | grep "[0-9].*")
         flushtime=$(echo ${flushtime:5})
         ssdwrites=$(grep 'disk' ${flush_file} | sort --unique | wc -l)
@@ -95,6 +95,7 @@ do
         then
             curr_exp="mem_final"
             cp .sea_flushlist_mem ${SEA_HOME}/.sea_flushlist
+            cp .sea_evictlist_mem ${SEA_HOME}/.sea_evictlist
             echo "mem_final ${experiments[0]}"
             launch_exp "$curr_exp" 39
 

### 2020-07-22 bin/sea_flusher.sh
added eviction to in-memory

@@ -9,15 +9,16 @@ sources_arr=()
 n_sources=0
 conf_file=${SEA_HOME}/sea.ini
 flush_file=${SEA_HOME}/.sea_flushlist
+evict_file=${SEA_HOME}/.sea_evictlist
 tmpfile=$(mktemp /tmp/.sea_flush.inprog.XXXXXX)
 
-mv_old () {
+fe_old () {
 
     found_files=$1 # files that currently exist in source mounts
     OLDTIME=$2 # user specified expiration time
-    task=$3 # the name of the calling task (either "process" or "cleanup")
+    task=$3 # mv / cp / rm (depending on which list)
+    flist="" # list containing files for flush/eviction
 
-    flushlist=""
     for f in $found_files
     do
         
@@ -41,30 +42,45 @@ mv_old () {
 
         # Check if file older
         if [ $TIMEDIFF -ge $OLDTIME ]; then
-            echo "flushing and evicting $f to $base_source"
+            echo "$task $f to $base_source"
 
             # some string formatting
-            if [[ "$flushlist" != "" ]]
+            if [[ "$flist" != "" ]]
             then
-                flushlist+=" "
+                flist+=" "
             fi
 
-            flushlist+=$f
+            flist+=$f
            
         fi
     done
 
-    # if we have some files that older that OLDTIME, move them
-    if [[ ${flushlist} != "" ]]
+    # if we have some files that older that OLDTIME, flush/evict them
+    if [[ ${flist} != "" ]]
     then
 
-        # mv each image individually (as opposed to all at once)
-        for f in $flushlist
-        do
-            ([ ! -f $f ] || mv $f ${base_source} || true) &
-        done
+        # process each image individually (as opposed to all at once)
+        if [[ $task == "mv" ]]
+        then 
+            for f in $flist
+            do
+                ([ ! -f $f ] || mv $f ${base_source} || true) &
+            done
+        elif [[ $task == "cp" ]]
+        then 
+            for f in $flist
+            do
+                ([ ! -f $f ] || cp $f ${base_source} || true) &
+            done
+        elif [[ $task == "rm" ]]
+        then 
+            for f in $flist
+            do
+                ([ ! -f $f ] || rm $f || true) &
+            done
+        fi
 
-        # wait for all images to be moved
+        # wait for all images to be processed
         wait
 
[…]

### 2020-07-23 experiments/incrementation/notebook/figures.py
removed some code from notebook and placed it in a script

@@ -1,0 +1,243 @@
+#!/usr/bin/env python3
+
+from bokeh.io import output_notebook, show
+from bokeh.models import ColumnDataSource, HoverTool, Legend, LegendItem, Whisker
+from bokeh.plotting import figure
+from datetime import datetime
+import pandas as pd
+
+
+colours = ["#c9d9d3", "#718dbf", "#e84d60"]
+
+
+def bokeh_gantt(df):
+
+    output_notebook()
+    y_range = sorted(df["pid"].unique())
+    source_read = ColumnDataSource(df[df["action"].str.contains("read")])
+    source_increment = ColumnDataSource(df[df["action"].str.contains("inc")])
+    source_write = ColumnDataSource(df[df["action"].str.contains("save")])
+
+    p = figure(toolbar_location="below", x_range=(0, 130), y_range=y_range)
+    p.hbar(
+        y="pid",
+        left="start",
+        right="end",
+        height=0.4,
+        fill_color="color",
+        source=source_read,
+        legend_label="read",
+    )
+    p.hbar(
+        y="pid",
+        left="start",
+        right="end",
+        height=0.4,
+        fill_color="color",
+        source=source_increment,
+        legend_label="increment",
+    )
+    p.hbar(
+        y="pid",
+        left="start",
+        right="end",
+        height=0.4,
+        fill_color="color",
+        source=source_write,
+        legend_label="write",
+    )
+
+    p.ygrid.grid_line_color = None
+    p.xaxis.axis_label = "Time (seconds)"
+    p.yaxis.axis_label = "PID"
+
+    p.legend.click_policy = "hide"
+
+    p.outline_line_color = None
+    p.add_tools(
+        HoverTool(
+            show_arrow=False,
+            line_policy="next",
+            tooltips=[
+                ("Task", "@action"),
+                ("Image", "@img"),
+                ("Start", "@start"),
+                ("End", "@end"),
+                ("Duration", "@duration"),
+            ],
+        )
+    )
+
+    show(p)
+
+
+def bokeh_stacked(df):
+
+    output_notebook()
+
+    actions = ["read", "increment", "write"]
+    source = ColumnDataSource(df)
+
+    p = figure(
+        x_range=df["type"],
+        toolbar_location="below",
+        tools="hover",
+        tooltips=[("Task", "$name"), ("Duration", "@$name")],
+    )
+
+    p.vbar_stack(
+        actions,
[…]

### 2020-07-23 src/sea.cpp
fixed test case

@@ -67,7 +67,7 @@ int sea_checkpath(const char* path) {
     struct utimbuf times;
     times.actime = time(0);
     times.modtime = buf.st_mtime;    
-    int out = utime(path, &times);
+    utime(path, &times);
     unset_internal();
     return 1;
 }

### 2020-07-23 tests/test_custom.cpp

@@ -122,8 +122,8 @@ TEST(Passthrough, GetRealPath) {
 TEST(Sea, InitializeSea) {
 
     initialize_sea();
-    // 2 mounts * 2 ('.' & '..') + 2 subdirs * 2 mounts * 3 (dirname & '.' & '..') + 3 files
-    ASSERT_EQ(sea_files.size(), 19);
+    // 2 mounts * 2 ('.' & '..') + 2 subdirs * 2 mounts * 3 (dirname & '.' & '..') + 5 files
+    ASSERT_EQ(sea_files.size(), 21);
 
     initialize_passthrough_if_necessary();
     config sea_config = get_sea_config();
@@ -135,8 +135,8 @@ TEST(Sea, InitializeSea) {
     ((funcptr_mkdir)libc_mkdir)(added_dir, 0666);
     sea_files.clear();
     initialize_sea();
-    // 19 from before + 6 ((dirname, ., ..)*2)
-    ASSERT_EQ(sea_files.size(), 25);
+    // 21 from before + 6 ((dirname, ., ..)*2)
+    ASSERT_EQ(sea_files.size(), 27);
 
 
 }

### 2020-07-23 paper/paper.tex
some comments on the paper
changes to flusher and launch script and minor updates to paper

@@
 \begin{document}
-\title{Sea: A hierarchical filesystem for processing Big Neuroimaging Data on HPC}
+\title{Sea: A hierarchical filesystem for Big-Data Scientific Computing}
+% remove neuroimaging and replace with scientific computing  add user-space, switch HPC to computing clusters
 
@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-\IEEEPARstart{N}{euroimaging} datasets have significantly grown in size in the past decade,
+\IEEEPARstart{S}{cientific} computing comes with a unique set of constraints. Such
+applications may be composed of heterogeneous workflows, with tasks individual tasks
+varying from data to compute intensive, and are typically executed on
+shared High Performance Computing (HPC) infrastructure. Scientific applications may consist
+of a series of well-established tools. These tools may have initially been developed at
+a time when Big Data played a less important role in Scientific computing, and are now
+paying a penalty as the amount of data has vastly increased in size. Significant effort
+
+datasets have significantly grown in size in the past decade,
 resulting in Petabytes of data. As a result, processing pipelines have now shifted from compute to 
@@
 
+% add need for userspace / intelligent pipeline flushing
+% there is no one-size fits all with data intensive applications
+% due to io patterns, so we decided with userspace.
+
 Goal - leverage local storage to speed up computation/Sea
 
+% add related work section
 \subsection{Lustre}
-\subsubsection{Burst Buffers}
+% change lustre to shared filesystem
 \subsubsection{Other Luster optimizations}
@@
 \subsubsection{Kernel-space file systems}
+\subsubsection{Burst Buffers}
+\subsubsection{Unionfs}
 \subsubsection{FUSE}
@@
 \subsubsection{LD\_PRELOAD}
+xtreemfs/xtreemos
 

### 2020-07-23 bin/sea_launch.sh

@@ -16,7 +16,7 @@ do
     mkdir -p ${source} 
 done
 
-${SEA_HOME}/bin/sea_flusher.sh 1 &
+${SEA_HOME}/bin/sea_flusher.sh &
 pid=$!
 echo "flusher pid" ${pid}
 
@@ -25,6 +25,5 @@ echo "flusher pid" ${pid}
 echo "Launching script $1"
 bash $1
 echo "killing flush" && kill $pid
-${SEA_HOME}/bin/sea_flusher.sh 2
 
 

### 2020-07-23 bin/sea_flusher.sh
changes to flusher and launch script and minor updates to paper

@@ -21,18 +21,12 @@ fe_old () {
 
     for f in $found_files
     do
-        
-        # check if file exists otherwise continue (in case it has been flushed by another process)
-        if [ ! -f $f ]
-        then
-            continue
-        fi
         # code adapted from here https://stackoverflow.com/a/28341234
         # Get current and file times
         CURTIME=$(date +%s)
         FILETIME=$(stat $f -c %Y || true && continue)
 
-        if [ $FILETIME == "" ]
+        if [[ $FILETIME == "" ]]
         then
             continue
         fi
@@ -62,30 +56,17 @@ fe_old () {
         # process each image individually (as opposed to all at once)
         if [[ $task == "mv" ]]
         then 
-            for f in $flist
-            do
-                ([ ! -f $f ] || mv $f ${base_source} || true) &
-            done
+            mv $flist ${base_source} || true
         elif [[ $task == "cp" ]]
         then 
-            for f in $flist
-            do
-                ([ ! -f $f ] || cp $f ${base_source} || true) &
-            done
+            cp $f ${base_source} || true
         elif [[ $task == "rm" ]]
         then 
-            for f in $flist
-            do
-                ([ ! -f $f ] || rm $f || true) &
-            done
+            rm $f || true
         fi
 
         # wait for all images to be processed
-        wait
-
-        #(mv $flushlist ${base_source} || true) &
-        #pid=$!
-        #wait $pid # wait until subprocess completes
+        #wait
     fi
 }
 
@@ -268,6 +249,8 @@ flush_process () {
         wait $!
     done
     echo "flush process terminated"
+    echo "starting cleanup"
+    time flush 0 "cleanup"
 }
 
 cleanup () {
@@ -305,11 +288,5 @@ get_sources () {
 
 trap '[ -f $tmpfile ] && rm $tmpfile' SIGTERM SIGINT EXIT
 
-if [[ $1 == 1 ]]
-then
-    get_sources
-    flush_process 5
-else
-    get_sources
-    cleanup
-fi
+get_sources
+flush_process 5

### 2020-08-27 paper/paper.tex
redid intro added contents to methods section

@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-\IEEEPARstart{S}{cientific} computing comes with a unique set of constraints. Such
-applications may be composed of heterogeneous workflows, with tasks individual tasks
-varying from data to compute intensive, and are typically executed on
-shared High Performance Computing (HPC) infrastructure. Scientific applications may consist
-of a series of well-established tools. These tools may have initially been developed at
-a time when Big Data played a less important role in Scientific computing, and are now
-paying a penalty as the amount of data has vastly increased in size. Significant effort
+\IEEEPARstart{E}{ffecient} data-management strategies have become imperative in 
+mitigating the costs of processing large datasets. These strategies come in the form
+of data locality, which ensures that computation is performed where the data resides
+rather than relocating the data, and in-memory computing, which ensures that the data is
+maintained in RAM, whenever possible. Big Data frameworks, such as the Hadoop MapReduce, Spark and Dask,
+have evolved to support either or both of these strategies.
+However, while many domains have adopted the use of Big Data framework and filesystems to ensure effective data
+management within their applications, many scientific applications are developed without such strategies in mind.
+% add speedup results
 
-datasets have significantly grown in size in the past decade,
-resulting in Petabytes of data. As a result, processing pipelines have now shifted from compute to 
-data intensive. Some efforts have been made to reduce the data transfer costs in neuroimaging by
-introducing new methods to generate pipelines, such as through the adoption of Big Data frameworks
-like Apache Spark \cite{Spark} for the pipeline definition \cite{thunder} \cite{other papers}. 
-However, such pipelines would have to be entirely rewritten for the framework of choice and Big Data
-frameworks like Spark and Dask were found to be difficult to use in describing neuroimaging pipelines \cite{mehta}.
-Furthermore, the neuroimaging community relies heavily on the use of well-established command-line tools \cite{fsl, afni, freesurfer}, resulting in
-data transfers being inevitable, even with a Big Data framework.
+Scientific applications may omit leveraging Big Data frameworks, despite having Big Data workloads,
+for many reasons. One such reason may be the data format. BigData frameworks were originally developed
+for text-processing. Scientific applications, in contrast, may rely on image processing for their
+analyses. The fundamental differences between these two data formats made leveraging BigData frameworks for
+scientific image processing non-trivial. While recent efforts have been made to support data-intensive
+image processing workloads, there remains some work to be done.
 
-High Performance Computing (HPC) clusters are often the infrastructure of choice for researchers looking to processing their data.
-While there are many benefits to using such clusters, such as large amounts of resources available at little to no cost to
-the researchers, the resources are shared amongst users concurrently.
+Scientific applications are usually composed of numerous well-established command-line tools. These tools
+may have been developed at a time when the processing would have been compute intensive, thus not requiring any
+data management strategies, yet remain scientifically relevant. While newer applications
+that leverage such tools may be written using Big Data frameworks, such frameworks are not designed to transfer
+data in-memory to command-line applications, and thus would require some instrumentation.
+
+While various infrastructures are available to researchers, including local workstations, cloud and 
+High Performance Computing (HPC) clusters, researchers commonly rely on HPC clusters due to their low cost and available
+resources. Such clusters typically rely on a network-based parallel file system, such as Lustre, Ceph and GlusterFS.
+Since data storage and compute are maintained as separate entities on HPC clusters, processing BigData workflows on them
+may be particularly costly, particularly when no BigData framework is applied. Furthermore, while it is possible to execute
+workloads using BigData frameworks on HPC, these frameworks are typically incompatible with the native resource managers (e.g. SLURM, PBS),
+and therefore, and overlay cluster must be used.
+
+In this paper we introduce Sea, a user-space hierarchical file system for scientific workloads. The aim of Sea
+is to bring BigData performance to scientific workloads through the incorporation of data locality and
+in-memory computing. Unlike Big Data frameworks, Sea can be utilized alongside existing workloads without
+the need to entirely reinstrument the existing workload. Furthermore, as it executes entirely in userspace,
+it does not require elevated privileges to be installed and executed.
 
@@
 
-Goal - leverage local storage to speed up computation/Sea
-
 % add related work section
+\section{Related Work}
 \subsection{Lustre}
@@
 \subsection{Sea design and implementation}
+\subsubsection{Overview}
+
+Sea is a C++ library that redirects files written to a user-specified mount point, to the
+appropriate storage devices. It achieves this through the interception of libc calls using
+the LD\_PRELOAD functionality. Within an intercepted call the file will be written to or read
+from the best available storage device. At minimum, a user must specify at least two storage devices, a fast
+temporary one and a slower long-term storage one. This could be RAM and SSD if working on a single node, or a compute-local SSD
+and a shared parallel file system, in the case of an HPC cluster. Ideally, a user will provide a multitude of short-term storage
+devices to improve Sea's efficacy. Furthermore, to maximize usage of the fastest available
+storage devices, Sea will allow the user to outline which files can be removed from short-term storage in addition
+to which files need to be materialized onto long-term storage.
+
+In this subsection, we will discuss and outline the various design and implementation details made. Furthermore,
+we will go into detail into the functionality of Sea and how it should and should not be used.
+
 \subsubsection{Requirements and Assumptions}
+
+Sea is primarily designed to work with workloads that generate copious amount of intermediate data.
+Since data must be read from the slower storage device and final output data most likely would have to
+be written to the same device, the use of Sea with workloads that do not produce intermediate data would
+bring no speedup and may introduce some overheads.
+
+Sea provides two main modes, that are dictated by how the user specifies flushing and eviction: In-memory computing and flush-all.
+In-memory computing is achieved when intermediate need not be materialized to long-term storage. In these scenarios, Sea can fully
+leverage local (short-term) storage performance to maximize performance. This can be coupled with eviction to limit the amount of
+data that will be written to slower, but larger local storage devices. With Sea in memory, we can expect performance comparable to
+that of using BigData frameworks.
+
+Occasionally, it is desired to materialize even intermediate data to long-term storage devices, for potential future use. The
+performance gain from using Sea in these scenarios, however, is largely bounded by data size. In the instance where the workloads is more compute intensive,
+we can gain speedup by masking all or part of the I/O through long-term storage by executing the flush during a compute task. In data-intensive scenarios, the 
+opposite can be seen: compute time can be masked by I/O (add figure + model). Despite the performance gain being limited in
+such a scenario, a performance gain can still be observed through the use of Sea.
+
+One important assumption that must be made is that the amount of data produced by the workload far exceeds the amount of page cache space available and utilized
+by the different filesystems. In the case where all workload data can fit into page cache memory, the benefit of Sea may be
+greatly diminished if not negligible.
+
+In order to redirect data to different filesystems, Sea uses LD\_PRELOAD to intercept libc calls. Depending on the
+version of libc used, function calls may differ. Sea is currently compatible with libc version X-X.
+
 \subsubsection{Benefits}
-User does not need to create logic to determine what file systems to write to / implement flushing
+Sea alleviates the user of the need to implement the logic to redirect data to the appropriate
+storage locations. Furthermore, it implements logic to ensure that files are written to the best possible
+location at any given time. It is both pipeline and infrastructure agnostic.
+
+As it executes in user-space, root permissions are not required, enabling users to use Sea on most linux-based systems
+available to them. This differs from Big Data filesystems which may require elevated permissions to install.
+The overhead of intercepting libc calls
+is minimal, and negligible compared to system call interception and file systems such as FUSE.
+
+Unlike BigData frameworks, Sea does not require reinstrumentation of the existing pipelines, allowing
+users to gain an instant performance boost.
 
 \subsubsection{libc interception}
-\subsubsection{Flushing}
+
+libc interception is achieved by writing wrappers to existing libc functions. Most importantly, every
+libc function accepting a file path needs to be wrapped. The wrappers' job is to take any input filepath
+that is located within the user-provided Sea mountpoint and convert it a filepath pointing to the best available storage device.
+
+\subsubsection{Flushing and eviction}
+There are four different available usage scenarios: flush, evict, flush and evict and do nothing.
+A flush operation represents a copy operation that duplicates the data located on short-term storage to long-term storage.
+In contrast, an evict represents a rm. That is 
 \subsubsection{Configuration File}
@@
 \subsubsection{Limitations}
+behaviour unspecified in mixed static/dynamic workloads
 

### 2020-08-29 experiments/incrementation/parse_json.sh
modifications to experiment launch script

@@ -1,0 +1,163 @@
+#!/bin/bash
+
+cond_file=$1
+
+out_fldr="/mnt/lustre/vhs/output"
+sea_mount="/mnt/lustre/vhs/mount"
+rdir="./results"
+library="/home/vhs/Sea/build/sea.so"
+executable="python increment.py"
+
+mkdir -p ${rdir}
+ncond=$(jq ". | length" ${cond_file})
+
+
+main () {
+    echo "Preparing experiment launch"
+
+    for ((i=0;i<ncond;i++))
+    do
+        name=$(jq -r ".[$i] .name" ${cond_file})
+        ndisks=$(jq ".[$i] .ndisks" ${cond_file})
+        nnodes=$(jq ".[$i] .nnodes" ${cond_file})
+        nthreads=$(jq ".[$i] .nthreads" ${cond_file})
+        nfiles=$(jq ".[$i] .nfiles" ${cond_file})
+        in_fldr=$(jq -r ".[$i] .in_fldr" ${cond_file})
+        niterations=$(jq ".[$i] .niterations" ${cond_file})
+        strategy=$(jq -r ".[$i] .strategy" ${cond_file})
+
+        echo "Experiment 1 Run 0 conditions
+name: $name
+strategy: $strategy
+file input folder: ${in_fldr}
+number of disks: $ndisks
+number of nodes: $nnodes
+number of threads: $nthreads
+number of files: $nfiles
+number of iterations: ${niterations}"
+        exp_fldr="${rdir}/${name}"
+        rm -rf $exp_fldr/*
+        echo -e "Experiment directory: ${exp_fldr}\n"
+
+        fnode=$(($nfiles / $nnodes))
+        rem=$(($nfiles % $nnodes))
+
+        if [[ $strategy == "lustre" ]]
+        then
+            prefix=$executable
+            out_mount=$out_fldr
+
+        else
+            prefix="LD_PRELOAD=$library $executable"
+            out_mount=$sea_mount
+        fi
+
+        fcount=0
+        node=0
+        nprev=-1
+        echo "Files to be executed per node: $fnode  remainder: $rem"
+
+        job_dir="${exp_fldr}/jobs"
+        mkdir -p ${job_dir}
+
+        sl_script="${exp_fldr}/sbatch_${strategy}.sh"
+
+        echo '#!/bin/bash
+#SBATCH --time=90:55:00
+#SBATCH --account=vhs
[…]

### 2020-08-29 experiments/incrementation/parse_json.sh
added modifying sea config

@@ -2,14 +2,68 @@
 
 cond_file=$1
 
+sea_home=/home/vhs/Sea
 out_fldr="/mnt/lustre/vhs/output"
 sea_mount="/mnt/lustre/vhs/mount"
 rdir="./results"
 library="/home/vhs/Sea/build/sea.so"
 executable="python increment.py"
+alldisks=("/disk0/vhs/seatmp" "/disk1/vhs/seatmp" "/disk2/vhs/seatmp" "/disk3/vhs/seatmp" "/disk4/vhs/seatmp" "/disk5/vhs/seatmp")
 
 mkdir -p ${rdir}
 ncond=$(jq ". | length" ${cond_file})
+
+
+format_config () {
+    ndisks=$1
+    nthreads=$2
+    fsize=$3
+    exp_fldr=$4
+    strategy=$5
+    niterations=$6
+
+    conf=${exp_fldr}/sea.ini
+    flushlist=${exp_fldr}/.sea_flushlist
+    evictlist=${exp_fldr}/.sea_evictlist
+
+    disks=""
+
+    for ((x=0;x<ndisks;x++))
+    do
+        if [[ $x -gt 0 ]]
+        then
+            disks+=","
+        fi
+        disks+="${alldisks[$x]}"
+    done
+    echo '# Sea configuration                                                      
+[Sea]                                                                    
+mount_dir = '"${sea_mount}"' ;           
+n_levels = 3 ;                                                          
+source_0 = /dev/shm/seatmp ;           
+source_1 = '"$disks"' ;
+source_2 = '"$out_fldr"' ;
+log_level = 0 ; # 4 crashes tests                                        
+log_file = /home/vhs/passthrough/sea.log ;
+max_fs = '"$fsize"' ;
+n_threads = '"$nthreads"' ;' > $conf
+
+    cp $conf ${sea_home}
+
+    final_it=$(($niterations-1))
+
+    if [[ $strategy == "cp" ]]
+    then
+        echo "*" > $flushlist
+    elif [[ $strategy == "mem" ]]
+    then
+        echo "${final_it}*.nii" > $flushlist
+    fi
+    echo "${final_it}*.nii" > $evictlist
+
+    cp $flushlist $evictlist ${sea_home}
+
+}
 
 
 main () {
@@ -24,6 +78,7 @@ main () {
[…]

### 2020-08-29 experiments/incrementation/parse_json.sh
added experiment execution functionality

@@ -1,6 +1,7 @@
 #!/bin/bash
 
 cond_file=$1
+results_file=$2
 
 sea_home=/home/vhs/Sea
 out_fldr="/mnt/lustre/vhs/output"
@@ -8,11 +9,14 @@ sea_mount="/mnt/lustre/vhs/mount"
 rdir="./results"
 library="/home/vhs/Sea/build/sea.so"
 executable="python increment.py"
+flush_out=flushlist.out
+bench_out=benchmarks.out
 alldisks=("/disk0/vhs/seatmp" "/disk1/vhs/seatmp" "/disk2/vhs/seatmp" "/disk3/vhs/seatmp" "/disk4/vhs/seatmp" "/disk5/vhs/seatmp")
 
 mkdir -p ${rdir}
 ncond=$(jq ". | length" ${cond_file})
 
+echo "experiment,repetition,runtime,max_mem,flush_time,disk_files,total_flush" > ${results_file}
 
 format_config () {
     ndisks=$1
@@ -65,8 +69,54 @@ n_threads = '"$nthreads"' ;' > $conf
 
 }
 
+launch_exp () {
+    name=$1
+    exp_dir=$2
+    sfile=$3
+    run=$4
+
+    # launch experiment and obtain job id
+    slurm_id=$(sbatch ${sfile} | grep -Eo "[0-9]*")
+    echo "Launched job ${name} with id ${slurm_id}"
+
+    # wait until experiment completes
+    running=$(squeue --job ${slurm_id} | wc -l)
+    while [[ ${running} == 2 ]]
+    do
+        sleep 10
+        running=$(squeue --job ${slurm_id} | wc -l)
+    done 
+
+    echo "Slurm job complete"
+
+    all_out="slurm-$name-${slurm_id}.out"
+    all_file="${exp_dir}/${all_out}"
+    flush_file="${exp_dir}/${flush_out}"
+    bench_file="${exp_dir}/${bench_out}"
+
+    grep -E "start,|end," ${all_file} > ${bench_file}
+
+    res=$(grep -i runtime ${all_file}| grep -Eo "[0-9]+\.[0-9]+" )
+    flushtime=0
+    sddwrites=0
+    totalflush=0
+
+    if [[ $name != "lustre" ]]
+    then
+        grep -Ei "mv|cp|rm" ${all_file} > ${flush_file}
+        flushtime=$(grep 'real' ${all_file} | grep "[0-9].*")
+        flushtime=$(echo ${flushtime:5})
[…]

### 2020-08-30 bin/sea_flusher.sh
fixes to json experiment launch

@@ -140,7 +140,6 @@ flush () {
             all_files+=$(find ${rgx} -type f -follow 2> /dev/null || true)
         done
 
-
         # if .sea_flushlist file contains regex
         for rgx in $re_flush
         do
@@ -256,9 +255,6 @@ flush_process () {
 
 cleanup () {
     echo "Cleaning up Sea mount"
-    #sp=$(jobs -p)
-    #[[ $sp != "" ]] && kill -INT $sp
-    #ps aux | grep vhs
     time flush 0 "cleanup"
 }
 

### 2020-08-30 bin/sea_launch.sh

@@ -26,4 +26,5 @@ echo "Launching script $1"
 bash $1
 echo "killing flush" && kill $pid
 
+wait $pid
 

### 2020-08-30 experiments/incrementation/parse_json.sh
fixes to json experiment launch

@@ -2,6 +2,7 @@
 
 cond_file=$1
 results_file=$2
+runs=$3
 
 sea_home=/home/vhs/Sea
 out_fldr="/mnt/lustre/vhs/output"
@@ -77,7 +78,7 @@ launch_exp () {
 
     # launch experiment and obtain job id
     slurm_id=$(sbatch ${sfile} | grep -Eo "[0-9]*")
-    echo "Launched job ${name} with id ${slurm_id}"
+    echo "Launched job ${name} run ${run} with id ${slurm_id}"
 
     # wait until experiment completes
     running=$(squeue --job ${slurm_id} | wc -l)
@@ -103,7 +104,7 @@ launch_exp () {
 
     if [[ $name != "lustre" ]]
     then
-        grep -Ei "mv|cp|rm" ${all_file} > ${flush_file}
+        grep -Ei "mv |cp |rm " ${all_file} > ${flush_file}
         flushtime=$(grep 'real' ${all_file} | grep "[0-9].*")
         flushtime=$(echo ${flushtime:5})
         ssdwrites=$(grep 'disk' ${flush_file} | sort --unique | wc -l)
@@ -116,10 +117,12 @@ launch_exp () {
 main () {
 
     run=$1
+    declare -a EXP_ORDER=( $(for((i=0;i<ncond;i++)); do echo $i; done;) )
 
     echo "Preparing experiment launch"
 
-    for ((i=0;i<ncond;i++))
+    EXP_ORDER=( $(shuf -e "${EXP_ORDER[@]}") )
+    for i in "${EXP_ORDER[@]}"
     do
         name=$(jq -r ".[$i] .name" ${cond_file})
         ndisks=$(jq ".[$i] .ndisks" ${cond_file})
@@ -141,7 +144,7 @@ number of threads: $nthreads
 number of files: $nfiles
 file size in bytes: $fsize
 number of iterations: ${niterations}"
-        exp_fldr="${rdir}/${name}"
+        exp_fldr="${rdir}/run-${run}/${name}"
         rm -rf $exp_fldr/*
         echo -e "Experiment directory: ${exp_fldr}\n"
 
@@ -237,7 +240,7 @@ srun -N'"$nnodes"' mkdir /disk0/vhs/seatmp /disk1/vhs/seatmp /disk2/vhs/seatmp /
 
                 if [[ ${strategy} == "lustre" ]]
                 then
[…]

### 2020-08-30 experiments/incrementation/run_bench.sh

@@ -21,6 +21,13 @@ time_params="%e,%K"
 curr_exp=""
 
 
+create_template() {
+    nnodes=$1
+    ncpus=$2
+
+}
+
+
 launch_exp () {
     curr_exp=$1
     num_files=$2
@@ -97,8 +104,9 @@ do
         then
             curr_exp="mem_all"
             cp .sea_flushlist_all ${SEA_HOME}/.sea_flushlist
+            echo "" > ${SEA_HOME}/.sea_evictlist
             echo "mem_all ${experiments[0]}" 
-            launch_exp "$curr_exp" 195
+            launch_exp "$curr_exp" 390
 
         elif [[ $e == 1 ]]
         then
@@ -112,7 +120,7 @@ do
         then
             curr_exp="lustre"
             echo "lustre ${experiments[1]}"
-            launch_exp "$curr_exp" 195
+            launch_exp "$curr_exp" 390
 
         else
             curr_exp="disk"

### 2021-06-16 paper/paper.tex
added paper updates

@@
 \documentclass[10pt,journal,compsoc]{IEEEtran}
+\usepackage{xcolor}
+\usepackage{hyperref}
+\usepackage{ulem}
+\usepackage{graphicx}
+\usepackage{rotating}
+\usepackage{setspace}
+\usepackage{longtable}
+\usepackage[binary-units=true]{siunitx}
+\usepackage{multirow}
+\usepackage{subcaption}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{amsmath}
+
+\algnewcommand\algorithmicforeach{\textbf{for each}}
+\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
+\algblock{Input}{EndInput}
+\algnotext{EndInput}
+\algblock{Output}{EndOutput}
+\algnotext{EndOutput}
+\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
+\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}: #1}}}
+
+% benchmark on beluga, prefreesurfer with RUIS
+% use more than 64 subjects for prefreesurfer --> few 100 subjects
+% think more about how to benchmark I/O
+% try distributed jobs with Sea
 
@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-\IEEEPARstart{E}{ffecient} data-management strategies have become imperative in 
-mitigating the costs of processing large datasets. These strategies come in the form
+\IEEEPARstart{E}{fficient} data-management strategies have become essential in 
+mitigating the costs of Big Data processing. These strategies come in the form
 of data locality, which ensures that computation is performed where the data resides
 rather than relocating the data, and in-memory computing, which ensures that the data is
-maintained in RAM, whenever possible. Big Data frameworks, such as the Hadoop MapReduce, Spark and Dask,
-have evolved to support either or both of these strategies.
-However, while many domains have adopted the use of Big Data framework and filesystems to ensure effective data
-management within their applications, many scientific applications are developed without such strategies in mind.
+maintained in RAM, whenever possible. These strategies are built-into popular Big Data Frameworks
+such as Hadoop MapReduce~\cite{mapreduce}, Spark~\cite{spark} and Dask~\cite{dask}.
+Despite recent widespread adoption of Big Data frameworks and filesystems to ensure effective data
+management within applications, many widely-used scientific applications forgo such optimization for various reasons.
+Sea aims to enable Big Data optimization performance to existing scientific applications without reinstrumentation.
 % add speedup results
 
-Scientific applications may omit leveraging Big Data frameworks, despite having Big Data workloads,
-for many reasons. One such reason may be the data format. BigData frameworks were originally developed
-for text-processing. Scientific applications, in contrast, may rely on image processing for their
-analyses. The fundamental differences between these two data formats made leveraging BigData frameworks for
-scientific image processing non-trivial. While recent efforts have been made to support data-intensive
-image processing workloads, there remains some work to be done.
+Many scientific applications, such as those found in geography and neuroscience, perform
+image processing. However,  BigData frameworks were originally developed
+for text-processing and have only recently incorporated components that could facilitate
+the processing of images. The fundamental differences between these two data formats make leveraging
+BigData frameworks for
+scientific image processing non-trivial, even considering recent advances.
 
 Scientific applications are usually composed of numerous well-established command-line tools. These tools
-may have been developed at a time when the processing would have been compute intensive, thus not requiring any
-data management strategies, yet remain scientifically relevant. While newer applications
+may have been developed at a time when processing would have been compute intensive, thus do not incorporate
+data management strategies. While newer applications
 that leverage such tools may be written using Big Data frameworks, such frameworks are not designed to transfer
-data in-memory to command-line applications, and thus would require some instrumentation.
+data in-memory to command-line applications, making it difficult to incorporate these tools without additional
+overhead.
 
-While various infrastructures are available to researchers, including local workstations, cloud and 
+Although various infrastructures are available to researchers, such as local workstations, the cloud (AWS, ..etc) and 
 High Performance Computing (HPC) clusters, researchers commonly rely on HPC clusters due to their low cost and available
@@
 
+%multiple disk support
+% pipeline agnostic and pipeline access patterns
+% don't leverage workflows
+% leverage existing filesystems
+% replace storage devices with storage location.
+% use applications
+% memory management paragraph: PAGE CACHE
+% shared filesystems are slow and metadata management can be slow
+
 % add need for userspace / intelligent pipeline flushing
@@
 \section{Related Work}
-\subsection{Lustre}
-% change lustre to shared filesystem
-\subsubsection{Other Luster optimizations}
+\subsection{HPC Infrastructure}
+      The general structure of HPC clusters may complicate the deployment of Big Data frameworks.
+      Typical HPC clusters consist of distinct storage and compute nodes. While the compute
+      nodes may also have local storage, there is no distributed file system like HDFS or
+      Alluxio to facilitate data locality. Furthermore, access to these compute nodes is
+      temporary, with allocation duration enforced by a batch scheduler. Therefore, any
+      data written to a compute node is inaccessible after the allocation is terminated.
+
+      The storage layer found on HPC clusters consists of a high-performance parallel
+      distributed file system such as Lustre~\cite{lustre}. This file system is made
+      accessible to every compute node within the cluster, connected to each node via
+      high-performance network interconnect such as InfiniBand~\cite{infiniband}. Although
+      it is a high performance network, it nevertheless can be a source of bottlenecks as
+      the network is shared between all users on the cluster. Furthermore, there are
+      additional overheads that can be incurred. In the case of Lustre,
+      there are many data nodes, known as Object Storage Servers (OSS) which contain
+      several storage devices, known as Object Storage Targets (OST). All file metadata
+      is maintained within a separate node known as the Metadata Server (MDS) and stored within
+      a device referred to as the Metadata Target (MDT). Although data transfers can be
+      communicated directly to the corresponding OST, the clients need to first communicate
+      with the MDS to determine which OST to communicate with. This can result
+      in data transfer overheads, particularly when making numerous requests to the metadata server.
+      Occasionally, HPC clusters provide a faster intermediate storage layer, known as a Burst Buffer,
+      which can improve the overall performance of I/O operations to Lustre.
+
+      \subsubsection{Burst Buffers}
+
 \subsection{Big Data Frameworks and File systems}
-HDFS Alluxio Hadoop Spark Dask
-\subsubsection{What are they/how do they work}
-\subsubsection{Why are the suboptimal for neuroimaging/on HPC}
+
+      Strategies to facilitate the processing of such large datasets is required. In text-based
+      processing, frameworks such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}
+      have implemented and popularized strategies to minimize the need for data transfers during processing.
+      These strategies are known as data locality and in-memory computing.
+
+      Data locality is the process by which compute tasks are scheduled nearest to where
+      the data is located. Traditionally, storage and compute were kept separate,
+      having to transfer data over the network to compute. Rather than having to transfer large amounts of data over the
+      network to the compute tasks, which could incur significant overheads, Big Data frameworks ensure that data is
+      stored directly on the compute nodes. When a compute task requires access to specific
+      data, the scheduler sends the task to the nearest available node to the data, thereby
+      minimizing any cost of network-related data transfers. This strategy is not only
+      used by Big Data Frameworks such as Hadoop MapReduce, Apache Spark, and Dask~\cite{dask},
+      but also enabled by file systems such as the Hadoop Distributed File System (HDFS)~\cite{hdfs}
+      and Alluxio~\cite{alluxio}.
+
 \subsection{The linux page cache}
-\subsubsection{What is it/how does it work}
-\subsubsection{Why do we still need Sea}
+ While scientific applications uncommonly use Big Data frameworks and must perform
+      large over-the-network data transfers, they may still benefit from in-memory
+      computing and data locality through the Linux Page Cache~\cite{pagecache}. Similarly to other
+      filesystems, Lustre leverages the client node page cache to reduce I/O overheads.
+      System memory is composed of two components: 1) anonymous memory and 2) page cache.
+      Anonymous memory consists of all application-related objects, whereas page cache
+      consists of recently accessed file data. When a file is read, that file is loaded
+      up into the page cache to be flagged for eviction based on a least recently used (LRU)
+      policy. That means subsequent accesses to that data may be done entirely
+      in memory so long as that data has not already been evicted. Similarly, for writes
+      to a file system with writeback cache enable, the file will be written to memory
+      completing the write operation once the file has been written entirely to memory.
+      That file will then be flushed asynchronously to the appropriate storage device.
+      As system memory may get overloaded with too many write requests, there is a limit to the amount of written data that
+      can exist in memory, known as the \texttt{dirtyi\_ratio}. Furthermore, applications
+      producing too many write requests may be throttled by the system.
 \subsection{File System implementations}
-\subsubsection{Kernel-space file systems}
-\subsubsection{Burst Buffers}
+      Due to the architecture of conventional HPC clusters, network based parallel
+      files systems are favoured over Big Data distributed filesystems. Such filesystems
+      require super user access, preventing users from deploying their own own on-the-fly cluster.
+      While HDFS can be loaded in user-space by mounting its File System in Userspace (FUSE) implementation,
+      FUSE-based filesystems can perform significantly worse than desired, depending on the application~\cite{tofuse}.
+      Even with a userspace version of HDFS, there would
+      be no mechanism to transfer all the data back to the parallel filesystem to be accessible
+      post resource requests. Furthermore, such filesystems are not POSIX-compliant, which
+      would be required for scientific computing applications.
+
+      There are, however, BigData filesystems, such as XtreemFS~\cite{xtreemfs} that
+      exist entirely within user space, are POSIX-compliant, and can use alternative
+      methods to FUSE to function. XtreemFS leverages libc interception using the LD\_PRELOAD
+      trick to intercept filesystem calls made to the libc library. There are
+      limitations to using libc interception, such as there needs to be libc call to intercept and the
+      application making the calls must be dynamically linked.
+      There are alternatives to libc intercept that can bypass these issues, such as
+      system call interception, but they have greater overheads as compared to libc interception~\cite{quinson}.
+      Furthermore, most applications are dynamically linked and use libc to interact
+      with the filesystem. Characteristics that make XtreemFS unlikely to be deployed
+      on-the-fly include complex filesytem deployment and configuration properties. It
+      also does not provide a simple way to leverage different classes of storage
+      devices, nor ensure that required data will be copied to the cluster's parallel file system.
+      %include triple-H
+%\subsubsection{Kernel-space file systems} add comments in methods
 \subsubsection{Unionfs}
 \subsubsection{FUSE}
+% add number w/o plot
 \subsubsection{System call interception with ptrace}
@@
 \subsubsection{Overview}
-
-Sea is a C++ library that redirects files written to a user-specified mount point, to the
-appropriate storage devices. It achieves this through the interception of libc calls using
-the LD\_PRELOAD functionality. Within an intercepted call the file will be written to or read
+\todo{mention that it's unix compatible}
+Sea is a Linux-compatible C++ library that redirects files accessed from a user-specified mount point to the
+appropriate storage devices. In order to effectively redirect all data transfers, it leverages the LIBC\_PRELOAD functionality which
+enables the interception of libc calls. Within an intercepted call the file will be written to or read
 from the best available storage device. At minimum, a user must specify at least two storage devices, a fast
@@
 be written to the same device, the use of Sea with workloads that do not produce intermediate data would
-bring no speedup and may introduce some overheads.
+bring no speedup and may even introduce some overheads.
 
-Sea provides two main modes, that are dictated by how the user specifies flushing and eviction: In-memory computing and flush-all.
+Sea provides two main modes that are dictated by how the user specifies flushing and eviction: In-memory computing and flush-all.
 In-memory computing is achieved when intermediate need not be materialized to long-term storage. In these scenarios, Sea can fully
-leverage local (short-term) storage performance to maximize performance. This can be coupled with eviction to limit the amount of
+leverage local (short-term) storage to maximize performance. This can be coupled with eviction to limit the amount of
 data that will be written to slower, but larger local storage devices. With Sea in memory, we can expect performance comparable to
@@
 Occasionally, it is desired to materialize even intermediate data to long-term storage devices, for potential future use. The
-performance gain from using Sea in these scenarios, however, is largely bounded by data size. In the instance where the workloads is more compute intensive,
+performance gain from using Sea in these scenarios, however, is largely bounded by data size. In the instances where the workloads is more compute intensive,
 we can gain speedup by masking all or part of the I/O through long-term storage by executing the flush during a compute task. In data-intensive scenarios, the 
@@
 
-In order to redirect data to different filesystems, Sea uses LD\_PRELOAD to intercept libc calls. Depending on the
-version of libc used, function calls may differ. Sea is currently compatible with libc version X-X.
+Depending on the version of libc used, function calls may differ. Sea is currently compatible with libc version X-X.
+\todo{depending on the version of libc, the application static/dynamic linking, version of the function}
 
@@
 
+Sea currently does not support the partitioning of files across multiple devices. Since it cannot predict the size of the outputs to ensure the
+existence of sufficient space on storage devices, the user must provide within the Sea configuration file the maximum file size produced by the workflow.
+Together with the specified amount of parallel processes, Sea calculates the minimum space required on a storage device to write the file to it \todo{should include an equation here}.
+Sea will then go through the hierarchy of available storage devices an select the fastest storage device with sufficient available space.
+
+While this process may result in not utilizing all available space in fast storage devices, it will still allow users to gain a performance speedup on the pipeline given that the
+number of threads multiplied by the file size does not exceed storage space.
+
+\todo{minimum storage size calculations is thread-safe}
+\todo{talk about testing}
+
 \subsubsection{Flushing and eviction}
+\todo{add diagrams here?}
 There are four different available usage scenarios: flush, evict, flush and evict and do nothing.
-A flush operation represents a copy operation that duplicates the data located on short-term storage to long-term storage.
-In contrast, an evict represents a rm. That is 
+A Sea flush is synonymous to a file copy in that files that are stored within Sea but have not yet been materialized
+to the long-term storage device will be copied to long-term storage. Flushing generally required for all intermediate data
+that are required for future analysis.
+
+Eviction, on the other hand, depicts a remove operation. Any file that is evicted from Sea will be removed from any of the storage mount
+points located in Sea.
+
 \subsubsection{Configuration File}
@@
 
+\subsection{The Sea and Lustre model}\label{model}
+
+      \begin{table}
+      \centering
+      \begin{tabular}{|p{0.05\linewidth}|p{0.6\linewidth}|p{0.1\linewidth}|}
+       \hline
+       \multicolumn{3}{|c|}{Makespans} \\
+       \hline
+       $M_{n}$ & I/O to storage device level $n$ & Eq.~\ref{eq:sea}\\
+       $M_{l}$ & I/O to Lustre w/o page cache & Eq.~\ref{eq:lustrenpc}\\
+       $M_{c}$ & I/O to page cache & Eq.~\ref{eq:cache}, \ref{eq:lustrepc}, \ref{eq:sea}\\
+       $M_{lc}$ & I/O to Lustre w/ page cache & Eq.~\ref{eq:lustrepc}\\
+       $M_{s}$ &  I/O to Sea w/o page cache & Eq.~\ref{eq:sea}\\
+       $M_{sl}$ & I/O to the Lustre component of Sea & Eq.~\ref{eq:snc}, \ref{eq:msl} \\
+       $M_{sd}$ & I/O to the local disk component of Sea & Eq.~\ref{eq:snc}, \ref{eq:msd} \\
+       $M_{st}$ & I/O to the tmpfs component of Sea & Eq.~\ref{eq:snc}, \ref{eq:mst} \\
+       $M_{sc}$ & I/O to Sea with page cache & Eq.~\ref{eq:msc} \\
+       \hline
+       \multicolumn{3}{|c|}{Data size} \\
+       \hline
+       $D_{r}$ & Read data & Eq.~\ref{eq:lustrenpc}\\
+       $D_{w}$ & Written data & Eq.~\ref{eq:lustrenpc}\\
+       $D_{cr}$ & Cached read data & Eq.~\ref{eq:cache}\\
+       $D_{cw}$ & Cached written data & Eq.~\ref{eq:cache}\\
+       $D_{I}$ & Input data & Eq.~\ref{eq:lustrepc}, \ref{eq:sea}\\
+       $D_{tr}$ & Sea: tmpfs read data & Eq.~\ref{eq:mst} \\
+       $D_{tw}$ & Sea: tmpfs written data & Eq.~\ref{eq:mst} \\
+       $D_{dr}$ & Sea: local disk read data & Eq.~\ref{eq:msd} \\
+       $D_{dw}$ & Sea: local disk written data & Eq.~\ref{eq:msd} \\
+       $D_{lr}$ & Sea: Lustre read data & Eq.~\ref{eq:msl} \\
+       $D_{lw}$ & Sea: Lustre written data & Eq.~\ref{eq:msl} \\
+       $D_{m}$ & Intermediate data & \ref{eq:msc} \\
+       $D_{f}$ & Final output data & \\
+       $F$ & File size & \\
+       \hline
+              \multicolumn{3}{|c|}{Bandwidths} \\                                         
+       \hline                                                                      
+       $B_{lr}$ & Perceived Lustre read bandwidth & Eq.~\ref{eq:lustrenpc}, \ref{eq:lustrepc}, \ref{eq:sea}, \ref{eq:msl}, \ref{eq:msc}\\
+       $B_{lw}$ & Perceived Lustre write bandwidth & Eq.~\ref{eq:lustrenpc}, \ref{eq:msl}\\
+       $B_{n}$ & Network bandwidth & \\              
+       $B_{or}$ & Lustre OST read bandwidth & \\     
+       $B_{ow}$ & Lustre OST write bandwidth & \\    
+       $B_{mr}$ & Memory read bandwidth (same as for tmpfs) & Eq.\ref{eq:cache}, \ref{eq:mst}, \ref{eq:msc}\\
+       $B_{mw}$ & Memory write bandwidth (same as for tmpfs) & Eq.~\ref{eq:cache}, \ref{eq:mst}, \ref{eq:msc}\\
+       $B_{dr}$ & Local disk read bandwidth & Eq.~\ref{eq:msd}\\                   
+       $B_{dw}$ & Local disk write bandwidth & Eq.~\ref{eq:msd}\\                  
+       \hline                                                                      
+       \multicolumn{3}{|c|}{Nodes} \\                                              
+       \hline                                                                      
+       $N_{c}$ & Number of compute nodes & Eq.~\ref{eq:cache}, \ref{eq:mst}, \ref{eq:msd}, \ref{eq:msc}\\
+       $N_{d}$ & Number of data nodes & \\           
+       $N_{t}$ & Number of threads per compute node & \\
+       \hline                                                                      
+       \multicolumn{3}{|c|}{Storage} \\                                            
+       \hline                                                                      
+       $O$ & Number of Lustre OSTs & \\              
+       $d$ & Number of local disks & \\              
+       $S_{t}$ & tmpfs storage space & \\
+       $S_{d}$ & Local disk storage space & \\
+       \hline                                                                      
+      \end{tabular}                                                                
+      \caption{Lustre and Sea model symbols}                                       
+      \label{table:1}                                                              
+      \end{table}
+
+      To effectively be able to predict in which scenarios Sea will provide speedup
+      over the baseline solution, we require a model. Since different parallel file
+      systems may operate differently, our baseline model will be based on Lustre which
+      is commonly used on HPC infrastructure.
+
+      For data intensive use cases, the makespan models for both Lustre and Sea can be broken
+      down into two components: The amount of time it takes read the data and the amount
+      of time it takes to write the data. With more heterogeneous applications (some components
+      are compute intensive whereas others are data intensive), a third component, comprising
+      of compute time, can be added. Furthermore, latency make also play a significant role
+      application makespan, particularly in scenarios with large amounts of small files.
+      We choose to ignore latency costs in our model and make the assumption that the
+      application bottleneck is the bandwidth, however, for more accurate estimates we
+      might consider the addition of file system latency, as a fourth model component.
+
+      A simplified version of the Lustre makespan model can be defined as follows:
+
+      \begin{equation}\label{eq:lustrenpc}
+          M_{l} =  \frac{D_{r}}{B_{lr}} + \frac{D_{w}}{B_{lw}}
+      \end{equation}
+
+      %\spacing{1}
+      %Where, \\
+      %$M_{l}$ is the Lustre makespan \\
+      %$D_{r}$ is the amount of data read \\
+      %$B_{r}$ is the read bandwidth \\
+      %$D_{w}$ is the amount of data written \\
+      %$B_{w}$ is the write bandwidth \\
+
+      %\spacing{1.5}
+      To determine the Lustre bandwidth, one must consider the three components involved:
+  1) the network bandwidth of the compute nodes, 2) the network bandwidth of the data nodes,
+  and 3) the collective bandwidth of the Lustre storage devices. Depending on each component's
+  respective values, either of the three may be the source of a bottleneck. The
+  Lustre bandwidth read and write models can therefore be described as follows:
+
+      \begin{equation} %\label{eq:blr}
+          B_{lr} = \min{(B_{n}N_{c}, B_{n}N_{d}, B_{or}\min{(O, N_{c}N_{t})})}
+      \end{equation}
+
+      and
+
+      \begin{equation}%\label{eq:blw}
+          B_{lw} = \min{(B_{n}N_{c}, B_{n}N_{d}, B_{ow}\min{(O, N_{c}N_{t})})}
+      \end{equation}
+
+      %\spacing{1}
+      %Where, \\
+      %$B_{lr}$ is the read bandwidth of Lustre \\
+      %$B_{lw}$ is the write bandwidth of Lustre \\
+      %$B_{or}$ is read bandwidth of the Lustre \gls{ost}s \\
+      %$B_{ow}$ is the write bandwith of the Lustre \gls{ost}s\\
+      %$B_{n}$ is the network bandwidth \\
+      %$N_{c}$ is the number of compute nodes \\
+      %$N_{d}$ is the number of data nodes \\
+      %$n$ is the number of threads per compute node \\
+
+      %\spacing{1.5}
+      For the sake of simplicity, the above models assume that the network bandwidth
+      between the compute nodes and data nodes is the same. This, however, may not
+      necessarily be the case. Furthermore, the model also assumes that each file can
+      only be located on a single OST, meaning that the parallel bandwidth can at maximum be
+      as fast as all OSTs combined and as slow as the minimum number of compute threads
+      reading and writing files.
+
+      As with many file systems, page cache plays an important role in the speed of
+      application read and writes in Lustre. Since the effect of page cache may be
+      non-negligible given amount of memory available and the data accessed during the
+      execution of the application, it is important to include it in our model. The makespan
+      of an application I/O to and from page cache can be described as the following:
+      Equation~\ref{eq:lustrenpc} where it is assumed that none of the data is written or read from page cache.
+
+      \begin{equation}\label{eq:cache}
+          M_{c} = \frac{D_{cr}}{B_{mr}N_{c}} + \frac{D_{cw}}{B_{mw}N_{c}}
+      \end{equation}
+
+      %\spacing{1}
+      %Where, \\
+      %$M_{c}$ is the makespan of writing to cache \\
+      %$D_{cr}$ is the amount of data read from cache \\
+      %$D_{cw}$ is the amount of data written to cache \\
+      %$N_{c}$ is the number of compute nodes \\
+      %$B_{mr}$ is the memory read bandwidth \\
+      %$B_{mw}$ is the memory write bandwidth \\
+
+      %\spacing{1.5}
+       As each individual compute node has its own set of memory, we treat the total
+      memory bandwidth as the sum of the individual memory bandwidth of each compute node.
+
+      Page cache is difficult to summarize accurately and effectively within a model.
+      For one, we must not only consider available memory and anonymous memory used by
+      the application, but we must also consider which pages are candidates for eviction
+      and which files they belong to. In addition, in the case of writes, we must consider
+      asynchronous flushing and the throttling that may occur as a consequence of surpassing
+      the \texttt{dirty\_ratio}. Furthermore, Lustre also has its own user-defined settings
+      for how it interacts with the cache that would add additional complexities to the model.
+      As a result, we assume two possible scenarios, one in which page cache is never used (Equation~\ref{eq:lustrenpc})
+      and one in which all application I/O occurs within page cache with the exception of the
+      first read which must occur on Lustre (Equation~\ref{eq:lustrepc}). These two models allow
+      us to define the bounds of Lustre's performance.
+
+      \begin{equation}\label{eq:lustrepc}
+          M_{lc} = \frac{D_{I}}{B_{lr}} + M_{c}
+      \end{equation}
+
+      %\spacing{1}
+      %Where, \\
+      %$M_{lc}$ is the Lustre makespan with page cache \\
+      %$D_{i}$ is the amount of input data \\
+      %$B_{lr}$ is the Lustre bandwidth \\
+      %$M_{c}$ is the makespan of the I/O to page cache \\
+
+      %\spacing{1.5}
+      Sea's model is more complex than Lustre's as there can be several
+      layers of different devices. For instance, Sea's model can be defined as:
+
+      \begin{equation}\label{eq:sea}
+          M_{s} = \frac{D_{I}}{B_{lr}} + M_{1} + \cdots + M_{n}
+      \end{equation}
+
+      Here, $M_{n}$ represents the makespans of the different possible storage levels
+      (e.g. tmpfs, NVMe, SSD, HDD, Lustre). For our model, we will assume 3 storage layers:
+      1) fast tmpfs, 2) intermediate local SSD storage, and 3) slow parallel file system layer.
+
+      Since the modelling of page cache is even more challenging with Sea due to the additional tmpfs
+      and SSD layer, we will will model the upper and lower performance bounds, as we did with Lustre.
+      Using the three layers and disregarding any possible effects of caching, we can redefine the
+      Sea model to be:
+
+      \begin{equation}\label{eq:snc}
+          M_{s} = M_{sl} + M_{sd} + M_{st}
+      \end{equation}
+
+         Where $M_{sl}$ represents the Lustre component of the Sea makespan, and
+      $M_{sd}$ and $M_{st}$ represent the disk and tmpfs component of the Sea
+      makespan, respectively.
+
+      The tmpfs component of the Sea makespan can be defined as the amount of
+      data that can be written ($D_{tw}$) to and read ($D_{tr}$) from tmpfs
+      over its respective bandwidths ($B_{mr}$ and $B_{mw}$). In other words:
+
+      \begin{equation}\label{eq:mst}
+          M_{st} = \frac{D_{tr}}{B_{mr}N_{c}} + \frac{D_{tw}}{B_{mw}N_{c}}
+      \end{equation}
+
+      \begin{equation*}\label{eq:dtr}
+          D_{tr} = \min\left(D_{m}, \max{\left(N_{c}(S_{t} - FN_{t}), 0 \right)} \right)
+      \end{equation*}
+      \begin{equation*}\label{eq:dtw}
+          D_{tw} = \min\left(D_{m} + D_{f}, \max{\left(N_{c}(S_{t} - FN_{t}), 0 \right)} \right)
+      \end{equation*}
+
+      In an optimal scenario all intermediate data ($D_{m}$) and final output
+      data ($D_{f}$) would fit in tmpfs. This would provide an application using
+      Sea in-memory performance. However, due to limited tmpfs storage space ($S_{t}$), it is unlikely to be the case. In addition, Sea may further restrict
+      available storage space to prevent exceeding tmpfs storage by ensuring that
+      there is at least sufficient space for $n$ threads to each write a file of
+      size $F$.
+
+      The local disk  makespan model is similar to the tmpfs makespan model, however, we
+      must ensure to disregard any data that has already been written to tmpfs.
+      Furthermore, in Sea, it is possible to leverage however many disk-based
+      file systems are available for use ($d$). For our model, we assume that the
+      size of each device is identical.
+      The makespan model can be defined as follows:
+
+      \begin{equation}\label{eq:msd}
+          M_{sd} =  \frac{D_{dr}}{B_{dr}dN_{c}} + \frac{D_{dw}}{B_{dw}dN_{c}}
+      \end{equation}
+
+      \begin{equation*}\label{eq:ddr}
+          D_{dr} = \min{(D_{m} - D_{tr}, \max{(N_{c}(S_{d}d - FN_{t}),0)})}
+      \end{equation}
+
+      \begin{equation*}\label{eq:ddw}
+          D_{dw} = \min{(D_{m} + D_{f} - D_{tw}, \max{(N_{c}(S_{d}d - FN_{t}),0)})}
+      \end{equation}
+
+            The final component of the Sea model is the Lustre component (Eq.~\ref{eq:msl}). Sea's Lustre
+      makespan model consists of the initial read from Lustre and includes and
+      data that must be written to Lustre due to insufficient space on local
+      storage and the makespan to read the intermediate data from Lustre.
+
+      \begin{equation}\label{eq:msl}
+          M_{sl} = \frac{D_{I}}{B_{lr}} + \frac{D_{lr}}{B_{lr}} + \frac{D_{lw}}{B_{lw}}
+      \end{equation}
+      \begin{equation*}\label{eq:dlr}
+          D_{lr} = D_{m} - D_{dr} - D_{tr}
+      \end{equation*}
+      \begin{equation*}\label{eq:dlw}
+          D_{lw} = D_{m} + D_{f} - D_{dw} - D_{tw}
+      \end{equation*}
+
+      Sea and Lustre have an identical lower bound. That is, ideally, both must
+      perform the first read from Lustre, but all subsequent data accesses can
+      be performed entirely within the page cache. The page cache model for Sea
+      can be defined as the following:
+
+      \begin{equation}\label{eq:msc}
+          M_{sc} = \frac{D_{I}}{B_{lr}} + \frac{D_{m}}{B_{mr}N_{c}} + \frac{D_{m} + D_{f}}{B_{mw}N_{c}}
+      \end{equation}
 \subsection{Experiments}
+\subsubsection{Pipelines}
+      To evaluate the Lustre and Sea models defined in Section~\ref{model} and the real
+      performance of both file systems with data intensive applications, we wrote a simple
+      Python application based off of Algorithm~\ref{alg:incrementation}.          
+      Using this application, we can easily control how much intermediate data is produced
+      by altering the amount of iterations required. Although our model should be able
+      to support images of different sizes, we wanted to minimize any possible     
+      scheduling effects from our experiments. Therefore, each application         
+      processes the same amount of input data and performs the same amount of computation.
+
+      As with the experiments performed in Chapter~\ref{chp:eval}, we will use the BigBrain
+      as a representative scientific datasets. For all our experiments, we utilize the
+      \SI{20}{\micro\meter} dataset, which totals to approximately \SI{603}{\gibi\byte}.
+      The dataset was broken down into 1000 files each consisting of \SI{617}{\mebi\byte}
+      of data.
+
+      We evaluate Sea using 4 different experimental conditions highlighted in Table~\ref{table:cond}:
+      1) varying the number of nodes, 2) varying the number of disks, 3) varying the number of threads and 4)
+      varying the number of iterations. Experimental condition 1 tests the effects 
+      of increasing concurrent accesses to Lustre while fixing disk parallel threads. Condition 2
+      varies disk contention while fixing contention to Lustre, whereas 3 tests the effects
+      of contention on both Lustre and local storage. Experimental condition       
+      4 varies the total amount of intermediate data produced by the application.
 \subsubsection{Infrastructure}
-\subsubsection{Pipelines}
-Big Brain incrementation
-fmri processing?
+\begin{table}
+      \centering
+      \begin{tabular}{@{}|c|c|c|c|c|@{}}
+       \hline
+       Condition & Nodes & Disks & Threads & Iterations \\
+       \hline
+       \multirow{5}{*}{\setword{1}{exp:nodes}} & 1 & 6 & 6 & 10 \\
+       & 2 & 6 & 6 & 10 \\
+       & 3 & 6 & 6 & 10 \\
+       & 4 & 6 & 6 & 10 \\
+       & 5 & 6 & 6 & 10 \\
+       \hline
+       \multirow{6}{*}{\setword{2}{exp:disks}} & 4 & 1 & 6 & 10 \\
+       & 4 & 2 & 6 & 10 \\
+       & 4 & 3 & 6 & 10 \\
+       & 4 & 4 & 6 & 10 \\
+       & 4 & 5 & 6 & 10 \\
+       & 4 & 6 & 6 & 10 \\
+       \hline
+       \multirow{5}{*}{\setword{3}{exp:threads}} & 4 & 6 & 1 & 10 \\
+       & 4 & 6 & 6 & 10 \\
+       & 4 & 6 & 12 & 10 \\
+       & 4 & 6 & 24 & 10 \\
+       & 4 & 6 & 48 & 10 \\
+       \hline
+       \multirow{5}{*}{\setword{4}{exp:iterations}} & 4 & 6 & 6 & 1 \\
+       & 4 & 6 & 6 & 5 \\
+       & 4 & 6 & 6 & 10 \\
+       & 4 & 6 & 6 & 15 \\
+       & 4 & 6 & 6 & 20 \\
+       \hline
+
+      \end{tabular}
+      \caption{Experimental conditions}
+      \label{table:cond}
+      \end{table}
+
+      \begin{table}
+      \centering
+      \begin{tabular}{@{}|c|c|c|@{}}
+       \hline
+       Storage layer & Action & Average bandwidth (MiB/s) \\
+              \hline
+       \multirow{3}{*}{tmpfs} & read & 6676.48 \\
+       & cached read & 6318.08  \\
+       & write & 2560.00 \\
+       \hline
+       \multirow{3}{*}{local disk} & read & 501.70  \\
+       & cached read & 7034.88 \\
+       & write & 426.00 \\
+       \hline
+       \multirow{3}{*}{Lustre} & read & 1381.14 \\
+       & cached read & 6103.04  \\
+       & write & 121.00  \\
+
+       \hline
+
+      \end{tabular}
+      \caption{Storage benchmarks}
+      \label{table:fs}
+      \end{table}
+            Our experiments were executed on a Centos 8.1 (Linux kernel 4.18.0) cluster with 8 compute nodes,
+      a 4 data node Lustre server with 1 dedicated metadata node. Each compute node
+      is equipped it two Intel(R) Xeon(R) Gold 6130 CPUs, \SI{250}{\gibi\byte} of memory with
+      \SI{126}{\gibi\byte} of tmpfs space and 6 \SI{447}{\gibi\byte} Intel SSDSC2KG480G8R SSDs.
+      The data nodes each contain 11 \SI{10}{\tera\byte} HGST HUH721010AL HDD OSTs
+      and \SI{62}{\gibi\byte} memory. The metadata server contains a Toshiba KPM5XVUG960G \SI{960}{\giga\byte} MDT.
+      The network bandwidth is \SI{25}{\giga\bit}E and uses tcp for communcation.
+      Jobs are scheduled on the cluster from a controller node using Slurm with cgroups. Swapping is
+      disabled.
+
+      Each file system was benchmarked using \texttt{dd} using 5 repetitions. The average bandwidths are
+      reported in Table~\ref{table:fs}
+
+%Big Brain incrementation
+%fmri processing?
 
@@
 
+      For the most part, the model appears to correctly encapsulate the
+      performance range of Sea and Lustre, often times with the real data being
+      situated at the halfway point between the min and max performance bounds.
+      However, there are some notable exceptions.
+      In Figure~\ref{fig:disks}, the model appears to adequately describe the
+      performance of 1 and 2 disks, but the real results progressively surpass
+      makespan estimates as the number of disks increases further. Model estimates
+      for Lustre, however, encapsulate the results obtained. Interestingly
+      enough, curve of the mid point line seems to parallel that of the results.
+
+      Model estimates appear to be inaccurate to for Experiment~\ref{exp:threads}, as seen in
+      Figure~\ref{fig:threads}. For 1 and 6 threads, the results occur within
+      the bounds of the model, however, at 12 threads, the real results start
+      to exceed model estimates and then plateau at somewhere around 24 threads.
+
+      In Figure~\ref{fig:iterations}, although model and results appear to
+      generally be in accordance with each other, the model captures neither
+      Sea's or Lustre's actual makespan. Furthermore, it is unclear from the
+      results if the trend is actually linear.
+
+      The greatest overall speedup occurred in Experiment~\ref{exp:threads} at 24 threads. In
+      this experiment, Sea was 4.3~x faster than the same conditions on Lustre.
+      The second highest speedup also occurred in Experiment~\ref{exp:threads}, when using 48
+      threads. Sea was approximately 3.3~x faster than its Lustre counterpart.
+      Otherwise, the maximum speedup obtained with Sea was 1.5~x in Experiment~\ref{exp:nodes},
+      1.4~x in Experiment~\ref{exp:disks} and 1.9~x in Experiment~\ref{exp:iterations}.
+
+      In general, Sea's performance was not worse than that of Lustre's. The only
+      time there was a noticeable decrease in performance occurred in Experiment~\ref{exp:disks}
+      when the number of disks used was below 3. This is a result of there being
+      a greater contention on local storage with a reduced number of disks. While
+      the local storage device bandwidths exceed that of Lustre OSTs (Table~\ref{table:fs}), Lustre has significantly more disks available that the overall
+      attainable bandwidth is greater than that of local storage. Furthermore,
+      the network bandwidth to Lustre is also superior to the combined local
+      storage bandwidth.
+
+\section{Discussion}
+          Overall, our results show that Sea performs better than Lustre when there is 
+      increased contention on Lustre as compared to the local storage devices. We  
+      expect that contention to Lustre will be significantly greater on a          
+      production cluster where many users are accessing Lustre at the same time.   
+      As this may be the case, we might even expect Sea to outperform Lustre when  
+      processing data in a sequential application. However, as can be seen in      
+      Figure~\ref{fig:disks}, when local storage contention is greater than that   
+      of Lustre's, it is not recommended to use Sea. To circumvent this issue      
+      such that a user would not need prior knowledge on file system usage trends  
+      before deciding whether or not to use Sea, Sea could periodically benchmark  
+      the different file systems to ensure that the file is written to the most    
+      appropriate location.
+
+      The model accuracy generally appeared to adequately capture Sea and Lustre   
+      performance trends, however, failed in some cases.                           
+      For Figure~\ref{fig:disks}, the inaccuracy of the model might be due to poor 
+      SSD bandwidth estimation as the trend appears to be similar to real results.
+
+      The model predictions for Experiment~\ref{exp:threads} may be off for Lustre as a result
+      of metadata contention. Whereas we have many OSTs and data nodes for         
+      data storage, we only have a single disk dedicated to metadata storage.      
+      This normally would not have been a problem, except the MDS can only support 
+      32 threads at a time and at 8 concurrent threads per node, we reach this limit.
+      The model, at the moment, does not consider latency for metadata requests, and
+      therefore, is unable to properly describe this behaviour.
+
+      Similarly to Lustre, model prediction error for Sea in Experiment~\ref{exp:threads}
+      may arise from unaccounted for factors, such as compute time, which would otherwise
+      be insignificant with a greater number of threads. What is interesting is perhaps
+      the closeness in proximity that Sea's makespan is to Lustre's at a single thread.
+      There may be overheads caused by Sea's asynchronous flusher processes that are more
+      apparent with a single thread, particularly if the asynchronous flushing is occurring
+      from the same file system that is currently being read to.
+
+      Sea's overheads may also explain the                                         
+      model prediction errors found in Experiment~\ref{exp:iterations} when at a single iteration.
+      As all the data in both cases would end up on Lustre, it is normal that Sea would experience increased
+      overheads as there is are multiple layers of filesystems involved.  
 \section{Conclusion}
-The conclusion goes here.
+     While the preliminary results of Sea are promising, there still remains some work
+      to be done. We would like to integrate bandwidth estimation into Sea such that
+      Sea writes to the most efficient filesystem at the given point in time rather
+      than always referring to the hierarchy. Furthermore, we intend on improving the benchmarks with
+      repetitions to be able to account for system variability, in addition to measuring
+      the bandwidth of Sea with relevant Big Data neuroimaging applications.
 

### 2021-06-17 src/passthrough.cpp
fixed bug in make_file_name_canonical (#10)

@@ -210,7 +210,7 @@ char *make_file_name_canonical(char const *file_path)
     }
     canonical_file_path = ((funcptr_realpath)libc_realpath)(file_path, NULL);
 
-    if (canonical_file_path == NULL && errno == ENOENT)
+    if (canonical_file_path == NULL && (errno == ENOENT || errno == ENOTDIR))
     {
       // The file was not found. Back up to a segment which exists,
       // and append the remainder of the path to it.

### 2021-06-17 src/passthrough.cpp
removed errno verification in make_file_name_canonical

@@ -210,7 +210,7 @@ char *make_file_name_canonical(char const *file_path)
     }
     canonical_file_path = ((funcptr_realpath)libc_realpath)(file_path, NULL);
 
-    if (canonical_file_path == NULL && (errno == ENOENT || errno == ENOTDIR))
+    if (canonical_file_path == NULL) //&& (errno == ENOENT || errno == ENOTDIR))
     {
       // The file was not found. Back up to a segment which exists,
       // and append the remainder of the path to it.

### 2021-06-17 src/passthrough.cpp
added statx wrapper

@@ -66,6 +66,7 @@ void *libc_dirfd;
 void *libc_access;
 void *libc_faccessat;
 void *libc_stat;
+void *libc_statx;
 void *libc_fstat;
 void *libc_fstatat;
 void *libc_lstat;
@@ -471,6 +472,7 @@ void initialize_functions()
   libc_access = dlsym(libc, "access");
   libc_faccessat = dlsym(libc, "faccessat");
   libc_stat = dlsym(libc, "stat");
+  libc_statx = dlsym(libc, "statx");
   libc_lstat = dlsym(libc, "lstat");
   libc_lstat64 = dlsym(libc, "lstat64");
   libc_fstat = dlsym(libc, "fstat");

### 2021-06-18 src/sea.cpp
Fixed #12

@@ -10,6 +10,8 @@
 #include <sys/statvfs.h>
 #include <sys/types.h>
 #include <utime.h>
+#include <unistd.h>
+#include <sys/stat.h>
 
 int sea_internal;
 
@@ -55,25 +57,7 @@ int unset_internal()
  */
 int sea_checkpath(const char *path)
 {
-    struct stat buf;
-
-    errno = 0;
-    //printf("checkpath %s\n", path);
-    set_internal();
-    int ret = 0;
-    if ((ret = __xstat(_STAT_VER_LINUX, path, &buf)) != 0)
-    {
-        //printf("errno %d\n", errno);
-        unset_internal();
-        return 0;
-    }
-
-    struct utimbuf times;
-    times.actime = time(0);
-    times.modtime = buf.st_mtime;
-    utime(path, &times);
-    unset_internal();
-    return 1;
+    return (access(path, F_OK) == 0);
 }
 
 /**

### 2021-06-18 src/sea.cpp
Fixed #12

@@ -11,7 +11,6 @@
 #include <sys/types.h>
 #include <utime.h>
 #include <unistd.h>
-#include <sys/stat.h>
 
 int sea_internal;
 

### 2021-06-19 src/passthrough.cpp
minor fix to make_file_name_canonical

@@ -248,7 +248,9 @@ char *make_file_name_canonical(char const *file_path)
             // to a canonical form of the existing path.
             char *combined_file_path = (char *)malloc(strlen(canonical_file_path) + strlen(file_path_copy + char_idx + 1) + 2);
             strcpy(combined_file_path, canonical_file_path);
-            strcat(combined_file_path, "/");
+
+            if (combined_file_path[strlen(combined_file_path) - 1] == '/')
+              strcat(combined_file_path, "/");
             strcat(combined_file_path, file_path_copy + char_idx + 1);
             free(canonical_file_path);
             canonical_file_path = combined_file_path;

### 2021-06-19 src/passthrough.cpp
fixed bug introduced in previous commit

@@ -249,7 +249,7 @@ char *make_file_name_canonical(char const *file_path)
             char *combined_file_path = (char *)malloc(strlen(canonical_file_path) + strlen(file_path_copy + char_idx + 1) + 2);
             strcpy(combined_file_path, canonical_file_path);
 
-            if (combined_file_path[strlen(combined_file_path) - 1] == '/')
+            if (combined_file_path[strlen(combined_file_path) - 1] != '/')
               strcat(combined_file_path, "/");
             strcat(combined_file_path, file_path_copy + char_idx + 1);
             free(canonical_file_path);

### 2021-06-19 tests/test_custom.cpp
fixed single sea source error and added a test case for it

@@ -322,8 +322,8 @@ int main(int argc, char **argv)
     ::testing::InitGoogleTest(&argc, argv);
 
     initialize_passthrough_if_necessary();
-    ((funcptr_mkdir)libc_mkdir)("newdir", 0666);
-    ((funcptr_mkdir)libc_mkdir)("newdir/subdir", 0666);
+    ((funcptr_mkdir)libc_mkdir)("newdir", 0766);
+    ((funcptr_mkdir)libc_mkdir)("newdir/subdir", 0766);
 
     return RUN_ALL_TESTS();
 }

### 2021-06-19 tests/test_custom.cpp
changed file permissions to test folder, might change to rm in teardown

@@ -322,8 +322,8 @@ int main(int argc, char **argv)
     ::testing::InitGoogleTest(&argc, argv);
 
     initialize_passthrough_if_necessary();
-    ((funcptr_mkdir)libc_mkdir)("newdir", 0766);
-    ((funcptr_mkdir)libc_mkdir)("newdir/subdir", 0766);
+    ((funcptr_mkdir)libc_mkdir)("newdir", 0777);
+    ((funcptr_mkdir)libc_mkdir)("newdir/subdir", 0777);
 
     return RUN_ALL_TESTS();
 }

### 2021-06-19 src/passthrough.cpp
added back flags to make_file_name_canonical

@@ -211,7 +211,7 @@ char *make_file_name_canonical(char const *file_path)
     }
     canonical_file_path = ((funcptr_realpath)libc_realpath)(file_path, NULL);
 
-    if (canonical_file_path == NULL) //&& (errno == ENOENT || errno == ENOTDIR))
+    if (canonical_file_path == NULL && (errno == ENOENT || errno == ENOTDIR))
     {
       // The file was not found. Back up to a segment which exists,
       // and append the remainder of the path to it.

### 2021-10-07 paper/paper_neuro.tex
start to paper

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{longtable}
\usepackage[binary-units=true]{siunitx}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{amsmath}

\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}: #1}}}

% benchmark on beluga, prefreesurfer with RUIS
% use more than 64 subjects for prefreesurfer --> few 100 subjects
% think more about how to benchmark I/O
% try distributed jobs with Sea

\begin{document}
\title{Automated neuroimaging data management with the Sea filesystem}
% remove neuroimaging and replace with scientific computing  add user-space, switch HPC to computing clusters

\author{Val\'erie Hayot-Sasson and Tristan Glatard}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
\IEEEtitleabstractindextext{%
\begin{abstract}
    Neuroimaging open-data initiatives have led to the availability of large scientific datasets. These voluminous
    datasets provide researchers with new insights into the human brain across diverse populations. Difficulties with
    managing such large data has partially hindered the advancement of scientific studies analyzing these datasets. Many
    software tools and strategies have emerged to facilitate the processing. Many open datasets have been stored on cloud storage
    services, such as AWS, enabling rapid transfer of the data at a cost. Software tools such as Datalad has facilitated both the
    sharing and versioning of data. Despite these advancements in-analysis data-management remains limited in neuroimaging workflows.

    Neuroimaging workflows, particularly preprocessing ones, may lead to a magnification in output sizes due to <add reason>. They also produce
    significant amounts of intermediary data due to being composed of a multitude of steps. 
\end{abstract}
}

% make the title area
\maketitle

\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{}{}

Neuroimaging datasets continue to grow in size resulting in new challenges related
to data management. The current largest neuroimaging datasets, such as the Human Connectome Project (HCP)~\ref{HCP}
and the UK Biobank~\ref{ukbiobank} reach up to Petabytes of data. Big Data in neuroimaging can be present in two formats: 1)
very large files, such as those found in the BigBrain~\ref{bigbrain}, and 2) large datasets made up of very small files,
such as those typically found in fMRI datasets.

When processing, large datasets can result in longer processing times even when compute resources are ample.
This is a direct result of the underlying storage used by the applications to read and write data. Since large
datasets require a significant amount of available storage space, more often than not, slower but larger storage is
selected, resulting in longer data transfer times during processing. With the Sea filesystem, we aim to
leverage all available storage in order to reduce overall data transfer time during pipeline execution.

Many researchers rely on one of two systems to meet their storage and computing needs: 1) High Performance 
Computing (HPC) clusters and 2) the cloud. Whereas the cloud simplifies data sharing and gives researchers
access to a wide variety of infrastructures, HPC clusters are a cost-effective solution to accessing a wide
array of resources for researchers. In this paper, we will focus on HPC processing.

HPC systems rely on scalable network-based parallel filesystems (e.g. Lustre) for storage. While such file
systems offer excellent performance, they are shared between all the users on the cluster. Meaning that users
with data-intensive can effectively deteriorate the performance of the shared file system for all users on
the cluster. Solutions for improving shared file system performance include throttling the data intensive workloads
or recommending the use of Burst Buffers~\cite{bb} (e.g. reserving a compute node for storage or leveraging
local compute storage during processing). The latter, however, requiring that the user manages their data to
and from the Burst Buffer.

Leveraging local storage to improve data intensive workload performance has long since existed in Big Data
frameworks such as Apache Spark~\cite{spark} and Dask~\cite{dask}. While these frameworks have been used to
process neuroimaging data~\cite{manypapers}, they remain seldom used as their require rewriting existing
standard neuroimaging tools (e.g FSL, AFNI, SPM) for the framework. Although it is possible to use the
tools within the Big Data frameworks, optimizations like in-memory computing would not be leveraged due to 
the fact that neuroimaging tools are meant to be used as command-line tools and do not provide interfaces that 
enable the data to be transferred in-memory.

Frameworks used in neuroimaging, such as NiPype~\cite{nipype} and Joblib~\cite{joblib} instead focus on reducing compute times of workloads.
This is because even with large datasets, neuroimaging data processing remains split between compute and data
intensive components. Although these frameworks do not prohibit the use of intelligent data management, it is not directly integrated
into the workflow. In order to give neuroimaging applications data management capabilities, the applications must interact with a
file system that can do so.

In order for a file system to be usable by the average researcher on an HPC system, they must be loadable without administrative
privileges. Furthermore, as the applications are typically made to interact with POSIX-based file systems, the new file system must
also be compliant to the format. One method to ensure that these conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick
enables the interception of library calls by redefining them.

In this paper, we present Sea, a file system designed to automate data management of neuroimaging pipelines running on HPC systems.

\section{Related Work}

Put in intro.
summary of computing landscape for neuroimaging.
HCP, CBRAIN, nipype, joblib. sea can be used in all these contexts.

\section{Materials and Methods}
\subsection{Datasets}
Functional magnetic resonance imaging datasets can vary greatly in total number
of images and number of volumes within each image. To adequately capture the 
diversity of datasets and when Sea could be pertinent, we selected three datasets of varying
sizes: 1) OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD  dataset~\cite{preventad}, and
3) the HCP dataset~\cite{hcp}. The ds001545 dataset is a total of 45.94~$GB$ (1778 files) and consists of data collected
from 30 subjects in a single session. The PREVENT-AD dataset is a 255.0~$GB$ dataset (53061 files) consisting of data collected
from 308 subjects. The HCP dataset is the largest of the three datasets at 85.4~$TB$ (x files) consisting of data collected from
1113 subjects.

%just for example run
%dataset size, resolution, resolution of voxels
%matrix size pixdim1X2
%number of slices pixdim
%number of volumes pixdim4
%inplane resolution pixsize1x2
%slice thickness pizsize3 in mm
%repetition time pixsize4 in s

%HCP resting state + x stats

%# files written and output size, duration and memory consumption.
\subsection{fMRI Preprocessing Pipelines}

Similarly to datasets, preprocessing pipelines can also vary in duration as a result of methodological differences.
Thus, we preprocessed each dataset using four standard preprocessing pipelines: FSL~\cite{fsl}, AFNI~\cite{AFNI},
SPM~\cite{SPM}, fMRIPrep~\cite{fmriprep}. Table~\ref{tb1} provides an reference of the differences in computation and data-intensivity
of the different pipelines on a single subject of the ds001545 dataset.

Each tool was set to only run the preprocessing pipeline. Scripts used to execute each pipeline can be found at~\ref{github.com}.
For the FSL preprocessing pipeline, the default settings were maintained with the exception of slice-timing correction (interleaved),
intensity normalization and non-linear registration to standard space.

%% Add SPM here

For AFNI preprocessing, the

%%{provide subject #/session#} of the ds001545 dataset.

\subsection{Controls}
\subsection{Infrastructure}

\section{Results}

\section{Discussion}
talk about testing

Flushing - rely on modification time
\section{Conclusion}
%

\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% biography section

\begin{IEEEbiography}{Val\'erie Hayot-Sasson}
\end{IEEEbiography}
\begin{IEEEbiography}{Tristan Glatard}
\end{IEEEbiography}
\end{document}

### 2021-10-08 paper/paper_neuro.tex
part of sea intro

@@
 privileges. Furthermore, as the applications are typically made to interact with POSIX-based file systems, the new file system must
-also be compliant to the format. One method to ensure that these conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick
-enables the interception of library calls by redefining them.
+also be compliant to the format. One method to ensure that these conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick is
+used to intercept select library calls and redefine their behaviour. It has been used in many projects and to create lightweight versions
+of filesystems~\cite{xtreemfs}.
 
 In this paper, we present Sea, a file system designed to automate data management of neuroimaging pipelines running on HPC systems.
+The goal of Sea is to complement existing neuroimaging applications such that both efficient compute and data management can be achieved
+executed on HPC without significant user intervention. As a secondary goal, Sea aims to ensure the fair sharing of cluster usage
+by alleviating the impacts of heavy writers on the shared file systems as a whole.
 
-\section{Related Work}
+% \section{Related Work}
 
-Put in intro.
-summary of computing landscape for neuroimaging.
-HCP, CBRAIN, nipype, joblib. sea can be used in all these contexts.
+% Put in intro.
+% summary of computing landscape for neuroimaging.
+% HCP, CBRAIN, nipype, joblib. sea can be used in all these contexts.
 
 \section{Materials and Methods}
-\subsection{Datasets}
+
+\subsection{The Sea filesystem}
+
+The Sea file system is an on-demand file system that leverages the \texttt{LD\_PRELOAD} trick to intercept POSIX file system (more specifically, glibc)
+calls on Linux systems. This enables Sea to redirect write calls aimed for a slower storage device to a faster device whenever
+possible. Similarly, when intercepting read calls, Sea can choose to read from a faster device if a copy is available on that device.
+Sea decides which storage location it can write to based on the detailed provided in an initialization file called \texttt{sea.ini}. 
+This file informs Sea of which locations it can use to read and write to, and well as their order of priority. An example of the initialization
+file can be seen in ~\ref{seaini}.
+
+As neuroimaging pipeline results are typically required post-processing and HPC compute-local resources are only accessible during the
+reserved duration, Sea provides functionality to flush and evict data to persistent shared storage. This is accomplished via a separate
+thread (known as the ``flusher'') that moves data from the caches to long-term storage. A separate thread is used, in this case, to avoid
+interrupting on going processing with data management operations.
+Users must inform Sea of files that need to be
+persisted to storage within a file called \texttt{.sea\_flushlist}, and temporary files which can be removed from cache within a file
+called the \texttt{.sea\_evictlist}. Both these files can be populated using regular expressions to denote the paths to flush and evict.
+If a file occurs in both the \texttt{.sea\_flushlist} and \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move} operation
+and simply move the file from local to persistent storage rather than copying it. Files that will be reused by the application should only
+be flushed rather than flushed and evicted, as files can benefit from speedups related to reading from fast rather than slow storage.
+Sea currently also provides a rudimentary prefetch thread that can move files located within the Sea filesystem to the fastest available cache.
+To use Sea's prefetch capabilities, a file called \texttt{.sea\_prefetchlist} needs to be populated using regular expressions like the
+flushing and eviction files.
+
+To interact with the Sea file system, a mountpoint is used. The mountpoint is an empty directory that behaves as a view to
+all the files and directories stored within Sea. In order to keep track of the locations of the files within the mountpoint, Sea
+mirrors the file structure of each storage location across all storage locations. That means, it is generally advisable to provide
+empty storage locations for Sea to write to as the mirroring of large directories can take some time. When prefetching, it is
+best to provide Sea with a path that only points to that data the requires prefetching as mirroring the paths of full datasets may 
+extend processing time.
+
+Sea can easily be launched directly using the available containers on the GitHub registry, or can be compiled via source using Make.
+It requires a version of glibc greater than X. Once compiled, Sea can be executed using the `sea-app.sh' binary.
+
+\subsection{Performance analysis of Sea for neuroimaging}
+
+To determine the performance gain Sea brings to neuroimaging analyses, we must evaluate the value of Sea on a variety of neuroimaging
+applications. For our analysis, we selected different fMRI preprocessing applications as fMRI processing has some of the most
+well-established tools for neuroimaging and some of the largest datasets. Of course, different modalities an tools may result in
+vastly different data access patterns and compute times, we use fMRI prepreprocessing as our baseline for the benefits that can be
+obtained with Sea.
+
+\subsubsection{Datasets}
 Functional magnetic resonance imaging datasets can vary greatly in total number
@@
 %# files written and output size, duration and memory consumption.
-\subsection{fMRI Preprocessing Pipelines}
+\subsubsection{fMRI Preprocessing Pipelines}
 
@@
 
-For AFNI preprocessing, the
+For AFNI preprocessing, the pipeline was configured to perform slice timing, alignment to Talaraich space, registration, smoothing
+and brain masking. While AFNI does have some parallelized components that can be controlled via environment variables, we chose to let 
+the pipeline use as much parallelism as required for its execution.
+
+We set fMRIPrep to run with \texttt{--fs-no-reconall} flag set and set the \texttt{--bids-database-dir} to a tmpfs location. 
+Furthermore, the seeds were fixed to ensure reproducible results. Unlike the other tools where only a single fMRI image was preprocessed
+per process, it was not possible to do so with fMRIPrep. Thus, fMRIPrep preprocessed entire subject directories at a time
 

### 2021-10-08 paper/paper_neuro.tex
typos

@@
 \algnotext{EndOutput}
+\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:} #1\color{black}\xspace}
 \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
@@
 
-    Neuroimaging workflows, particularly preprocessing ones, may lead to a magnification in output sizes due to <add reason>. They also produce
+    Neuroimaging workflows, particularly preprocessing ones, may lead to a magnification in output sizes due to the presence of multiple intermediary processing steps. They also produce
     significant amounts of intermediary data due to being composed of a multitude of steps. 
@@
 systems offer excellent performance, they are shared between all the users on the cluster. Meaning that users
-with data-intensive can effectively deteriorate the performance of the shared file system for all users on
+with data-intensive analyses can effectively deteriorate the performance of the shared file system for all users on
 the cluster. Solutions for improving shared file system performance include throttling the data intensive workloads
@@
 
-As neuroimaging pipeline results are typically required post-processing and HPC compute-local resources are only accessible during the
+As neuroimaging pipeline results typically require post-processing and HPC compute-local resources are only accessible during the
 reserved duration, Sea provides functionality to flush and evict data to persistent shared storage. This is accomplished via a separate

### 2021-10-09 src/config.cpp
changed name of source to cache and fixed bug with duplicate files

@@ -55,7 +55,7 @@ void create_config_file(char *config_file)
             "log_file = %s/sea.log ;\n"
             "log_level = 1 ; # DEBUG=4, INFO=3, WARNING=2, ERROR=1, NONE=0 (use <= 3 tests)\n"
             "n_levels = 1 ;\n"
-            "source_0 = %s/source ; # use absolute paths\n"
+            "cache_0 = %s/source ; # use absolute paths\n"
             "max_fs = 100"
             "\n",
             sea_home, sea_home, sea_home);
@@ -84,7 +84,7 @@ void parse_config()
     }
     if ((sea_config.n_levels = iniparser_getint(config_dict, "sea:n_levels", 0)) == 0)
     {
-        printf("Missing n_sources in config file %s\n", config_file);
+        printf("Missing n_caches in config file %s\n", config_file);
         exit(1);
     }
     std::vector<std::string> all_sources;
@@ -96,7 +96,7 @@ void parse_config()
     {
         std::vector<std::string> current_lvl;
         char source_name[15];
-        sprintf(source_name, "sea:source_%d", i);
+        sprintf(source_name, "sea:cache_%d", i);
         char *lvl_str;
         if ((lvl_str = (char *)iniparser_getstring(config_dict, source_name, NULL)) == 0)
         {

### 2021-10-09 src/functions.cpp

@@ -23,7 +23,7 @@
         if (d != NULL)                                                                                                                             \
         {                                                                                                                                          \
             log_msg(DEBUG, "sea_readnext%s: read entry %s", #VERSION, d->d_name);                                                                  \
-            if (sd->curr_index > 0 && d->d_type == DT_DIR)                                                                                         \
+            if (!sd->seen_files.insert(std::string(d->d_name)).second || (sd->curr_index > 0 && d->d_type == DT_DIR))                              \
             {                                                                                                                                      \
                 log_msg(DEBUG, "sea_readnext%s: Curr dir is a directory, reading next inode.", #VERSION);                                          \
                 d = sea_readnext##VERSION(d, sea_conf, sd);                                                                                        \
@@ -58,6 +58,11 @@
                 {                                                                                                                       \
                     log_msg(DEBUG, "readdir%s: reading sea dir current idx = %d", #VERSION, sd->curr_index);                            \
                     d = ((funcptr_readdir##VERSION)libc_readdir##VERSION)(sd->dirp);                                                    \
+                                                                                                                                        \
+                    if (d != NULL)                                                                                                      \
+                    {                                                                                                                   \
+                        sd->seen_files.insert(std::string(d->d_name));                                                                  \
+                    }                                                                                                                   \
                 }                                                                                                                       \
                                                                                                                                         \
                 log_msg(DEBUG, "readdir%s: curr_idx %d n_sources %d d NULL %d", #VERSION, sd->curr_index + 1, sd->total_dp, d == NULL); \
@@ -70,6 +75,10 @@
                     }                                                                                                                   \
                     log_msg(DEBUG, "readdir%s: launching sea_readnext %d %d", #VERSION, sd->curr_index, sd->total_dp);                  \
                     d = sea_readnext##VERSION(d, sea_conf, sd);                                                                         \
+                }                                                                                                                       \
+                if (d == NULL)                                                                                                          \
+                {                                                                                                                       \
[…]

### 2021-10-09 src/sea.h

@@ -28,6 +28,7 @@ struct SEA_DIR
     char **dirnames;
     int curr_index;
     int total_dp;
+    std::set<std::string> seen_files; // bandaid solution to not listing duplicate files in mount
 };
 
 #endif

### 2021-10-10 bin/sea_flusher.sh
fixes to flusher added prefetch and test cases for both

@@ -4,6 +4,8 @@
 set -u
 #set -x
 
+run=$1
+
 base_source=""
 sources_arr=()
 n_sources=0
@@ -13,6 +15,12 @@ evict_file=${SEA_HOME}/.sea_evictlist
 tmpfile=$(mktemp /tmp/.sea_flush.inprog.XXXXXX)
 
 fe_old () {
+
+    # if we have no cache directories, flusher should terminate
+    if [[ ${sources_arr[@]} == "" ]]
+    then
+        exit
+    fi
 
     found_files=$1 # files that currently exist in source mounts
     OLDTIME=$2 # user specified expiration time
@@ -47,20 +55,40 @@ fe_old () {
                 fi
             done
 
-            echo "$task $f to ${base_source}${subpath}"
+            if [[ -f ${base_source}${subpath} ]]
+            then
+                BASETIME=$(stat ${base_source}${subpath} -c %Y || true && continue)
+
+                if [[ $BASETIME != NULL ]]
+                then
+                    TIMEDIFF=$((${FILETIME} - ${BASETIME}))
+
+                    if [[ ${TIMEDIFF} == 0 ]]
+                    then
+                        continue 
+                    fi
+                fi
+
+            fi
+
+            if [[ $task != "rm" ]]
+            then
+                echo "[Sea] Flusher: $task $f to ${base_source}${subpath}"
+            fi
             # makedirectory in case it does not exist
             mkdir -p $(dirname ${base_source}${subpath})
             # if we have some files that older that OLDTIME, flush/evict them
             # process each image individually (as opposed to all at once)
             if [[ $task == "mv" ]]
             then 
-                mv $f ${base_source}${subpath} || true
+                mv $f ${base_source}${subpath}
             elif [[ $task == "cp" ]]
             then 
-                cp $f ${base_source}${subpath} || true
+                cp $f ${base_source}${subpath}
             elif [[ $task == "rm" ]]
             then 
-                rm  $f || true
+                echo "[Sea] Flusher: $task $f"
+                rm $f || true
             fi
            
         fi
@@ -84,12 +112,12 @@ get_rgx () {
[…]

### 2021-10-10 bin/sea_launch.sh

@@ -7,25 +7,29 @@ trap "exit" INT TERM
 #trap "kill 0" EXIT
 
 conf_file=${SEA_HOME}/sea.ini
-curr_dir=`dirname "$(readlink -f "$0")"`
+curr_dir=$(dirname "$(readlink -f "$0")")
+parent_dir=$(dirname ${curr_dir})
 
 n_sources="$(($(awk -F "=" '/n_sources/ {print $2}' ${conf_file} | tr -d ' ;') - 1))"
 
 for ((i=0;i<${n_sources};i++))
 do
-    source=($(cat ${conf_file} | grep "source_$i" | cut -d "=" -f 2 | tr -d ' ;'))
+    source=($(cat ${conf_file} | grep "cache_$i" | cut -d "=" -f 2 | tr -d ' ;'))
     mkdir -p ${source} 
 done
 
-${curr_dir}/sea_flusher.sh &
+${parent_dir}/build/mirror
+${curr_dir}/sea_prefetch.sh 1 &
+${curr_dir}/sea_flusher.sh 1 &
+
 pid=$!
-echo "flusher pid" ${pid}
+echo "[Sea] Launch: flusher pid" ${pid}
 
 trap 'trap - SIGTERM && kill $pid' SIGINT SIGTERM #EXIT
 
-echo "Launching script $@"
-bash $@
-echo "killing flush" && kill $pid
+echo "[Sea] Launch: Launching script $@"
+LD_PRELOAD=${parent_dir}/build/sea.so bash -c "$@"
+echo "[Sea] Launch: Killing flush" && kill $pid
 
 wait $pid
 

### 2021-10-10 bin/sea_prefetch.sh

@@ -1,0 +1,122 @@
+#!/bin/bash
+
+set -u
+
+run=$1
+
+base_source=""
+sources_arr=()
+n_sources=0
+conf_file=${SEA_HOME}/sea.ini
+prefetch_file=${SEA_HOME}/.sea_prefetchlist
+
+get_rgx () {
+    SOURCE=$1
+    FILE=$2
+    rgx=""
+    
+    if [ -f $FILE ]
+    then
+        while IFS="" read -r re || [ -n "$re" ]
+        do
+            if [[ $rgx != "" ]]
+            then
+                rgx+=" "
+            fi
+
+            rgx+="$SOURCE/$re"
+
+        done < $FILE
+
+    fi
+    echo "${rgx}"
+}
+
+prefetch () {
+
+        re_prefetch="" # regex of files to prefetch
+        prefetch_files="" # files to prefetch
+        re_all="${base_source}/*" # regex representing all files in directory
+        all_files="$(find -L ${re_all} -type f,l 2> /dev/null || true)" # all files in source mounts
+        sources_str="" # string containing all source paths
+
+        # load all potential regex from file (concatenate the regex to existing source mounts)
+        # load rgx of files that need to be prefetched
+        if [[ $re_prefetch != "" ]]
+        then
+            re_prefetch+=" "
+        fi
+        re_prefetch+=$(get_rgx ${base_source} $prefetch_file)
+
+        # if .sea_prefetchlist file contains regex
+        for rgx in $re_prefetch
+        do
+            if [[ $prefetch_files != "" ]]
+            then
+                prefetch_files+=" "
+            fi
+            prefetch_files+=$(echo $all_files | tr " " "\n" | grep -Eo $rgx)
+        done
+
+        # check to see if there are any candidate files for flush/eviction and process them
+        if [[ $prefetch_files != "" ]]
+        then
+            for f in $prefetch_files
+            do
+                echo ${f} ${sources_arr[0]}
+                subpath=${f#${base_source}}
+
+                for ((i=0;i<${n_lvls};i++)) {
+                    FILESIZE=$(stat -c%s "$f")
+                    AVAILSPACE=$(df --output=avail -B 1 "${sources_arr[0]}" | tail -n 1)
+
+                    if [[ ${FILESIZE} -le ${AVAILSPACE} ]]
+                    then
[…]

### 2021-10-10 tests/test_prefetch.sh

@@ -1,0 +1,72 @@
+#!/bin/bash
+
+export MOUNT="$PWD/mount"
+export SOURCE="$PWD/source"
+export SOURCE_1="$PWD/source_1"
+export SEA_LOG_FILE="${PWD}/sea.log"
+export SEA_HOME=${PWD}
+
+gen_conf() {
+cat > ${SEA_HOME}/sea.ini << DOC
+# Sea configuration
+[Sea]
+mount_dir = ${MOUNT} ;
+n_levels = 2 ;
+cache_0 = ${SOURCE} ;
+cache_1 = ${SOURCE_1}
+log_level = 0 ; # 4 crashes tests
+log_file = ${SEA_HOME}/sea.log ;
+max_fs = 1048576 ;
+DOC
+}
+
+setup () {
+	gen_conf
+}
+
+@test "get_sources" {
+	. bin/sea_prefetch.sh 0
+	get_sources
+	[[ ${sources_arr[@]} == ${SOURCE} ]]
+	[[ ${base_source} == ${SOURCE_1} ]]
+}
+
+@test "get_rgx" {
+	. bin/sea_prefetch.sh 0
+	echo ".*prefetch.*" > ${SEA_HOME}/.sea_prefetchlist
+	output=$(get_rgx ${SOURCE} ${SEA_HOME}/.sea_prefetchlist )
+
+	echo ${output}
+	[[ ${output} == "${SOURCE}/.*prefetch.*" ]]
+}
+
+@test "prefetch" {
+	. bin/sea_prefetch.sh 0
+	get_sources
+	f="${SOURCE_1}"/prefetch_f1.txt
+	f2="${SOURCE_1}"/bin/prefetch_f2.sh
+	f3="${SOURCE_1}"/subdir/prefetch_f3.txt
+	touch ${f} ${f2} ${f3}
+	
+	prefetch_process
+
+	[[ -f "${SOURCE}"/prefetch_f1.txt  && -f "${SOURCE}"/bin/prefetch_f2.sh && -f "${SOURCE}"/subdir/prefetch_f3.txt ]]
+}
+
+@test "prefetch symlinks" {
+	. bin/sea_prefetch.sh 0
+	get_sources
+
+	mkdir -p inputs
+	f=inputs/file1.txt
+	f2=inputs/file2.txt
+	echo "to prefetch" > $f
+	echo "to prefetch" > $f2
+	
[…]

### 2021-10-10 bin/sea_flusher.sh
some fixes to tests to make work with other OS, flush/evict/prefetch still needs work

@@ -27,12 +27,12 @@ fe_old () {
     task=$3 # mv / cp / rm (depending on which list)
     flist="" # list containing files for flush/eviction
 
-    for f in $found_files
+    for f in ${found_files}
     do
         # code adapted from here https://stackoverflow.com/a/28341234
         # Get current and file times
         CURTIME=$(date +%s)
-        FILETIME=$(stat $f -c %Y || true && continue)
+        FILETIME=$(stat $f -c %Y || true)
 
         if [[ $FILETIME == "" ]]
         then
@@ -51,17 +51,19 @@ fe_old () {
                 if [[ $f == *$s* ]]
                 then
                     subpath=${subpath#$s}
+                    subpath=$(echo $subpath | sed 's/.*\//\//' )
                     break
                 fi
             done
 
             if [[ -f ${base_source}${subpath} ]]
             then
-                BASETIME=$(stat ${base_source}${subpath} -c %Y || true && continue)
+                BASETIME=$(stat ${base_source}${subpath} -c %Y || true)
+                BASETIME=$(stat ${base_source}${subpath} -c %Y || true)
 
                 if [[ $BASETIME != NULL ]]
                 then
-                    TIMEDIFF=$((${FILETIME} - ${BASETIME}))
+                    TDIFF=$((${FILETIME} - ${BASETIME}))
 
                     if [[ ${TIMEDIFF} == 0 ]]
                     then
@@ -137,55 +139,62 @@ flush () {
         for s in "${sources_arr[@]}"
         do
             # load rgx of files that need to be flushed
-            if [[ $re_flush != "" ]]
+            if [[ "${re_flush}" != "" ]]
             then
                 re_flush+=" "
             fi
-            re_flush+=$(get_rgx $s $flush_file)
+            re_flush+="$(get_rgx $s ${flush_file})"
 
             # do the same for files that need to be pinned
-            if [[ $re_evict != "" ]]
+            if [[ "${re_evict}" != "" ]]
             then
                 re_evict+=" "
             fi
-            re_evict+=$(get_rgx $s $evict_file)
+            re_evict+="$(get_rgx $s ${evict_file})"
 
-            if [[ $re_all != "" ]]
+            if [[ "${re_all}" != "" ]]
             then
                 re_all+=" "
             fi
-            re_all+=$s/*
+            re_all+="$s/*"
         done
 
-        for rgx in $re_all
+        for rgx in ${re_all}
         do
-            if [[ $all_files != "" ]]
+            if [[ ${all_files} != "" ]]
             then
[…]

### 2021-10-10 bin/sea_prefetch.sh

@@ -37,31 +37,32 @@ prefetch () {
         re_prefetch="" # regex of files to prefetch
         prefetch_files="" # files to prefetch
         re_all="${base_source}/*" # regex representing all files in directory
-        all_files="$(find -L ${re_all} -type f,l 2> /dev/null || true)" # all files in source mounts
+        all_files="$(find -L ${re_all} -type f 2> /dev/null || true)" # all files in source mounts
+        all_files+="$(find -L ${re_all} -type l 2> /dev/null || true)" # all files in source mounts
         sources_str="" # string containing all source paths
 
         # load all potential regex from file (concatenate the regex to existing source mounts)
         # load rgx of files that need to be prefetched
-        if [[ $re_prefetch != "" ]]
+        if [[ "${re_prefetch}" != "" ]]
         then
             re_prefetch+=" "
         fi
-        re_prefetch+=$(get_rgx ${base_source} $prefetch_file)
+        re_prefetch+="$(get_rgx ${base_source} ${prefetch_file})"
 
         # if .sea_prefetchlist file contains regex
-        for rgx in $re_prefetch
+        for rgx in ${re_prefetch}
         do
-            if [[ $prefetch_files != "" ]]
+            if [[ "${prefetch_files}" != "" ]]
             then
                 prefetch_files+=" "
             fi
-            prefetch_files+=$(echo $all_files | tr " " "\n" | grep -Eo $rgx)
+            prefetch_files+=$(echo ${all_files} | tr " " "\n" | grep -Eo "${rgx}")
         done
 
         # check to see if there are any candidate files for flush/eviction and process them
-        if [[ $prefetch_files != "" ]]
+        if [[ "${prefetch_files}" != "" ]]
         then
-            for f in $prefetch_files
+            for f in ${prefetch_files}
             do
                 echo ${f} ${sources_arr[0]}
                 subpath=${f#${base_source}}

### 2021-10-10 bin/sea_flusher.sh
fix to flusher substitution

@@ -51,15 +51,15 @@ fe_old () {
                 if [[ $f == *$s* ]]
                 then
                     subpath=${subpath#$s}
-                    subpath=$(echo $subpath | sed 's/.*\//\//' )
+                    subpath=$(echo $subpath | cut -f 2- -d '/' )
                     break
                 fi
             done
 
-            if [[ -f ${base_source}${subpath} ]]
+            if [[ -f ${base_source}/${subpath} ]]
             then
-                BASETIME=$(stat ${base_source}${subpath} -c %Y || true)
-                BASETIME=$(stat ${base_source}${subpath} -c %Y || true)
+                BASETIME=$(stat ${base_source}/${subpath} -c %Y || true)
+                BASETIME=$(stat ${base_source}/${subpath} -c %Y || true)
 
                 if [[ $BASETIME != NULL ]]
                 then
@@ -75,18 +75,18 @@ fe_old () {
 
             if [[ $task != "rm" ]]
             then
-                echo "[Sea] Flusher: $task $f to ${base_source}${subpath}"
+                echo "[Sea] Flusher: $task $f to ${base_source}/${subpath}"
             fi
             # makedirectory in case it does not exist
-            mkdir -p $(dirname ${base_source}${subpath})
+            mkdir -p $(dirname ${base_source}/${subpath})
             # if we have some files that older that OLDTIME, flush/evict them
             # process each image individually (as opposed to all at once)
             if [[ $task == "mv" ]]
             then 
-                mv $f ${base_source}${subpath}
+                mv $f ${base_source}/${subpath}
             elif [[ $task == "cp" ]]
             then 
-                cp $f ${base_source}${subpath}
+                cp $f ${base_source}/${subpath}
             elif [[ $task == "rm" ]]
             then 
                 echo "[Sea] Flusher: $task $f"

### 2021-10-10 bin/sea_flusher.sh
fixed bug in test cases and flusher script

@@ -65,7 +65,7 @@ fe_old () {
                 then
                     TDIFF=$((${FILETIME} - ${BASETIME}))
 
-                    if [[ ${TIMEDIFF} == 0 ]]
+                    if [[ ${TDIFF} == 0 ]]
                     then
                         continue 
                     fi
@@ -254,6 +254,9 @@ flush () {
 
         flush_files=${tmp_flush}
         evict_files=${tmp_evict}
+
+        #echo "FLUSH ${flush_files}"
+        #echo "EVICT ${evict_files}"
 
         # check to see if there are any candidate files for flush/eviction and process them
         if [[ ${fe_files} != "" ]]

### 2021-10-10 bin/sea_flusher.sh
fixed some bugs

@@ -299,7 +299,6 @@ get_sources () {
     # refers to the last index and not the total number of sources
     n_lvls="$(($(awk -F "=" '/n_levels/ {print $2}' ${SEA_HOME}/sea.ini | tr -d ' ;') - 1))"
 
-    echo $n_lvls
     for ((i=0;i<${n_lvls};i++))
     do
         source_lvl=$(cat ${conf_file} | grep "^\s*cache_$i" | cut -d "=" -f 2 | tr -d ' ;')

### 2021-10-10 bin/sea_launch.sh

@@ -28,7 +28,8 @@ echo "[Sea] Launch: flusher pid" ${pid}
 trap 'trap - SIGTERM && kill $pid' SIGINT SIGTERM #EXIT
 
 echo "[Sea] Launch: Launching script $@"
-LD_PRELOAD=${parent_dir}/build/sea.so bash -c "$@"
+export LD_PRELOAD=${parent_dir}/build/sea.so
+eval "$@"
 echo "[Sea] Launch: Killing flush" && kill $pid
 
 wait $pid

### 2021-10-10 bin/sea_prefetch.sh

@@ -94,7 +94,6 @@ get_sources () {
     # refers to the last index and not the total number of sources
     n_lvls="$(($(awk -F "=" '/n_levels/ {print $2}' ${SEA_HOME}/sea.ini | tr -d ' ;') - 1))"
 
-    echo $n_lvls
     for ((i=0;i<${n_lvls};i++))
     do
         source_lvl=$(cat ${conf_file} | grep "^\s*cache_$i" | cut -d "=" -f 2 | tr -d ' ;')

### 2021-10-11 bin/sea_flusher.sh
fixed flusher/prefetch and added logging to stdout only when level 3+ is specified

@@ -13,6 +13,14 @@ conf_file=${SEA_HOME}/sea.ini
 flush_file=${SEA_HOME}/.sea_flushlist
 evict_file=${SEA_HOME}/.sea_evictlist
 tmpfile=$(mktemp /tmp/.sea_flush.inprog.XXXXXX)
+logging=0
+
+log () {
+    if [[ logging -ge 3 ]]
+    then
+        echo "[Sea] $@"
+    fi
+}
 
 fe_old () {
 
@@ -75,7 +83,7 @@ fe_old () {
 
             if [[ $task != "rm" ]]
             then
-                echo "[Sea] Flusher: $task $f to ${base_source}/${subpath}"
+                log "Flusher: $task $f to ${base_source}/${subpath}"
             fi
             # makedirectory in case it does not exist
             mkdir -p $(dirname ${base_source}/${subpath})
@@ -89,7 +97,7 @@ fe_old () {
                 cp $f ${base_source}/${subpath}
             elif [[ $task == "rm" ]]
             then 
-                echo "[Sea] Flusher: $task $f"
+                log "Flusher: $task $f"
                 rm $f || true
             fi
            
@@ -126,9 +134,9 @@ flush () {
         OLDTIME=$1 # user specified expiration time
         task=$2 # the name of the calling task. either "process" or "cleanup"
 
-        re_flush="" # regex of files to flush
-        re_evict="" # regex of files to pin to local disk (don't flush or evict)
-        re_all="" # regex for all files
+        re_flush=() # regex of files to flush
+        re_evict=() # regex of files to pin to local disk (don't flush or evict)
+        re_all=() # regex for all files
         flush_files="" # files to flush only
         evict_files="" # files to evict only
         fe_files="" # files to flush and evict
@@ -139,27 +147,21 @@ flush () {
         for s in "${sources_arr[@]}"
         do
             # load rgx of files that need to be flushed
-            if [[ "${re_flush}" != "" ]]
-            then
-                re_flush+=" "
-            fi
-            re_flush+="$(get_rgx $s ${flush_file})"
+            tmp_reflush=("$(get_rgx $s ${flush_file})")
+            reflush_arr=()
+            IFS=' ' read -a reflush_arr <<< "${tmp_reflush}"
+            re_flush=("${re_flush[@]}" "${reflush_arr[@]}")
 
[…]

### 2021-10-11 bin/sea_prefetch.sh

@@ -9,6 +9,14 @@ sources_arr=()
 n_sources=0
 conf_file=${SEA_HOME}/sea.ini
 prefetch_file=${SEA_HOME}/.sea_prefetchlist
+logging=0
+
+log () {
+    if [[ logging -ge 3 ]]
+    then
+        echo "[Sea] $@"
+    fi
+}
 
 get_rgx () {
     SOURCE=$1
@@ -34,7 +42,7 @@ get_rgx () {
 
 prefetch () {
 
-        re_prefetch="" # regex of files to prefetch
+        re_prefetch=() # regex of files to prefetch
         prefetch_files="" # files to prefetch
         re_all="${base_source}/*" # regex representing all files in directory
         all_files="$(find -L ${re_all} -type f 2> /dev/null || true)" # all files in source mounts
@@ -43,14 +51,13 @@ prefetch () {
 
         # load all potential regex from file (concatenate the regex to existing source mounts)
         # load rgx of files that need to be prefetched
-        if [[ "${re_prefetch}" != "" ]]
-        then
-            re_prefetch+=" "
-        fi
-        re_prefetch+="$(get_rgx ${base_source} ${prefetch_file})"
+        tmp_reprefetch=("$(get_rgx ${base_source} ${prefetch_file})")
+        reprefetch_arr=()
+        IFS=' ' read -a reprefetch_arr <<< "${tmp_reprefetch}"
+        re_prefetch=("${re_prefetch[@]}" "${reprefetch_arr[@]}")
 
         # if .sea_prefetchlist file contains regex
-        for rgx in ${re_prefetch}
+        for rgx in "${re_prefetch[@]}"
         do
             if [[ "${prefetch_files}" != "" ]]
             then
@@ -64,7 +71,6 @@ prefetch () {
         then
             for f in ${prefetch_files}
             do
-                echo ${f} ${sources_arr[0]}
                 subpath=${f#${base_source}}
 
                 for ((i=0;i<${n_lvls};i++)) {
@@ -73,7 +79,7 @@ prefetch () {
 
                     if [[ ${FILESIZE} -le ${AVAILSPACE} ]]
                     then
-                        echo "[Sea] Prefetch: Prefetching $f to ${sources_arr[i]}${subpath}"
+                        log "Prefetch: Prefetching $f to ${sources_arr[i]}${subpath}"
[…]

### 2021-10-11 bin/sea_flusher.sh
bugfix in prefetch test and made flusher and prefetch backwards compatible

@@ -150,13 +150,13 @@ flush () {
             tmp_reflush=("$(get_rgx $s ${flush_file})")
             reflush_arr=()
             IFS=' ' read -a reflush_arr <<< "${tmp_reflush}"
-            re_flush=("${re_flush[@]}" "${reflush_arr[@]}")
+            re_flush=(${re_flush[@]+"${re_flush[@]}"} ${reflush_arr[@]+"${reflush_arr[@]}"})
 
             # do the same for files that need to be pinned
             tmp_reevict="$(get_rgx $s ${evict_file})"
             reevict_arr=()
             IFS=' ' read -a reevict_arr <<< "${tmp_reevict}"
-            re_evict=("${re_evict[@]}" "${reevict_arr[@]}")
+            re_evict=(${re_evict[@]+"${re_evict[@]}"} ${reevict_arr[@]+"${reevict_arr[@]}"})
 
             re_all+=("$s/*")
         done

### 2021-10-11 bin/sea_prefetch.sh

@@ -54,11 +54,12 @@ prefetch () {
         tmp_reprefetch=("$(get_rgx ${base_source} ${prefetch_file})")
         reprefetch_arr=()
         IFS=' ' read -a reprefetch_arr <<< "${tmp_reprefetch}"
-        re_prefetch=("${re_prefetch[@]}" "${reprefetch_arr[@]}")
+        re_prefetch=(${re_prefetch[@]+"${re_prefetch[@]}"} ${reprefetch_arr[@]+"${reprefetch_arr[@]}"})
 
         # if .sea_prefetchlist file contains regex
         for rgx in "${re_prefetch[@]}"
         do
+            echo REGEX ${rgx}
             if [[ "${prefetch_files}" != "" ]]
             then
                 prefetch_files+=" "

### 2021-10-11 bin/sea_flusher.sh
another bugfix for empty arrays

@@ -172,7 +172,7 @@ flush () {
         done
 
         # if .sea_flushlist file contains regex
-        for rgx in "${re_flush[@]}"
+        for rgx in ${re_flush[@]+"${re_flush[@]}"}
         do
             if [[ ${flush_files} != "" ]]
             then
@@ -184,7 +184,7 @@ flush () {
         #echo "flush_files ${flush_files}"
 
         # if .sea_evictlist file contains regex
-        for rgx in "${re_evict[@]}"
+        for rgx in ${re_evict[@]+"${re_evict[@]}"}
         do
             if [[ ${evict_files} != "" ]]
             then

### 2021-10-11 bin/sea_prefetch.sh

@@ -57,7 +57,7 @@ prefetch () {
         re_prefetch=(${re_prefetch[@]+"${re_prefetch[@]}"} ${reprefetch_arr[@]+"${reprefetch_arr[@]}"})
 
         # if .sea_prefetchlist file contains regex
-        for rgx in "${re_prefetch[@]}"
+        for rgx in ${re_prefetch[@]+"${re_prefetch[@]}"}
         do
             echo REGEX ${rgx}
             if [[ "${prefetch_files}" != "" ]]

### 2021-10-11 bin/sea_flusher.sh
took regex array generation outside of for loop

@@ -15,6 +15,11 @@ evict_file=${SEA_HOME}/.sea_evictlist
 tmpfile=$(mktemp /tmp/.sea_flush.inprog.XXXXXX)
 logging=0
 
+# Regex arrays
+re_flush=() # regex of files to flush
+re_evict=() # regex of files to pin to local disk (don't flush or evict)
+re_all=() # regex for all files
+
 log () {
     if [[ logging -ge 3 ]]
     then
@@ -130,154 +135,155 @@ get_rgx () {
     echo "${rgx}"
 }
 
+assign_rgx() {
+
+    # load all potential regex from file (concatenate the regex to existing source mounts)
+    for s in "${sources_arr[@]}"
+    do
+        # load rgx of files that need to be flushed
+        tmp_reflush=("$(get_rgx $s ${flush_file})")
+        reflush_arr=()
+        IFS=' ' read -a reflush_arr <<< "${tmp_reflush}"
+        re_flush=(${re_flush[@]+"${re_flush[@]}"} ${reflush_arr[@]+"${reflush_arr[@]}"})
+
+        # do the same for files that need to be pinned
+        tmp_reevict="$(get_rgx $s ${evict_file})"
+        reevict_arr=()
+        IFS=' ' read -a reevict_arr <<< "${tmp_reevict}"
+        re_evict=(${re_evict[@]+"${re_evict[@]}"} ${reevict_arr[@]+"${reevict_arr[@]}"})
+
+        re_all+=("$s/*")
+    done
+}
+
 flush () {
-        OLDTIME=$1 # user specified expiration time
-        task=$2 # the name of the calling task. either "process" or "cleanup"
+    OLDTIME=$1 # user specified expiration time
+    task=$2 # the name of the calling task. either "process" or "cleanup"
 
-        re_flush=() # regex of files to flush
-        re_evict=() # regex of files to pin to local disk (don't flush or evict)
-        re_all=() # regex for all files
-        flush_files="" # files to flush only
-        evict_files="" # files to evict only
-        fe_files="" # files to flush and evict
-        all_files="" # all files in source mounts
-        sources_str="" # string containing all source paths
+    flush_files="" # files to flush only
+    evict_files="" # files to evict only
+    fe_files="" # files to flush and evict
+    all_files="" # all files in source mounts
+    sources_str="" # string containing all source paths
 
-        # load all potential regex from file (concatenate the regex to existing source mounts)
-        for s in "${sources_arr[@]}"
-        do
[…]

### 2021-10-11 bin/sea_prefetch.sh

@@ -11,6 +11,10 @@ conf_file=${SEA_HOME}/sea.ini
 prefetch_file=${SEA_HOME}/.sea_prefetchlist
 logging=0
 
+
+re_prefetch=() # regex of files to prefetch
+re_all=() # regex representing all files in directory
+
 log () {
     if [[ logging -ge 3 ]]
     then
@@ -40,26 +44,26 @@ get_rgx () {
     echo "${rgx}"
 }
 
+assign_rgx () {
+    re_all="${base_source}/*"
+
+    # load all potential regex from file (concatenate the regex to existing source mounts)
+    # load rgx of files that need to be prefetched
+    tmp_reprefetch=("$(get_rgx ${base_source} ${prefetch_file})")
+    reprefetch_arr=()
+    IFS=' ' read -a reprefetch_arr <<< "${tmp_reprefetch}"
+    re_prefetch=(${re_prefetch[@]+"${re_prefetch[@]}"} ${reprefetch_arr[@]+"${reprefetch_arr[@]}"})
+}
+
 prefetch () {
 
-        re_prefetch=() # regex of files to prefetch
         prefetch_files="" # files to prefetch
-        re_all="${base_source}/*" # regex representing all files in directory
         all_files="$(find -L ${re_all} -type f 2> /dev/null || true)" # all files in source mounts
         all_files+="$(find -L ${re_all} -type l 2> /dev/null || true)" # all files in source mounts
-        sources_str="" # string containing all source paths
-
-        # load all potential regex from file (concatenate the regex to existing source mounts)
-        # load rgx of files that need to be prefetched
-        tmp_reprefetch=("$(get_rgx ${base_source} ${prefetch_file})")
-        reprefetch_arr=()
-        IFS=' ' read -a reprefetch_arr <<< "${tmp_reprefetch}"
-        re_prefetch=(${re_prefetch[@]+"${re_prefetch[@]}"} ${reprefetch_arr[@]+"${reprefetch_arr[@]}"})
 
         # if .sea_prefetchlist file contains regex
         for rgx in ${re_prefetch[@]+"${re_prefetch[@]}"}
         do
-            echo REGEX ${rgx}
             if [[ "${prefetch_files}" != "" ]]
             then
                 prefetch_files+=" "
@@ -92,6 +96,7 @@ prefetch () {
 
 prefetch_process () {
     log "Prefetch: Starting prefetch thread"
+    assign_rgx
     prefetch
     log "Prefetch: Prefetch process completed"
 }

### 2021-10-11 bin/sea_flusher.sh
silence flush & prefetch errors

@@ -96,14 +96,14 @@ fe_old () {
             # process each image individually (as opposed to all at once)
             if [[ $task == "mv" ]]
             then 
-                mv $f ${base_source}/${subpath}
+                mv $f ${base_source}/${subpath} 2> /dev/null || true
             elif [[ $task == "cp" ]]
             then 
-                cp $f ${base_source}/${subpath}
+                cp $f ${base_source}/${subpath} 2> /dev/null || true
             elif [[ $task == "rm" ]]
             then 
                 log "Flusher: $task $f"
-                rm $f || true
+                rm $f 2> /dev/null || true
             fi
            
         fi

### 2021-10-11 bin/sea_prefetch.sh

@@ -85,7 +85,7 @@ prefetch () {
                     if [[ ${FILESIZE} -le ${AVAILSPACE} ]]
                     then
                         log "Prefetch: Prefetching $f to ${sources_arr[i]}${subpath}"
-                        cp -H $f ${sources_arr[i]}${subpath}
+                        cp -H $f ${sources_arr[i]}${subpath} 2> /dev/null || true
                         break
                     fi
                 }

### 2022-01-25 paper/sea-comp/paper.tex
paper updates

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{longtable}
\usepackage[binary-units=true]{siunitx}
\usepackage{multirow}
\usepackage{subcaption}
%\usepackage{algorithm}
\usepackage[ruled,norelsize]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{tabularx}

% for algorithm2e
\makeatletter
\newcommand{\removelatexerror}{\let\@latex@error\@gobble}
\makeatother

% \algnewcommand\algorithmicforeach{\textbf{for each}}
% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
% \algblock{Input}{EndInput}
% \algnotext{EndInput}
% \algblock{Output}{EndOutput}
% \algnotext{EndOutput}
% \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}: #1}}}

% benchmark on beluga, prefreesurfer with RUIS
% use more than 64 subjects for prefreesurfer --> few 100 subjects
% think more about how to benchmark I/O
% try distributed jobs with Sea

\begin{document}
\title{Sea: A lightweight data-management library for Big Data scientific computing}
% remove neuroimaging and replace with scientific computing  add user-space, switch HPC to computing clusters

\author{Val\'erie Hayot-Sasson, Mathieu Dugr\'e and Tristan Glatard}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
\IEEEtitleabstractindextext{%
\begin{abstract}
    The recent influx of scientific computing datasets have shifted scientific
computing from compute intensive to data intensive. Whereas many Big Data
Frameworks exist that minimize the cost of data transfers, few scientific
applications adopt data-management strategies as the effort required to update
the existing applications is significant.

We developed Sea as a means to enable data-management strategies for scientific
applications executing on HPC clusters without needing to update the underlying
applications. Sea leverages GNU C library interception to intercept
POSIX-compliant file system calls made by the applications. We designed a
performance model and evaluated the performance of Sea on a synthetic
data-intensive application processing a representative neuroimaging dataset (the
Big Brain). Our results demonstrate that Sea significantly improves performance,
up to a factor of 3$\times$.

\end{abstract}
}

% make the title area
\maketitle

\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{E}{fficient} data-management
strategies have become essential in minimizing the costs of processing Big Data.
Such strategies include data locality and in-memory computing, which minimize
transfers by ensuring compute occurs where the data is located and maintain data
in the fastest available storage (i.e. memory). Big Data frameworks, such as
MapReduce~\cite{dean2008mapreduce}, Apache Spark~\cite{zaharia2016apache} and
Dask~\cite{rocklin2015dask}, provide such data management mechanisms by default.

Until the recent surge in publicly available scientific data, scientific
workflows had been regarded as compute intensive. As a result, standardized
scientific tools typically lack these data-management mechanisms. The sheer
effort required to update the applications to leverage Big Data frameworks (e.g.
changing framework, programming language, training staff) can be a significant
deterrent. Furthermore, Big Data frameworks still require significant work to be
well adapted for the processing of different scientific data, such as imaging
data~\cite{mehta2017comparative}.

%multiple disk support
% pipeline agnostic and pipeline access patterns
% don't leverage workflows
% leverage existing filesystems
% replace storage devices with storage location.
% use applications
% memory management paragraph: PAGE CACHE
% shared filesystems are slow and metadata management can be slow

% add need for userspace / intelligent pipeline flushing
% there is no one-size fits all with data intensive applications
% due to io patterns, so we decided with userspace.

% add related work section

Our goal is to provide data-management strategies to scientific tools. We focus
on well-established scientific applications that operate as command-line tools.
These applications are written in a variety of different programming languages
and communicate data between tasks via a POSIX-compliant shared file system.

Scientific tools are typically platform-agnostic with the exception that the
file system must be POSIX-compliant. That is, they can run on local
workstations, high performance computing (HPC) clusters and cloud
infrastructure, with the exception that they cannot communicate with a
cloud-based storage natively. We focus our efforts on HPC clusters as they are
widely used in scientific computing and are currently used for the processing of
scientific Big Data.

Our goal is to provide data-management strategies to scientific tools. We focus
on well-established scientific applications that operate as command-line tools.
These applications are written in a variety of different programming languages
and communicate data between tasks via a POSIX-compliant shared file system.

Scientific tools are typically platform-agnostic with the exception that the
file system must be POSIX-compliant. That is, they can run on local
workstations, high performance computing (HPC) clusters and cloud
infrastructure, with the exception that they cannot communicate with a
cloud-based storage natively. We focus our efforts on HPC clusters as they are
widely used in scientific computing and are currently used for the processing of
scientific Big Data.

\section{Related Work}
\subsection{HPC Infrastructure}
      The general structure of HPC clusters may complicate the deployment of Big
      Data frameworks. Typical HPC clusters consist of distinct storage and
      compute nodes. While the compute nodes may also have local storage, there
      is no distributed file system like the Hadoop Distributed File System
      (HDFS)~\cite{shvachko2010hadoop} or Alluxio~\cite{alluxio} to facilitate
      data locality. Furthermore, access to these compute nodes is temporary,
      with allocation duration enforced by a batch scheduler. Therefore, any
      data written to a compute node is inaccessible after the allocation is
      terminated.

      The storage layer found on HPC clusters consists of a high-performance
      parallel file system (PFS) (e.g. Lustre~\cite{lustre}). These file systems
      are shared amongst all compute nodes within a cluster, typically connected
      to each node via high-performance network interconnect such as
      InfiniBand~\cite{infiniband}. I/O overheads of shared PFS can be incurred
      in many places, from the shared network, to the bandwidth of the storage
      disks.  With Lustre specifically, there typically are many data nodes, known as
      Object Storage Servers (OSS) which contain several storage devices, known
      as Object Storage Targets (OST). All file metadata is maintained within a
      separate node known as the Metadata Server (MDS) and stored within a
      device referred to as the Metadata Target (MDT). Although data transfers
      can be communicated directly to the corresponding OST, the clients need to
      first communicate with the MDS to determine which OST to communicate with.
      This can result in major data transfer overheads, particularly when making
      numerous requests to the metadata server.

      HPC clusters may provide a faster additional storage layer, known as a
      Burst Buffer, aimed at improving I/O performance. A Burst Buffer typically
      consists of fast storage (SSD and memory) and can be local to the compute
      node or as a dedicated I/O node. It was introduced as a response to offset
      the impacts of large-scale application checkpointing on the PFS~\cite{bb}. The
      application would write a checkpoint to the Burst Buffer and resume
      processing, while the checkpoint would be asynchronously flushed to the
      PFS.

\subsection{Big Data Frameworks and File systems}

      Strategies mitigate the impacts of data transfers in data-intensive workloads is
      required. In text-based processing, frameworks such as
      MapReduce~\cite{dean2008mapreduce} and Apache
      Spark~\cite{zaharia2016apache} have implemented and popularized strategies
      to minimize the need for data transfers during processing. These
      strategies are known as data locality and in-memory computing.

      Data locality is the process by which compute tasks are scheduled nearest
      to where the data is located. Traditionally, storage and compute were kept
      separate, having to transfer data over the network to compute. Rather than
      having to transfer large amounts of data over the network to the compute
      tasks, which could incur significant overheads, Big Data frameworks ensure
      that data is stored directly on the compute nodes. When a compute task
      requires access to specific data, the scheduler sends the task to the
      nearest available node to the data, thereby minimizing any cost of
      network-related data transfers. This strategy is not only used by Big Data
      Frameworks such as Hadoop MapReduce, Apache Spark, and
      Dask~\cite{rocklin2015dask}, but also enabled by file systems such as HDFS
      and Alluxio.

\subsection{The Linux page cache}
 While scientific applications uncommonly use Big Data frameworks and must
      perform large over-the-network data transfers, they may still benefit from
      in-memory computing and data locality through the Linux Page
      Cache~\cite{pagecache}. Similarly to other file systems, Lustre leverages
      the client node page cache to reduce I/O overheads. System memory is
      composed of two components: 1) anonymous memory and 2) page cache.
      Anonymous memory consists of all application-related objects, whereas page
      cache consists of recently accessed file data. When a file is read, that
      file is loaded up into the page cache to be flagged for eviction based on
      a least recently used (LRU) policy. That means subsequent accesses to that
      data may be done entirely in memory so long as that data has not already
      been evicted. Similarly, for writes to a file system with writeback cache
      enabled, the file will be written to memory completing the write operation
      once the file has been written entirely to memory. That file will then be
      flushed asynchronously to the appropriate storage device. As system memory
      may get overloaded with too many write requests, there is a limit to the
      amount of written data that can exist in memory, known as the
      \texttt{dirty\_ratio}. Furthermore, applications producing too many write
      requests may be throttled by the system.

\subsection{File System implementations}
      Due to the architecture of conventional HPC clusters, network-based
      parallel files systems are favoured over Big Data distributed file
      systems. Such file systems require super-user access, preventing users
      from deploying their own on-the-fly cluster. While HDFS can be loaded
      in user-space by mounting its File System in Userspace (FUSE)
      implementation, FUSE-based file systems can perform significantly worse
      than desired, depending on the application~\cite{tofuse}. Even with a
      user-space version of HDFS, there would be no mechanism to transfer all
      the data back to the parallel file system to be accessible post resource
      requests. Furthermore, such file systems are not POSIX-compliant, which
      would be required for scientific computing applications.

      There are, however, Big Data file systems, such as
      XtreemFS~\cite{xtreemfs} that exist entirely within user space, are
      POSIX-compliant, and can use alternative methods to FUSE to function.
      XtreemFS uses the LD\_PRELOAD trick to intercept file system calls made to
      the GNU C library (glibc). There are limitations to using glibc
      interception. For instance, the application must be dynamically linked and
      make glibc calls. However, statically-linked applications are uncommon and
      applications interacting with POSIX-compliant file systems on Linux
      machines, the predominant OS of HPC clusters, will make glibc calls.
      Nevertheless, there are alternatives to glibc intercept that can bypass
      the aforementioned issues, such as system call interception, but they
      result in greater overheads~\cite{quinson}.

      Characteristics that make XtreemFS unlikely to be deployed on-the-fly
      include complex file system deployment and configuration properties. It
      also does not provide a simple way to leverage different classes of
      storage devices, nor ensure that required data will be copied to the
      cluster's parallel file system.

      BurstFS~\cite{burstfs} and GekkoFS~\cite{gekkofs} are two user-space file
      systems that are specifically designed for Burst Buffers. That is, both of
      these file systems exist only during the time-window in which the
      resources are allocated. Furthermore, they both exist entirely within
      user-space using glibc interception. While these implementations simplify
      usage of burst buffers for users, their installation and configuration would be out of the
      scope of a typical HPC user and thus require sysadmin intervention.
\section{Materials and Methods}

\subsection{Sea design and implementation}

\begin{figure*}

    \centering
    \includegraphics[width=\columnwidth]{figures/seacomp-diagram.pdf}%
\caption{The Sea data-management library. Sea intercepts I/O-related glibc calls issued by the application and
translates the path provided if located within the Sea mountpoint. Sea leverages a user-defined storage hierarchy to help select the most appropriate path and the redirects the call to glibc to process the request on the translated path. }
\label{fig:sea-comp:diagram}
\end{figure*}

Sea\footnote{\url{https://github.com/valhayot/sea}} is an open-source user-space
data-management C++ library for applications running on HPC systems. Its main
aim is to reduce application data transfer costs by leveraging node-local
storage (Figure~\ref{fig:sea-comp:diagram}). Sea redirects files accessed from a user-specified mount point to the
appropriate storage devices using glibc interception. Within an intercepted call
the file will be written to or read from the fastest storage device available.

At minimum, Sea requires the user to specify at least two storage devices, a
fast temporary one (i.e. cache) and a slower long-term storage one. This could be RAM and SSD
if working on a single node, or a compute-local SSD and a shared parallel file
system, in the case of an HPC cluster. Ideally, a user will provide a multitude
of short-term storage devices to improve Sea's efficacy. Furthermore, to
maximize usage of the fastest available storage devices, Sea will allow the user
to outline which files can be removed from short-term storage in addition to
which files need to be materialized onto long-term storage. Eviction of files is
important as this will allow Sea to maximize usage of short-term storage.

Sea was meant as a data-management service to be employed at the discretion of
the users on scientific applications. When coming up with the idea for Sea, we
observed the design differences between Big Data and scientific, specifically
neuroimaging, frameworks and applications. Scientific frameworks tend to favour
ease-of-use, reproducibility, portability and parallelism over minimizing data
transfers. It was important for us to ensure that Sea would preserve the design
decisions of the underlying tools. Sea achieves this by having few dependencies,
is lightweight, and requires minimal configuration.

In this subsection, we will discuss and outline the various design and
implementation details made. Furthermore, we will go into detail into the
functionality of Sea.

\subsubsection{Requirements and assumptions}

Sea is primarily designed to work with workloads that generate large amounts of
intermediate data. Since data must be read from the slower storage device and
final output data most likely would have to be written to the same device, the
use of Sea with workloads that do not produce intermediate data would result in
limited speedup and may even introduce some overheads.

Sea provides two main modes based on flushing specification: In-memory computing
and flush-all. In-memory computing is achieved when intermediate need not be
materialized to long-term storage (i.e. no flushing). With Sea in-memory, we can
expect speedups comparable to Big Data frameworks.

As the name implies, Sea flush-all is achieved when all data must be
materialized to long-term storage (i.e. flush everything). This can be coupled
with eviction to limit the amount of data that will be written to slower, but
larger local storage devices. While this option will have overheads, due to the
need to flush all application data, performance can be improved when data
transfers and compute time are similar.

% Occasionally, it is desired to materialize even intermediate data to long-term
% storage devices, for potential future use. The performance gain from using Sea
% in these scenarios, however, is largely bounded by data size. In the instances
% where the workloads is more compute intensive, we can gain speedup by masking
% all or part of the I/O through long-term storage by executing the flush during
% a compute task. In data-intensive scenarios, the opposite can be seen: compute
% time can be masked by I/O (add figure + model). Despite the performance gain
% being limited in such a scenario, a performance gain can still be observed
% through the use of Sea.

One important assumption that must be made is that the amount of data produced
by the workload far exceeds the amount of page cache space available and
utilized by the different file systems. In the case where all workload data can
fit into page cache memory, the benefit of Sea may be greatly diminished if not
negligible.

Naturally, the amount of data written may not fit entirely into a single local
storage location. Currently, the user cannot specify the maximum amount of space
to use in a given mountpoint explicitly. Sea queries all the available file
systems directly to determine the amount of available space. It relies on
user-specified information on the maximum file size produced by the pipeline and
the number of concurrent pipeline parallel threads to determine if there is
sufficient space to write to the file system without issue. While it prevents
overheads due to locking, it makes the assumption that the user is alone writing
to the storage device. The user can reduce the amount of local storage space
used by increasing the maximum file size of the application.

%\todo{depending on the version of libc, the application static/dynamic linking,
%version of the function}

\subsubsection{Benefits}
Sea removes the need to implement the logic to redirect data to the appropriate
storage locations. Furthermore, it implements logic to ensure that files are
written to the best possible location at any given time. It is both pipeline and
infrastructure agnostic.

As it executes in user-space, root permissions are not required, enabling users
to use Sea on most Linux-based systems available to them. This differs from Big
Data file systems which may require elevated permissions to install. The
overhead of intercepting glibc calls is minimal, and negligible compared to
system call interception and file systems such as FUSE.

Unlike Big Data frameworks, Sea does not require reinstrumentation of the
existing pipelines, allowing users to gain an instant performance boost.

\subsubsection{Glibc interception}

Glibc interception is achieved by writing wrappers to existing glibc functions.
Most importantly, every glibc function accepting a file path needs to be
wrapped. The wrappers take any input filepath that is located within the
user-provided Sea mountpoint and convert it a filepath pointing to the best
available storage device.

Sea does not currently support the partitioning of files across multiple
devices. Since it cannot predict the size of the outputs to ensure the existence
of sufficient space on storage devices, the user must provide within the Sea
configuration file the maximum file size produced by the workflow. Together with
the specified amount of parallel processes, Sea calculates the minimum space
required on a storage device to write the file to it.
%\todo{should include an equation here}
Sea will then go through the hierarchy of available storage devices and select
the fastest storage device with sufficient available space. While this process
may result in not utilizing all available space in fast storage devices, it will
still allow users to gain a performance speedup on the pipeline given that the
number of threads multiplied by the file size does not exceed storage space.

\subsection{Testing}
To work effectively, Sea has to intercept every glibc function that interacts with files or
directories. Failure to intercept some of these function may result in the whole application
crashing. This is because only Sea can translate the Sea mountpoint file paths to their
respective real locations. Due to an absence of standardized glibc testing suites,
we developed our own integration tests.

Our tests have been executed on three different operating systems: Ubuntu (xenial, bionic, focal),
Fedora (32, 33, 34) and Centos (6, 7). Using this test matrix, the Sea tests were able to span seven glibc
versions ranging from 2.17 to 2.33. As glibc file system functions may change over time, Sea behaviour with
newer and older versions of glibc is undefined. 
% ldd (GNU libc) 2.17
% ldd (Ubuntu GLIBC 2.23-0ubuntu11.3) 2.23
% ldd (Ubuntu GLIBC 2.27-3ubuntu1.4) 2.27
% ldd (GNU libc) 2.28
% ldd (Ubuntu GLIBC 2.31-0ubuntu9.2) 2.31
% ldd (GNU libc) 2.31
% ldd (GNU libc) 2.32
% ldd (GNU libc) 2.33

\subsubsection{Flushing and eviction modes}
\begin{figure*}

    \centering
    \includegraphics[scale=0.8]{figures/sea-flushevict.pdf}%
\caption{The four Sea flush and eviction modes }
\label{fig:sea-comp:fe}
\end{figure*}
%\todo{add diagrams here?}
Flushing and eviction mechanisms give way to four different modes: flush-only,
evict-only, flush-and-evict and keep. A flush-only is synonymous with a
file copy in that a file that already exists within a Sea cache will be copied
to persistent storage. This operation is particularly useful when a file will be
reused by the application, and thus benefits from remaining in cache, but also
needs to be communicated between other nodes or needs to be persisted for
post-processing analysis. To flush-only a file, the file must only be specified
in the \texttt{.sea\_flushlist}.

Evict-only is equivalent to a remove operation. This mode is for files that are located
within a Sea cache, but do not need to be persisted and will not be reused by
the pipeline and therefore do not need to occupy cache space. Evict-only is
useful for removing unnecessary application log files. To evict-only a file, the
file must only be specified in the \texttt{.sea\_evictlist}.

Similar to the flush-only, the flush-and-evict mode copies data from cache to
persistent storage. However, the assumption here is that these files will not be
reused by the application and therefore need not occupy space in cache.
Flush-and-evict is therefore synonymous with a move operation and is invoked on
files that are specified in both the \texttt{.sea\_flushlist} and \texttt{.sea\_evictlist}.

Keep is useful for when file needs to be in cache as it will be reused by
the application, but is not needed in post-processing analysis or by other nodes.

%\subsubsection{Configuration File} \subsubsection{Program execution}
%\subsubsection{Limitations} behaviour unspecified in mixed static/dynamic
%workloads

\subsection{The Sea and Lustre model}\label{ss:sea-comp:model}

      Experimental results can be prone to error and variability, particularly in instances
      where there a many factors at play, such as in cluster environments. To improve
      confidence in our results, we developed a simplified performance model that
      could help describe the behaviours found in our experimental results. Since different
      parallel file systems may operate differently, our baseline model will be
      based on Lustre which is commonly used on HPC infrastructure.

      For data intensive use cases, the makespan models for both Lustre and Sea
      can be broken down into two components: The amount of time it takes read
      the data and the amount of time it takes to write the data. With more
      heterogeneous applications (some components are compute intensive whereas
      others are data intensive), a third component, consisting of compute time,
      can be added. Furthermore, latency may also play a significant role
      application makespan, particularly in scenarios with large amounts of
      small files. We choose to ignore latency costs in our model and make the
      assumption that the application bottleneck is the bandwidth, however, for
      more accurate estimates we might consider the addition of file system
      latency, as a fourth model component.

      A simplified version of the Lustre makespan model can be defined as
      follows:

      \begin{equation}\label{eq:sea-comp:lustrenpc}
          M_{l} =  \frac{D_{r}}{L_{r}} + \frac{D_{w}}{L_{w}}
      \end{equation}

      {\noindent}
      Where, \\
      $M_{l}$ is the application makespan using Lustre \\
      $D_{r}$ is the amount of data read \\
      $D_{w}$ is the amount of data written \\
      $L_{r}$ is Lustre's read bandwidth \\
      $L_{w}$ is Lustre's write bandwidth \\

    %   Where the makespan of an application reading and writing to Lustre,
    %   $M_{l}$, is the sum of the time it takes to read data of size $D_{r}$
    %   given the Lustre read bandwidth, $L_{r}$, and the time it takes to write
    %   the data of size $D_{w}$ given the Lustre write bandwidth, $L_{w}$.

      %\spacing{1} Where, \\
      %$M_{l}$ is the Lustre makespan \\
      %$D_{r}$ is the amount of data read \\
      %$B_{r}$ is the read bandwidth \\
      %$D_{w}$ is the amount of data written \\
      %$B_{w}$ is the write bandwidth \\

      %\spacing{1.5}
      To determine the Lustre bandwidth, one must consider the three components
  involved: 1) the network bandwidth of the compute nodes, 2) the network
  bandwidth of the data nodes, and 3) the collective bandwidth of the Lustre
  storage devices. Depending on each component's respective values, either of
  the three may be the source of a bottleneck. The Lustre bandwidth read and
  write models can therefore be described as follows:

    %   \begin{equation} %\label{eq:blr} L_{lr} = \min{(B_{n}N_{c}, B_{n}N_{d},
    %       B_{or}\min{(O, N_{c}N_{t})})} \end{equation}
    \begin{equation} %\label{eq:blr}
        L_{r} = \min{(cN, sN, d_{r}\min{(d, cp)})}
    \end{equation}

    and

    % \begin{equation}%\label{eq:blw} B_{lw} = \min{(B_{n}N_{c}, B_{n}N_{d},
    %     B_{ow}\min{(O, N_{c}N_{t})})} \end{equation}

    \begin{equation}%\label{eq:blw}
        L_{w} = \min{(cN, sN, d_{w}\min{(d, cp)})}
    \end{equation}

    {\noindent}
    Where, \\
    $N$ is the network bandwidth \\
    $c$ is the number of compute nodes used \\
    $s$ is the number of Lustre storage nodes \\
    $p$ is the number of parallel application processes \\
    $d_{r}$ is the read bandwidth of a single Lustre storage disk \\
    $d_{w}$ is the write bandwidth of a single Lustre storage disk \\

    % The network bandwidth of the compute node can be described as the number
    % of compute nodes available, $c$, multiplied by the bandwidth of the
    % network $N$. The network bandwidth of the storage nodes is calculated
    % similarly where the number of storage nodes, $s$, is multiplied by the
    % network bandwidth.

    % The collective bandwidth of the Lustre storage devices can be computed as
    % the bandwidth of a single Lustre disk ($d_{r}$ for read bandwidth and
    % $d_{w}$ for write bandwidth) multiplied by either the total number of
    % parallel processes writing to Lustre, $cp$, or the number of Lustre
    % storage disks, depending on which is limiting.

      %\spacing{1} Where, \\
      %$B_{lr}$ is the read bandwidth of Lustre \\
      %$B_{lw}$ is the write bandwidth of Lustre \\
      %$B_{or}$ is read bandwidth of the Lustre \gls{ost}s \\
      %$B_{ow}$ is the write bandwith of the Lustre \gls{ost}s\\
      %$B_{n}$ is the network bandwidth \\
      %$N_{c}$ is the number of compute nodes \\
      %$N_{d}$ is the number of data nodes \\
      %$n$ is the number of threads per compute node \\

      %\spacing{1.5}
      For the sake of simplicity, the above models assume that the network
      bandwidth between the compute and data nodes is the same. This,
      however, may not necessarily be the case. Furthermore, the model also
      assumes that each file can only be located on a single disk, meaning that
      the parallel bandwidth can at maximum be as fast as all Lustre disks
      combined and as slow as the minimum number of compute threads reading and
      writing files.

      As with many file systems, page cache plays an important role in the speed
      of application read and writes in Lustre. Since the effects of page cache
      may be non-negligible given amount of memory available and the data
      accessed during the execution of the application, it is important to
      include it in our model. The makespan of an application I/O to and from
      page cache, $M_{c}$, can be described as the following:
      Equation~\ref{eq:sea-comp:lustrenpc} where it is assumed that none of the
      data is written or read from page cache.

      \begin{equation}\label{eq:sea-comp:cache}
          M_{c} = \frac{D_{cr}}{cC_{r}} + \frac{D_{cw}}{cC_{w}}
      \end{equation}

      {\noindent}
      Where, \\
      $M_{c}$ is the makespan of application I/O to page cache \\
    %   $c$ is the number of compute nodes \\
      $D_{cr}$ is the amount of data read from cache \\
      $D_{cw}$ is the amount of data written to cache \\
      $C_{r}$ is the page cache read bandwidth \\
      $C_{w}$ is the page cache write bandwidth \\

    %   Or in other words, the makespan of page cache I/O is the sum of the
    %   amount time it takes to read from data cache of size $D_{cr}$ given the
    %   parallel cache read bandwidth $C_{r}$ of all compute nodes $c$, and the
    %   amount of time to write to cache data of size $D_{cw}$ given the
    %   parallel cache write bandwidth $C_{w}$ across all compute nodes.

      %\spacing{1} Where, \\
      %$M_{c}$ is the makespan of writing to cache \\
      %$D_{cr}$ is the amount of data read from cache \\
      %$D_{cw}$ is the amount of data written to cache \\
      %$N_{c}$ is the number of compute nodes \\
      %$B_{mr}$ is the memory read bandwidth \\
      %$B_{mw}$ is the memory write bandwidth \\

      %\spacing{1.5}
       As each individual compute node has its own set of memory, we treat the
      total memory bandwidth as the sum of the individual memory bandwidth of
      each compute node.

      Page cache is difficult to summarize accurately and effectively within a
      model. For one, we must not only consider available memory and anonymous
      memory used by the application, but we must also consider which pages are
      candidates for eviction and which files they belong to. In addition, in
      the case of writes, we must consider asynchronous flushing and the
      throttling that may occur as a consequence of surpassing the
      \texttt{dirty\_ratio}. Furthermore, Lustre also has its own user-defined
      settings for how it interacts with the cache that would add additional
      complexities to the model. As a result, we assume two possible scenarios,
      one in which page cache is never used
      (Equation~\ref{eq:sea-comp:lustrenpc}) and one in which all application
      I/O occurs entirely within page cache, excluding the first read which must
      occur on Lustre (Equation~\ref{eq:sea-comp:lustrepc}). These two models
      allow us to define the bounds of Lustre's performance.

      \begin{equation}\label{eq:sea-comp:lustrepc}
          M_{lc} = \frac{D_{I}}{L_{r}} + M_{c}
      \end{equation}

      {\noindent}
      Where, \\
      $M_{lc}$ is the application makespan using Lustre with page cache \\
      $D_{I}$ is the amount of input data \\
      $L_{r}$ is the Lustre read bandwidth \\
      $M_{c}$ is the application makespan on page cache \\

    %   Where the makespan of an application using Lustre with page cache,
    %   $M_{lc}$, is the sum of the amount of time it takes to read the initial
    %   data $D_{I}$ given the Lustre bandwidth $L_{r}$ with the makespan of the
    %   time spent reading and writing to page cache $M_{c}$.

      %\spacing{1} Where, \\
      %$M_{lc}$ is the Lustre makespan with page cache \\
      %$D_{i}$ is the amount of input data \\
      %$B_{lr}$ is the Lustre bandwidth \\
      %$M_{c}$ is the makespan of the I/O to page cache \\

      %\spacing{1.5}
      Sea's model is more complex than Lustre's as there can be several layers
      of different devices. For instance, Sea's model can be defined as:

      \begin{equation}\label{eq:sea-comp:sea}
          M_{S} = \frac{D_{I}}{L_{r}} + M_{1} + \cdots + M_{n}
      \end{equation}

      Here, $M_{n}$ represents the makespans of the different possible storage
      levels (e.g. tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
      makespan of an application reading and writing to Sea. For our model, we
      will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
      storage, and 3) slow parallel file system layer.

      Since the modelling of page cache is even more challenging with Sea due to
      the additional tmpfs and SSD layer, we will will model the upper and lower
      performance bounds, as we did with Lustre. Using the three layers and
      disregarding any possible effects of caching, we can redefine the Sea
      model to be:

      \begin{equation}\label{eq:sea-comp:snc}
          M_{S} = M_{SL} + M_{Sg} + M_{St}
      \end{equation}

      Where, \\
      $M_{S}$ is the makespan of an application using Sea \\
      $M_{SL}$ is the makespan of the Sea I/O to Lustre \\
      $M_{Sg}$ is the makespan of the Sea I/O to local disks \\
      $M_{St}$ is the makespan of the Sea I/O to tmpfs \\

    %      Where $M_{Sl}$ represents the Lustre component of the Sea makespan,
    %   and $M_{Sg}$ and $M_{St}$ represent the disk and tmpfs component of the
    %   Sea makespan, respectively.

      The tmpfs component of the Sea makespan can be defined as the amount of
      data that can be written ($D_{tw}$) to and read ($D_{tr}$) from tmpfs over
      its respective bandwidths ($C_{r}$ and $C_{w}$). In other words:

      \begin{equation}\label{eq:sea-comp:mst}
          M_{st} = \frac{D_{tr}}{cC_{r}} + \frac{D_{tw}}{cC_{w}}
      \end{equation}

      \begin{equation*}\label{eq:sea-comp:dtr}
          D_{tr} = \min\left(D_{m}, \max{\left(c(t - pF), 0 \right)} \right)
      \end{equation*}
      \begin{equation*}\label{eq:sea-comp:dtw}
          D_{tw} = \min\left(D_{m} + D_{f}, \max{\left(c(t - pF), 0 \right)} \right)
      \end{equation*}

      Where,\\
    %   $M_{st}$ is the makespan of Sea I/O to tmpfs \\
      $D_{tr}$ is the amount of data read from tmpfs \\
      $D_{tw}$ is the amount of data written to tmpfs \\
      $D_{m}$ is the amount of intermediate data \\
      $D_{f}$ is the amount of final output data \\
    %   $C_{r}$ is the tmpfs read bandwidth \\
    %   $C_{w}$ is the tmpfs write bandwidth \\
    %   $c$ is the number of compute nodes \\
      $t$ is the amount of available space on tmpfs \\
    %   $p$ is the number of parallel application processes \\
      $F$ is the size of a single file \\

      In an optimal scenario all intermediate data and final output data would
      fit in tmpfs. This would provide an application using Sea in-memory
      performance. However, due to limited tmpfs storage space, it is unlikely
      to be the case. In addition, Sea may further restrict available storage
      space to prevent exceeding tmpfs storage by ensuring that there is at
      least sufficient space for all processes to each write a file
      concurrently.

      The local disk  makespan model is similar to the tmpfs makespan model,
      however, we must ensure to disregard any data that has already been
      written to tmpfs. Furthermore, in Sea, it is possible to leverage however
      many disk-based file systems are available for use ($g$). For our model,
      we assume that the size of each device is identical. The makespan model
      can be defined as follows:

      \begin{equation}\label{eq:sea-comp:msd}
          M_{Sg} =  \frac{D_{gr}}{gcG_{r}} + \frac{D_{gw}}{gcG_{w}}
      \end{equation}

      \begin{equation*}\label{eq:sea-comp:ddr}
          D_{gr} = \min{(D_{m} - D_{tr}, \max{(c(gr - pF),0)})}
      \end{equation*}

      \begin{equation*}\label{eq:sea-comp:ddw}
          D_{gw} = \min{(D_{m} + D_{f} - D_{tw}, \max{(c(gr - pF),0)})}
      \end{equation*}

      Where, \\
    %   $M_{Sg}$ is the makespan of the Sea disk component \\
      $D_{gr}$ is the amount of input data read from local disk \\
      $D_{gw}$ is the amount of data written to local disk \\
    %   $D_{m}$ is the amount of intermediate data \\
    %   $D_{f}$ is the final output data \\ $D_{tr}$ is the amount of data read
    %   from tmpfs \\
    %   $D_{tw}$ is the amount of data written to tmpfs \\
      $g$ is the number of local disks on a compute node \\
      $r$ is the amount of disk space available on a given disk \\
    %   $c$ is the number of compute nodes \\
    %   $p$ is the number of application parallel processes \\
    %   $F$ is the size of a single output file \\
      $G_{r}$ is the read bandwidth of the disks \\
      $G_{w}$ is the write bandwidth of the disks \\

      The final component of the Sea model is the Lustre component
      (Eq.~\ref{eq:sea-comp:msl}). Sea's Lustre makespan model consists of the
      initial read from Lustre and includes and data that must be written to
      Lustre due to insufficient space on local storage and the makespan to read
      the intermediate data from Lustre.

      \begin{equation}\label{eq:sea-comp:msl}
          M_{SL} = \frac{D_{I}}{L_{r}} + \frac{D_{Lr}}{L_{r}} + \frac{D_{Lw}}{L_{w}}
      \end{equation}
      \begin{equation*}\label{eq:sea-comp:dlr}
          D_{lr} = D_{m} - D_{gr} - D_{tr}
      \end{equation*}
      \begin{equation*}\label{eq:sea-comp:dlw}
          D_{lw} = D_{m} + D_{f} - D_{gw} - D_{tw}
      \end{equation*}

      Where, \\
    %   $M_{SL}$ is the makespan of Sea's I/O to Lustre \\
    %   $D_{I}$ is the amount of application input data \\
      $D_{Lr}$ is the amount of data read to Lustre by Sea \\
      $D_{Lw}$ is the amount of data written to Lustre by Sea \\
    %   $D_{m}$ is the amount of intermediate data \\
    %   $D_{f}$ is the final output data \\
    %   $D_{gr}$ is the amount of data read from local disk by Sea \\
    %   $D_{gw}$ is the amount of data written to local disk by Sea \\
    %   $D_{tr}$ is the amount of data read from tmpfs by Sea \\
    %   $D_{tw}$ is the amount of data written to tmpfs by Sea \\
    %   $L_{r}$ is Lustre's read bandwidth \\
    %   $L_{w}$ is Lustre's write bandwidth \\

      Sea and Lustre have an identical lower bound. That is, ideally, both must
      perform the first read from Lustre, but all subsequent data accesses can
      be performed entirely within the page cache. The page cache model for Sea
      can be defined as the following:

      \begin{equation}\label{eq:sea-comp:msc}
          M_{Sc} = \frac{D_{I}}{L_{r}} + \frac{D_{m}}{cC_{r}} + \frac{D_{m} + D_{f}}{cC_{w}}
      \end{equation}
\subsection{Experiments}
\subsubsection{Application}

      To evaluate the Lustre and Sea models defined in
      Subsection~\ref{ss:sea-comp:model} and the real performance of both file
      systems with data intensive applications, we wrote a simple Python
      application based off of Algorithm~\ref{alg:sea-comp:incrementation}.

      Using this application, we can easily control how much intermediate data
      is produced by altering the amount of iterations required. Although our
      model should be able to support images of different sizes, we wanted to
      minimize any possible scheduling effects from our experiments. Therefore,
      each application processes the same amount of input data, produces the
      same amount of intermediate and output data and performs the same amount
      of computation.

      \begin{figure}[!t]
        \removelatexerror
        \begin{algorithm}[H]
          \caption{Incrementation}\label{alg:sea-comp:incrementation}
          \SetAlgoLined 
            \SetKwInOut{Input}{Input}
            \Input{$n$ number of iterations;\\
                $C$ a set of image chunks;\\
                $fs$ the file system to write to (Lustre or Sea)\\
            }
            \ForEach{$chunk \in C$}
            { 
              read $chunk$ from Lustre 
              \For{$i \in [1,n]$}
              { $chunk\gets chunk+1$ save $chunk$ to $fs$ }
            }
        \end{algorithm}
      \end{figure}

      We use the BigBrain~\cite{amunts2013bigbrain}, a one-of-a-kind histology
      dataset of the human brain, as a representative scientific dataset. For
      all our experiments, we utilize the \SI{20}{\micro\meter} dataset, which
      totals to approximately \SI{603}{\gibi\byte}. The dataset was broken down
      into 1000 files each consisting of \SI{617}{\mebi\byte} of data.
      %of data as well as
      %5000 blocks each \SI{124}{\mebi\byte} in size.

      We evaluated Sea using 4 different experimental conditions: 1) varying the
      number of nodes, 2) varying the number of disks, 3) varying the number of
      threads, and 4) varying the number of iterations. Experimental condition 1
      tests the effects of increasing concurrent accesses to Lustre while fixing
      disk parallel threads. Condition 2 varies disk contention while fixing
      contention to Lustre, whereas 3 tests the effects of contention on both
      Lustre and local storage. Experimental condition 4 varies the total amount
      of intermediate data produced by the application. The fixed conditions for
      the experiment were 5 nodes, 6 processes, 6 disks, 10 iterations and 1000
      blocks. Sea in-memory (flushing and eviction of only the last iteration of
      files) strategy was used for all conditions.

      We also evaluated Sea flush-all (flush all files to Lustre) using 5 nodes,
      64 processes, 6 disks and 5 iterations using the same incrementation
      application.

\subsubsection{Infrastructure}
      \begin{table}
      \centering
      \begin{tabular}{@{}|c|c|c|@{}}
       \hline
       Storage layer & Action & Avg. bandwidth (MiB/s) \\
              \hline
       \multirow{3}{*}{tmpfs} & read & 6676.48 \\
       & cached read & 6318.08  \\
       & write & 2560.00 \\
       \hline
       \multirow{3}{*}{local disk} & read & 501.70  \\
       & cached read & 7034.88 \\
       & write & 426.00 \\
       \hline
       \multirow{3}{*}{Lustre} & read & 1381.14 \\
       & cached read & 6103.04  \\
       & write & 121.00  \\

       \hline

      \end{tabular}
      \caption{Storage benchmarks}
      \label{table:sea-comp:fs}
      \end{table}
            Our experiments were executed on a Centos 8.1 (Linux kernel 4.18.0)
      cluster with 8 compute nodes, a 4 data node Lustre server with 1 dedicated
      metadata node. Each compute node is equipped it two Intel(R) Xeon(R) Gold
      6130 CPUs, \SI{250}{\gibi\byte} of memory with \SI{126}{\gibi\byte} of
      tmpfs space and 6 \SI{447}{\gibi\byte} Intel SSDSC2KG480G8R SSDs. The data
      nodes each contain 11 \SI{10}{\tera\byte} HGST HUH721010AL HDD OSTs and
      \SI{62}{\gibi\byte} memory. The metadata server contains a Toshiba
      KPM5XVUG960G \SI{960}{\giga\byte} MDT. The network bandwidth is
      \SI{25}{\giga\bit}E and uses TCP for communication. Jobs are scheduled on
      the cluster from a controller node using Slurm with cgroups. Swapping is
      disabled and Lustre dirty writes is limited to \SI{1}{\giga\byte} per OST.

      Each file system was benchmarked using \texttt{dd} using 5 repetitions.
      The average bandwidths are reported in Table~\ref{table:sea-comp:fs}

%Big Brain incrementation fmri processing?
\section{Results}

    \begin{figure*}

    \begin{subfigure}{\columnwidth}
        \centering
        \captionsetup{width=.85\linewidth}
        \includegraphics[width=\columnwidth]{figures/nodes.pdf}%
        \caption{Experiment 1: Varying the number of nodes, 10
        iterations}\label{fig:sea-comp:nodes}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \captionsetup{width=.85\linewidth}
        \includegraphics[width=\linewidth]{figures/disks.pdf}
        \caption{Experiment 2: Varying the number of disks, 5
        iterations}\label{fig:sea-comp:disks}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \captionsetup{width=.85\linewidth}
        \includegraphics[width=\linewidth]{figures/iterations.pdf}
        \caption{Experiment 3: Varying the number of
        iterations}\label{fig:sea-comp:iterations}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \captionsetup{width=.85\linewidth}
        \includegraphics[width=\linewidth]{figures/threads.pdf}
        \caption{Experiment 4: Varying the number of parallel processes, 5
        iterations}\label{fig:sea-comp:processes}
    \end{subfigure}
    \caption{Benchmarks comparing Lustre and Sea In-memory performance under
    various conditions using the incrementation application on the Big Brain
    image. All experiments were repeated 5$\times$. Model bounds represented by coloured regions}
    \label{fig:sea-comp:benchmarks}
    \end{figure*}

      \subsection{Sea achieves significant speedups}

      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea with the
      in-memory configuration speeds up execution the majority of the
      conditions. The largest speedups observed can be seen in
      Figure~\ref{fig:sea-comp:processes} at 32 processes, where the speedup is
      nearly 3$\times$. We believe that the speedup here is primarily due to the
      contention on Lustre. While each local disk has an average of 5 processes
      trying to write concurrently to it at 32 processes, each Lustre has around
      3 concurrent processes while to it. Whereas one might assume this means
      Lustre would have less contention overall and should perform better, there
      are significant bandwidth differences to account for between the local
      disks and Lustre (see Table~\ref{table:sea-comp:fs}) that would result in
      improved local disk performance despite increased contention. Furthermore,
      Lustre has a centralized metadata server that is responsible for
      determining which OST is assigned to each block, guaranteeing a certain
      amount of load-balance on the storage disks. This is in contrast with the
      local disks, who do not rely on a metadata server and are selected by Sea
      via a random shuffling.

      The experimental conditions with the next largest speedup can be seen in
      Figure~\ref{fig:sea-comp:iterations} at 10 iterations with a 2.6$\times$
      speedup. It is believed that Sea performs best at 10 iterations, rather
      than 15, because that is when the majority of the writes must be made to
      local disk. While Sea is not expected to surpass the Lustre makespan,
      Lustre does have a slight performance advantage in that is able to evict
      data once it is persisted to Lustre, allowing it to make more efficient
      use of memory whenever possible.

      When varying the number of nodes (Figure~\ref{fig:sea-comp:nodes}), we
      achieved the greatest speedup at 5 nodes, with a speedup of 2.4$\times$.
      Similar to our experiments with multiprocessing, the speedup here appears
      to be due to increased contention on Lustre. However, there is one main
      difference here, and that is that only Sea is experiencing increased
      contention, as the contention within the compute nodes is fixed. Both the
      model and experimental results state that the speedup is approaching a
      plateau, however, it is expected that once the number of threads writing
      to Lustre exceeds the number of OSTs (e.g. 9 nodes), we will observe an
      even greater speedup from Sea due to the increased contention on Lustre.

      As expected, our results demonstrate that with an increase in local disk,
      we can achieve greater speedups. Our results
      (Figure~\ref{fig:sea-comp:disks}) show that at 6 disks, we achieve a
      speedup of 2$\times$. This is natural, due to the fact that the most disks
      we have, the less contention there will be on any given disk. In
      particular, in our case, each disk at 6 disks should optimally only have a
      single process writing to it.

      When Sea does not provide speedups, it either performs similarly to
      Lustre, as can be seen in Figure~\ref{fig:sea-comp:nodes} at one node and in
      Figure~\ref{fig:sea-comp:iterations} at 1 iteration, or sightly slows down
      execution, as can be seen in~\ref{fig:sea-comp:disks}. As expected, Sea at
      a single iteration can at best perform similarly or slightly worse than
      Lustre. This is because all the data is read from Lustre and written to
      Lustre, operating in the same way that Lustre would with page cache.

      Sea at a single node likely performs equivalently to Lustre because
      Lustre, in this case, has very good bandwidth, having at most 6 concurrent
      writes to the whole file system. Furthermore, due to the limited number of
      concurrent writes on Lustre at any given time, Lustre needs to wait less
      for writes to flush to disk, making better use of available page cache
      space. Sea's performance, in contrast, is negatively impacted as there is
      more data being written to local storage and the combined bandwidth of the
      local storage is far less superior to that of Lustre's, due to only having
      6 disks available.

      The contention on the disk at Sea operation with only a single local disk
      is likely the cause of the decline in performance in
      Figure~\ref{fig:sea-comp:disks}. As Lustre has significantly less
      contention on the file system, it is able to exhibit a superior
      performance to Sea. As we can see, increasing the number of local disks
      improves Sea's performance beyond what Lustre can achieve.

      \subsection{Model accurately predicts performance trends}

      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model correctly describes The
      bounds for all but two experiments: Experiment 3
      (Figure~\ref{fig:sea-comp:iterations}) and Experiment 4
      (Figure~\ref{fig:sea-comp:processes}). In Experiment 3, the model
      incorrectly represents the bounds for 1 iteration. Since, in this case, all
      the data should be able to fit in page cache for both Sea and Lustre, it
      is possible that memory bandwidth has been overestimated, resulting in
      incorrect model predictions.

      We believe that Lustre exceeds the model bounds in Experiment 4 due to
      increased contention on the file system. While our model predicts that the
      disk bandwidth will be the bottleneck, thus plateauing at 9 parallel
      processes per node, there are other Lustre bottlenecks that are not
      included in the model, like the metadata server. We expect that there were
      too many incoming requests to the server at 30+ parallel processes, that
      performance declined above model bounds.

    \begin{figure*}

        \centering
        \includegraphics{figures/flushall.pdf}%
        \caption{Comparison of the different Sea modes with Lustre on the
        incrementation application processing 1000 Big Brain blocks with 5
        nodes, 5 iterations, 6 disks and 6 parallel processes.}
    \label{fig:sea-comp:flush}
    \end{figure*}

      \subsection{Significant overheads with Flush-all mode}

      Figure~\ref{fig:sea-comp:flush} demonstrates the overhead that can be
      obtained by using flush-all when there is no compute to mask the flushing.
      Not only is Sea flush-all 3.5$\times$ slower than Sea in-memory, it is
      1.3$\times$ slower than Lustre. The reason for it being slower than Lustre
      is that Sea flush-all does not evict any files and had to copy files from
      local disk to Lustre, as a result. In contrast, Lustre does not make use
      of local disk space. It is able to evict data from memory once it is
      materialized to Lustre. Consequently, Lustre alone does not have the third
      overhead of writing the data to disk like Sea flush-all does. The
      performance loss is likely to not have been discernible if we had compute
      time that matched data transfer time, however, the performance gain might
      not have been as significant unless we were writing in parallel data that
      far exceeded page cache space as the application would not have to wait
      for data to be flushed to Lustre before proceeding.

\section{Discussion}

    \subsection{Lightweight data-management library for CLI applications}

    Through the interception of glibc calls, Sea successfully manages to
    redirect I/O to the different available storage devices on an HPC cluster.
    Since Sea is lightweight and requires minimal configuration, it preserves
    the characteristics that are important to scientific applications.

    At minimum, Sea requires the specification of a configuration file for it to
    work. A user would need to know details on the cluster storage that can be
    leveraged as well as approximate details on how much data an application
    execution can produce at any given time. Due to the simplicity of the
    configuration file, Sea maintains the ease-of-use requirement for scientific
    applications.

    Sea maintains whole files, as produced by the pipeline. In no instance does
    it modify or alter the data produced by the pipeline. Since there is no risk
    of Sea modifying the contents of the data, Sea preserves the reproducibility
    requirement.

    The flush-and-evict process may affect overall application parallelism. It
    would depend on how many Sea processes are started on the compute node. If
    only a single instance of Sea is called on a compute node, there will only
    be a single flush and evict process, and therefore, only a single flush and
    evict process will exist. However, if Sea is launched many times on a given
    node, there will be many flush and evict processes which may interfere with
    the application compute. However, our results do not indicate that
    performance is significantly impacted by the presence of a single flush and
    evict process.

    Sea cannot be used with operating systems
    that are not Linux-based, and therefore, the use of Sea limits application
    portability. However, HPC systems commonly use Linux operating systems,
    thus, it is not limiting for its intended use case. Furthermore, we provide
    publicly available containers to simplify the usage of Sea on various other
    operating systems.

    \subsection{In-memory performance with Sea}

    Our results indicate that Sea can significantly improve the performance in
    applications executing data-intensive workflows. In all experiments, we
    observed speedups of up to 3$\times$, with the majority of cases reaching a
    2$\times$ speedup. In very few cases did the use of Sea not result in any
    speedups. In two of the three scenarios, Sea performed identically to
    Lustre, either because Sea was issuing the same amount of data transfers to
    Lustre or Lustre bandwidth far exceeded what was available locally.

    On shared cluster environments, such as though found in high-performance
    computing clusters, it is less likely that the PFS would be so underused
    that it could achieve better performance than leveraging local storage, even
    if local storage was limited. We therefore believe that it is very unlikely
    obtain no speedups from using Sea on a production HPC cluster as long as the
    application is data intensive.

    For our experiments we relied on a synthetic data-intensive application,
    however, the I/O patterns exhibited by such an application do not adequately
    mimic the patterns of scientific applications. In our experiments, we
    demonstrate what the possible performance upper bound can look like and how
    it is affected by various different factors. Despite the fact that typical
    scientific workflows may not be as data-intensive, it is believed that
    scientific applications can still benefit a significant amount by limiting
    writes to a shared PFS that is experiencing traffic generated by many other
    users. Furthermore, given that scientific applications have more compute,
    Sea flush-all can be used with reduced overheads.

    \subsection{Performance boost with local storage availability}

    As expected, Sea's performance increases with the number of disks available.
    However, our results also demonstrate that it does not require that many
    disks to surpass Lustre's performance, even when Lustre is underutilized.
    Therefore, while Sea underperformed at a single disk, it is likely that this
    will be sufficient to experience speedups even in a production environment,
    where Lustre has to deal with user traffic across the cluster. This is
    important to note because it is not uncommon for HPC cluster compute nodes
    to only have a single disk available as burst buffer. However, should more
    disks be available, it would be best to include as many as possible in Sea.

    \subsection{Sea preferred when Lustre is overloaded}

    In instances where contention on Lustre exceeds that of local disks, Sea
    in-memory outperforms Lustre. In all cases, it did not take very much for
    Sea to outperform Lustre. It is expected that the resource requirements of
    scientific applications running on HPC systems could far exceed those of our
    current experiments, relying on 100s of nodes instead of just 5. Therefore,
    data-intensive scientific applications would benefit greatly from using Sea.

    When Lustre is not overloaded, there is little benefit to using Sea.
    However, we found that Sea performs similarly to Lustre in these cases.
    Since there is no real penalty to using Sea when Lustre is not overloaded,
    it is recommended to use Sea in all scenarios. Moreover, users on HPC
    clusters are unaware of what the PFS performance will be at the time when
    their experiments will be scheduled. Knowing that Sea does not incur
    significant overheads will allow users to freely execute Sea without
    hesitation.

    \subsection{Flush when necessary and evict often}

    The results demonstrate that flushing all the data incurs significant
    overheads with a data-intensive application. These overheads can be so
    significant that Sea's performance, in these cases, can be found to be
    inferior to that of Lustre. In applications where data is shared or when
    results are required for post-processing, there is no other option than to
    flush this data. Since the majority of the overhead appears to have arisen
    from writing to and flushing from local disk, it is recommended that
    lesser-used data be evicted from Sea, freeing up space for newer data.

    There is limited benefit to Sea when flushing all the data in a
    data-intensive scenario. Sea must perform the same number of I/O operations
    as Lustre in these cases. While the application itself can proceed to
    completion faster, as it only needs to wait for all the data to be written
    to local storage, the time required for the final flush of the data can be
    quite significant, particularly when flushing from disk to Lustre.
    Therefore, we recommend that flushing all the data is reserved for
    more compute-intensive applications.

    Increase in performance from eviction is not only limited to scenarios where
    all the data is flushed. Sea also benefits from eviction when using the
    in-memory option as not all data may fit in memory and thus need to be
    written to slower local storage. Sea currently cannot handle scenarios where
    the application is attempting to access a file that is in the process of
    being moved, and as a result, we were not able to use much eviction in our
    experiments. This would be an important feature to enable in future Sea
    releases, despite the potential slowdown that may be incurred from waiting
    for the data to be materialized to Lustre.

\section{Conclusion}
    We created Sea, a lightweight open-source data-management library for
    scientific applications. With the help of glibc interception, we were able
    to create a library that maintains qualities important to scientific computing
    (i.e. ease-of-use, reproducibility, portability and parallelism). Our
    results demonstrate that Sea is quite beneficial to reducing the data
    transfers overheads, particularly when using an in-memory computing
    configuration, producing speedups of up to 3$\times$. Sea's performance is,
    however, limited when the PFS is not being heavily used.

    Experimental results demonstrate that our performance model accurately
    depicts the bounds in the majority of the cases. The model overestimated
    performance when metadata calls were heavily impacting performance. This is
    because the model neglects to account for any kind of latency. More accurate
    predictions could be obtained with a more complex model, although, due to
    Lustre's complex functioning, a simulator would likely be more appropriate
    here.

    More complex functioning of Sea, such as splitting of individual files, as
    seen with the other burst buffer file systems, may be preferential,
    particularly in maximizing cache usage. However, the more we complicate
    these libraries, the harder they are for users to use. As a result, users
    must wait until the cluster guidelines are developed detailing how a user
    should use these libraries. The boundary between ease-of-use and performance
    needs to be further explored to determine what is best for users and
    applications.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

% % Can use something like this to put references on a page
% % by themselves when using endfloat and the captionsoff option.
% \ifCLASSOPTIONcaptionsoff
%   \newpage
% \fi
\bibliographystyle{plain}
\bibliography{biblio}

% biography section

\begin{IEEEbiography}{Val\'erie Hayot-Sasson}
\end{IEEEbiography}
\begin{IEEEbiography}{Tristan Glatard}
\end{IEEEbiography}
\end{document}

### 2022-01-25 paper/sea-neuro/paper_neuro.tex
paper updates

--- 2021-10-08 paper/paper_neuro.tex
@@
 \subsubsection{Datasets}
-Functional magnetic resonance imaging datasets can vary greatly in total number
+fMRI datasets may vary greatly in total number
 of images and number of volumes within each image. To adequately capture the 
-diversity of datasets and when Sea could be pertinent, we selected three datasets of varying
-sizes: 1) OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD  dataset~\cite{preventad}, and
-3) the HCP dataset~\cite{hcp}. The ds001545 dataset is a total of 45.94~$GB$ (1778 files) and consists of data collected
-from 30 subjects in a single session. The PREVENT-AD dataset is a 255.0~$GB$ dataset (53061 files) consisting of data collected
-from 308 subjects. The HCP dataset is the largest of the three datasets at 85.4~$TB$ (x files) consisting of data collected from
-1113 subjects.
+diversity of datasets and the applicability of Sea, we have selected three datasets of varying
+time and space resolutions (Table~\ref{table:data}): 1) OpenNeuro's ds001545 dataset~\cite{ds001545},
+2) the PREVENT-AD dataset~\cite{preventad}, and 3) the HCP dataset~\cite{hcp}.
+The ds001545 dataset is a total of 45.94~$GB$ (1778 files) and consists of data collected
+from 30 participants in a single session watching 3 different clips (i.e. three runs/participants) of the The Grand Budapest Hotel.
+
+The PREVENT-AD dataset is publicly available dataset comprising of data from 330 participants with a familial history
+of Alzheimer's Disease.
+At the time of the experiments, the dataset contained 255.0~$GB$ (53061 files) originating from the 308 subjects available on DataLad~\cite{}.
+
+The HCP project, whose aim is to characterize brain connectivity using data collected from 1200 subjects, is the largest of the three datasets.
+At the time of our experimments, the dataset obtained consisted 85.4~$TB$ (x files) of data collected from 1113 subjects.
+
+By experimenting with datasets of different spatial and temporal resolutions, the amount of compute may be affected.
+Naturally, with more data there may inevitably be more computations performed, but also, the code itself may
+chose to perform entirely different computations as a result of differences in resolution. Moreover, parallelism and program I/O behaviour
+may be impacted depending on the particular characteristics of the data. Capturing this potential
+variety in processing will provide us with a better understanding of when and where Sea can be useful.
+
+\begin{table*}[t]
+  \small\centering
+\begin{tabular}{|c c c c c|}
+  \hline
+  Dataset & Total Size (MB) & Total Number of images & Number of images processed & Total size of processed (MB) \\
+  \hline
+  \multirow{3}{*}{PREVENT-AD} & \multirow{3}{*}{289532} & \multirow{3}{*}{53061} & 1 & 51.26\\
+  & & & 8 & \\
+  & & & 16 & \\
+  \hline
+  \multirow{3}{*}{ds001545} & \multirow{3}{*}{27377} & \multirow{3}{*}{1778} & 1 & 280.27\\
+  & & & 8 & \\
+  & & & 16 & \\
+  \hline
+  \multirow{3}{*}{HCP} & \multirow{3}{*}{83140079} & \multirow{3}{*}{X} & 1 &  1299.93\\
+  & & & 8 & \\
+  & & & 16 & \\
+
+  \hline
+
+  \hline
+\end{tabular}\caption{Dataset characteristics}\label{table:data}
+\end{table*}
 
@@
 Thus, we preprocessed each dataset using four standard preprocessing pipelines: FSL~\cite{fsl}, AFNI~\cite{AFNI},
-SPM~\cite{SPM}, fMRIPrep~\cite{fmriprep}. Table~\ref{tb1} provides an reference of the differences in computation and data-intensivity
-of the different pipelines on a single subject of the ds001545 dataset.
+SPM~\cite{SPM}, fMRIPrep~\cite{fmriprep}. Table~\ref{table:pipelines} provides an reference of the differences in computation and data-intensivity
+of the different pipelines on a single subject of all three of the datasets.
 
-Each tool was set to only run the preprocessing pipeline. Scripts used to execute each pipeline can be found at~\ref{github.com}.
+Each tool was set to only run the preprocessing pipeline. Scripts used to execute each pipeline can be found at~\ref{github.com/valhayot/sea-slurm-scripts}.
+
 For the FSL preprocessing pipeline, the default settings were maintained with the exception of slice-timing correction (interleaved),
@@
 %% Add SPM here
+To preprocess the datasets with SPM12, we reused the template described in \cite{haitas2021}, with fieldmap correction removed.
+As the template was designed for a specific dataset, we updated the paths to the fMRI and anatomical images,
+as well as the dimensions, such as the number of volumes and slices and the TR and TA, for each dataset. Furthermore,
+we set the slice order to interleaved in all cases.
 
@@
 
+\begin{table*}[t]
+  \small\centering
+\begin{tabular}{|c c c c c c|}
+  \hline
+  Tool & Dataset &  Output Size (MB) & Total libc calls & Libc Lustre calls & Compute time (s) \\
+  \hline
+  \multirow{3}{1em}{AFNI} & PREVENT-AD & 539.6 & 272390 & 4348 & 103.25 \\
+  & ds001545 & 3063.4 & 281707 & 4570  & 280.30 \\
+  & HCP & 18720.4 & 305310 & 5367 & 816.16 \\
+  \hline
+  \multirow{3}{1em}{FSL Feat} & PREVENT-AD & 253.7 & &  & 1338.29 \\
+  & ds001545 & 550.6 & &  & 2145.96 \\
+  & HCP & 1607.9 & & & 6596.46 \\
+  \hline
+  \multirow{3}{1em}{SPM} & PREVENT-AD & 331.0 & 34868 & 10048 &  483.67 \\
+  & ds001545 & 744.4 & 47251 & 19568 & 446.53 \\
+  & HCP & 2082.5 & 54753 & 25226 & 715.43 \\
+
+  \hline
+\end{tabular}\caption{Pipeline execution characteristics}\label{table:pipelines}
+\end{table*}
 \subsection{Controls}
+Our objective is to measure the speedup brought by Sea on real neuroimaging pipelines.
+Although Sea is agnostic to the internals of the pipelines and their output results, incorrect parameter
+selection and resulting outputs may result in I/O patterns that would not necessarily be observed had the
+pipeline been executed properly.
+
+Visual QC on the pipeline to ensure it was correctly registered and skull-striped to the template brain (MNI template).
+
+To validate that the outputs produced were sufficiently correct, we overlayed the output on the input image.
+This allowed up to see if steps such as skull-stripping and registration were correctly performed on the data.
+We repeated this step for all datasets and all pipelines processing a single image. In addition, we compared output
+number of files and output size to ensure that the pipeline was not producing different outputs depending on the file
+system used.
+
+% compare md5sums of uncompressed files.
+
 \subsection{Infrastructure}
 
+Our experiments were exected on two different clusters: 1) a local cluster and 2) Compute Canada's Beluga cluster.
+The local cluster comprised of 8 compute nodes with 256GB memory and 125GB tmpfs, running Centos 8. Storage on the local
+cluster was comprised of 4 Lustre ZFS storage nodes with 44 HDD OSTs and 1 MDS with a single MDT. The
+storage nodes were connected to the compute nodes via a 20Gbps ethernet connection.
+The cluster's hardware is similar to that of Beluga.
+
+The Beluga cluster
+
+%Busy writers
 \section{Results}
@@
 \section{Discussion}
+\subsection{Significance of speedups with Sea}
+
+\subsection{Other use cases}
+writes a lot but it's actually more complicated. When dataset increases
+\subsection{Predicting speedups}
+ difficulties with summarizing speedups into guidelines.
+ variability is important and it's coming from the interaction between infrastructure and application.
+ Benefits of sea depend on parameters such as load of the filesystem which is multifold and the application compute and
+ file system configuration. 
+ page cache
+ examples:
+  - large dataset - > more computation
+  - data access patterns and bursty-ness
+  - small datasets + metadata accesses
+
+\subsection{Libc interception of neuroimaging pipelines}
+
+As demonstrated by our Sea experiments, we have found that libc interception is effective at capturing all file system
+calls executed by the applications. This means that Sea is compatible with the most prominently used neuroimaging
+preprocessing pipelines. Although we have not extensively against all possible parameters and toolbox applications, we
+assume that use of glibc within the toolboxes themselves are likely to be consistent.
+
+The ability to use libc interception over more common alternative methods such as kernel-based or fuse-based filesystems
+was essential in ensuring ease-of-use and minimal performance overheads.
+
+opportunities for other optimizations
+
+Sea may be beneficial
+\subsection{Should Sea always be used?}
+
+user
+
+sysadmin
+
+write intensive
+
+\subsection{Effective use of available cluster resources}
+\subsection{Prefetching, flushing and eviction with Sea}
+\subsection{Comparisons to best case scenario}
+
 talk about testing

### 2022-03-23 paper/sea-comp/paper.tex
updates to sea comp paper

@@
 \begin{abstract}
-    The recent influx of scientific computing datasets have shifted scientific
-computing from compute intensive to data intensive. Whereas many Big Data
+    The recent influx of scientific data has lead to scientific
+computing transitioning from compute intensive to data intensive. Whereas many Big Data
 Frameworks exist that minimize the cost of data transfers, few scientific
-applications adopt data-management strategies as the effort required to update
+applications adopt data-placement strategies as the efforts required to update
 the existing applications is significant.
 
-We developed Sea as a means to enable data-management strategies for scientific
-applications executing on HPC clusters without needing to update the underlying
-applications. Sea leverages GNU C library interception to intercept
+We developed Sea as a means to enable data-placement strategies for scientific
+applications executing on HPC clusters without the need to reinstrument workflows.
+Sea leverages GNU C library interception to intercept
 POSIX-compliant file system calls made by the applications. We designed a
@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-\IEEEPARstart{E}{fficient} data-management
+\IEEEPARstart{E}{fficient} data-placement
 strategies have become essential in minimizing the costs of processing Big Data.
-Such strategies include data locality and in-memory computing, which minimize
-transfers by ensuring compute occurs where the data is located and maintain data
-in the fastest available storage (i.e. memory). Big Data frameworks, such as
+Strategies, such as data locality and in-memory computing, improve application runtime by ensuring
+data access occur on the most efficient available storage. Existing Big Data frameworks, such as
 MapReduce~\cite{dean2008mapreduce}, Apache Spark~\cite{zaharia2016apache} and
-Dask~\cite{rocklin2015dask}, provide such data management mechanisms by default.
+Dask~\cite{rocklin2015dask}, improve runtime performance through the incorporation of these data-placement
+strategies.
 
 Until the recent surge in publicly available scientific data, scientific
-workflows had been regarded as compute intensive. As a result, standardized
-scientific tools typically lack these data-management mechanisms. The sheer
-effort required to update the applications to leverage Big Data frameworks (e.g.
-changing framework, programming language, training staff) can be a significant
-deterrent. Furthermore, Big Data frameworks still require significant work to be
-well adapted for the processing of different scientific data, such as imaging
-data~\cite{mehta2017comparative}.
+workflows have been regarded as compute intensive. As a result, standardized
+scientific tools typically lack data-placement mechanisms. In contrast, efforts have
+been placed on efficient storage representation, such as HDF5~\cite{hdf5} and Zarr~\cite{zarr}, and
+the sharing of data, through initiatives like DataLad~\cite{wagner2022fairly}.
+
+Data-placement strategies can complement existing solutions by selecting where data should be accessed at runtime,
+facilitating the transfer and processing of large datasets across different infrastructures, and placing the different
+file formats at the preferred storage layer to maximize efficiency. However, converting adapting existing scientific frameworks
+and applications to Big Data frameworks requires significant effort, necessitating the rewrite of many well established applications.
+Furthermore, as noted in \cite{mehta2017comparative}, Big Data frameworks are still difficult to use for the processing of certain
+scientific data, such as imaging data.
 
@@
 
-Our goal is to provide data-management strategies to scientific tools. We focus
-on well-established scientific applications that operate as command-line tools.
-These applications are written in a variety of different programming languages
-and communicate data between tasks via a POSIX-compliant shared file system.
-
-Scientific tools are typically platform-agnostic with the exception that the
-file system must be POSIX-compliant. That is, they can run on local
-workstations, high performance computing (HPC) clusters and cloud
-infrastructure, with the exception that they cannot communicate with a
-cloud-based storage natively. We focus our efforts on HPC clusters as they are
-widely used in scientific computing and are currently used for the processing of
-scientific Big Data.
-
-Our goal is to provide data-management strategies to scientific tools. We focus
-on well-established scientific applications that operate as command-line tools.
-These applications are written in a variety of different programming languages
-and communicate data between tasks via a POSIX-compliant shared file system.
-
-Scientific tools are typically platform-agnostic with the exception that the
-file system must be POSIX-compliant. That is, they can run on local
-workstations, high performance computing (HPC) clusters and cloud
-infrastructure, with the exception that they cannot communicate with a
-cloud-based storage natively. We focus our efforts on HPC clusters as they are
-widely used in scientific computing and are currently used for the processing of
-scientific Big Data.
+Our goal is to provide a data-placement add-on to standardized scientific computing applications executing on High Performance Computing (HPC)
+clusters. The tools of particular interest to us consist of various tasks that communicate to each other via a POSIX-compliant file system. Whereas
+they can be platform-agnostic, users typically leverage them within an HPC context, particularly when processing large volumes of data, due to their
+inherent reliance on POSIX-compliance.
+
+We developed Sea, a user-space data-placement library, that provides efficient data placement without reinstrumentation of the underlying application.
+Given a predefined storage hierarchy, Sea can prefetch, cache, flush and evict data to and from a permanent storage location. To evaluate Sea,
+we have designed a performance model, as well as evaluate its performance on a synthetic Big Data application running on a representative neuroimaging dataset.
 
@@
 Sea\footnote{\url{https://github.com/valhayot/sea}} is an open-source user-space
-data-management C++ library for applications running on HPC systems. Its main
+data-placement C++ library for applications running on HPC systems. Its main
 aim is to reduce application data transfer costs by leveraging node-local

### 2022-03-29 paper/sea-comp/paper.tex
updates to comp paper

@@
 % \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
-% \algblock{Input}{EndInput}
-% \algnotext{EndInput}
-% \algblock{Output}{EndOutput}
-% \algnotext{EndOutput}
-% \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
-\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}: #1}}}
-
-% benchmark on beluga, prefreesurfer with RUIS
-% use more than 64 subjects for prefreesurfer --> few 100 subjects
-% think more about how to benchmark I/O
-% try distributed jobs with Sea
+% \algblock{Input}{EndInput} \algnotext{EndInput} \algblock{Output}{EndOutput}
+% \algnotext{EndOutput} \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
+\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
+#1}}}
+
+% benchmark on beluga, prefreesurfer with RUIS use more than 64 subjects for
+% prefreesurfer --> few 100 subjects think more about how to benchmark I/O try
+% distributed jobs with Sea
 
 \begin{document}
-\title{Sea: A lightweight data-management library for Big Data scientific computing}
-% remove neuroimaging and replace with scientific computing  add user-space, switch HPC to computing clusters
+\title{Sea: A lightweight data-placement library for Big Data scientific computing}
+% remove neuroimaging and replace with scientific computing  add user-space,
+% switch HPC to computing clusters
 
@@
 \markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
-{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
-\IEEEtitleabstractindextext{%
+{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for
+Computer Society Journals} \IEEEtitleabstractindextext{%
 \begin{abstract}
-    The recent influx of scientific data has lead to scientific
-computing transitioning from compute intensive to data intensive. Whereas many Big Data
+    The recent influx of scientific data has lead to scientific computing
+transitioning from compute intensive to data intensive. Whereas many Big Data
 Frameworks exist that minimize the cost of data transfers, few scientific
@@
 We developed Sea as a means to enable data-placement strategies for scientific
-applications executing on HPC clusters without the need to reinstrument workflows.
-Sea leverages GNU C library interception to intercept
-POSIX-compliant file system calls made by the applications. We designed a
-performance model and evaluated the performance of Sea on a synthetic
-data-intensive application processing a representative neuroimaging dataset (the
-Big Brain). Our results demonstrate that Sea significantly improves performance,
-up to a factor of 3$\times$.
+applications executing on HPC clusters without the need to reinstrument
+workflows. Sea leverages GNU C library interception to intercept POSIX-compliant
+file system calls made by the applications. We designed a performance model and
+evaluated the performance of Sea on a synthetic data-intensive application
+processing a representative neuroimaging dataset (the Big Brain). Our results
+demonstrate that Sea significantly improves performance, up to a factor of
+3$\times$.
 
@@
 
-\IEEEdisplaynontitleabstractindextext
-\IEEEpeerreviewmaketitle
+\IEEEdisplaynontitleabstractindextext \IEEEpeerreviewmaketitle
 
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-\IEEEPARstart{E}{fficient} data-placement
-strategies have become essential in minimizing the costs of processing Big Data.
-Strategies, such as data locality and in-memory computing, improve application runtime by ensuring
-data access occur on the most efficient available storage. Existing Big Data frameworks, such as
-MapReduce~\cite{dean2008mapreduce}, Apache Spark~\cite{zaharia2016apache} and
-Dask~\cite{rocklin2015dask}, improve runtime performance through the incorporation of these data-placement
-strategies.
+\IEEEPARstart{E}{fficient} data-placement strategies have become essential in
+minimizing Big Data processing overheads. These strategies, such as data
+locality and in-memory computing, improve application runtime by ensuring data
+accesses occur on the most efficient available storage. Existing Big Data
+frameworks, such as MapReduce~\cite{dean2008mapreduce}, Apache
+Spark~\cite{zaharia2016apache} and Dask~\cite{rocklin2015dask}, improve runtime
+performance through the incorporation of these data-placement strategies.
 
@@
 workflows have been regarded as compute intensive. As a result, standardized
-scientific tools typically lack data-placement mechanisms. In contrast, efforts have
-been placed on efficient storage representation, such as HDF5~\cite{hdf5} and Zarr~\cite{zarr}, and
-the sharing of data, through initiatives like DataLad~\cite{wagner2022fairly}.
-
-Data-placement strategies can complement existing solutions by selecting where data should be accessed at runtime,
-facilitating the transfer and processing of large datasets across different infrastructures, and placing the different
-file formats at the preferred storage layer to maximize efficiency. However, converting adapting existing scientific frameworks
-and applications to Big Data frameworks requires significant effort, necessitating the rewrite of many well established applications.
-Furthermore, as noted in \cite{mehta2017comparative}, Big Data frameworks are still difficult to use for the processing of certain
-scientific data, such as imaging data.
-
-%multiple disk support
-% pipeline agnostic and pipeline access patterns
-% don't leverage workflows
-% leverage existing filesystems
-% replace storage devices with storage location.
-% use applications
-% memory management paragraph: PAGE CACHE
+scientific tools typically lack data-placement mechanisms. In contrast, efforts
+have been placed on efficient storage representation, such as HDF5~\cite{hdf5}
+and Zarr~\cite{zarr}, and the sharing of data, through initiatives like
+DataLad~\cite{wagner2022fairly}.
+
+Data-placement strategies can complement existing solutions by facilitating the
+transfer and processing of large datasets across different infrastructures, and
+placing the different file formats at the preferred storage layer to maximize
+efficiency. However, adapting existing scientific applications to Big Data
+frameworks requires considerable effort, necessitating the rewrite of many well
+established applications. Furthermore, as noted in \cite{mehta2017comparative},
+Big Data frameworks are not easy to use for the processing of certain scientific
+data, such as imaging data, due to the level of expertise required to rewrite
+reference implementations using the frameworks.
+
+%multiple disk support pipeline agnostic and pipeline access patterns don't
+% leverage workflows leverage existing filesystems replace storage devices with
+% storage location. use applications memory management paragraph: PAGE CACHE
 % shared filesystems are slow and metadata management can be slow
 
-% add need for userspace / intelligent pipeline flushing
-% there is no one-size fits all with data intensive applications
-% due to io patterns, so we decided with userspace.
+% add need for userspace / intelligent pipeline flushing there is no one-size
+% fits all with data intensive applications due to io patterns, so we decided
+% with userspace.
 
@@
 
-Our goal is to provide a data-placement add-on to standardized scientific computing applications executing on High Performance Computing (HPC)
-clusters. The tools of particular interest to us consist of various tasks that communicate to each other via a POSIX-compliant file system. Whereas
-they can be platform-agnostic, users typically leverage them within an HPC context, particularly when processing large volumes of data, due to their
-inherent reliance on POSIX-compliance.
-
-We developed Sea, a user-space data-placement library, that provides efficient data placement without reinstrumentation of the underlying application.
-Given a predefined storage hierarchy, Sea can prefetch, cache, flush and evict data to and from a permanent storage location. To evaluate Sea,
-we have designed a performance model, as well as evaluate its performance on a synthetic Big Data application running on a representative neuroimaging dataset.
+Our goal is to provide a data-placement solution to standardized scientific
+computing applications executing on High Performance Computing (HPC) clusters.
+The tools of particular interest to us consist of various tasks that communicate
+to each other via a POSIX-compliant file system. Whereas the tools can be
+platform-agnostic, we assume users will leverage them within an HPC context,
+particularly when processing large volumes of data, due to their inherent
+reliance on POSIX-compliance.
+
+We developed Sea, a user-space data-placement library, that provides efficient
+data placement without reinstrumentation of the underlying application. Given a
+predefined storage hierarchy, Sea can prefetch, cache, flush and evict data to
+and from a permanent storage location. To assist in providing a
+frame-of-reference for Sea's performance on HPC systems, we designed a
+performance model. We then evaluated Sea's performance on a synthetic Big Data
+application running on a representative neuroimaging dataset, comparing its
+performance with the bounds delineated by the model.
 
@@
 \subsection{HPC Infrastructure}
-      The general structure of HPC clusters may complicate the deployment of Big
+      The general structure of HPC clusters complicates the deployment of Big
       Data frameworks. Typical HPC clusters consist of distinct storage and
-      compute nodes. While the compute nodes may also have local storage, there
-      is no distributed file system like the Hadoop Distributed File System
+      compute nodes. While compute nodes may also have local storage, there is
+      no distributed file system like the Hadoop Distributed File System
       (HDFS)~\cite{shvachko2010hadoop} or Alluxio~\cite{alluxio} to facilitate
@@
       with allocation duration enforced by a batch scheduler. Therefore, any
-      data written to a compute node is inaccessible after the allocation is
-      terminated.
+      data written to compute nodes are rendered inaccessible after the
+      allocation is terminated.
 
@@
       InfiniBand~\cite{infiniband}. I/O overheads of shared PFS can be incurred
-      in many places, from the shared network, to the bandwidth of the storage
-      disks.  With Lustre specifically, there typically are many data nodes, known as
-      Object Storage Servers (OSS) which contain several storage devices, known
-      as Object Storage Targets (OST). All file metadata is maintained within a
-      separate node known as the Metadata Server (MDS) and stored within a
-      device referred to as the Metadata Target (MDT). Although data transfers
-      can be communicated directly to the corresponding OST, the clients need to
-      first communicate with the MDS to determine which OST to communicate with.
-      This can result in major data transfer overheads, particularly when making
-      numerous requests to the metadata server.
-
-      HPC clusters may provide a faster additional storage layer, known as a
+      in many places, from the shared network to the bandwidth and latency of
+      the storage devices.  With Lustre specifically, there typically are many
+      data nodes, known as Object Storage Servers (OSS) which contain several
+      storage devices, known as Object Storage Targets (OST). All file metadata
+      is maintained within a separate node known as the Metadata Server (MDS)
+      and stored within a device referred to as the Metadata Target (MDT).
+      Although data transfers can be communicated directly to the corresponding
+      OST, the clients need to first communicate with the MDS to determine which
+      OST to communicate with. This can result in major data transfer overheads,
+      particularly when making numerous requests to the metadata server.
+
+      HPC clusters may provide a faster intermediate storage layer, known as a
       Burst Buffer, aimed at improving I/O performance. A Burst Buffer typically
-      consists of fast storage (SSD and memory) and can be local to the compute
-      node or as a dedicated I/O node. It was introduced as a response to offset
-      the impacts of large-scale application checkpointing on the PFS~\cite{bb}. The
-      application would write a checkpoint to the Burst Buffer and resume
-      processing, while the checkpoint would be asynchronously flushed to the
-      PFS.
+      consists of higher-performance storage devices (SSD and memory) and can be
+      local to the compute node or as a dedicated I/O node. Burst Buffers were
+      introduced as a response to offset the impacts of large-scale application
+      checkpointing on the PFS~\cite{bb}. The application would write a
+      checkpoint to the Burst Buffer and resume processing, while the checkpoint
+      would be asynchronously flushed to the PFS.
 
@@
 
-      Strategies mitigate the impacts of data transfers in data-intensive workloads is
-      required. In text-based processing, frameworks such as
-      MapReduce~\cite{dean2008mapreduce} and Apache
-      Spark~\cite{zaharia2016apache} have implemented and popularized strategies
-      to minimize the need for data transfers during processing. These
-      strategies are known as data locality and in-memory computing.
-
-      Data locality is the process by which compute tasks are scheduled nearest
-      to where the data is located. Traditionally, storage and compute were kept
-      separate, having to transfer data over the network to compute. Rather than
-      having to transfer large amounts of data over the network to the compute
-      tasks, which could incur significant overheads, Big Data frameworks ensure
-      that data is stored directly on the compute nodes. When a compute task
-      requires access to specific data, the scheduler sends the task to the
-      nearest available node to the data, thereby minimizing any cost of
-      network-related data transfers. This strategy is not only used by Big Data
-      Frameworks such as Hadoop MapReduce, Apache Spark, and
-      Dask~\cite{rocklin2015dask}, but also enabled by file systems such as HDFS
-      and Alluxio.
+      Strategies to mitigate the impacts of data transfers in data-intensive
+      workloads are required to ensure optimal processing time. Early Big Data
+      frameworks, notably MapReduce~\cite{dean2008mapreduce} and Apache
+      Spark~\cite{zaharia2016apache} have implemented and popularized
+      data-management strategies to minimize data transfers at processing time.
+      These strategies are known as data locality and in-memory computing.
+
+      Data locality ensures that compute tasks are scheduled nearest to where
+      the data is located. Historically, storage and compute nodes were distinct
+      layers, data transfer to the compute location. Rather than transferring
+      large amounts of data over the network to compute tasks, which could incur
+      significant overheads, Big Data frameworks ensure that data is stored
+      directly on the compute nodes. When a compute task requires access to
+      specific data, the scheduler sends the task to the nearest available node
+      to the data, thereby minimizing any cost of network-related data
+      transfers. This strategy is not only used by Big Data Frameworks such as
+      Hadoop MapReduce, Apache Spark, and Dask~\cite{rocklin2015dask}, but also
+      enabled by file systems such as HDFS and Alluxio.
+
+      In-memory computing complements data locality by maximizing use of
+      available memory to maintain intermediate data. Whereas with data locality
+      alone, tasks would leverage local storage, typically in the form of HDDs
+      or SSD, to maintain task input and output data, with the addition of
+      in-memory computing, data would be stored in main memory to minimize
+      latency and maximize bandwidth.
+
+      While Big Data frameworks and filesystems are not commonly used in HPC,
+      HPC clusters are typically Linux-based and benefit from the Linux page
+      cache, which can provide data locality and in-memory computing at a
+      limited capacity. Furthermore, HPC architectures may also provide an
+      intermediate storage layer, known as a Burst Buffer, to improve
+      application I/O-related overheads. To facilitate an applications
+      interaction between application and storage layer, distributed file
+      systems have been developed.
 
 \subsection{The Linux page cache}
- While scientific applications uncommonly use Big Data frameworks and must
-      perform large over-the-network data transfers, they may still benefit from
-      in-memory computing and data locality through the Linux Page
-      Cache~\cite{pagecache}. Similarly to other file systems, Lustre leverages
-      the client node page cache to reduce I/O overheads. System memory is
-      composed of two components: 1) anonymous memory and 2) page cache.
-      Anonymous memory consists of all application-related objects, whereas page
-      cache consists of recently accessed file data. When a file is read, that
-      file is loaded up into the page cache to be flagged for eviction based on
-      a least recently used (LRU) policy. That means subsequent accesses to that
-      data may be done entirely in memory so long as that data has not already
-      been evicted. Similarly, for writes to a file system with writeback cache
-      enabled, the file will be written to memory completing the write operation
-      once the file has been written entirely to memory. That file will then be
-      flushed asynchronously to the appropriate storage device. As system memory
-      may get overloaded with too many write requests, there is a limit to the
-      amount of written data that can exist in memory, known as the
-      \texttt{dirty\_ratio}. Furthermore, applications producing too many write
-      requests may be throttled by the system.
+ While scientific applications do not customarily leverage data-placement
+      strategies and may perform large over-the-network data transfers, they may
+      still benefit from in-memory computing and data locality through the Linux
+      Page Cache~\cite{pagecache}. Similarly to other file systems, Lustre
+      leverages the compute node page cache to reduce I/O overheads. System
+      memory is composed of two components: 1) anonymous memory and 2) page
+      cache. Anonymous memory consists of all application-related objects,
+      whereas page cache consists of recently accessed file data. When a file is
+      read, that file is loaded up into the page cache to be flagged for
+      eviction based on a least recently used (LRU) policy. That means
+      subsequent accesses to that data may be done entirely in memory so long as
+      that data has not already been evicted. Similarly, for writes to a file
+      system with writeback cache enabled, the file will be written to memory
+      completing the write operation once the file has been written entirely to
+      memory. That file will then be flushed asynchronously to the appropriate
+      storage device. As system memory may get overloaded with too many write
+      requests, there is a limit to the amount of written data that can exist in
+      memory, known as the dirty\_ratio. Furthermore, applications producing too
+      many write requests may be throttled by the system.
 
@@
       systems. Such file systems require super-user access, preventing users
-      from deploying their own on-the-fly cluster. While HDFS can be loaded
-      in user-space by mounting its File System in Userspace (FUSE)
-      implementation, FUSE-based file systems can perform significantly worse
-      than desired, depending on the application~\cite{tofuse}. Even with a
-      user-space version of HDFS, there would be no mechanism to transfer all
-      the data back to the parallel file system to be accessible post resource
-      requests. Furthermore, such file systems are not POSIX-compliant, which
-      would be required for scientific computing applications.
-
-      There are, however, Big Data file systems, such as
-      XtreemFS~\cite{xtreemfs} that exist entirely within user space, are
-      POSIX-compliant, and can use alternative methods to FUSE to function.
-      XtreemFS uses the LD\_PRELOAD trick to intercept file system calls made to
-      the GNU C library (glibc). There are limitations to using glibc
-      interception. For instance, the application must be dynamically linked and
-      make glibc calls. However, statically-linked applications are uncommon and
+      from deploying their own on-the-fly cluster. While HDFS can be loaded in
+      user-space by mounting its FUSE (Filesystem in User Space) implementation,
+      FUSE-based file systems may perform significantly worse than desired,
+      depending on the application~\cite{tofuse}. Even with a user-space version
+      of HDFS, there would be no mechanism to transfer all the data back to the
+      PFS post-processing.
+      % Furthermore, such file systems are not POSIX-compliant, which would be
+      % required for scientific computing applications.
+
+      However, there exist Big Data file systems, such as
+      XtreemFS~\cite{xtreemfs}, that exist entirely within user space, are
+      POSIX-compliant, and can use alternative methods to FUSE to function. For
+      instance, XtreemFS uses the LD\_PRELOAD trick to intercept file system
+      calls made to the GNU C library (glibc). There are limitations to using
+      glibc interception; the application must be dynamically linked and make
+      glibc calls. However, statically-linked applications are uncommon and
       applications interacting with POSIX-compliant file systems on Linux
@@
 
-      Characteristics that make XtreemFS unlikely to be deployed on-the-fly
-      include complex file system deployment and configuration properties. It
-      also does not provide a simple way to leverage different classes of
-      storage devices, nor ensure that required data will be copied to the
-      cluster's parallel file system.
+      Due to complex file system deployment and configuration properties,
+      XtreemFS is unlikely to be deployed on an ephemeral burst buffer. It also
+      does not provide a simple way to leverage different classes of storage
+      devices, nor ensure that required data will be copied to the cluster's
+      parallel file system.
 
@@
       user-space using glibc interception. While these implementations simplify
-      usage of burst buffers for users, their installation and configuration would be out of the
-      scope of a typical HPC user and thus require sysadmin intervention.
+      usage of burst buffers for users, their installation and configuration is
+      similarly out-of-scope for a typical HPC user.
+
 \section{Materials and Methods}
@@
     \includegraphics[width=\columnwidth]{figures/seacomp-diagram.pdf}%
-\caption{The Sea data-management library. Sea intercepts I/O-related glibc calls issued by the application and
-translates the path provided if located within the Sea mountpoint.
-Sea leverages a user-defined storage hierarchy to help select the most appropriate path and the redirects the call to glibc to process the request on the translated path.
-}
+\caption{The Sea data-placement library. Sea intercepts I/O-related glibc calls
+issued by the application and translates the path provided if located within the
+Sea mountpoint. Sea leverages a user-defined storage hierarchy to help select
+the most appropriate path and the redirects the call to glibc to process the
+request on the translated path. }
 \label{fig:sea-comp:diagram}
@@
 
-Sea\footnote{\url{https://github.com/valhayot/sea}} is an open-source user-space
-data-placement C++ library for applications running on HPC systems. Its main
-aim is to reduce application data transfer costs by leveraging node-local
-storage (Figure~\ref{fig:sea-comp:diagram}). Sea redirects files accessed from a user-specified mount point to the
-appropriate storage devices using glibc interception. Within an intercepted call
-the file will be written to or read from the fastest storage device available.
-
-At minimum, Sea requires the user to specify at least two storage devices, a
-fast temporary one (i.e. cache) and a slower long-term storage one. This could be RAM and SSD
-if working on a single node, or a compute-local SSD and a shared parallel file
-system, in the case of an HPC cluster. Ideally, a user will provide a multitude
-of short-term storage devices to improve Sea's efficacy. Furthermore, to
-maximize usage of the fastest available storage devices, Sea will allow the user
-to outline which files can be removed from short-term storage in addition to
-which files need to be materialized onto long-term storage. Eviction of files is
-important as this will allow Sea to maximize usage of short-term storage.
-
-Sea was meant as a data-management service to be employed at the discretion of
-the users on scientific applications. When coming up with the idea for Sea, we
+Sea\footnote{\url{https://github.com/big-data-lab-team/sea}} is an open-source
+user-space data-placement C++ library for applications running on HPC systems.
+Its main aim is to reduce application data transfer costs by leveraging
+node-local storage (Figure~\ref{fig:sea-comp:diagram}). Sea redirects files
+accessed from a user-specified mount point to the appropriate storage devices
+using glibc interception. Within an intercepted call the file will be written to
+or read from the fastest storage device available.
+
+Sea requires minimal configuration for use. In order to work, Sea requires the
+user to specify at least two storage devices, a fast temporary one (i.e. cache)
+and a slower long-term storage one. This could be RAM and SSD if working on a
+single node, or a compute-local SSD and a shared parallel file system, in the
+case of an HPC cluster. Ideally, a user will provide a multitude of ephemeral
+storage devices to improve Sea's efficacy. Furthermore, to maximize usage of the
+fastest available storage devices, Sea will allow the user to outline which
+files can be removed from short-term storage in addition to which files need to
+be materialized onto long-term storage. Eviction of files is important as this
+will allow Sea to maximize usage of short-term storage.
+
+Sea was created as a data-placement service to be employed at the discretion of
+the users on scientific applications. When developing the idea for Sea, we
 observed the design differences between Big Data and scientific, specifically
-neuroimaging, frameworks and applications. Scientific frameworks tend to favour
-ease-of-use, reproducibility, portability and parallelism over minimizing data
-transfers. It was important for us to ensure that Sea would preserve the design
-decisions of the underlying tools. Sea achieves this by having few dependencies,
-is lightweight, and requires minimal configuration.
+neuroimaging, frameworks and applications. Scientific frameworks and
+applications tend to favour ease-of-use, reproducibility, portability and
+parallelism over minimizing data
+transfers\cite{jarecka2020pydra,esteban2019fmriprep,wagner2022fairly}. It was
+important for us to ensure that Sea would preserve the design decisions of the
+underlying tools. Sea achieves this by having few dependencies, is lightweight,
+and requires minimal configuration.
 
@@
 Sea is primarily designed to work with workloads that generate large amounts of
-intermediate data. Since data must be read from the slower storage device and
-final output data most likely would have to be written to the same device, the
+intermediate data. Since data must be read from the slower storage device and final outputs will
+also have to be written to the same device, the
 use of Sea with workloads that do not produce intermediate data would result in
@@
 Sea provides two main modes based on flushing specification: In-memory computing
-and flush-all. In-memory computing is achieved when intermediate need not be
+and copy-all. In-memory computing is achieved when intermediate need not be
 materialized to long-term storage (i.e. no flushing). With Sea in-memory, we can
@@
 
-As the name implies, Sea flush-all is achieved when all data must be
+As the name implies, Sea copy-all is applicable when all data must be
 materialized to long-term storage (i.e. flush everything). This can be coupled
 with eviction to limit the amount of data that will be written to slower, but
-larger local storage devices. While this option will have overheads, due to the
-need to flush all application data, performance can be improved when data
-transfers and compute time are similar.
+larger, local storage devices. While this option will have overheads, due to the
+need to flush all application data, overheads will be negligible when data transfer and compute time
+are similar.
 
@@
 \subsection{Testing}
-To work effectively, Sea has to intercept every glibc function that interacts with files or
-directories. Failure to intercept some of these function may result in the whole application
-crashing. This is because only Sea can translate the Sea mountpoint file paths to their
-respective real locations. Due to an absence of standardized glibc testing suites,
-we developed our own integration tests.
-
-Our tests have been executed on three different operating systems: Ubuntu (xenial, bionic, focal),
-Fedora (32, 33, 34) and Centos (6, 7). Using this test matrix, the Sea tests were able to span seven glibc
-versions ranging from 2.17 to 2.33. As glibc file system functions may change over time, Sea behaviour with
+To work effectively, Sea has to intercept every glibc function that interacts
+with files or directories. Failure to intercept some of these function may
+result in the whole application crashing. This is because only Sea can translate
+the Sea mountpoint file paths to their respective real locations. Due to an
+absence of standardized glibc testing suites, we developed our own integration
+tests.
+
+Our tests have been executed on three different operating systems: Ubuntu
+(xenial, bionic, focal), Fedora (32, 33, 34) and Centos (6, 7). Using this test
+matrix, the Sea tests were able to span seven glibc versions ranging from 2.17
+to 2.33. As glibc file system functions may change over time, Sea behaviour with
 newer and older versions of glibc is undefined. 
-% ldd (GNU libc) 2.17
-% ldd (Ubuntu GLIBC 2.23-0ubuntu11.3) 2.23
-% ldd (Ubuntu GLIBC 2.27-3ubuntu1.4) 2.27
-% ldd (GNU libc) 2.28
-% ldd (Ubuntu GLIBC 2.31-0ubuntu9.2) 2.31
-% ldd (GNU libc) 2.31
-% ldd (GNU libc) 2.32
-% ldd (GNU libc) 2.33
+% ldd (GNU libc) 2.17 ldd (Ubuntu GLIBC 2.23-0ubuntu11.3) 2.23 ldd (Ubuntu GLIBC
+% 2.27-3ubuntu1.4) 2.27 ldd (GNU libc) 2.28 ldd (Ubuntu GLIBC 2.31-0ubuntu9.2)
+% 2.31 ldd (GNU libc) 2.31 ldd (GNU libc) 2.32 ldd (GNU libc) 2.33
 
@@
 Flushing and eviction mechanisms give way to four different modes: flush-only,
-evict-only, flush-and-evict and keep. A flush-only is synonymous with a
-file copy in that a file that already exists within a Sea cache will be copied
-to persistent storage. This operation is particularly useful when a file will be
+evict-only, flush-and-evict and keep. A flush-only is synonymous with a file
+copy in that a file that already exists within a Sea cache will be copied to
+persistent storage. This operation is particularly useful when a file will be
 reused by the application, and thus benefits from remaining in cache, but also
@@
 
-Evict-only is equivalent to a remove operation. This mode is for files that are located
-within a Sea cache, but do not need to be persisted and will not be reused by
-the pipeline and therefore do not need to occupy cache space. Evict-only is
-useful for removing unnecessary application log files. To evict-only a file, the
-file must only be specified in the \texttt{.sea\_evictlist}.
+Evict-only is equivalent to a remove operation. This mode is for files that are
+located within a Sea cache, but do not need to be persisted and will not be
+reused by the pipeline and therefore do not need to occupy cache space.
+Evict-only is useful for removing unnecessary application log files. To
+evict-only a file, the file must only be specified in the
+\texttt{.sea\_evictlist}.
 
@@
 Flush-and-evict is therefore synonymous with a move operation and is invoked on
-files that are specified in both the \texttt{.sea\_flushlist} and \texttt{.sea\_evictlist}.
-
-Keep is useful for when file needs to be in cache as it will be reused by
-the application, but is not needed in post-processing analysis or by other nodes.
+files that are specified in both the \texttt{.sea\_flushlist} and
+\texttt{.sea\_evictlist}.
+
+Keep is useful for when file needs to be in cache as it will be reused by the
+application, but is not needed in post-processing analysis or by other nodes.
 
@@
 
-      Experimental results can be prone to error and variability, particularly in instances
-      where there a many factors at play, such as in cluster environments. To improve
-      confidence in our results, we developed a simplified performance model that
-      could help describe the behaviours found in our experimental results. Since different
-      parallel file systems may operate differently, our baseline model will be
-      based on Lustre which is commonly used on HPC infrastructure.
+      Experimental results can be prone to error and variability, particularly
+      in instances where there a many factors at play, such as in cluster
+      environments. To improve confidence in our results, we developed a
+      simplified performance model that could help describe the behaviours found
+      in our experimental results. Since different parallel file systems may
+      operate differently, our baseline model will be based on Lustre which is
+      commonly used on HPC infrastructure.
 
@@
 
-      {\noindent}
-      Where, \\
+      {\noindent} Where, \\
       $M_{l}$ is the application makespan using Lustre \\
@@
 
-    {\noindent}
-    Where, \\
+    {\noindent} Where, \\
     $N$ is the network bandwidth \\
@@
       For the sake of simplicity, the above models assume that the network
-      bandwidth between the compute and data nodes is the same. This,
-      however, may not necessarily be the case. Furthermore, the model also
-      assumes that each file can only be located on a single disk, meaning that
-      the parallel bandwidth can at maximum be as fast as all Lustre disks
-      combined and as slow as the minimum number of compute threads reading and
-      writing files.
+      bandwidth between the compute and data nodes is the same. This, however,
+      may not necessarily be the case. Furthermore, the model also assumes that
+      each file can only be located on a single disk, meaning that the parallel
+      bandwidth can at maximum be as fast as all Lustre disks combined and as
+      slow as the minimum number of compute threads reading and writing files.
 
@@
 
-      {\noindent}
-      Where, \\
+      {\noindent} Where, \\
       $M_{c}$ is the makespan of application I/O to page cache \\
@@
 
-      {\noindent}
-      Where, \\
+      {\noindent} Where, \\
       $M_{lc}$ is the application makespan using Lustre with page cache \\
@@
           \caption{Incrementation}\label{alg:sea-comp:incrementation}
-          \SetAlgoLined 
-            \SetKwInOut{Input}{Input}
-            \Input{$n$ number of iterations;\\
+          \SetAlgoLined \SetKwInOut{Input}{Input} \Input{$n$ number of
+          iterations;\\
                 $C$ a set of image chunks;\\
                 $fs$ the file system to write to (Lustre or Sea)\\
-            }
-            \ForEach{$chunk \in C$}
-            { 
-              read $chunk$ from Lustre 
-              \For{$i \in [1,n]$}
-              { $chunk\gets chunk+1$ save $chunk$ to $fs$ }
-            }
+            } \ForEach{$chunk \in C$} { read $chunk$ from Lustre \For{$i \in
+            [1,n]$} { $chunk\gets chunk+1$ save $chunk$ to $fs$ } }
         \end{algorithm}
@@
       into 1000 files each consisting of \SI{617}{\mebi\byte} of data.
-      %of data as well as
-      %5000 blocks each \SI{124}{\mebi\byte} in size.
+      %of data as well as 5000 blocks each \SI{124}{\mebi\byte} in size.
 
@@
     various conditions using the incrementation application on the Big Brain
-    image. All experiments were repeated 5$\times$. Model bounds represented by coloured regions}
+    image. All experiments were repeated 5$\times$. Model bounds represented by
+    coloured regions}
     \label{fig:sea-comp:benchmarks}
@@
       When Sea does not provide speedups, it either performs similarly to
-      Lustre, as can be seen in Figure~\ref{fig:sea-comp:nodes} at one node and in
-      Figure~\ref{fig:sea-comp:iterations} at 1 iteration, or sightly slows down
-      execution, as can be seen in~\ref{fig:sea-comp:disks}. As expected, Sea at
-      a single iteration can at best perform similarly or slightly worse than
-      Lustre. This is because all the data is read from Lustre and written to
-      Lustre, operating in the same way that Lustre would with page cache.
+      Lustre, as can be seen in Figure~\ref{fig:sea-comp:nodes} at one node and
+      in Figure~\ref{fig:sea-comp:iterations} at 1 iteration, or sightly slows
+      down execution, as can be seen in~\ref{fig:sea-comp:disks}. As expected,
+      Sea at a single iteration can at best perform similarly or slightly worse
+      than Lustre. This is because all the data is read from Lustre and written
+      to Lustre, operating in the same way that Lustre would with page cache.
 
@@
 
-      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model correctly describes The
-      bounds for all but two experiments: Experiment 3
+      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model correctly
+      describes The bounds for all but two experiments: Experiment 3
       (Figure~\ref{fig:sea-comp:iterations}) and Experiment 4
       (Figure~\ref{fig:sea-comp:processes}). In Experiment 3, the model
-      incorrectly represents the bounds for 1 iteration. Since, in this case, all
-      the data should be able to fit in page cache for both Sea and Lustre, it
-      is possible that memory bandwidth has been overestimated, resulting in
+      incorrectly represents the bounds for 1 iteration. Since, in this case,
+      all the data should be able to fit in page cache for both Sea and Lustre,
+      it is possible that memory bandwidth has been overestimated, resulting in
       incorrect model predictions.
@@
 
-    Sea cannot be used with operating systems
-    that are not Linux-based, and therefore, the use of Sea limits application
-    portability. However, HPC systems commonly use Linux operating systems,
-    thus, it is not limiting for its intended use case. Furthermore, we provide
-    publicly available containers to simplify the usage of Sea on various other
-    operating systems.
+    Sea cannot be used with operating systems that are not Linux-based, and
+    therefore, the use of Sea limits application portability. However, HPC
+    systems commonly use Linux operating systems, thus, it is not limiting for
+    its intended use case. Furthermore, we provide publicly available containers
+    to simplify the usage of Sea on various other operating systems.
 
@@
     quite significant, particularly when flushing from disk to Lustre.
-    Therefore, we recommend that flushing all the data is reserved for
-    more compute-intensive applications.
+    Therefore, we recommend that flushing all the data is reserved for more
+    compute-intensive applications.
 
@@
     scientific applications. With the help of glibc interception, we were able
-    to create a library that maintains qualities important to scientific computing
-    (i.e. ease-of-use, reproducibility, portability and parallelism). Our
-    results demonstrate that Sea is quite beneficial to reducing the data
+    to create a library that maintains qualities important to scientific
+    computing (i.e. ease-of-use, reproducibility, portability and parallelism).
+    Our results demonstrate that Sea is quite beneficial to reducing the data
     transfers overheads, particularly when using an in-memory computing
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
+% if have a single appendix: \appendix[Proof of the Zonklar Equations] or
+%\appendix  % for no appendix heading do not use \section anymore after
+%\appendix, only \section* is possibly needed
+
+% use appendices with more than one appendix then use \section to start each
+% appendix you must declare a \section before using any \subsection or using
+% \label (\appendices by itself starts a section numbered zero.)
 %
@@
 \appendices
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
+% you can choose not to have a title for an appendix if you want by leaving the
+% argument blank
 \section{}
@@
 
-% % Can use something like this to put references on a page
-% % by themselves when using endfloat and the captionsoff option.
-% \ifCLASSOPTIONcaptionsoff
-%   \newpage
+% % Can use something like this to put references on a page % by themselves when
+% using endfloat and the captionsoff option. \ifCLASSOPTIONcaptionsoff \newpage
 % \fi

### 2022-03-29 paper/sea-comp/paper.tex
Minor fixes

@@
 the existing applications is significant.
-
 We developed Sea as a means to enable data-placement strategies for scientific
@@
 
-Data-placement strategies can complement existing solutions by facilitating the
-transfer and processing of large datasets across different infrastructures, and
-placing the different file formats at the preferred storage layer to maximize
-efficiency. However, adapting existing scientific applications to Big Data
-frameworks requires considerable effort, necessitating the rewrite of many well
-established applications. Furthermore, as noted in \cite{mehta2017comparative},
-Big Data frameworks are not easy to use for the processing of certain scientific
-data, such as imaging data, due to the level of expertise required to rewrite
-reference implementations using the frameworks.
+Data-placement strategies can complement existing solutions by facilitating
+the transfer and processing of large datasets across different
+infrastructures, and placing the different file formats at the preferred
+storage layer to maximize efficiency. However, adapting existing scientific
+applications to Big Data frameworks requires considerable effort,
+necessitating the rewrite of many well-established applications.
+Furthermore, as noted in \cite{mehta2017comparative}, Big Data frameworks
+are not easy to use for the processing of certain scientific data, such as
+imaging data, due to the level of expertise required to rewrite reference
+implementations using the frameworks.
 
@@
 
-We developed Sea, a user-space data-placement library, that provides efficient
+We developed Sea, a user-space data-placement library that provides efficient
 data placement without reinstrumentation of the underlying application. Given a
-predefined storage hierarchy, Sea can prefetch, cache, flush and evict data to
-and from a permanent storage location. To assist in providing a
+predefined storage hierarchy, Sea can prefetch, cache, flush, and evict data to
+and from a permanent storage location. To provide a
 frame-of-reference for Sea's performance on HPC systems, we designed a

### 2022-03-30 paper/sea-comp/paper.tex
fixes to comp paper

--- 2022-03-29 paper/sea-comp/paper.tex
@@
 \documentclass[10pt,journal,compsoc]{IEEEtran}
-\usepackage{xcolor}
+\usepackage[table]{xcolor}
+\usepackage{booktabs}
 \usepackage{hyperref}
@@
 
+\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
 % \algnewcommand\algorithmicforeach{\textbf{for each}}
@@
 
+Sea removes the need to implement the logic to redirect data to the appropriate
+storage locations. Furthermore, it implements logic to ensure that files are
+written to the best possible location at any given time. It is both pipeline and
+infrastructure agnostic.
+
+As it executes in user-space, root permissions are not required, enabling users
+to use Sea on most Linux-based systems available to them. This differs from Big
+Data file systems which may require elevated permissions to install. The
+overhead of intercepting glibc calls is minimal, and negligible compared to
+system call interception and file systems such as FUSE.
+
+Unlike Big Data frameworks, Sea does not require reinstrumentation of the
+existing pipelines, allowing users to gain an instant performance boost.
+
 Sea is primarily designed to work with workloads that generate large amounts of
@@
 
-\subsubsection{Benefits}
-Sea removes the need to implement the logic to redirect data to the appropriate
-storage locations. Furthermore, it implements logic to ensure that files are
-written to the best possible location at any given time. It is both pipeline and
-infrastructure agnostic.
-
-As it executes in user-space, root permissions are not required, enabling users
-to use Sea on most Linux-based systems available to them. This differs from Big
-Data file systems which may require elevated permissions to install. The
-overhead of intercepting glibc calls is minimal, and negligible compared to
-system call interception and file systems such as FUSE.
-
-Unlike Big Data frameworks, Sea does not require reinstrumentation of the
-existing pipelines, allowing users to gain an instant performance boost.
-
 \subsubsection{Glibc interception}
@@
 Glibc interception is achieved by writing wrappers to existing glibc functions.
-Most importantly, every glibc function accepting a file path needs to be
+Most importantly, every glibc function containing accepting a file path as input parameter needs to be
 wrapped. The wrappers take any input filepath that is located within the
@@
 to 2.33. As glibc file system functions may change over time, Sea behaviour with
-newer and older versions of glibc is undefined. 
+newer and older versions of glibc is unknown. 
 % ldd (GNU libc) 2.17 ldd (Ubuntu GLIBC 2.23-0ubuntu11.3) 2.23 ldd (Ubuntu GLIBC
@@
 
-\subsubsection{Flushing and eviction modes}
-\begin{figure*}
-
+\subsubsection{Memory-management}
+      \begin{table}
       \centering
-    \includegraphics[scale=0.8]{figures/sea-flushevict.pdf}%
-\caption{The four Sea flush and eviction modes }
-\label{fig:sea-comp:fe}
-\end{figure*}
+      \begin{tabular}{ccc}
+        \toprule
+       Mode & .sea\_flushlist & .sea\_evictlist \\
+       \midrule
+       Copy & yes & no \\
+       Remove & no & yes \\
+       Move & yes & yes \\
+       Keep & no & no \\
+
+       \bottomrule
+
+      \end{tabular}
+      \caption{Sea's memory management modes}
+      \label{table:sea-comp:fe}
+      \end{table}
 %\todo{add diagrams here?}
-Flushing and eviction mechanisms give way to four different modes: flush-only,
+Memory-management is an integral component of Sea as deallocating data from cache
+makes room for future files to be stored within the higher-performance storage devices.
+Generalizing an efficient memory-management strategy across all applications is a challenging task.
+Since Sea works on a per-application basis, memory-management in Sea is application-specific, configured
+via two files: \texttt{.sea\_flushlist} and \texttt{.sea\_evictlist}.
+
+Memory management in Sea consists of four different modes, as described in Table~\ref{table:sea-comp:fe}: flush-only,
 evict-only, flush-and-evict and keep. A flush-only is synonymous with a file
@@
 
+A third text file, namely the \texttt{.sea\_prefetchlist}, exists that can be populated by input
+files to be prefetched. At this time, Sea cannot determine when prefetched files are no longer needed by the pipeline, and
+therefore should not be evicted. For files to be prefetched, they must be located within Sea's mountpoint at startup.
+
 %\subsubsection{Configuration File} \subsubsection{Program execution}
@@
       Experimental results can be prone to error and variability, particularly
-      in instances where there a many factors at play, such as in cluster
+      in instances where there a many factors at play, such as in distributed cluster
       environments. To improve confidence in our results, we developed a
-      simplified performance model that could help describe the behaviours found
-      in our experimental results. Since different parallel file systems may
+      simplified performance model that could help describe the behaviours observed in our experimental
+      results. Since different PFS may
       operate differently, our baseline model will be based on Lustre which is
@@
 
-      For data intensive use cases, the makespan models for both Lustre and Sea
+      For data-intensive use cases, the makespan models for both Lustre and Sea
       can be broken down into two components: The amount of time it takes read
       the data and the amount of time it takes to write the data. With more
-      heterogeneous applications (some components are compute intensive whereas
-      others are data intensive), a third component, consisting of compute time,
+      heterogeneous applications (some components are compute-intensive whereas
+      others are data-intensive), a third component, consisting of compute time,
       can be added. Furthermore, latency may also play a significant role
@@
       For the sake of simplicity, the above models assume that the network
-      bandwidth between the compute and data nodes is the same. This, however,
+      bandwidth between the compute and data nodes is the same. However, this
       may not necessarily be the case. Furthermore, the model also assumes that
@@
 
-      Page cache is difficult to summarize accurately and effectively within a
+      Page cache is difficult to summarize accurately within a
       model. For one, we must not only consider available memory and anonymous
@@
       throttling that may occur as a consequence of surpassing the
-      \texttt{dirty\_ratio}. Furthermore, Lustre also has its own user-defined
+      dirty\_ratio. Furthermore, Lustre also has its own user-defined
       settings for how it interacts with the cache that would add additional
@@
       will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
-      storage, and 3) slow parallel file system layer.
+      storage, and 3) slow parallel file system layer. It is important to note that
+      the model assumes that I/O never occurs in distinct storage layers in parallel.
+      Although Sea tries to exhaust the layers, starting from the fastest, it is Nevertheless
+      possible that Sea writes to two storage layers at the same time, for instance, when one becomes full
+      and another parallel thread needs to perform a write.
 
@@
       The local disk  makespan model is similar to the tmpfs makespan model,
-      however, we must ensure to disregard any data that has already been
+      although we must disregard any data that has already been
       written to tmpfs. Furthermore, in Sea, it is possible to leverage however
@@
       \centering
-      \begin{tabular}{@{}|c|c|c|@{}}
-       \hline
+      \begin{tabular}{ccc}
+       \toprule
        Storage layer & Action & Avg. bandwidth (MiB/s) \\
-              \hline
+       \midrule
       \multirow{3}{*}{tmpfs} & read & 6676.48 \\
@@
       & write & 2560.00 \\
-       \hline
+      \midrule
        \multirow{3}{*}{local disk} & read & 501.70  \\
@@
        & write & 426.00 \\
-       \hline
+       \midrule
        \multirow{3}{*}{Lustre} & read & 1381.14 \\
@@
 
-       \hline
+       \bottomrule
 
@@
             Our experiments were executed on a Centos 8.1 (Linux kernel 4.18.0)
-      cluster with 8 compute nodes, a 4 data node Lustre server with 1 dedicated
+      cluster with 8 compute nodes, a 4 data node Lustre server with 1
       metadata node. Each compute node is equipped it two Intel(R) Xeon(R) Gold
@@
       disabled and Lustre dirty writes is limited to \SI{1}{\giga\byte} per OST.
+
+      There was no background load on the cluster for the duration of the experiments.
 
@@
 
-      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea with the
+      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea using
       in-memory configuration speeds up execution the majority of the
@@
       Figure~\ref{fig:sea-comp:processes} at 32 processes, where the speedup is
-      nearly 3$\times$. We believe that the speedup here is primarily due to the
+      nearly 3$\times$. We believe that the speedup is primarily due to the
       contention on Lustre. While each local disk has an average of 5 processes
@@
       achieved the greatest speedup at 5 nodes, with a speedup of 2.4$\times$.
-      Similar to our experiments with multiprocessing, the speedup here appears
+      Similar to our experiments with multiprocessing, the speedup appears
       to be due to increased contention on Lustre. However, there is one main
-      difference here, and that is that only Sea is experiencing increased
+      difference, and that is that only Sea is experiencing increased
       contention, as the contention within the compute nodes is fixed. Both the
@@
 
-      \subsection{Model accurately predicts performance trends}
-
-      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model correctly
-      describes The bounds for all but two experiments: Experiment 3
+      \subsection{Performance trends described by the model}
+
+      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model can
+      describe the performance trends for all but two experiments: Experiment 3
       (Figure~\ref{fig:sea-comp:iterations}) and Experiment 4

### 2022-04-04 paper/sea-comp/paper.tex
fixed column width

@@
 
-    \begin{figure*}
+    \begin{figure}
 
         \centering
-        \includegraphics{figures/flushall.pdf}%
+        \includegraphics[width=\columnwidth]{figures/flushall.pdf}%
         \caption{Comparison of the different Sea modes with Lustre on the
@@
     \label{fig:sea-comp:flush}
-    \end{figure*}
+    \end{figure}
 

### 2022-04-13 paper/sea-comp/paper.tex

@@
 \usepackage{tabularx}
+\usepackage{xspace}
 
@@
 #1}}}
+\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
 
@@
       the data is located. Historically, storage and compute nodes were distinct
-      layers, data transfer to the compute location. Rather than transferring
+      layers, \TG{missing word?} data transfer to the compute location. Rather than transferring
       large amounts of data over the network to compute tasks, which could incur
@@
 Glibc interception is achieved by writing wrappers to existing glibc functions.
-Most importantly, every glibc function containing accepting a file path as input parameter needs to be
+Most importantly, every glibc function accepting a file path as input parameter needs to be
 wrapped. The wrappers take any input filepath that is located within the
-user-provided Sea mountpoint and convert it a filepath pointing to the best
+user-provided Sea mountpoint and convert it to a filepath pointing to the best
 available storage device.
@@
 
-\subsubsection{Memory-management}
+\subsection{Memory-management}
       \begin{table}
@@
       in instances where there a many factors at play, such as in distributed cluster
-      environments. To improve confidence in our results, we developed a
+      environments. To improve confidence in our experimental results, we developed a
       simplified performance model that could help describe the behaviours observed in our experimental
@@
       include it in our model. The makespan of an application I/O to and from
-      page cache, $M_{c}$, can be described as the following:
-      Equation~\ref{eq:sea-comp:lustrenpc} where it is assumed that none of the
+      page cache, $M_{c}$, can be described as in
+      Equation~\ref{eq:sea-comp:lustrenpc}, where it is assumed that none of the
       data is written or read from page cache.
@@
       the model assumes that I/O never occurs in distinct storage layers in parallel.
-      Although Sea tries to exhaust the layers, starting from the fastest, it is Nevertheless
+      Although Sea tries to exhaust the layers, starting from the fastest, it is nevertheless
       possible that Sea writes to two storage layers at the same time, for instance, when one becomes full
@@
       \begin{equation}\label{eq:sea-comp:mst}
-          M_{st} = \frac{D_{tr}}{cC_{r}} + \frac{D_{tw}}{cC_{w}}
+          M_{St} = \frac{D_{tr}}{cC_{r}} + \frac{D_{tw}}{cC_{w}}
       \end{equation}
@@
 
-      To evaluate the Lustre and Sea models defined in
-      Subsection~\ref{ss:sea-comp:model} and the real performance of both file
-      systems with data intensive applications, we wrote a simple Python
+      To evaluate the performance of Sea with data intensive applications, we wrote a simple Python
       application based off of Algorithm~\ref{alg:sea-comp:incrementation}.
@@
       trying to write concurrently to it at 32 processes, each Lustre has around
-      3 concurrent processes while to it. Whereas one might assume this means
+      3 concurrent processes while to it \TG{this sentence doesn't make sense, revise}. Whereas one might assume this means
       Lustre would have less contention overall and should perform better, there
@@
     be a single flush and evict process, and therefore, only a single flush and
-    evict process will exist. However, if Sea is launched many times on a given
+    evict process will exist \TG{this sentence is circular, revise}. However, if Sea is launched many times on a given
     node, there will be many flush and evict processes which may interfere with
@@
 
-    On shared cluster environments, such as though found in high-performance
+    On shared cluster environments, such as those found in high-performance
     computing clusters, it is less likely that the PFS would be so underused
@@
 
-\appendices
-% you can choose not to have a title for an appendix if you want by leaving the
-% argument blank
-\section{}
-Appendix two text goes here.
-
-% use section* for acknowledgment
-\ifCLASSOPTIONcompsoc
-  % The Computer Society usually uses the plural form
-  \section*{Acknowledgments}
-\else
-  % regular IEEE prefers the singular form
-  \section*{Acknowledgment}
-\fi
+% \appendices
+% % you can choose not to have a title for an appendix if you want by leaving the
+% % argument blank
+% \section{}
+% Appendix two text goes here.
+
+% % use section* for acknowledgment
+% \ifCLASSOPTIONcompsoc
+%   % The Computer Society usually uses the plural form
+%   \section*{Acknowledgments}
+% \else
+%   % regular IEEE prefers the singular form
+%   \section*{Acknowledgment}
+% \fi
 

### 2022-04-13 paper/sea-comp/paper.tex
addressed comments

@@
       the data is located. Historically, storage and compute nodes were distinct
-      layers, \TG{missing word?} data transfer to the compute location. Rather than transferring
+      layers, transferring data to the compute location whenever necessary. Rather than transferring
       large amounts of data over the network to compute tasks, which could incur
@@
       contention on Lustre. While each local disk has an average of 5 processes
-      trying to write concurrently to it at 32 processes, each Lustre has around
-      3 concurrent processes while to it \TG{this sentence doesn't make sense, revise}. Whereas one might assume this means
+      trying to write concurrently to it at 32 processes, each Lustre disk has around
+      3 concurrent processes writing to it. Whereas one might assume this means
       Lustre would have less contention overall and should perform better, there
@@
 
-    The flush-and-evict process may affect overall application parallelism. It
-    would depend on how many Sea processes are started on the compute node. If
+    The flush-and-evict process may affect overall application parallelism if multiple Sea instances
+    are launched on a given node. If
     only a single instance of Sea is called on a compute node, there will only
-    be a single flush and evict process, and therefore, only a single flush and
-    evict process will exist \TG{this sentence is circular, revise}. However, if Sea is launched many times on a given
+    be a single flush and evict process. However, if Sea is launched many times on a given
     node, there will be many flush and evict processes which may interfere with

### 2022-05-06 paper/sea-comp/paper.tex
additional fixes

@@
 \begin{abstract}
-    The recent influx of scientific data has lead to scientific computing
-transitioning from compute intensive to data intensive. Whereas many Big Data
-Frameworks exist that minimize the cost of data transfers, few scientific
-applications adopt data-placement strategies as the efforts required to update
-the existing applications is significant.
+    The recent influx of open scientific data has contributed to the
+    transitioning of scientific computing from compute intensive to data intensive.
+    Whereas many Big Data
+frameworks exist that minimize the cost of data transfers, few scientific
+applications integrate these frameworks or adopt data-placement strategies to mitigate the costs.
+Scientific applications commonly rely on well-established command-line tools that would require complete reinstrumentation
+in order to incorporate existing frameworks.
 We developed Sea as a means to enable data-placement strategies for scientific
@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-\IEEEPARstart{E}{fficient} data-placement strategies have become essential in
+\IEEEPARstart{E}{fficient} data-placement strategies have become essential to
 minimizing Big Data processing overheads. These strategies, such as data
@@
 Spark~\cite{zaharia2016apache} and Dask~\cite{rocklin2015dask}, improve runtime
-performance through the incorporation of these data-placement strategies.
+performance through the incorporation of data-placement strategies.
 
@@
 computing applications executing on High Performance Computing (HPC) clusters.
-The tools of particular interest to us consist of various tasks that communicate
-to each other via a POSIX-compliant file system. Whereas the tools can be
-platform-agnostic, we assume users will leverage them within an HPC context,
+The applications of interest to us consist of various tasks that communicate
+to each other via a POSIX-compliant file system. Whereas the application can be
+platform-agnostic, we assume users will leverage them within an linux-based HPC context,
 particularly when processing large volumes of data, due to their inherent
@@
 predefined storage hierarchy, Sea can prefetch, cache, flush, and evict data to
-and from a permanent storage location. To provide a
+and from a permanent storage location.
+
+To provide a
 frame-of-reference for Sea's performance on HPC systems, we designed a
@@
       (HDFS)~\cite{shvachko2010hadoop} or Alluxio~\cite{alluxio} to facilitate
-      data locality. Furthermore, access to these compute nodes is temporary,
+      data locality. Furthermore, access to these compute nodes is ephemeral,
       with allocation duration enforced by a batch scheduler. Therefore, any
@@
       are shared amongst all compute nodes within a cluster, typically connected
-      to each node via high-performance network interconnect such as
+      to the nodes via high-performance network interconnect like
       InfiniBand~\cite{infiniband}. I/O overheads of shared PFS can be incurred
@@
 
-      Strategies to mitigate the impacts of data transfers in data-intensive
-      workloads are required to ensure optimal processing time. Early Big Data
+Early Big Data
       frameworks, notably MapReduce~\cite{dean2008mapreduce} and Apache
@@
       These strategies are known as data locality and in-memory computing.
-
       Data locality ensures that compute tasks are scheduled nearest to where
@@
       intermediate storage layer, known as a Burst Buffer, to improve
-      application I/O-related overheads. To facilitate an applications
+      application I/O-related overheads. To facilitate an application's
       interaction between application and storage layer, distributed file
@@
       subsequent accesses to that data may be done entirely in memory so long as
-      that data has not already been evicted. Similarly, for writes to a file
+      the data has not already been evicted. Similarly, for writes to a file
       system with writeback cache enabled, the file will be written to memory
@@
       calls made to the GNU C library (glibc). There are limitations to using
-      glibc interception; the application must be dynamically linked and make
+      glibc interception; the application must be dynamically-linked and make
       glibc calls. However, statically-linked applications are uncommon and
@@
       machines, the predominant OS of HPC clusters, will make glibc calls.
-      Nevertheless, there are alternatives to glibc intercept that can bypass
+      Nevertheless, alternatives to glibc intercept exist that can bypass
       the aforementioned issues, such as system call interception, but they
@@
       BurstFS~\cite{burstfs} and GekkoFS~\cite{gekkofs} are two user-space file
-      systems that are specifically designed for Burst Buffers. That is, both of
-      these file systems exist only during the time-window in which the
-      resources are allocated. Furthermore, they both exist entirely within
-      user-space using glibc interception. While these implementations simplify
-      usage of burst buffers for users, their installation and configuration is
-      similarly out-of-scope for a typical HPC user.
+      systems that are specifically designed with HPC Burst Buffers. That is, both of
+      these file systems are designed to support the ephemeral nature of HPC compute nodes and their associated storage.
+      Additionally, these implementations exist entirely in user-space leveraging the LD\_PRELOAD trick to intercept
+      necessary library calls. Whereas these libraries incorporate a client-server architecture to aid in transferring data
+      between nodes, Sea opts for a more lightweight approach ensuring decentralization, statelessness and leveraging underlying
+      file systems for communication and consistency.
 
@@
 Sea requires minimal configuration for use. In order to work, Sea requires the
-user to specify at least two storage devices, a fast temporary one (i.e. cache)
-and a slower long-term storage one. This could be RAM and SSD if working on a
-single node, or a compute-local SSD and a shared parallel file system, in the
+user to specify at least two storage devices, a fast temporary device used as cache
+and a slower long-term storage device. This could be RAM and SSD if working on a
+single node, or a compute-local SSD and a shared PFS, in the
 case of an HPC cluster. Ideally, a user will provide a multitude of ephemeral
@@
 %\todo{add diagrams here?}
-Memory-management is an integral component of Sea as deallocating data from cache
+Memory-management is an integral component of Sea as de-allocating data from cache
 makes room for future files to be stored within the higher-performance storage devices.
@@
 
-Memory management in Sea consists of four different modes, as described in Table~\ref{table:sea-comp:fe}: flush-only,
-evict-only, flush-and-evict and keep. A flush-only is synonymous with a file
-copy in that a file that already exists within a Sea cache will be copied to
+Memory management in Sea consists of four different modes, as described in Table~\ref{table:sea-comp:fe}: copy,
+remove, move and keep. A copy copies a file located within the Sea cache to
 persistent storage. This operation is particularly useful when a file will be
@@
 needs to be communicated between other nodes or needs to be persisted for
-post-processing analysis. To flush-only a file, the file must only be specified
+post-processing analysis. To copy a file, the file must only be specified
 in the \texttt{.sea\_flushlist}.
 
-Evict-only is equivalent to a remove operation. This mode is for files that are
+Remove is for files that are
 located within a Sea cache, but do not need to be persisted and will not be
 reused by the pipeline and therefore do not need to occupy cache space.
-Evict-only is useful for removing unnecessary application log files. To
-evict-only a file, the file must only be specified in the
+Remove is useful for removing unnecessary application log files. To
+remove a file, the file must only be specified in the
 \texttt{.sea\_evictlist}.
 
-Similar to the flush-only, the flush-and-evict mode copies data from cache to
+Similar to the copy, the move mode copies data from cache to
 persistent storage. However, the assumption here is that these files will not be
 reused by the application and therefore need not occupy space in cache.
-Flush-and-evict is therefore synonymous with a move operation and is invoked on
+Move is therefore synonymous with a copy-and-remove operation and is invoked on
 files that are specified in both the \texttt{.sea\_flushlist} and
@@
       makespan of an application reading and writing to Sea. For our model, we
-      will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
+      will assume three storage layers: 1) fast tmpfs, 2) intermediate local SSD
       storage, and 3) slow parallel file system layer. It is important to note that

### 2022-05-06 paper/sea-neuro/paper-neuro.tex
updates to neuro intro

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\usepackage[binary-units]{siunitx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\linenumbers

\title{Automated neuroimaging data-management with Sea}

\author{Val\'erie Hayot-Sasson}
\author{Tristan Glatard}
\affil{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}

\begin{abstract}

	Neuroimaging open-data initiatives have led to increased availability of large
  scientific datasets. Whereas these datasets are shifting the processing
  bottleneck from compute-intensive to data-intensive, current standardized
  analysis tools lag behind in the adoption of strategies to mitigate the costs
  associated with data transfers. A major challenge in adapting neuroimaging
  applications for data-intensive processing is that they must be entirely
  rewritten.

  To facilitate data management for standardized neuroimaging tools, we
  developed Sea, a library that intercepts and redirects application read and
  write calls to minimize data transfer time. In this paper, we investigate the
  performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
  SPM and AFNI), using three neuroimaging datasets of different sizes
  (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
  computing clusters. Our results demonstrate that given the features of the
  applications and datasets, Sea provides large speedups (up to 32$\times$) when
  the shared file system's (e.g. Lustre) performance is deteriorated. When the
  shared file system is not overburdened by other users, performance is
  unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
  where it is not useful. Overall, Sea is beneficial, even when performance gain
  is minimal, as it can be used to limit files on PFS.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
%  Click the title above to edit the author information and abstract

\thispagestyle{empty}

%\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists or tables should be included. Structure of the main text is provided below.

  \section{Introduction}\label{sec:sea_neuro:introduction}

    The recent explosion in publicly available neuroimaging data has lead to new data
    management challenges, from storage infrastructure application processing times. To meet the storage, accessibility and
    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers.
    Standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids}, have been implemented to facilitate the sharing
    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide versioning and provenance capture of data.
    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches.
    While all these solutions to Big Data-related data management exist, certain aspects, such as processing-related data-transfer overheads
    have received limited attention.

    The current largest neuroimaging datasets,
    the Human Connectome Project (HCP)~\cite{HCP} and the UK
    Biobank~\cite{ukbiobank}, reach up to Petabytes of data. For datasets of this
    volume, parallel high-capacity remote storage solutions are commonly relied
    upon as the persistent storage layer during processing. 
    Despite their respective high-scalability, these remote file systems cannot typically
    compete with the performance of compute-local storage (e.g. DRAM and SSDs). Should the pipelines 
    need to transfer large amounts of data to these remote storage locations, significant data
    processing overheads may occur.

    Big Data in neuroimaging can be present in two formats: 1) very large files
    (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
    population of individuals, such as those found in MRI datasets, that
    collectively amount to very large datasets. Differences in the number and
    size of files result in different data transfer overheads. Whereas large
    file I/O is predominantly impacted by the underlying storage bandwidth,
    small file I/O results in a non-negligible latency overhead.

    Big Data frameworks have addressed the performance limitations arising from large data transfers via the
    implementation of two strategies: data locality and in-memory computing~\cite{zaharia2016apache, rocklin2015dask}.
    The collective aim of these strategies is to reduce overheads related to large data
    transfers by leveraging compute-local storage as cache.
    While these frameworks have been used
    to process neuroimaging data~\cite{rokem2021pan,thunder,boubela2016big},
    their popularity within the domain remains low due to efforts required to
    rewrite standardized neuroimaging toolboxes to utilize the frameworks.
    % Although it is possible to use the tools within the Big Data frameworks,
    % optimizations like in-memory computing would not be leveraged due to the
    % fact that neuroimaging tools are meant to be used as command-line tools
    % and do not provide interfaces that enable the data to be transferred
    % in-memory.

    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
    also do not benefit from processing-related data management.
    Although engines such as Nipype~\cite{nipype} and
    Joblib~\cite{joblib} do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into their resulting workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system or library that enable the strategies.

    % Data management-specific scientific frameworks, such as
    % DataLad~\cite{halchenko2021datalad} aim to increase performance by focusing
    % on reproducibility, thereby reducing the need to recompute
    % results~\cite{wagner2021fairly}. We aim to complement frameworks such as
    % these, by reducing data transfer times when computation is required.

    %Big Data in neuroimaging can be present in two formats: 1) very large
    %files, such as those found in the BigBrain~\cite{bigbrain}, and
    %2)\todo{what is that 2?} large datasets made up of small files, such as
    %those typically found in fMRI datasets.

    % When processing, large datasets can result in longer processing times even
    % when compute resources are ample. This is a direct result of the
    % underlying storage used by the applications to read and write data. Since
    % large datasets require a significant amount of available storage space,
    % more often than not, slower but larger storage is selected, resulting in
    % longer data transfer times during processing. With the Sea filesystem, we
    % aim to leverage all available storage in order to reduce overall data
    % transfer time during pipeline execution.

    Many researchers rely on one of two systems to meet their storage and
    computing needs: 1) High Performance Computing (HPC) clusters and 2) the
    cloud. Whereas the cloud simplifies data sharing and gives researchers
    access to a wide variety of infrastructures, HPC clusters are a
    cost-effective solution to accessing a wide array of resources for
    researchers. %In this paper, we will focus on HPC processing.

    HPC systems rely on scalable network-based parallel file systems (e.g.
    Lustre) for storage. While such file systems offer excellent performance,
    they are shared between all the users on the cluster, meaning that users
    with data-intensive applications can effectively deteriorate the performance
    of the shared file system for all users on the cluster. Solutions for
    improving shared file system performance include
    throttling~\cite{huang2020ooops} the data-intensive workloads or
    recommending the use of Burst Buffers~\cite{bb} (e.g. reserving a compute
    node for storage or leveraging local compute storage during processing). The
    latter, however, requires that the user manages their data to and from the
    Burst Buffer, unless a Burst Buffer file system, such as
    BurstFS~\cite{burstfs} or GekkoFS~\cite{gekkofs} is installed.

    In order for a file system to be mounted by an HPC cluster user, it must be
    loadable without administrative privileges. As the applications are
    typically made to interact with POSIX-based file systems, the new file
    system must also be compliant to the format. One method to ensure that these
    conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick is
    used to intercept select library calls and redefine their behaviour. It has
    been used in many projects and to create lightweight userspace versions of
    file systems~\cite{xtreemfs,burstfs,gekkofs}.

    We present Sea~\cite{sea-comp}, a data-management library designed to reduce processing-related overheads. We view Sea as a complement
    to the existing stack of neuroimaging tools and standards meant to facilitate the processing of neuroimaging Big Data.    
    % \tristan{Intro logic is good but writing could be polished}
    % \tristan{Outlining the content of the paper would be useful}

    % \section{Related Work}

    % Put in intro. summary of computing landscape for neuroimaging. HCP,
    % CBRAIN, nipype, joblib. sea can be used in all these contexts.

    \section{Materials and Methods}

    \subsection{Sea}

\begin{figure*}

    \centering
    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
\caption{The Sea data-management library overview}
\label{fig:seaneuro:diagram}
\end{figure*}
    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system (more specifically, the GNU C library,
    glibc) calls on Linux systems. This enables Sea to redirect write calls
    aimed for a slower storage device to a faster device whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the detailed provided in an
    initialization file called \texttt{sea.ini}. This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority.

    We define all locations in which Sea can redirect writes to as a cache. Such
    caches include any available storage located on the compute nodes, such as
    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs).

    As neuroimaging pipeline results are typically required post-processing and
    HPC compute-local resources are only accessible during the reserved
    duration, Sea provides functionality to flush and evict data to persistent
    shared storage. This is accomplished via a separate thread (known as the
    ``flusher'') that moves data from the caches to long-term storage. A
    separate thread is used, in this case, to avoid interrupting ongoing
    processing with data management operations. Users must inform Sea of files
    that need to be persisted to storage within a file called
    \texttt{.sea\_flushlist}, and temporary files which can be removed from
    cache within a file called the \texttt{.sea\_evictlist}. Both these files
    can be populated using regular expressions to denote the paths to flush and
    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
    operation and simply move the file from local to persistent storage rather
    than copying it. Files that will be reused by the application should only be
    flushed rather than flushed and evicted, as files can benefit from speedups
    related to reading from fast rather than slow storage. Sea currently also
    provides a rudimentary prefetch thread that can move files located within
    Sea to the fastest available cache. To use Sea's prefetch
    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
    populated using regular expressions like the flushing and eviction files.

    To interact with Sea, a mountpoint is used. The mountpoint
    is an empty directory that behaves as a view to all the files and
    directories stored within Sea. In order to keep track of the locations of
    files within the mountpoint, Sea mirrors the directory structure of each
    storage location across all caches. It is generally advisable to provide
    empty storage locations for Sea to use as the mirroring of large directories
    can take some time.

    Sea can easily be executed directly using the available containers on the
    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.

    \subsection{Datasets and Pipelines}

    To determine the performance gain Sea brings to neuroimaging analyses, we
    must evaluate the value of Sea on a variety of neuroimaging applications.
    For our analysis, we selected different fMRI preprocessing applications as
    fMRI processing has some of the most well-established tools for neuroimaging
    and some of the largest datasets. Of course, different modalities and tools
    may result in vastly different data access patterns and compute times, we
    use fMRI pre-processing as our baseline for the benefits that can be
    obtained with Sea.

    \subsubsection{Datasets}
    fMRI datasets may vary greatly in total number of images and number of
    volumes within each image. To adequately capture the diversity of datasets
    and the applicability of Sea, we have selected three datasets of varying
    time and space resolutions (Table~\ref{table:sea-neuro:data}): 1)
    OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD
    dataset~\cite{preventad}, and 3) the HCP dataset~\cite{HCP}. The ds001545
    dataset is a total of \SI{46}{\giga\byte} (1,778 files) and consists of data
    collected from 30 participants in a single session watching 3 different
    clips (i.e. three runs/participants) of The Grand Budapest Hotel.

    The PREVENT-AD dataset is an open dataset consisting of data from 330
    participants with a familial history of Alzheimer's Disease. At the time of
    the experiments, the dataset contained \SI{255}{\giga\byte} (53,061 files)
    originating from the 308 subjects available on
    DataLad~\cite{halchenko2021datalad}.

    The HCP project, whose aim is to characterize brain connectivity using data
    collected from 1,200 subjects, is the largest of the three datasets. At the
    time of our experiments, the dataset obtained consisted \SI{85}{\tera\byte}
    (15,716,060 files) of data collected from 1,113 subjects.

    By experimenting with datasets of different spatial and temporal
    resolutions, the amount of compute may be affected. Naturally, with more
    data there may inevitably be more computations performed, but also, the code
    itself may perform entirely different computations as a result of
    differences in resolution. Moreover, parallelism and program I/O behaviour
    may be impacted depending on the particular characteristics of the data.
    Capturing this potential variety in processing will provide us with a better
    understanding of when and where Sea can be useful.

    \begin{table*}[t]
      \small\centering
    \resizebox{\textwidth}{!}{\begin{tabular}{|c c c c c|}
      \hline
      Dataset & Total Size (MB) & Total Number of images & Number of images per experiment & Total compressed size processed (MB) \\
      \hline
      \multirow{3}{*}{PREVENT-AD} & \multirow{3}{*}{289,532} &
      \multirow{3}{*}{53,061} & 1 & 52\\
      & & & 8 & 402\\
      & & & 16 & 732\\
      \hline
      \multirow{3}{*}{ds001545} & \multirow{3}{*}{27,377} &
      \multirow{3}{*}{1,778} & 1 & 282\\
      & & & 8 & 2115\\
      & & & 16 & 4167\\
      \hline
      \multirow{3}{*}{HCP} & \multirow{3}{*}{83,140,079} &
      \multirow{3}{*}{15,716,060*} & 1 &  1301\\
      & & & 8 & 5998\\
      & & & 16 & 8328\\

      \hline

      \hline
    \end{tabular}} \footnotesize{*Number estimated based on data we had access
    to}\\
    \caption{Dataset characteristics. Compressed size is listed, since all
    applications, except for SPM, processed compressed
    data}\label{table:sea-neuro:data}
    \end{table*}

    %just for example run dataset size, resolution, resolution of voxels matrix
    %size pixdim1X2 number of slices pixdim number of volumes pixdim4 inplane
    %resolution pixsize1x2 slice thickness pizsize3 in mm repetition time
    %pixsize4 in s

    %HCP resting state + x stats

    %# files written and output size, duration and memory consumption.
    \subsubsection{fMRI Preprocessing Pipelines}

    Similarly to datasets, preprocessing pipelines can also vary in duration as
    a result of methodological differences. Thus, we preprocessed each dataset
    using three standard preprocessing pipelines available within the FSL~\cite{fsl},
    AFNI~\cite{cox1996afni}, and SPM~\cite{spm} toolboxes.
    Table~\ref{table:seaneuro:pipelines} provides a reference of the differences
    in computation and data-intensiveness of the different pipelines on a single
    subject for all three of the datasets.

    Each tool was set to only run the preprocessing pipeline. Scripts used to
    execute each pipeline can be found
    at~\url{github.com/valhayot/sea-slurm-scripts}.

    For the FSL preprocessing pipeline, we used FEAT configured with many of
    the default parameters. We set the following parameters: slice-timing
    correction (interleaved), intensity normalization and non-linear
    registration to standard space. FSL version 6.0.4 was used.

    %% Add SPM here
    To preprocess the datasets with SPM12, we reused the template described in
    \cite{haitas2021age}, with fieldmap correction removed. As the template was
    designed for a specific dataset, we updated the paths to the fMRI and
    anatomical images, as well as the dimensions, such as the number of volumes
    and slices and the TR and TA, for each dataset. Furthermore, we set the
    slice order to interleaved in all cases.

    For AFNI preprocessing, the pipeline was configured to perform slice timing,
    alignment to Talaraich space, registration, smoothing and brain masking.
    While AFNI does have some parallelized components that can be controlled via
    environment variables, we chose to let the pipeline use as much parallelism
    as required for its execution. AFNI version 21.1.02 was used.

    All the application Dockerfiles with Sea pre-installed are available
    at~\url{https://github.com/ValHayot/Sea/tree/master/containers/applications}.

    % We set fMRIPrep to run with \texttt{--fs-no-reconall} flag set and set the
    % \texttt{--bids\-database-dir} to a tmpfs location. Furthermore, the seeds
    % were fixed to ensure reproducible results. Unlike the other tools where
    % only a single fMRI image was preprocessed per process, it was not possible
    % to do so with fMRIPrep. Thus, fMRIPrep preprocessed entire subject
    % directories at a time

    %%{provide subject #/session#} of the ds001545 dataset.

    \begin{table*}[t]
      \small\centering
      \resizebox{\textwidth}{!}{\begin{tabular}{|c c c c c c|}
      \hline
      Tool & Dataset &  Output Size (MB) & Total glibc calls & Glibc Lustre calls
      & Compute time (s)* \\
      \hline
      \multirow{3}{1em}{AFNI} & PREVENT-AD & 540 & 272,342 & 4,118 & 103.25 \\
      & ds001545 & 3,063 & 281,660 & 4,340 & 280.30 \\
      & HCP & 18,720 & 305,555 & 5,137 & 816.16 \\
      \hline
      \multirow{3}{1em}{FSL Feat} & PREVENT-AD & 254 & 191,148 & 28,099 &
      1,338.29 \\
      & ds001545 & 551 & 192,404 & 28,371 & 2,145.96 \\
      & HCP & 1,608 & 192,445 & 28,997 & 6,596.46 \\
      \hline
      \multirow{3}{1em}{SPM} & PREVENT-AD & 331.0 & 42,329 & 18,257 &  483.67 \\
      & ds001545 & 744 & 54,481 & 27,770 & 446.53 \\
      & HCP & 2,083 & 62,234 & 33,477 & 715.43 \\

      \hline
    \end{tabular}}
    \footnotesize{$*$ Measurements taken on the dedicated cluster}
    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}\label{table:seaneuro:pipelines}
    \end{table*}

    \subsection{Controls}
    Our objective is to measure the speedup brought by Sea on real neuroimaging
    pipelines. Although Sea is agnostic to the internals of the pipelines and
    their output results, incorrect parameter selection and resulting outputs
    may result in I/O patterns that would not be observed had the pipeline been
    executed properly.

    To validate that the outputs produced were sufficiently correct, we
    performed visual quality control of the preprocessed images to ensure that
    they were correctly skull-stripped and registered to the template brain (MNI
    template). We repeated this step for all datasets and all pipelines
    processing a single image. In addition, we compared output number of files
    and output size to ensure that the pipeline was not producing different
    outputs depending on the file system used.

    % compare md5sums of uncompressed files.

    \subsection{Infrastructure}

    Our experiments were executed on two different clusters: 1) a dedicated cluster
    and 2) and a production cluster. We chose this experimental setup as the dedicated
    cluster allowed us to understand Sea's performance without activity of other users.
    The production cluster gave us insight into what performance typical users might
    be able to expect on a shared cluster.

    The dedicated cluster was composed of 8
    compute nodes with \SI{256}{\gibi\byte} memory and \SI{125}{\gibi\byte}
    tmpfs, running Centos 8.1.1911.
    Storage on the dedicated cluster consisted of 4 Lustre ZFS storage nodes with 44
    HDD object storage targets (OSTs) and 1 metadata server (MDS) with a single
    metadata target (MDT). The storage nodes were connected to the compute nodes
    via a 20Gbps ethernet connection. The cluster's hardware was similar to that
    of Beluga.

    Compute Canada's Beluga cluster was used as the production cluster. It consisted of a total of 977 compute nodes, of which we
    used up to 16 general compute nodes at a time with \SI{186}{\gibi\byte} of
    available memory, \SI{480}{\gibi\byte} of local scratch SSD space, 2 Intel
    Gold 6148 Skylake @ 2.4 GHz CPUs and Centos 7 installed. Each compute node
    was connected to the Lustre scratch file system via an Mellanox Infiniband
    EDR (100 Gbps) network interconnect. The Lustre scratch file system consists
    of a total \SI{2.6}{\pebi\byte} of space, with 38 OSTs of
    \SI{69.8}{\tebi\byte} each and two MDTs of \SI{3.3}{\tebi\byte}.

    On the dedicated cluster, we ran the experiments with and without
    Lustre degradation, which was induced by ``busy writers''. Without busy
    writers, the fMRI preprocessing pipelines have exclusive access to the
    cluster. With busy writers, the fMRI preprocessing pipelines are executed
    alongside a Spark application that continuously reads and writes approximately
    1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
    and writes. In our experiments, we either had 6 nodes each executing the
    Spark application or no busy writers.

    Our experiments generally did not use prefetching, flushing or eviction. The
    only pipeline configured to prefetch was SPM, as the input data is otherwise
    read through a memmap (i.e. only loading necessary portions of the file to memory) with Lustre. Flushing would normally be necessary with
    preprocessing as the user would always require the output data. To
    investigate the impacts of flushing on these pipelines, we performed a
    separate set of experiments where AFNI and SPM would flush all data
    produced, on the production cluster.

    For all experiments, we compared the performance of Sea with that of Lustre
    alone (Baseline). To ensure that system performance was equivalent between
    Sea and our Baseline, we executed Sea and Baseline pipelines together.

    All experiments were executed using 1, 8 and 16 processes. Each process consisted of a single application call processing a single fMRI image.

    %Busy writers
    \section{Results}

% \tristan{I find this section lacks a clear focus. What are the key points from
% these results? You should focus on them and possibly name your paragraphs
% accordingly, for improved clarity. My takeaways would be the following:
% \begin{enumerate} \item In controlled conditions, important speed-ups are
% observed. Mention min, average and max speed-up for all conditions. \item
% Speed-up is correlated with baseline makespan \item Other determinants of
% speed-up are unclear \item Overhead is minor \item Speed-ups are also observed
% in production cluster \end{enumerate}
% }
\begin{figure*}

\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
\label{fig:seaneuro:slashbin}
\end{figure*}

\subsection{Speedups observed in controlled environment}

In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
variable, speedups are obtained by using Sea when Lustre
performance has been degraded by busy writers. On average, a single SPM
process preprocessing a single HCP image produced the greatest speedup
(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process was the next fastest pipeline, with an average speedup of
(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that
AFNI would be the application that would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
AFNI performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affects both Sea and
Baseline executions an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. AFNI still
obtained an average of 2$\times$ speedup across all experimental conditions.

FSL Feat, in contrast, appears to be the most compute-bound of the three
applications. Not only did it spend an extensive amount of time computing, the
amount of output data generated was least of the three. Due to the
compute-intensive nature of FSL Feat, it is expected that it resulted in the
least significant speedups.

Our results indicate that HCP, on average, obtained the greatest speedups with
Sea. Out of the three datasets, HCP has the largest images (see
Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
single image) results in the next largest speedups, followed by PREVENT-AD, the
dataset with the smallest functional images.

For both the pipelines and the datasets, we noticed that obtained speedups
decrease with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process is likely attempting to utilize all cores resulting in longer wait times
with the increase in threads. With this unintentional additional wait time that
is added to the applications due to compute contention, the benefits of using
Sea becomes more limited.

\subsection{Speedup correlated with Lustre degradation}

Baseline performance without busy writers is comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
busy writers, Baseline performance may deviate dramatically from the identity
line. We observe that the more Baseline deviates from this identity line, the
greater the speedup.

We do occasionally observe slowdowns from using Sea. These slowdowns
may arise from the initial read of the data, as they appear to occur less
frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.

\subsection{Overhead of Sea is minimal}

\begin{figure*}
\centering
\includegraphics[width=\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
\label{fig:seaneuro:tmpfs}
\end{figure*}

To determine the overhead of Sea, we compared the average performance of Sea with
the performance of pipeline executing entirely within memory
(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
behave more or less identically. Given these results, we conclude that the
overhead of Sea is minimal and that the differences in makespan may be
due to Sea's initial reading of the data from Lustre.

\subsection{Speedups observed in production environments}

\begin{figure*}

\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafull}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
\label{fig:seaneuro:beluga-noflush}
\end{figure*}

\begin{figure*}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}

  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
  shows us that although performance in production clusters is highly variable,
  we can still obtain important speedups. It is very likely that when we
  executed our experiments without flushing
  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
  resulting in Sea and Baseline performing
  quite similarly to each other.

  Our results with flushing enabled confirm this theory. When running with
  flushing enabled, which would have the additional overhead of having to
  ensure that all the data is copied to Lustre, we notice that occasionally, we
  obtain very large speedups with both SPM and AFNI, suggesting that Lustre
  performance was degraded. The
  maximum speedup observed when running on the production cluster with flushing enabled was
  11$\times$ with AFNI processing 1 HCP image. The maximum speedup for SPM was
  3.2$\times$, processing 8 ds001545 images.

    \section{Discussion}
    \subsection{Significance of speedups with Sea}

    Our results show that the speedups obtained from using Sea can be quite
    substantial (up to 32$\times$) depending on the degradation of the Lustre
    file system. The highest speed-ups observed on a production cluster were
    found to be 11$\times$.

    While on the production cluster, Sea and Baseline generally performed similarly,
    it is impossible for cluster users to know if Lustre's
    performance will be deteriorated during the lifetime of their application
    execution, nor to predict how such a deterioration would impact the
    application. The fact that Sea and Baseline perform similarly in the average
    case is an indication that runtime is generally not deteriorated by Sea, as
    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
    is safe to use in the event where Lustre performance is sufficiently
    degraded.

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation is less than that of possible speedups
    obtained. Our controlled experiments demonstrate that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
    less important that the potential speedups that can arise.

    \subsection{Limitations of using Sea}

    Sea really shines when it is executed on data-intensive pipelines. This
    means, the data is written at a rate which exceeds the rate at which the
    compute node's page cache can flush results to Lustre. In our experiments,
    the pipeline which consisted of both the shortest duration and largest
    output size was 16 AFNI processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
    approximately 100~GB of data, the controlled cluster would have had
    approximately 44~GB of dirty cache available per OST and a maximum amount of
    100~GB of page cache. Since all writes would not have occurred at the same
    time and computation time was extended as a result of CPU contention, even
    our controlled experiments do not showcase the benefits of Sea in a highly
    data-intensive scenario. However, HCP is one of the largest fMRI datasets,
    thus the only way to augment data intensiveness of the applications is
    through increased parallelism which is limited by the number of available
    cores. As a result, we can only demonstrate the speedups brought by Sea when
    Lustre performance has been degraded.

    Sea cannot speedup applications that are very compute-intensive, as is the case with FSL. This is because the bottleneck is the
    CPU time, which Sea does not address. Moreover, depending on the
    application's I/O patterns, it is possible that there will be sufficient
    page cache space and time to flush all the results asynchronously
    to Lustre. However, even with minimal speedups, Sea can still be useful in
    these cases to limit the number of new files created on Lustre, which can be
    particularly useful in instances where users have strict file limits.

    %\subsection{Sea uses more memory} \subsection{Other use cases} writes a lot
    %but it's actually more complicated. When dataset increases
    \subsection{Predicting speedups}

    Our results expose the complexities of performance predictions on HPC clusters. We present results on a dedicated cluster with which we have
    control over the global cluster usage and are able to demonstrate that
    significant speedups can be obtained when all the Lustre storage devices are
    busy writing other data concurrently. However, even these experiments are
    limited in demonstrating potential speedups with Sea as there are many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster has a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment were executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.

    Users cannot control or predict the cluster status at the time of their
    experiments.
    Our experiments show that when parallelism is optimal (no competing threads
    and minimized compute time), we obtain speedups with Sea if Lustre
    performance is degraded. Otherwise, we should experience the same
    performance without Sea. The more data the application writes, the bigger
    the speedups, since what cannot be written to page cache can still benefit from
    being written to local storage.

    The frequency and the size of the user application I/O also play an
    important role in determining speedups. An application which writes a lot of
    data, but spreads out the writes evenly, may result in no speedups from
    using Sea, whereas an application which writes large amounts of data in
    bursts may benefit from significant speedups. Furthermore, applications
    which create many small files may also experience larger speedups with Sea
    as the these small files may overburden the Lustre metadata server.

    %  difficulties with summarizing speedups into guidelines. variability is
    %  important and it's coming from the interaction between infrastructure and
    %  application. Benefits of sea depend on parameters such as load of the
    %  filesystem which is multifold and the application compute and file system
    %  configuration. page cache examples:
    %   - large dataset - > more computation
    %   - data access patterns and bursty-ness
    %   - small datasets + metadata accesses

    \subsection{Prefetching, flushing and eviction with Sea}

    Sea provides prefetching, flushing and eviction mechanisms to further
    improve the speedups that can be obtained. Prefetching allows input data to
    be moved to and manipulated from memory. All of our experiments with SPM
    used prefetching as SPM updates the input data via the use of a memory map.
    Without prefetching, updates to the input files would have been performed
    directly on Lustre, thus exhibiting a less important speedup. It was
    favorable to proceed with prefetching, in this case, as the input files
    themselves were not large compared to available memory space. However, in
    cases with very large input files, it may be preferable to maintain the
    memory map to Lustre.

    Whereas flushing and eviction were not used in the controlled experiments,
    they were enabled for the production cluster experiments. We made the assumption, that
    users would want access to all available intermediate data, although this
    would likely not be the case with the processing of full datasets. Thus,
    eviction was mostly used to ensure that files eventually deleted by the
    pipelines would not be copied to Lustre. Slowdowns experienced with Sea may
    be a result of flushing occurring outside of the computation. We assume that
    in large scale analyses, users will not need access to all intermediate
    results, leading to better usage of tmpfs and also minimized writes to
    Lustre. Such a scenario should lead to performance results more akin to
    those obtained using Big Data frameworks.

    \subsection{Glibc interception of neuroimaging pipelines}

    As demonstrated by our Sea experiments, we have found that glibc interception
    is effective at capturing all file system calls executed by the
    applications. This means that Sea is compatible with three of the most
    prominently used neuroimaging preprocessing pipelines. Although we have not
    tested it extensively against all possible parameters and toolbox
    applications, we assume that use of glibc within the toolboxes themselves are
    likely to be consistent. Despite not testing against all parameters and
    toolbox applications, Sea has been extensively tested using many integration
    tests on 3 different Linux operating systems (Ubuntu, CentOS and Fedora),
    and a total of 8 different OS versions.

    The ability to use glibc interception over more common alternative methods
    such as kernel-based or FUSE-based file systems was essential in ensuring
    ease-of-use and minimal performance overheads. As both our controlled and
    production environment experiments demonstrate, Sea's overhead is minimal
    and the magnitude of the speedups to be gained by the use of Sea is
    greater than the magnitude of the slowdowns.

    %opportunities for other optimizations Sea may be beneficial
    \subsection{Should Sea always be used ?}

    Our results show that performance with Sea is generally hard to predict as
    users do not have control over the variables outside the application they
    are executing. We have previously discussed the conditions in which Sea may
    not be favorable to use, however, differences in the execution environment
    can still shift performance results to be favorable to Sea. As we cannot
    guarantee performance with Sea, we break down this discussion from the
    perspective of a cluster administrator and that of a user.

    From the perspective of a cluster administrator, Sea is considered useful to
    maintaining the global health of the system. By using available local
    resources, Sea is able to alleviate strain on globally shared cluster
    resources, thus improving the experience for all users. Sea is able to do
    this through the use of eviction which ensures that data that is not
    required, never makes it to Lustre.

    From the perspective of a user, Sea provides speedups when
    Lustre performance is degraded and ensures limited overhead. It is most efficient
    when the compute time and non-Lustre I/O overhead is minimal. Even in
    instances where Sea does not improve performance, it helps reduce and ensure
    that the number of files that make it to Lustre are the exact ones required,
    therefore not exceeding any predetermined quota a user might have on the
    shared file system.

    %user

    %sysadmin

    %write intensive

    %\subsection{Effective use of available cluster resources}
    %\subsection{Comparisons to best case scenario}

    %talk about testing

    %Flushing - rely on modification time
    \section{Conclusion and Future work}

    We developed a lightweight data-management library, Sea, to seamlessly
    provide data management capabilities to existing neuroimaging workflows. We
    have tested our library on three of the most commonly used fMRI
    preprocessing libraries and found that Sea was able to intercept and
    redirect all application I/O calls without issue.

    When benchmarking the performance of Sea compared to Baseline, it was found that Sea provided speedups
    when Lustre performance was degraded. We were able to obtain up to a
    32$\times$ in the controlled environment and 11$\times$ in a production
    cluster. It was also found that speedups were greatest with a larger
    filesize, as long as the compute time and non-Lustre I/O-related overheads
    were also minimal.

    It is believed that Sea's benefits are not only limited to performance as
    users can also benefit from limiting the number of files created on Lustre
    with Sea. Archiving of the output directory on Lustre with Sea to further
    reduce number of files may be an interesting addition to Sea.

    \bibliography{biblio}

\end{document}

### 2022-05-10 paper/sea-neuro/paper-neuro.tex
reorganized text and resized figures
Comments in intro

@@
 \usepackage{subcaption}
+\usepackage{xspace}
+
+\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
 
@@
     The recent explosion in publicly available neuroimaging data has lead to new data
-    management challenges, from storage infrastructure application processing times. To meet the storage, accessibility and
-    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers.
+    management challenges, from storage infrastructure application processing times \TG{missing word?}. To meet the storage, accessibility and
+    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers \TG{mentioning clouds so early in the paper may give the impression that the paper will focus on them.}.
     Standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids}, have been implemented to facilitate the sharing
     of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide versioning and provenance capture of data.
-    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches.
+    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches \TG{cite fastsurfer, hd-bet}.
     While all these solutions to Big Data-related data management exist, certain aspects, such as processing-related data-transfer overheads
@@
     Although engines such as Nipype~\cite{nipype} and
-    Joblib~\cite{joblib} do not prohibit the use of data-management
+    Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.} do not prohibit the use of data-management
     strategies, they do not facilitate the integration of these strategies into their resulting workflow. To give

### 2022-05-18 paper/sea-comp/paper.tex
updated acknowdlegements and bibliographies

@@
 % The paper headers
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
-{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for
-Computer Society Journals} \IEEEtitleabstractindextext{%
+% \markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
+% {Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for
+% Computer Society Journals} 
+\IEEEtitleabstractindextext{%
 \begin{abstract}
@@
 
+\section*{Acknowlegments}
+Val\'erie Hayot-Sasson was funded by the Canadian Open Neuroscience Platform (CONP) Scholar Award.
+This work was also supported by the Canada Foundation for Innovation as well as the Canada Research Chairs program.
+
 % % use section* for acknowledgment
@@
 
-\begin{IEEEbiography}{Val\'erie Hayot-Sasson}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{biopics/vhs.pdf}}]{Val\'erie Hayot-Sasson} is a PhD student at Concordia University in the Big Data for Neuroinformatics lab.
+Her research interests focus on studying the effects of data transfers on Big Data scientific application.
 \end{IEEEbiography}
-\begin{IEEEbiography}{Tristan Glatard}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{biopics/md}}]{Mathieu Dugr\'e} is a Ph.D.
+student at the Big Data Infrastructure for Neuroinformatics lab at Concordia University, Montreal, Canada.
+His research interests are in Big Data performance techniques for neuroimaging applications.
+
+\end{IEEEbiography}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{biopics/tg}}]{Tristan Glatard} is Associate Professor in the
+Department of Computer Science and Software
+Engineering at Concordia University in Montreal,
+and Canada Research Chair (Tier II) on Big
+Data Infrastructures for Neuroinformatics. Before
+that, he was research scientist at the French Na-
+tional Centre for Scientific Research and Visiting
+Scholar at McGill University.
 \end{IEEEbiography}

### 2022-05-18 paper/sea-neuro/paper-neuro.tex

--- 2022-05-06 paper/sea-neuro/paper-neuro.tex
@@
 
-    The recent explosion in publicly available neuroimaging data has lead to new data
-    management challenges, from storage infrastructure application processing times. To meet the storage, accessibility and
-    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers.
-    Standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids}, have been implemented to facilitate the sharing
-    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide versioning and provenance capture of data.
-    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches.
-    While all these solutions to Big Data-related data management exist, certain aspects, such as processing-related data-transfer overheads
-    have received limited attention.
+    Neuroimaging datasets continue to grow in size resulting in new data
+    management challenges. The current largest neuroimaging datasets, such as
+    the Human Connectome Project (HCP)~\cite{HCP} and the UK
+    Biobank~\cite{ukbiobank} reach up to Petabytes of data. For datasets of this
+    volume, there is no choice but to store the data in slower remote storage,
+    which can incur important data transfer costs when processing. Should the
+    processed data be of a similar magnitude in volume, application writes will
+    also need to be made to compute-remote storage thus further prolonging
+    application makespan.
 
-    The current largest neuroimaging datasets,
-    the Human Connectome Project (HCP)~\cite{HCP} and the UK
-    Biobank~\cite{ukbiobank}, reach up to Petabytes of data. For datasets of this
-    volume, parallel high-capacity remote storage solutions are commonly relied
-    upon as the persistent storage layer during processing. 
-    Despite their respective high-scalability, these remote file systems cannot typically
-    compete with the performance of compute-local storage (e.g. DRAM and SSDs). Should the pipelines 
-    need to transfer large amounts of data to these remote storage locations, significant data
-    processing overheads may occur.
+    To minimize the costs of data transfers, Big Data frameworks, such as Apache
+    Spark~\cite{zaharia2016apache} and Dask~\cite{rocklin2015dask} leverage
+    memory and local storage as a cache. While these frameworks have been used
+    to process neuroimaging data~\cite{rokem2021pan,thunder,boubela2016big},
+    their popularity within the domain remains low due to the effort required to
+    rewrite standardized neuroimaging toolboxes for the
+    frameworks. 
+    % Although it is possible to use the tools within the Big Data frameworks,
+    % optimizations like in-memory computing would not be leveraged due to the
+    % fact that neuroimaging tools are meant to be used as command-line tools
+    % and do not provide interfaces that enable the data to be transferred
+    % in-memory.
+
+    Workflow engines used in neuroimaging, such as Nipype~\cite{nipype} and
+    Joblib~\cite{joblib} instead focus on reducing compute time of workloads via
+    parallelization of tasks.
+    %This is because even with large datasets, neuroimaging data processing
+    %remains split between compute and data intensive components.
+    Although these workflow engines do not prohibit the use of data-management
+    strategies, it is not directly integrated into the workflow. To give
+    neuroimaging applications data management capabilities, the applications
+    must interact with a file system that can do so.
+
+    Data management-specific scientific frameworks, such as
+    DataLad~\cite{halchenko2021datalad} aim to increase performance by focusing
+    on reproducibility, thereby reducing the need to recompute
+    results~\cite{wagner2021fairly}. We aim to complement frameworks such as
+    these, by reducing data transfer times when computation is required.
 
@@
     file I/O is predominantly impacted by the underlying storage bandwidth,
-    small file I/O results in a non-negligible latency overhead.
-
-    Big Data frameworks have addressed the performance limitations arising from large data transfers via the
-    implementation of two strategies: data locality and in-memory computing~\cite{zaharia2016apache, rocklin2015dask}.
-    The collective aim of these strategies is to reduce overheads related to large data
-    transfers by leveraging compute-local storage as cache.
-    While these frameworks have been used
-    to process neuroimaging data~\cite{rokem2021pan,thunder,boubela2016big},
-    their popularity within the domain remains low due to efforts required to
-    rewrite standardized neuroimaging toolboxes to utilize the frameworks.
-    % Although it is possible to use the tools within the Big Data frameworks,
-    % optimizations like in-memory computing would not be leveraged due to the
-    % fact that neuroimaging tools are meant to be used as command-line tools
-    % and do not provide interfaces that enable the data to be transferred
-    % in-memory.
-
-    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
-    also do not benefit from processing-related data management.
-    Although engines such as Nipype~\cite{nipype} and
-    Joblib~\cite{joblib} do not prohibit the use of data-management
-    strategies, they do not facilitate the integration of these strategies into their resulting workflow. To give
-    neuroimaging applications data management capabilities, the applications
-    must interact with a file system or library that enable the strategies.
-
-    % Data management-specific scientific frameworks, such as
-    % DataLad~\cite{halchenko2021datalad} aim to increase performance by focusing
-    % on reproducibility, thereby reducing the need to recompute
-    % results~\cite{wagner2021fairly}. We aim to complement frameworks such as
-    % these, by reducing data transfer times when computation is required.
+    small file I/O results in a non-negligible latency overhead. Addressing both
+    of these issues in neuroimaging is important, however, for this paper we
+    focus on fMRI datasets, which are currently more likely to experience a
+    latency overhead. We chose to focus on these types of datasets due to the
+    number of well-established tools to process them.
 
@@
 
-    We present Sea~\cite{sea-comp}, a data-management library designed to reduce processing-related overheads. We view Sea as a complement
-    to the existing stack of neuroimaging tools and standards meant to facilitate the processing of neuroimaging Big Data.    
+    In this paper, we evaluate Sea (Chapter~\ref{chapter:sea-comp}), a data-management library for HPC, on three neuroimaging preprocessing pipelines
+    executed on three different datasets. We structure our analysis of Sea for the data-management of
+    neuroimaging pipelines by first investigating whether Sea can effectively intercept the calls of these neuroimaging pipelines.
+    We then investigate the presence of speedup by using Sea alongside neuroimaging pipelines and try to understand the conditions
+    which result in greater speedups.
+
     % \tristan{Intro logic is good but writing could be polished}

### 2022-05-18 paper/sea-comp/paper.tex
[no ci] Minor comments.

--- 2022-05-06 paper/sea-comp/paper.tex
@@
 \newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
+\newcommand{\MD}[1]{\color{magenta}\textsc{From Mathieu}: #1\xspace\color{black}}
 
@@
       The storage layer found on HPC clusters consists of a high-performance
-      parallel file system (PFS) (e.g. Lustre~\cite{lustre}). These file systems
+      parallel file system (PFS) (e.g., Lustre~\cite{lustre}). These file systems
       are shared amongst all compute nodes within a cluster, typically connected
@@
 
-Unlike Big Data frameworks, Sea does not require reinstrumentation of the
+Unlike most Big Data frameworks, Sea does not require reinstrumentation of the
 existing pipelines, allowing users to gain an instant performance boost.
@@
       others are data-intensive), a third component, consisting of compute time,
-      can be added. Furthermore, latency may also play a significant role
+      can be added. Furthermore, latency may also play a significant role in an
       application makespan, particularly in scenarios with large amounts of
@@
       Here, $M_{n}$ represents the makespans of the different possible storage
-      levels (e.g. tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
+      levels (e.g., tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
       makespan of an application reading and writing to Sea. For our model, we
@@
       each application processes the same amount of input data, produces the
-      same amount of intermediate and output data and performs the same amount
+      same amount of intermediate and output data, and performs the same amount
       of computation.
@@
       all our experiments, we utilize the \SI{20}{\micro\meter} dataset, which
-      totals to approximately \SI{603}{\gibi\byte}. The dataset was broken down
+      totals to approximately \SI{603}{\gibi\byte} \MD{Are you sure it's 603? I had calculated 606~GiB for my Spark \& Dask paper}. The dataset was broken down
       into 1000 files each consisting of \SI{617}{\mebi\byte} of data.
@@
       number of nodes, 2) varying the number of disks, 3) varying the number of
-      threads, and 4) varying the number of iterations. Experimental condition 1
+      threads, and 4) varying the number of iterations. Experimental condition 1 \MD{Is the number of thread constant when varying nodes? i.e., 64 threads total for either 8 nodes or 4 nodes}
       tests the effects of increasing concurrent accesses to Lustre while fixing
@@
 
-      Each file system was benchmarked using \texttt{dd} using 5 repetitions.
+      Each file system was benchmarked using \texttt{dd} using 5 repetitions. \MD{What dd command was used? In particular, the block size and the total data size transfered}
       The average bandwidths are reported in Table~\ref{table:sea-comp:fs}
@@
     \end{figure*}
+    \MD{The y-lim on figure 2-d could be set to 35000, or even 25000, to make the rest of the plot easier to see}
 
@@
       plateau, however, it is expected that once the number of threads writing
-      to Lustre exceeds the number of OSTs (e.g. 9 nodes), we will observe an
+      to Lustre exceeds the number of OSTs (e.g., 9 nodes), we will observe an
       even greater speedup from Sea due to the increased contention on Lustre.
@@
     Sea cannot be used with operating systems that are not Linux-based, and
-    therefore, the use of Sea limits application portability. However, HPC
+    therefore, the use of Sea limits application portability \MD{Does Sea really limits portability?
+The application could still be used on other platform easily.
+i.e., Sea is limited to Posix-based system but doesn't limit application by itself, since it no rewrite of the application is required.}.
+However, HPC
     systems commonly use Linux operating systems, thus, it is not limiting for
@@
     that it could achieve better performance than leveraging local storage, even
-    if local storage was limited. We therefore believe that it is very unlikely
+    if local storage was limited. We therefore believe that it is very unlikely to
     obtain no speedups from using Sea on a production HPC cluster as long as the

### 2022-05-20 paper/sea-comp/paper.tex
reverted requested modif back

@@
 
-Unlike most Big Data frameworks, Sea does not require reinstrumentation of the
+Unlike Big Data frameworks, Sea does not require reinstrumentation of the
 existing pipelines, allowing users to gain an instant performance boost.

### 2022-07-04 paper/sea-comp/paper.tex
notebook and text updates

@@
 \usepackage{xspace}
+\usepackage{footmisc}
 
@@
     $p$ is the number of parallel application processes \\
+    $d$ is the number of Lustre storage disks \\
     $d_{r}$ is the read bandwidth of a single Lustre storage disk \\
@@
       all our experiments, we utilize the \SI{20}{\micro\meter} dataset, which
-      totals to approximately \SI{603}{\gibi\byte} \MD{Are you sure it's 603? I had calculated 606~GiB for my Spark \& Dask paper}. The dataset was broken down
+      totals to approximately \SI{603}{\gibi\byte}. The dataset was broken down
       into 1000 files each consisting of \SI{617}{\mebi\byte} of data.
@@
       number of nodes, 2) varying the number of disks, 3) varying the number of
-      threads, and 4) varying the number of iterations. Experimental condition 1 \MD{Is the number of thread constant when varying nodes? i.e., 64 threads total for either 8 nodes or 4 nodes}
+      threads, and 4) varying the number of iterations. Experimental condition 1
       tests the effects of increasing concurrent accesses to Lustre while fixing
-      disk parallel threads. Condition 2 varies disk contention while fixing
-      contention to Lustre, whereas 3 tests the effects of contention on both
+      disk parallel threads (i.e., the number of threads per node is constant).
+      Condition 2 varies disk contention while fixing
+      contention to Lustre, whereas condition 3 evaluates the effects of contention on both
       Lustre and local storage. Experimental condition 4 varies the total amount
@@
 
-      Each file system was benchmarked using \texttt{dd} using 5 repetitions. \MD{What dd command was used? In particular, the block size and the total data size transfered}
+      Each file system was benchmarked 5 times with the \texttt{dd} command \footnote{\url{https://github.com/big-data-lab-team/Sea/blob/master/benchmarks/scripts/bench_disks.sh}}.
       The average bandwidths are reported in Table~\ref{table:sea-comp:fs}
@@
     Sea cannot be used with operating systems that are not Linux-based, and
-    therefore, the use of Sea limits application portability \MD{Does Sea really limits portability?
-The application could still be used on other platform easily.
-i.e., Sea is limited to Posix-based system but doesn't limit application by itself, since it no rewrite of the application is required.}.
-However, HPC
-    systems commonly use Linux operating systems, thus, it is not limiting for
-    its intended use case. Furthermore, we provide publicly available containers
+    therefore, the use of an application with Sea has limited portability. However, Linux-based operating systems
+    are ubiquitous in HPC, and therefore, Sea is not limiting for
+    its intended use case. Furthermore, we provide publicly available containers~\footnote{\url{https://github.com/orgs/big-data-lab-team/packages?repo_name=Sea}}
     to simplify the usage of Sea on various other operating systems.

### 2023-06-28 paper/sea-neuro/paper-neuro.tex
Minor updates to text

--- 2022-05-10 paper/sea-neuro/paper-neuro.tex
@@
 
-\title{Automated neuroimaging data-management with Sea}
+\title{Ephemeral hierarchical data-management for neuroimaging applications}
 
@@
 	Neuroimaging open-data initiatives have led to increased availability of large
-  scientific datasets. Whereas these datasets are shifting the processing
+  scientific datasets. While these datasets are shifting the processing
   bottleneck from compute-intensive to data-intensive, current standardized
-  analysis tools lag behind in the adoption of strategies to mitigate the costs
-  associated with data transfers. A major challenge in adapting neuroimaging
+  analysis tools have yet to adopt strategies that mitigate the costs
+  associated with large data transfers. A major challenge in adapting neuroimaging
   applications for data-intensive processing is that they must be entirely
@@
   unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
-  where it is not useful. Overall, Sea is beneficial, even when performance gain
+  where its benefits are limited. Overall, Sea is beneficial, even when performance gain
   is minimal, as it can be used to limit files on PFS.
@@
 
-    The recent explosion in publicly available neuroimaging data has lead to new data
-    management challenges, from storage infrastructure application processing times \TG{missing word?}. To meet the storage, accessibility and
-    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers \TG{mentioning clouds so early in the paper may give the impression that the paper will focus on them.}.
-    Standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids}, have been implemented to facilitate the sharing
-    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide versioning and provenance capture of data.
-    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches \TG{cite fastsurfer, hd-bet}.
-    While all these solutions to Big Data-related data management exist, certain aspects, such as processing-related data-transfer overheads
+    The ever-growing surge in publicly available neuroimaging data has lead to new data
+    management challenges, from storage infrastructure to application processing times.
+    Many tools have been developed by the community to address to various needs in 
+    neuroimaging data management.
+    %To meet the storage, accessibility and security demands of neuroimaging data,
+    %large datasets have been stored in object stores provided by cloud storage providers
+    %\TG{mentioning clouds so early in the paper may give the impression that the paper will 
+    % focus on them.}.
+    For instance, standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids},
+    have been implemented to facilitate the sharing
+    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide 
+    versioning and provenance capture of data.
+    Furthermore, recent developments in neuroimaging pipelines have addressed computation time 
+    limitations by adopting machine-learning approaches \TG{cite fastsurfer, hd-bet}.
+    While all these solutions to Big Data-related data management exist, certain aspects, 
+    such as processing-related data-transfer overheads
     have received limited attention.
@@
     Despite their respective high-scalability, these remote file systems cannot typically
-    compete with the performance of compute-local storage (e.g. DRAM and SSDs). Should the pipelines 
+    compete with the performance of compute-local storage (e.g., tmpfs and SSDs). Should the pipelines 
     need to transfer large amounts of data to these remote storage locations, significant data
@@
 
-    Big Data in neuroimaging can be present in two formats: 1) very large files
+    Big Data in neuroimaging max exist in two formats: 1) very large files
     (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
@@
     also do not benefit from processing-related data management.
-    Although engines such as Nipype~\cite{nipype} and
-    Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.} do not prohibit the use of data-management
-    strategies, they do not facilitate the integration of these strategies into their resulting workflow. To give
+    Although engines such as Nipype~\cite{nipype}
+    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
+    do not prohibit the use of data-management
+    strategies, they do not facilitate the integration of these strategies into
+    their resulting workflow. To give
     neuroimaging applications data management capabilities, the applications
@@
 
-    HPC systems rely on scalable network-based parallel file systems (e.g.
+    HPC systems rely on scalable network-based parallel file systems (e.g.,
     Lustre) for storage. While such file systems offer excellent performance,
@@
     throttling~\cite{huang2020ooops} the data-intensive workloads or
-    recommending the use of Burst Buffers~\cite{bb} (e.g. reserving a compute
+    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
     node for storage or leveraging local compute storage during processing). The
@@
     system must also be compliant to the format. One method to ensure that these
-    conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick is
+    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
     used to intercept select library calls and redefine their behaviour. It has
@@
 
-    We present Sea~\cite{sea-comp}, a data-management library designed to reduce processing-related overheads. We view Sea as a complement
-    to the existing stack of neuroimaging tools and standards meant to facilitate the processing of neuroimaging Big Data.    
+    In this paper, we present Sea~\cite{sea-comp}, a data-management library designed to reduce
+    transfer-related overheads. We view Sea as a complement
+    to the existing stack of neuroimaging tools and standards meant to facilitate 
+    the processing of neuroimaging Big Data.   
     % \tristan{Intro logic is good but writing could be polished}
@@
 
-    \section{Materials and Methods}
+    \section{Results}
 
@@
 \end{figure*}
-    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages the \texttt{LD\_PRELOAD}
+    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
+    the \texttt{LD\_PRELOAD}
     trick to intercept POSIX file system (more specifically, the GNU C library,
     glibc) calls on Linux systems. This enables Sea to redirect write calls
-    aimed for a slower storage device to a faster device whenever possible.
+    aimed at slower storage devices to a faster devices whenever possible.
     Similarly, when intercepting read calls, Sea can choose to read from a
     faster device if a copy is available on that device. Sea decides which
-    storage location it can write to based on the detailed provided in an
+    storage location it can write to based on the details provided in an
     initialization file called \texttt{sea.ini}. This file informs Sea of which
     locations it can use to read and write to, and well as their order of
-    priority.
+    priority. It is important to note that Sea is not a filesystem, but rather a
+    lightweight tool that redirects files to more optimal storage devices.
 
@@
     compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.
+
+% \tristan{I find this section lacks a clear focus. What are the key points from
+% these results? You should focus on them and possibly name your paragraphs
+% accordingly, for improved clarity. My takeaways would be the following:
+% \begin{enumerate} \item In controlled conditions, important speed-ups are
+% observed. Mention min, average and max speed-up for all conditions. \item
+% Speed-up is correlated with baseline makespan \item Other determinants of
+% speed-up are unclear \item Overhead is minor \item Speed-ups are also observed
+% in production cluster \end{enumerate}
+% }
+\begin{figure*}
+
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \captionsetup{width=.85\linewidth}
+    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
+    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \captionsetup{width=.85\linewidth}
+    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
+    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
+\end{subfigure}
+\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
+\label{fig:seaneuro:slashbin}
+\end{figure*}
+
+\subsection{Speedups observed in controlled environment}
+
+In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
+variable, speedups are obtained by using Sea when Lustre
+performance has been degraded by busy writers. On average, a single SPM
+process preprocessing a single HCP image produced the greatest speedup
+(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
+using a single process was the next fastest pipeline, with an average speedup of
+(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
+1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
+excellent speedups, which is likely due to a mix of prefetching the initial
+input files and the I/O patterns of the application. While it was expected that
+AFNI would benefit the most from Sea due to its
+minimal compute time and the amount of data it generates, it did not benefit
+from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
+AFNI performs a very high number of glibc calls. While the overhead of an
+individual call is likely very minimal, it is possible that the culmination of
+all these calls results in a large enough bottleneck that affects both Sea and
+Baseline executions an extensive amount. Moreover, the number of glibc calls
+aimed at interacting with data on the Lustre storage are very few. AFNI still
+obtained an average of 2$\times$ speedup across all experimental conditions.
+
+FSL Feat, in contrast, appears to be the most compute-bound of the three
+applications. Not only did it spend an extensive amount of time computing, the
+amount of output data generated was least of the three. Due to the
+compute-intensive nature of FSL Feat, it is expected that it resulted in the
+least significant speedups.
+
+Our results indicate that HCP, on average, obtained the greatest speedups with
+Sea. Out of the three datasets, HCP has the largest images (see
+Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
+occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
+the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
+single image) results in the next largest speedups, followed by PREVENT-AD, the
+dataset with the smallest functional images.
+
+For both the pipelines and the datasets, we noticed that obtained speedups
+decrease with increased parallelism. We did not modify the default parallelism
+level of the applications in our experiments. As a result, each application
+process is likely attempting to utilize all cores resulting in longer wait times
+with the increase in threads. With this unintentional additional wait time that
+is added to the applications due to compute contention, the benefits of using
+Sea becomes more limited.
+
+\subsection{Speedup correlated with Lustre degradation}
+
+Baseline performance without busy writers is comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
+busy writers, Baseline performance may deviate dramatically from the identity
+line. We observe that the more Baseline deviates from this identity line, the
+greater the speedup.
+
+We do occasionally observe slowdowns from using Sea. These slowdowns
+may arise from the initial read of the data, as they appear to occur less
+frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.
+
+\subsection{Overhead of Sea is minimal}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.7\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
+\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
+\label{fig:seaneuro:tmpfs}
+\end{figure*}
+
+To determine the overhead of Sea, we compared the average performance of Sea with
+the performance of pipeline executing entirely within memory
+(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
+behave more or less identically. Given these results, we conclude that the
+overhead of Sea is minimal and that the differences in makespan may be
+due to Sea's initial reading of the data from Lustre.
+
+\subsection{Speedups observed in production environments}
+
+\begin{figure*}
+
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \captionsetup{width=.85\linewidth}
+    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
+    \caption{Complete results}\label{fig:seaneuro:belugafull}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \captionsetup{width=.85\linewidth}
+    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
+    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
+\end{subfigure}
+\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
+\label{fig:seaneuro:beluga-noflush}
+\end{figure*}
+
+\begin{figure*}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \captionsetup{width=.85\linewidth}
+    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
+    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \captionsetup{width=.85\linewidth}
+    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
+    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
+\end{subfigure}
+\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
+\label{fig:seaneuro:beluga-wflush}
+\end{figure*}
+
+  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
+  shows us that although performance in production clusters is highly variable,
+  we can still obtain important speedups. It is very likely that when we
+  executed our experiments without flushing
+  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
+  resulting in Sea and Baseline performing
+  quite similarly to each other.
+
+  Our results with flushing enabled confirm this theory. When running with
+  flushing enabled, which would have the additional overhead of having to
+  ensure that all the data is copied to Lustre, we notice that occasionally, we
+  obtain very large speedups with both SPM and AFNI, suggesting that Lustre
+  performance was degraded. The
+  maximum speedup observed when running on the production cluster with flushing enabled was
+  11$\times$ with AFNI processing 1 HCP image. The maximum speedup for SPM was
+  3.2$\times$, processing 8 ds001545 images.
+
+    \section{Discussion}
+    \subsection{Significance of speedups with Sea}
+
+    Our results show that the speedups obtained from using Sea can be quite
+    substantial (up to 32$\times$) depending on the degradation of the Lustre
+    file system. The highest speed-ups observed on a production cluster were
+    found to be 11$\times$.
+
+    While on the production cluster, Sea and Baseline generally performed similarly,
+    it is impossible for cluster users to know if Lustre's
+    performance will be deteriorated during the lifetime of their application
+    execution, nor to predict how such a deterioration would impact the
+    application. The fact that Sea and Baseline perform similarly in the average
+    case is an indication that runtime is generally not deteriorated by Sea, as
+    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
+    is safe to use in the event where Lustre performance is sufficiently
+    degraded.
+
+    Although Sea usage did result in performance degradation at times, the magnitude
+    of the degradation is less than that of possible speedups
+    obtained. Our controlled experiments demonstrate that using Sea with a
+    degraded Lustre file system can result in average speedups of up to
+    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
+    less important that the potential speedups that can arise.
+
+    \subsection{Limitations of using Sea}
+
+    Sea really shines when it is executed on data-intensive pipelines. This
+    means, the data is written at a rate which exceeds the rate at which the
+    compute node's page cache can flush results to Lustre. In our experiments,
+    the pipeline which consisted of both the shortest duration and largest
+    output size was 16 AFNI processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
+    approximately 100~GB of data, the controlled cluster would have had
+    approximately 44~GB of dirty cache available per OST and a maximum amount of
+    100~GB of page cache. Since all writes would not have occurred at the same
+    time and computation time was extended as a result of CPU contention, even
+    our controlled experiments do not showcase the benefits of Sea in a highly
+    data-intensive scenario. However, HCP is one of the largest fMRI datasets,
+    thus the only way to augment data intensiveness of the applications is
+    through increased parallelism which is limited by the number of available
+    cores. As a result, we can only demonstrate the speedups brought by Sea when
+    Lustre performance has been degraded.
+
+    Sea cannot speedup applications that are very compute-intensive, as is the case with FSL. This is because the bottleneck is the
+    CPU time, which Sea does not address. Moreover, depending on the
+    application's I/O patterns, it is possible that there will be sufficient
+    page cache space and time to flush all the results asynchronously
+    to Lustre. However, even with minimal speedups, Sea can still be useful in
+    these cases to limit the number of new files created on Lustre, which can be
+    particularly useful in instances where users have strict file limits.
+
+    %\subsection{Sea uses more memory} \subsection{Other use cases} writes a lot
+    %but it's actually more complicated. When dataset increases
+    \subsection{Predicting speedups}
+
+    Our results expose the complexities of performance predictions on HPC clusters. We present results on a dedicated cluster with which we have
+    control over the global cluster usage and are able to demonstrate that
+    significant speedups can be obtained when all the Lustre storage devices are
+    busy writing other data concurrently. However, even these experiments are
+    limited in demonstrating potential speedups with Sea as there are many
+    controlled variables, some of which may vary from cluster to cluster. For
+    example, the production cluster has a total of 977 nodes and only 38 disks associated with
+    their Lustre scratch file system. We can imagine that if 900 of the nodes
+    are busy writing to the scratch file system, we would observe a speedup
+    larger than what has been reported on our dedicated cluster with only 8 compute
+    nodes and 44 disks associated to Lustre. However, these are not the only
+    factors that can affect reported results, others can include when the
+    experiment were executed, what kind of applications were being executed
+    alongside the experiment, scratch file system usage during experiments, and
+    what were the application read and write patterns.
+
+    Users cannot control or predict the cluster status at the time of their
+    experiments.
+    Our experiments show that when parallelism is optimal (no competing threads
+    and minimized compute time), we obtain speedups with Sea if Lustre
+    performance is degraded. Otherwise, we should experience the same
+    performance without Sea. The more data the application writes, the bigger
+    the speedups, since what cannot be written to page cache can still benefit from
+    being written to local storage.
+
+    The frequency and the size of the user application I/O also play an
+    important role in determining speedups. An application which writes a lot of
+    data, but spreads out the writes evenly, may result in no speedups from
+    using Sea, whereas an application which writes large amounts of data in
+    bursts may benefit from significant speedups. Furthermore, applications
+    which create many small files may also experience larger speedups with Sea
+    as the these small files may overburden the Lustre metadata server.
+
+    %  difficulties with summarizing speedups into guidelines. variability is
+    %  important and it's coming from the interaction between infrastructure and
+    %  application. Benefits of sea depend on parameters such as load of the
+    %  filesystem which is multifold and the application compute and file system
+    %  configuration. page cache examples:
+    %   - large dataset - > more computation
+    %   - data access patterns and bursty-ness
+    %   - small datasets + metadata accesses
+
+    \subsection{Prefetching, flushing and eviction with Sea}
+
+    Sea provides prefetching, flushing and eviction mechanisms to further
+    improve the speedups that can be obtained. Prefetching allows input data to
+    be moved to and manipulated from memory. All of our experiments with SPM
+    used prefetching as SPM updates the input data via the use of a memory map.
+    Without prefetching, updates to the input files would have been performed
+    directly on Lustre, thus exhibiting a less important speedup. It was
+    favorable to proceed with prefetching, in this case, as the input files
+    themselves were not large compared to available memory space. However, in
+    cases with very large input files, it may be preferable to maintain the
+    memory map to Lustre.
+
+    Whereas flushing and eviction were not used in the controlled experiments,
+    they were enabled for the production cluster experiments. We made the assumption, that
+    users would want access to all available intermediate data, although this
+    would likely not be the case with the processing of full datasets. Thus,
+    eviction was mostly used to ensure that files eventually deleted by the
+    pipelines would not be copied to Lustre. Slowdowns experienced with Sea may
+    be a result of flushing occurring outside of the computation. We assume that
+    in large scale analyses, users will not need access to all intermediate
+    results, leading to better usage of tmpfs and also minimized writes to
+    Lustre. Such a scenario should lead to performance results more akin to
+    those obtained using Big Data frameworks.
+
+    \subsection{Glibc interception of neuroimaging pipelines}
+
+    As demonstrated by our Sea experiments, we have found that glibc interception
+    is effective at capturing all file system calls executed by the
+    applications. This means that Sea is compatible with three of the most
+    prominently used neuroimaging preprocessing pipelines. Although we have not
+    tested it extensively against all possible parameters and toolbox
+    applications, we assume that use of glibc within the toolboxes themselves are
+    likely to be consistent. Despite not testing against all parameters and
+    toolbox applications, Sea has been extensively tested using many integration
+    tests on 3 different Linux operating systems (Ubuntu, CentOS and Fedora),
+    and a total of 8 different OS versions.
+
+    The ability to use glibc interception over more common alternative methods
+    such as kernel-based or FUSE-based file systems was essential in ensuring
+    ease-of-use and minimal performance overheads. As both our controlled and
+    production environment experiments demonstrate, Sea's overhead is minimal
+    and the magnitude of the speedups to be gained by the use of Sea is
+    greater than the magnitude of the slowdowns.
+
+    %opportunities for other optimizations Sea may be beneficial
+    \subsection{Should Sea always be used ?}
+
+    Our results show that performance with Sea is generally hard to predict as
+    users do not have control over the variables outside the application they
+    are executing. We have previously discussed the conditions in which Sea may
+    not be favorable to use, however, differences in the execution environment
+    can still shift performance results to be favorable to Sea. As we cannot
+    guarantee performance with Sea, we break down this discussion from the
+    perspective of a cluster administrator and that of a user.
+
+    From the perspective of a cluster administrator, Sea is considered useful to
+    maintaining the global health of the system. By using available local
+    resources, Sea is able to alleviate strain on globally shared cluster
+    resources, thus improving the experience for all users. Sea is able to do
+    this through the use of eviction which ensures that data that is not
+    required, never makes it to Lustre.
+
+    From the perspective of a user, Sea provides speedups when
+    Lustre performance is degraded and ensures limited overhead. It is most efficient
+    when the compute time and non-Lustre I/O overhead is minimal. Even in
+    instances where Sea does not improve performance, it helps reduce and ensure
+    that the number of files that make it to Lustre are the exact ones required,
+    therefore not exceeding any predetermined quota a user might have on the
+    shared file system.
+
+    %user
+
+    %sysadmin
+
+    %write intensive
+
+    %\subsection{Effective use of available cluster resources}
+    %\subsection{Comparisons to best case scenario}
+
+    %talk about testing
+
+    %Flushing - rely on modification time
+    \section{Materials and Methods}
 
@@
     %Busy writers
-    \section{Results}
-
-% \tristan{I find this section lacks a clear focus. What are the key points from
-% these results? You should focus on them and possibly name your paragraphs
-% accordingly, for improved clarity. My takeaways would be the following:
-% \begin{enumerate} \item In controlled conditions, important speed-ups are
-% observed. Mention min, average and max speed-up for all conditions. \item
-% Speed-up is correlated with baseline makespan \item Other determinants of
-% speed-up are unclear \item Overhead is minor \item Speed-ups are also observed
-% in production cluster \end{enumerate}
-% }
-\begin{figure*}
-
-\begin{subfigure}{\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
-    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
-    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
-\end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
-\label{fig:seaneuro:slashbin}
-\end{figure*}
-
-\subsection{Speedups observed in controlled environment}
-
-In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
-variable, speedups are obtained by using Sea when Lustre
-performance has been degraded by busy writers. On average, a single SPM
-process preprocessing a single HCP image produced the greatest speedup
-(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
-using a single process was the next fastest pipeline, with an average speedup of
-(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
-1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
-excellent speedups, which is likely due to a mix of prefetching the initial
-input files and the I/O patterns of the application. While it was expected that
-AFNI would be the application that would benefit the most from Sea due to its
-minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
-AFNI performs a very high number of glibc calls. While the overhead of an
-individual call is likely very minimal, it is possible that the culmination of
-all these calls results in a large enough bottleneck that affects both Sea and
-Baseline executions an extensive amount. Moreover, the number of glibc calls
-aimed at interacting with data on the Lustre storage are very few. AFNI still
-obtained an average of 2$\times$ speedup across all experimental conditions.
-
-FSL Feat, in contrast, appears to be the most compute-bound of the three
-applications. Not only did it spend an extensive amount of time computing, the
-amount of output data generated was least of the three. Due to the
-compute-intensive nature of FSL Feat, it is expected that it resulted in the
-least significant speedups.
-
-Our results indicate that HCP, on average, obtained the greatest speedups with
-Sea. Out of the three datasets, HCP has the largest images (see
-Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
-occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
-the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
-single image) results in the next largest speedups, followed by PREVENT-AD, the
-dataset with the smallest functional images.
-
-For both the pipelines and the datasets, we noticed that obtained speedups
-decrease with increased parallelism. We did not modify the default parallelism
-level of the applications in our experiments. As a result, each application
-process is likely attempting to utilize all cores resulting in longer wait times
-with the increase in threads. With this unintentional additional wait time that
-is added to the applications due to compute contention, the benefits of using
-Sea becomes more limited.
-
-\subsection{Speedup correlated with Lustre degradation}
-
-Baseline performance without busy writers is comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
-busy writers, Baseline performance may deviate dramatically from the identity
-line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup.
-
-We do occasionally observe slowdowns from using Sea. These slowdowns
-may arise from the initial read of the data, as they appear to occur less
-frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.
-
-\subsection{Overhead of Sea is minimal}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
-\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
-\label{fig:seaneuro:tmpfs}
-\end{figure*}
-
-To determine the overhead of Sea, we compared the average performance of Sea with
-the performance of pipeline executing entirely within memory
-(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
-behave more or less identically. Given these results, we conclude that the
-overhead of Sea is minimal and that the differences in makespan may be
-due to Sea's initial reading of the data from Lustre.
-
-\subsection{Speedups observed in production environments}
-
-\begin{figure*}
-
-\begin{subfigure}{\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
-    \caption{Complete results}\label{fig:seaneuro:belugafull}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
-    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
-\end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
-\label{fig:seaneuro:beluga-noflush}
-\end{figure*}
-
-\begin{figure*}
-\begin{subfigure}{\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
-    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
-    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
-\end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
-\label{fig:seaneuro:beluga-wflush}
-\end{figure*}
-
-  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
-  shows us that although performance in production clusters is highly variable,
-  we can still obtain important speedups. It is very likely that when we
-  executed our experiments without flushing
-  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
-  resulting in Sea and Baseline performing
-  quite similarly to each other.
-
-  Our results with flushing enabled confirm this theory. When running with
-  flushing enabled, which would have the additional overhead of having to
-  ensure that all the data is copied to Lustre, we notice that occasionally, we
-  obtain very large speedups with both SPM and AFNI, suggesting that Lustre
-  performance was degraded. The
-  maximum speedup observed when running on the production cluster with flushing enabled was
-  11$\times$ with AFNI processing 1 HCP image. The maximum speedup for SPM was
-  3.2$\times$, processing 8 ds001545 images.
-
-    \section{Discussion}
-    \subsection{Significance of speedups with Sea}
-
-    Our results show that the speedups obtained from using Sea can be quite
-    substantial (up to 32$\times$) depending on the degradation of the Lustre
-    file system. The highest speed-ups observed on a production cluster were
-    found to be 11$\times$.
-
-    While on the production cluster, Sea and Baseline generally performed similarly,
-    it is impossible for cluster users to know if Lustre's
-    performance will be deteriorated during the lifetime of their application
-    execution, nor to predict how such a deterioration would impact the
-    application. The fact that Sea and Baseline perform similarly in the average
-    case is an indication that runtime is generally not deteriorated by Sea, as
-    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
-    is safe to use in the event where Lustre performance is sufficiently
-    degraded.
-
-    Although Sea usage did result in performance degradation at times, the magnitude
-    of the degradation is less than that of possible speedups
-    obtained. Our controlled experiments demonstrate that using Sea with a
-    degraded Lustre file system can result in average speedups of up to
-    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
-    less important that the potential speedups that can arise.
-
-    \subsection{Limitations of using Sea}
-
-    Sea really shines when it is executed on data-intensive pipelines. This
-    means, the data is written at a rate which exceeds the rate at which the
-    compute node's page cache can flush results to Lustre. In our experiments,
-    the pipeline which consisted of both the shortest duration and largest
-    output size was 16 AFNI processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
-    approximately 100~GB of data, the controlled cluster would have had
-    approximately 44~GB of dirty cache available per OST and a maximum amount of
-    100~GB of page cache. Since all writes would not have occurred at the same
-    time and computation time was extended as a result of CPU contention, even
-    our controlled experiments do not showcase the benefits of Sea in a highly
-    data-intensive scenario. However, HCP is one of the largest fMRI datasets,
-    thus the only way to augment data intensiveness of the applications is
-    through increased parallelism which is limited by the number of available
-    cores. As a result, we can only demonstrate the speedups brought by Sea when
-    Lustre performance has been degraded.
-
-    Sea cannot speedup applications that are very compute-intensive, as is the case with FSL. This is because the bottleneck is the
-    CPU time, which Sea does not address. Moreover, depending on the
-    application's I/O patterns, it is possible that there will be sufficient
-    page cache space and time to flush all the results asynchronously
-    to Lustre. However, even with minimal speedups, Sea can still be useful in
-    these cases to limit the number of new files created on Lustre, which can be
-    particularly useful in instances where users have strict file limits.
-
-    %\subsection{Sea uses more memory} \subsection{Other use cases} writes a lot
-    %but it's actually more complicated. When dataset increases
-    \subsection{Predicting speedups}
-
-    Our results expose the complexities of performance predictions on HPC clusters. We present results on a dedicated cluster with which we have
-    control over the global cluster usage and are able to demonstrate that
-    significant speedups can be obtained when all the Lustre storage devices are
-    busy writing other data concurrently. However, even these experiments are
-    limited in demonstrating potential speedups with Sea as there are many
-    controlled variables, some of which may vary from cluster to cluster. For
-    example, the production cluster has a total of 977 nodes and only 38 disks associated with
-    their Lustre scratch file system. We can imagine that if 900 of the nodes
-    are busy writing to the scratch file system, we would observe a speedup
-    larger than what has been reported on our dedicated cluster with only 8 compute
-    nodes and 44 disks associated to Lustre. However, these are not the only
-    factors that can affect reported results, others can include when the
-    experiment were executed, what kind of applications were being executed
-    alongside the experiment, scratch file system usage during experiments, and
-    what were the application read and write patterns.
-
-    Users cannot control or predict the cluster status at the time of their
-    experiments.
-    Our experiments show that when parallelism is optimal (no competing threads
-    and minimized compute time), we obtain speedups with Sea if Lustre
-    performance is degraded. Otherwise, we should experience the same
-    performance without Sea. The more data the application writes, the bigger
-    the speedups, since what cannot be written to page cache can still benefit from
-    being written to local storage.
-
-    The frequency and the size of the user application I/O also play an
-    important role in determining speedups. An application which writes a lot of
-    data, but spreads out the writes evenly, may result in no speedups from
-    using Sea, whereas an application which writes large amounts of data in
-    bursts may benefit from significant speedups. Furthermore, applications
-    which create many small files may also experience larger speedups with Sea
-    as the these small files may overburden the Lustre metadata server.
-
-    %  difficulties with summarizing speedups into guidelines. variability is
-    %  important and it's coming from the interaction between infrastructure and
-    %  application. Benefits of sea depend on parameters such as load of the
-    %  filesystem which is multifold and the application compute and file system
-    %  configuration. page cache examples:
-    %   - large dataset - > more computation
-    %   - data access patterns and bursty-ness
-    %   - small datasets + metadata accesses
-
-    \subsection{Prefetching, flushing and eviction with Sea}
-
-    Sea provides prefetching, flushing and eviction mechanisms to further
-    improve the speedups that can be obtained. Prefetching allows input data to
-    be moved to and manipulated from memory. All of our experiments with SPM
-    used prefetching as SPM updates the input data via the use of a memory map.
-    Without prefetching, updates to the input files would have been performed
-    directly on Lustre, thus exhibiting a less important speedup. It was
-    favorable to proceed with prefetching, in this case, as the input files
-    themselves were not large compared to available memory space. However, in
-    cases with very large input files, it may be preferable to maintain the
-    memory map to Lustre.
-
-    Whereas flushing and eviction were not used in the controlled experiments,
-    they were enabled for the production cluster experiments. We made the assumption, that
-    users would want access to all available intermediate data, although this
-    would likely not be the case with the processing of full datasets. Thus,
-    eviction was mostly used to ensure that files eventually deleted by the
-    pipelines would not be copied to Lustre. Slowdowns experienced with Sea may
-    be a result of flushing occurring outside of the computation. We assume that
-    in large scale analyses, users will not need access to all intermediate
-    results, leading to better usage of tmpfs and also minimized writes to
-    Lustre. Such a scenario should lead to performance results more akin to
-    those obtained using Big Data frameworks.
-
-    \subsection{Glibc interception of neuroimaging pipelines}
-
-    As demonstrated by our Sea experiments, we have found that glibc interception
-    is effective at capturing all file system calls executed by the
-    applications. This means that Sea is compatible with three of the most
-    prominently used neuroimaging preprocessing pipelines. Although we have not
-    tested it extensively against all possible parameters and toolbox
-    applications, we assume that use of glibc within the toolboxes themselves are
-    likely to be consistent. Despite not testing against all parameters and
-    toolbox applications, Sea has been extensively tested using many integration
-    tests on 3 different Linux operating systems (Ubuntu, CentOS and Fedora),
-    and a total of 8 different OS versions.
-
-    The ability to use glibc interception over more common alternative methods
-    such as kernel-based or FUSE-based file systems was essential in ensuring
-    ease-of-use and minimal performance overheads. As both our controlled and
-    production environment experiments demonstrate, Sea's overhead is minimal
-    and the magnitude of the speedups to be gained by the use of Sea is
-    greater than the magnitude of the slowdowns.
-
-    %opportunities for other optimizations Sea may be beneficial
-    \subsection{Should Sea always be used ?}
-
-    Our results show that performance with Sea is generally hard to predict as
-    users do not have control over the variables outside the application they
-    are executing. We have previously discussed the conditions in which Sea may
-    not be favorable to use, however, differences in the execution environment
-    can still shift performance results to be favorable to Sea. As we cannot
-    guarantee performance with Sea, we break down this discussion from the
-    perspective of a cluster administrator and that of a user.
-
-    From the perspective of a cluster administrator, Sea is considered useful to
-    maintaining the global health of the system. By using available local
-    resources, Sea is able to alleviate strain on globally shared cluster
-    resources, thus improving the experience for all users. Sea is able to do
-    this through the use of eviction which ensures that data that is not
-    required, never makes it to Lustre.
-
-    From the perspective of a user, Sea provides speedups when
-    Lustre performance is degraded and ensures limited overhead. It is most efficient
-    when the compute time and non-Lustre I/O overhead is minimal. Even in
-    instances where Sea does not improve performance, it helps reduce and ensure
-    that the number of files that make it to Lustre are the exact ones required,
-    therefore not exceeding any predetermined quota a user might have on the
-    shared file system.
-
-    %user
-
-    %sysadmin
-
-    %write intensive
-
-    %\subsection{Effective use of available cluster resources}
-    %\subsection{Comparisons to best case scenario}
-
-    %talk about testing
-
-    %Flushing - rely on modification time
     \section{Conclusion and Future work}

### 2023-06-28 paper/sea-neuro/paper-neuro.tex
Minor updates to text

@@
 \usepackage{xspace}
+\usepackage{todonotes}
 
-\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
+\newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}
 
@@
   scientific datasets. While these datasets are shifting the processing
+  scientific datasets. While these datasets are shifting the processing
   bottleneck from compute-intensive to data-intensive, current standardized
@@
   associated with large data transfers. A major challenge in adapting neuroimaging
+  analysis tools have yet to adopt strategies that mitigate the costs
+  associated with large data transfers. A major challenge in adapting neuroimaging
   applications for data-intensive processing is that they must be entirely
-  rewritten.
-
-  To facilitate data management for standardized neuroimaging tools, we
+  rewritten. To facilitate data management for standardized neuroimaging tools, we
   developed Sea, a library that intercepts and redirects application read and
@@
     their resulting workflow. To give
+    Although engines such as Nipype~\cite{nipype}
+    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
+    do not prohibit the use of data-management
+    strategies, they do not facilitate the integration of these strategies into
+    their resulting workflow. To give
     neuroimaging applications data management capabilities, the applications
@@
     HPC systems rely on scalable network-based parallel file systems (e.g.,
+    HPC systems rely on scalable network-based parallel file systems (e.g.,
     Lustre) for storage. While such file systems offer excellent performance,
@@
     throttling~\cite{huang2020ooops} the data-intensive workloads or
+    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
     recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
@@
     system must also be compliant to the format. One method to ensure that these
+    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
     conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
@@
     storage location it can write to based on the details provided in an
+    storage location it can write to based on the details provided in an
     initialization file called \texttt{sea.ini}. This file informs Sea of which
@@
 
-    As neuroimaging pipeline results are typically required post-processing and
+    As neuroimaging pipeline results typically require post-processing and
     HPC compute-local resources are only accessible during the reserved
     duration, Sea provides functionality to flush and evict data to persistent
-    shared storage. This is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. A
-    separate thread is used, in this case, to avoid interrupting ongoing
-    processing with data management operations. Users must inform Sea of files
+    shared storage. To avoid interrupting ongoing
+    processing with data management operations, this is accomplished via a separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
     that need to be persisted to storage within a file called
@@
 \end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
+\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
+denotes the total time between execution launch and completion of the last computing task.}
 \label{fig:seaneuro:slashbin}
@@
 
-In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
-variable, speedups are obtained by using Sea when Lustre
-performance has been degraded by busy writers. On average, a single SPM
+We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
+(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB) with 3 pipelines 
+from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
+dedicated to our experiments. We deployed ``busy writers'' in the cluster
+that simulated the load from other user applications by writing to the shared file system at a controlled rate.
+
+Speedups were obtained by using Sea when Lustre
+performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
+although performance was variable. On average, a single SPM pipeline
 process preprocessing a single HCP image produced the greatest speedup
-(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
+(13$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
 using a single process was the next fastest pipeline, with an average speedup of
-(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
-1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
+(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
+1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
 excellent speedups, which is likely due to a mix of prefetching the initial
@@
 minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
-AFNI performs a very high number of glibc calls. While the overhead of an
+from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
+AFNI pipeline performs a very high number of glibc calls. While the overhead of an
 individual call is likely very minimal, it is possible that the culmination of
-all these calls results in a large enough bottleneck that affects both Sea and
-Baseline executions an extensive amount. Moreover, the number of glibc calls
-aimed at interacting with data on the Lustre storage are very few. AFNI still
+all these calls results in a large enough bottleneck that affected both Sea and
+Baseline executions to an extensive amount. Moreover, the number of glibc calls
+aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
 obtained an average of 2$\times$ speedup across all experimental conditions.
 
-FSL Feat, in contrast, appears to be the most compute-bound of the three
+The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
 applications. Not only did it spend an extensive amount of time computing, the
 amount of output data generated was least of the three. Due to the
-compute-intensive nature of FSL Feat, it is expected that it resulted in the
+compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
 least significant speedups.
 
-Our results indicate that HCP, on average, obtained the greatest speedups with
+The HCP dataset, on average, obtained the greatest speedups with
 Sea. Out of the three datasets, HCP has the largest images (see
@@
 For both the pipelines and the datasets, we noticed that obtained speedups
-decrease with increased parallelism. We did not modify the default parallelism
+decreased with increased parallelism. We did not modify the default parallelism
 level of the applications in our experiments. As a result, each application
-process is likely attempting to utilize all cores resulting in longer wait times
-with the increase in threads. With this unintentional additional wait time that
+process was likely attempting to utilize all cores, resulting in longer wait times
+with the increase in parallelism. With this unintentional additional wait time that
 is added to the applications due to compute contention, the benefits of using
@@
 
-Baseline performance without busy writers is comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
-busy writers, Baseline performance may deviate dramatically from the identity
+Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
+busy writers, Baseline performance deviated dramatically from the identity
 line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup.
+greater the speedup. We did occasionally observe slowdowns from using Sea. These slowdowns
+may arise from the initial read of the data, as they appeared to occur less
+frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.
 
-We do occasionally observe slowdowns from using Sea. These slowdowns
-may arise from the initial read of the data, as they appear to occur less
-frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.
-
-\subsection{Overhead of Sea is minimal}
+\subsection{Overhead of Sea was minimal}
 
@@
 the performance of pipeline executing entirely within memory
-(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
-behave more or less identically. Given these results, we conclude that the
+(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
+behave mostly identically. Given these results, we conclude that the
 overhead of Sea is minimal and that the differences in makespan may be
@@
 \end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
+\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
 \label{fig:seaneuro:beluga-wflush}
@@
 
+  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
+  shared with other users,
+  using the same datasets and pipelines as for the controlled cluster. 
   Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
-  shows us that although performance in production clusters is highly variable,
-  we can still obtain important speedups. It is very likely that when we
+  show that although performance in production clusters was highly variable,
+  we could still obtain important speedups by using Sea. It is very likely that when we
   executed our experiments without flushing
@@
 
-  Our results with flushing enabled confirm this theory. When running with
-  flushing enabled, which would have the additional overhead of having to
-  ensure that all the data is copied to Lustre, we notice that occasionally, we
-  obtain very large speedups with both SPM and AFNI, suggesting that Lustre
+  Our results with flushing enabled confirmed this theory. When running with
+  flushing enabled, which had the additional overhead of having to
+  ensure that all the data was copied to Lustre, we noticed that occasionally, we
+  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
   performance was degraded. The
   maximum speedup observed when running on the production cluster with flushing enabled was
-  11$\times$ with AFNI processing 1 HCP image. The maximum speedup for SPM was
+  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
   3.2$\times$, processing 8 ds001545 images.
@@
 
-    Our results show that the speedups obtained from using Sea can be quite
-    substantial (up to 32$\times$) depending on the degradation of the Lustre
+    Our results show that the speedups obtained from using Sea can be 
+    substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
     file system. The highest speed-ups observed on a production cluster were
@@
     Although Sea usage did result in performance degradation at times, the magnitude
-    of the degradation is less than that of possible speedups
-    obtained. Our controlled experiments demonstrate that using Sea with a
+    of the degradation was less than that of possible speedups
+    obtained. Our controlled experiments demonstrated that using Sea with a
     degraded Lustre file system can result in average speedups of up to
@@
 
-    Sea really shines when it is executed on data-intensive pipelines. This
-    means, the data is written at a rate which exceeds the rate at which the
+    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
+     when the data is written at a rate which exceeds the rate at which the
     compute node's page cache can flush results to Lustre. In our experiments,
     the pipeline which consisted of both the shortest duration and largest
-    output size was 16 AFNI processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
+    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
     approximately 100~GB of data, the controlled cluster would have had
-    approximately 44~GB of dirty cache available per OST and a maximum amount of
+    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
     100~GB of page cache. Since all writes would not have occurred at the same
     time and computation time was extended as a result of CPU contention, even
-    our controlled experiments do not showcase the benefits of Sea in a highly
-    data-intensive scenario. However, HCP is one of the largest fMRI datasets,
-    thus the only way to augment data intensiveness of the applications is
+    our controlled experiments did not showcase the benefits of Sea in a highly
+    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
+    the only way to augment data intensiveness of the applications is
     through increased parallelism which is limited by the number of available
-    cores. As a result, we can only demonstrate the speedups brought by Sea when
-    Lustre performance has been degraded.
+    cores. As a result, we could only demonstrate the speedups brought by Sea when
+    Lustre performance had been degraded.
 
-    Sea cannot speedup applications that are very compute-intensive, as is the case with FSL. This is because the bottleneck is the
+    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
+    In such cases, the performance bottleneck is the
     CPU time, which Sea does not address. Moreover, depending on the
@@
 
-    Our results expose the complexities of performance predictions on HPC clusters. We present results on a dedicated cluster with which we have
-    control over the global cluster usage and are able to demonstrate that
-    significant speedups can be obtained when all the Lustre storage devices are
-    busy writing other data concurrently. However, even these experiments are
-    limited in demonstrating potential speedups with Sea as there are many
+    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
+    where we had
+    control over the global cluster usage and were able to demonstrate that
+    significant speedups could be obtained when all the Lustre storage devices were
+    busy writing other data concurrently. However, even these experiments were
+    limited in demonstrating potential speedups with Sea as there were many
     controlled variables, some of which may vary from cluster to cluster. For
-    example, the production cluster has a total of 977 nodes and only 38 disks associated with
+    example, the production cluster had a total of 977 nodes and only 38 disks associated with
     their Lustre scratch file system. We can imagine that if 900 of the nodes
@@
     improve the speedups that can be obtained. Prefetching allows input data to
-    be moved to and manipulated from memory. All of our experiments with SPM
+    be moved to and manipulated from memory. All of our experiments with the SPM pipeline
     used prefetching as SPM updates the input data via the use of a memory map.
@@
     ease-of-use and minimal performance overheads. As both our controlled and
-    production environment experiments demonstrate, Sea's overhead is minimal
+    production environment experiments demonstrated, Sea's overhead is minimal
     and the magnitude of the speedups to be gained by the use of Sea is
@@
     this through the use of eviction which ensures that data that is not
-    required, never makes it to Lustre.
+    required never makes it to Lustre.
 
@@
     and some of the largest datasets. Of course, different modalities and tools
-    may result in vastly different data access patterns and compute times, we
-    use fMRI pre-processing as our baseline for the benefits that can be
-    obtained with Sea.
+    may result in vastly different data access patterns and compute times.
 
@@
     AFNI~\cite{cox1996afni}, and SPM~\cite{spm} toolboxes.
-    Table~\ref{table:seaneuro:pipelines} provides a reference of the differences
+    Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
     in computation and data-intensiveness of the different pipelines on a single
@@
 
-    Each tool was set to only run the preprocessing pipeline. Scripts used to
-    execute each pipeline can be found
-    at~\url{github.com/valhayot/sea-slurm-scripts}.
-
-    For the FSL preprocessing pipeline, we used FEAT configured with many of
+    Each tool was set to only run the functional preprocessing pipeline. For the FSL preprocessing pipeline, 
+    we used FEAT configured with many of
     the default parameters. We set the following parameters: slice-timing
@@
     %% Add SPM here
-    To preprocess the datasets with SPM12, we reused the template described in
+    To preprocess the datasets with SPM, we reused the template described in
     \cite{haitas2021age}, with fieldmap correction removed. As the template was
@@
     and slices and the TR and TA, for each dataset. Furthermore, we set the
-    slice order to interleaved in all cases.
+    slice order to interleaved in all cases. SPM version 12 was used.
 
@@
     as required for its execution. AFNI version 21.1.02 was used.
-
-    All the application Dockerfiles with Sea pre-installed are available
-    at~\url{https://github.com/ValHayot/Sea/tree/master/containers/applications}.
 
@@
     \footnotesize{$*$ Measurements taken on the dedicated cluster}
-    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}\label{table:seaneuro:pipelines}
+    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
+    \label{table:seaneuro-pipelines}
   \end{table*}
@@
     \subsection{Controls}
-    Our objective is to measure the speedup brought by Sea on real neuroimaging
+    Our objective was to measure the speedup brought by Sea on real neuroimaging
     pipelines. Although Sea is agnostic to the internals of the pipelines and
@@
 
-    Our experiments were executed on two different clusters: 1) a dedicated cluster
+    Our experiments were executed on two different clusters: 1) a controlled, dedicated cluster
     and 2) and a production cluster. We chose this experimental setup as the dedicated
@@
     via a 20Gbps ethernet connection. The cluster's hardware was similar to that
-    of Beluga.
+    of Compute Canada's Beluga cluster.
 
-    Compute Canada's Beluga cluster was used as the production cluster. It consisted of a total of 977 compute nodes, of which we
+    Compute Canada's Beluga cluster was used as the production cluster.
+    It consisted of a total of 977 compute nodes, of which we
     used up to 16 general compute nodes at a time with \SI{186}{\gibi\byte} of
@@
     Lustre degradation, which was induced by ``busy writers''. Without busy
-    writers, the fMRI preprocessing pipelines have exclusive access to the
-    cluster. With busy writers, the fMRI preprocessing pipelines are executed
-    alongside a Spark application that continuously reads and writes approximately
+    writers, the fMRI preprocessing pipelines had exclusive access to the
+    cluster. With busy writers, the fMRI preprocessing pipelines were executed
+    alongside an Apache Spark application that continuously read and wrote approximately
     1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
@@
     Our experiments generally did not use prefetching, flushing or eviction. The
-    only pipeline configured to prefetch was SPM, as the input data is otherwise
-    read through a memmap (i.e. only loading necessary portions of the file to memory) with Lustre. Flushing would normally be necessary with
+    only pipeline configured to prefetch was SPM, as the input data was otherwise
+    read through a memmap (i.e. only loading necessary portions of the file to memory) 
+    with Lustre. Flushing would normally be necessary with
     preprocessing as the user would always require the output data. To
@@
 
-    All experiments were executed using 1, 8 and 16 processes. Each process consisted of a single application call processing a single fMRI image.
+    All experiments were executed using 1, 8 and 16 processes.
+    Each process consisted of a single application call processing a single fMRI image.
 
@@
 
-    It is believed that Sea's benefits are not only limited to performance as
+     Sea's benefits are not only limited to performance as
     users can also benefit from limiting the number of files created on Lustre
@@
 
+    \section{Code Availability}
+
+    Scripts used to
+    execute each pipeline can be found
+    at~\url{https://github.com/valhayot/sea-slurm-scripts}. 
+    Dockerfiles containing AFNI, SPM, and FSL with Sea pre-installed are available
+    at~\url{https://github.com/ValHayot/Sea/tree/master/containers/applications}.
+
+    \section{Data Availability}
+
+    The ds001545 dataset is openly available at \url{https://openneuro.org/datasets/ds001545/versions/1.1.1} in BIDS format.
+    The PREVENT-AD dataset is available at \url{https://openpreventad.loris.ca} under registered access.
+    The Human Connectome Project (HCP) dataset is available at \url{https://www.humanconnectome.org} under registered access.
+
     \bibliography{biblio}

### 2023-07-05 paper/sea-neuro/paper-neuro.tex
minor edits and comments

--- 2023-06-28 paper/sea-neuro/paper-neuro.tex
@@
 
-\title{Ephemeral hierarchical data-management for neuroimaging applications}
+\title{Ephemeral hierarchical data-management for neuroimaging applications \TG{Update title as discussed previously}}
 
@@
   analysis tools have yet to adopt strategies that mitigate the costs
-  associated with large data transfers. A major challenge in adapting neuroimaging
+  associated with large data transfers \TG{use a word better than transfers}. A major challenge in adapting neuroimaging
   applications for data-intensive processing is that they must be entirely
@@
     In this paper, we present Sea~\cite{sea-comp}, a data-management library designed to reduce
-    transfer-related overheads. We view Sea as a complement
+    transfer-related overheads. \TG{add a few sentences to summarize the main features of Sea} We view Sea as a complement
     to the existing stack of neuroimaging tools and standards meant to facilitate 
@@
     the \texttt{LD\_PRELOAD}
-    trick to intercept POSIX file system (more specifically, the GNU C library,
-    glibc) calls on Linux systems. This enables Sea to redirect write calls
+    trick to intercept POSIX file system calls (more specifically, file access calls to the GNU C library,
+    glibc) on Linux systems. This enables Sea to redirect write calls
     aimed at slower storage devices to a faster devices whenever possible.
@@
 
-    As neuroimaging pipeline results are typically required post-processing and
-    HPC compute-local resources are only accessible during the reserved
+    As neuroimaging pipeline results typically require post-processing and
+    HPC \TG{It'd be useful to add a sentence before explaining that neuroimaging applications typically run on HPC systems} compute-local resources are only accessible during the reserved
     duration, Sea provides functionality to flush and evict data to persistent
-    shared storage. This is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. A
-    separate thread is used, in this case, to avoid interrupting ongoing
-    processing with data management operations. Users must inform Sea of files
+    shared storage. To avoid interrupting ongoing
+    processing with data management operations, this is accomplished via a separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
     that need to be persisted to storage within a file called
@@
 \subsection{Speedups observed in controlled environment}
+\TG{I think the methods need to be summarized in 2-3 sentences here}
+\TG{I would also say "The SPM pipeline" rather than "SPM" (same for all pipelines)}
 
-In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
-variable, speedups are obtained by using Sea when Lustre
-performance has been degraded by busy writers. On average, a single SPM
+Speedups are obtained by using Sea when Lustre
+performance has been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}), although performance is variable. On average, a single SPM
 process preprocessing a single HCP image produced the greatest speedup
@@
 minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
+from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro-pipelines},
 AFNI performs a very high number of glibc calls. While the overhead of an
@@
 all these calls results in a large enough bottleneck that affects both Sea and
-Baseline executions an extensive amount. Moreover, the number of glibc calls
+Baseline executions to an extensive amount. Moreover, the number of glibc calls
 aimed at interacting with data on the Lustre storage are very few. AFNI still
@@
 line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup.
-
-We do occasionally observe slowdowns from using Sea. These slowdowns
+greater the speedup. We do occasionally observe slowdowns from using Sea. These slowdowns
 may arise from the initial read of the data, as they appear to occur less
@@
 (Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
-behave more or less identically. Given these results, we conclude that the
+behave mostly identically. Given these results, we conclude that the
 overhead of Sea is minimal and that the differences in makespan may be
@@
   shows us that although performance in production clusters is highly variable,
-  we can still obtain important speedups. It is very likely that when we
+  we can still obtain important speedups by using Sea. It is very likely that when we
   executed our experiments without flushing
@@
     Our results show that the speedups obtained from using Sea can be quite
-    substantial (up to 32$\times$) depending on the degradation of the Lustre
+    substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
     file system. The highest speed-ups observed on a production cluster were
@@
     execute each pipeline can be found
-    at~\url{github.com/valhayot/sea-slurm-scripts}.
+    at~\url{github.com/valhayot/sea-slurm-scripts}. \TG{I'd group some of these short paragraphs together}
 

### 2023-07-05 paper/sea-neuro/paper-neuro.tex
minor edits and comments

--- 2023-06-28 paper/sea-neuro/paper-neuro.tex
@@
   analysis tools have yet to adopt strategies that mitigate the costs
-  associated with large data transfers. A major challenge in adapting neuroimaging
-  analysis tools have yet to adopt strategies that mitigate the costs
-  associated with large data transfers. A major challenge in adapting neuroimaging
+  associated with large data transfers \TG{use a word better than transfers}. A major challenge in adapting neuroimaging
   applications for data-intensive processing is that they must be entirely
@@
     As neuroimaging pipeline results typically require post-processing and
-    HPC compute-local resources are only accessible during the reserved
+    HPC \TG{It'd be useful to add a sentence before explaining that neuroimaging applications typically run on HPC systems} compute-local resources are only accessible during the reserved
     duration, Sea provides functionality to flush and evict data to persistent
+    shared storage. To avoid interrupting ongoing
+    processing with data management operations, this is accomplished via a separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
     shared storage. To avoid interrupting ongoing
@@
 \subsection{Speedups observed in controlled environment}
+\TG{I think the methods need to be summarized in 2-3 sentences here}
+\TG{I would also say "The SPM pipeline" rather than "SPM" (same for all pipelines)}
 
-We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
-(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB) with 3 pipelines 
-from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
-dedicated to our experiments. We deployed ``busy writers'' in the cluster
-that simulated the load from other user applications by writing to the shared file system at a controlled rate.
-
-Speedups were obtained by using Sea when Lustre
-performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
-although performance was variable. On average, a single SPM pipeline
+Speedups are obtained by using Sea when Lustre
+performance has been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}), although performance is variable. On average, a single SPM
 process preprocessing a single HCP image produced the greatest speedup
@@
 minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
-AFNI pipeline performs a very high number of glibc calls. While the overhead of an
+from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro-pipelines},
+AFNI performs a very high number of glibc calls. While the overhead of an
 individual call is likely very minimal, it is possible that the culmination of
-all these calls results in a large enough bottleneck that affected both Sea and
+all these calls results in a large enough bottleneck that affects both Sea and
 Baseline executions to an extensive amount. Moreover, the number of glibc calls
-aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
+aimed at interacting with data on the Lustre storage are very few. AFNI still
 obtained an average of 2$\times$ speedup across all experimental conditions.
@@
 line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup. We did occasionally observe slowdowns from using Sea. These slowdowns
-may arise from the initial read of the data, as they appeared to occur less
-frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.
+greater the speedup. We do occasionally observe slowdowns from using Sea. These slowdowns
+may arise from the initial read of the data, as they appear to occur less
+frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.
 
@@
 the performance of pipeline executing entirely within memory
-(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
+(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
 behave mostly identically. Given these results, we conclude that the
@@
   Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
-  show that although performance in production clusters was highly variable,
-  we could still obtain important speedups by using Sea. It is very likely that when we
+  shows us that although performance in production clusters is highly variable,
+  we can still obtain important speedups by using Sea. It is very likely that when we
   executed our experiments without flushing
@@
 
-    Our results show that the speedups obtained from using Sea can be 
+    Our results show that the speedups obtained from using Sea can be quite
     substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
@@
     Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
+    Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
     in computation and data-intensiveness of the different pipelines on a single
@@
 
-    Each tool was set to only run the functional preprocessing pipeline. For the FSL preprocessing pipeline, 
-    we used FEAT configured with many of
+    Each tool was set to only run the preprocessing pipeline. Scripts used to
+    execute each pipeline can be found
+    at~\url{github.com/valhayot/sea-slurm-scripts}. \TG{I'd group some of these short paragraphs together}
+
+    For the FSL preprocessing pipeline, we used FEAT configured with many of
     the default parameters. We set the following parameters: slice-timing
@@
     \footnotesize{$*$ Measurements taken on the dedicated cluster}
+    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
+    \label{table:seaneuro-pipelines}
+  \end{table*}
     \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
@@
      Sea's benefits are not only limited to performance as
+     Sea's benefits are not only limited to performance as
     users can also benefit from limiting the number of files created on Lustre

### 2023-08-08 paper/sea-neuro/paper-neuro.tex
Edits up to section 2.2

@@
 
-\title{Ephemeral hierarchical data-management for neuroimaging applications \TG{Update title as discussed previously}}
+\title{Hierarchical storage management for neuroimaging applications in user space}
 
@@
   analysis tools have yet to adopt strategies that mitigate the costs
-  associated with large data transfers \TG{use a word better than transfers}. A major challenge in adapting neuroimaging
+  associated with large data transfers. A major challenge in adapting neuroimaging
   applications for data-intensive processing is that they must be entirely
@@
   where its benefits are limited. Overall, Sea is beneficial, even when performance gain
-  is minimal, as it can be used to limit files on PFS.
+  is minimal, as it can be used to limit the number of files created on parallel file systems.
 \end{abstract}
@@
     management challenges, from storage infrastructure to application processing times.
-    Many tools have been developed by the community to address to various needs in 
+    Many tools have been developed by the community to address various needs in 
     neuroimaging data management.
@@
     % focus on them.}.
-    For instance, standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids},
+    For instance, standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{gorgolewski2016brain},
     have been implemented to facilitate the sharing
-    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide 
+    of the datasets, and data management tools such as DataLad~\cite{halchenko2021datalad} have been developed to provide 
     versioning and provenance capture of data.
     Furthermore, recent developments in neuroimaging pipelines have addressed computation time 
-    limitations by adopting machine-learning approaches \TG{cite fastsurfer, hd-bet}.
+    limitations by adopting machine-learning approaches~\cite{henschel2020fastsurfer,hoffmann2021synthmorph}.
     While all these solutions to Big Data-related data management exist, certain aspects, 
@@
     Despite their respective high-scalability, these remote file systems cannot typically
-    compete with the performance of compute-local storage (e.g., tmpfs and SSDs). Should the pipelines 
+    compete with the performance of compute-local storage such as in-memory file systems (tmpfs) or solid-state drives (SSDs). Should the pipelines 
     need to transfer large amounts of data to these remote storage locations, significant data
@@
     Big Data in neuroimaging max exist in two formats: 1) very large files
-    (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
+    (e.g., BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
     population of individuals, such as those found in MRI datasets, that
@@
 
-    In this paper, we present Sea~\cite{sea-comp}, a data-management library designed to reduce
-    transfer-related overheads. \TG{add a few sentences to summarize the main features of Sea} We view Sea as a complement
+    In this paper, we present Sea~\cite{hayot2022sea}, a data-management library designed to reduce
+    transfer-related overheads in user space. Sea leverages the \texttt{LD\_PRELOAD} trick to redirect applications read and write calls to 
+    local or remote storage transparently. We view Sea as a complement
     to the existing stack of neuroimaging tools and standards meant to facilitate 
@@
     locations it can use to read and write to, and well as their order of
-    priority. It is important to note that Sea is not a filesystem, but rather a
+    priority. Importantly, Sea is not a filesystem, but rather a
     lightweight tool that redirects files to more optimal storage devices.
@@
     As neuroimaging pipeline results typically require post-processing and
-    HPC \TG{It'd be useful to add a sentence before explaining that neuroimaging applications typically run on HPC systems} compute-local resources are only accessible during the reserved
+    HPC compute-local resources are only accessible during the reserved
     duration, Sea provides functionality to flush and evict data to persistent

### 2023-08-08 paper/sea-neuro/paper-neuro.tex
Edits up to section 2.2

--- 2023-07-05 paper/sea-neuro/paper-neuro.tex
@@
   analysis tools have yet to adopt strategies that mitigate the costs
-  associated with large data transfers \TG{use a word better than transfers}. A major challenge in adapting neuroimaging
+  associated with large data transfers. A major challenge in adapting neuroimaging
   applications for data-intensive processing is that they must be entirely
-  rewritten.
-
-  To facilitate data management for standardized neuroimaging tools, we
+  rewritten. To facilitate data management for standardized neuroimaging tools, we
   developed Sea, a library that intercepts and redirects application read and

### 2023-08-08 paper/sea-neuro/paper-neuro.tex
Edits until the end

--- 2023-08-08 paper/sea-neuro/paper-neuro.tex
@@
 
-\title{Hierarchical storage management for neuroimaging applications in user space}
+\title{Hierarchical storage management in user space for neuroimaging applications}
 
@@
 \subsection{Speedups observed in controlled environment}
-\TG{I think the methods need to be summarized in 2-3 sentences here}
-\TG{I would also say "The SPM pipeline" rather than "SPM" (same for all pipelines)}
 
-Speedups are obtained by using Sea when Lustre
-performance has been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}), although performance is variable. On average, a single SPM
+We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
+(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB) with 3 pipelines 
+from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
+dedicated to our experiments. We deployed ``busy writers'' in the cluster
+that simulated the load from other user applications by writing to the shared file system at a controlled rate.
+
+Speedups were obtained by using Sea when Lustre
+performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
+although performance was variable. On average, a single SPM pipeline
 process preprocessing a single HCP image produced the greatest speedup
+(13$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
 (13$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
@@
 1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
+(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
+1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
 excellent speedups, which is likely due to a mix of prefetching the initial
-input files and the I/O patterns of the application. While it was expected that
-AFNI would benefit the most from Sea due to its
+input files and the I/O patterns of the application. While it was expected that the
+AFNI pipeline would benefit the most from Sea due to its
 minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro-pipelines},
-AFNI performs a very high number of glibc calls. While the overhead of an
+from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
+AFNI pipeline performs a very high number of glibc calls. While the overhead of an
 individual call is likely very minimal, it is possible that the culmination of
-all these calls results in a large enough bottleneck that affects both Sea and
+all these calls results in a large enough bottleneck that affected both Sea and
 Baseline executions to an extensive amount. Moreover, the number of glibc calls
-aimed at interacting with data on the Lustre storage are very few. AFNI still
+aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
 obtained an average of 2$\times$ speedup across all experimental conditions.
 
+The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
 The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
@@
 compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
+compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
 least significant speedups.
 
+The HCP dataset, on average, obtained the greatest speedups with
 The HCP dataset, on average, obtained the greatest speedups with
@@
 decreased with increased parallelism. We did not modify the default parallelism
+decreased with increased parallelism. We did not modify the default parallelism
 level of the applications in our experiments. As a result, each application
+process was likely attempting to utilize all cores, resulting in longer wait times
+with the increase in parallelism. With this unintentional additional wait time that
 process was likely attempting to utilize all cores, resulting in longer wait times
@@
 busy writers, Baseline performance deviated dramatically from the identity
+Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
+busy writers, Baseline performance deviated dramatically from the identity
 line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup. We do occasionally observe slowdowns from using Sea. These slowdowns
-may arise from the initial read of the data, as they appear to occur less
-frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.
+greater the speedup. We did occasionally observe slowdowns from using Sea. These slowdowns
+may arise from the initial read of the data, as they appeared to occur less
+frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.
 
@@
 the performance of pipeline executing entirely within memory
-(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrate that Sea and tmpfs
+(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
 behave mostly identically. Given these results, we conclude that the
@@
 \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
+\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
 \label{fig:seaneuro:beluga-wflush}
@@
   using the same datasets and pipelines as for the controlled cluster. 
+  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
+  shared with other users,
+  using the same datasets and pipelines as for the controlled cluster. 
   Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
-  shows us that although performance in production clusters is highly variable,
-  we can still obtain important speedups by using Sea. It is very likely that when we
+  show that although performance in production clusters was highly variable,
+  we could still obtain important speedups by using Sea. It is very likely that when we
   executed our experiments without flushing
@@
   obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
+  Our results with flushing enabled confirmed this theory. When running with
+  flushing enabled, which had the additional overhead of having to
+  ensure that all the data was copied to Lustre, we noticed that occasionally, we
+  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
   performance was degraded. The
   maximum speedup observed when running on the production cluster with flushing enabled was
+  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
   11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
@@
 
-    Our results show that the speedups obtained from using Sea can be quite
+    Our results show that the speedups obtained from using Sea can be 
     substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
@@
     obtained. Our controlled experiments demonstrated that using Sea with a
+    of the degradation was less than that of possible speedups
+    obtained. Our controlled experiments demonstrated that using Sea with a
     degraded Lustre file system can result in average speedups of up to
@@
      when the data is written at a rate which exceeds the rate at which the
+    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
+     when the data is written at a rate which exceeds the rate at which the
     compute node's page cache can flush results to Lustre. In our experiments,
@@
     output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
+    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
     approximately 100~GB of data, the controlled cluster would have had
+    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
     approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
@@
     the only way to augment data intensiveness of the applications is
+    our controlled experiments did not showcase the benefits of Sea in a highly
+    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
+    the only way to augment data intensiveness of the applications is
     through increased parallelism which is limited by the number of available
@@
     Lustre performance had been degraded.
+    cores. As a result, we could only demonstrate the speedups brought by Sea when
+    Lustre performance had been degraded.
 
+    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
+    In such cases, the performance bottleneck is the
     Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
@@
     limited in demonstrating potential speedups with Sea as there were many
+    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
+    where we had
+    control over the global cluster usage and were able to demonstrate that
+    significant speedups could be obtained when all the Lustre storage devices were
+    busy writing other data concurrently. However, even these experiments were
+    limited in demonstrating potential speedups with Sea as there were many
     controlled variables, some of which may vary from cluster to cluster. For
+    example, the production cluster had a total of 977 nodes and only 38 disks associated with
     example, the production cluster had a total of 977 nodes and only 38 disks associated with
@@
     be moved to and manipulated from memory. All of our experiments with the SPM pipeline
+    be moved to and manipulated from memory. All of our experiments with the SPM pipeline
     used prefetching as SPM updates the input data via the use of a memory map.
@@
     production environment experiments demonstrated, Sea's overhead is minimal
+    production environment experiments demonstrated, Sea's overhead is minimal
     and the magnitude of the speedups to be gained by the use of Sea is
@@
     this through the use of eviction which ensures that data that is not
+    required never makes it to Lustre.
     required never makes it to Lustre.
@@
     and some of the largest datasets. Of course, different modalities and tools
+    may result in vastly different data access patterns and compute times.
     may result in vastly different data access patterns and compute times.
@@
 
-    Each tool was set to only run the preprocessing pipeline. Scripts used to
-    execute each pipeline can be found
-    at~\url{github.com/valhayot/sea-slurm-scripts}. \TG{I'd group some of these short paragraphs together}
-
-    For the FSL preprocessing pipeline, we used FEAT configured with many of
+    Each tool was set to only run the functional preprocessing pipeline. For the FSL preprocessing pipeline, 
+    we used FEAT configured with many of
     the default parameters. We set the following parameters: slice-timing
@@
     To preprocess the datasets with SPM, we reused the template described in
+    To preprocess the datasets with SPM, we reused the template described in
     \cite{haitas2021age}, with fieldmap correction removed. As the template was
@@
     and slices and the TR and TA, for each dataset. Furthermore, we set the
+    slice order to interleaved in all cases. SPM version 12 was used.
     slice order to interleaved in all cases. SPM version 12 was used.
@@
     Our objective was to measure the speedup brought by Sea on real neuroimaging
+    Our objective was to measure the speedup brought by Sea on real neuroimaging
     pipelines. Although Sea is agnostic to the internals of the pipelines and
@@
     Our experiments were executed on two different clusters: 1) a controlled, dedicated cluster
+    Our experiments were executed on two different clusters: 1) a controlled, dedicated cluster
     and 2) and a production cluster. We chose this experimental setup as the dedicated
@@
     of Compute Canada's Beluga cluster.
+    of Compute Canada's Beluga cluster.
 
+    Compute Canada's Beluga cluster was used as the production cluster.
+    It consisted of a total of 977 compute nodes, of which we
     Compute Canada's Beluga cluster was used as the production cluster.
@@
     alongside an Apache Spark application that continuously read and wrote approximately
+    writers, the fMRI preprocessing pipelines had exclusive access to the
+    cluster. With busy writers, the fMRI preprocessing pipelines were executed
+    alongside an Apache Spark application that continuously read and wrote approximately
     1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
@@
     Our experiments generally did not use prefetching, flushing or eviction. The
+    only pipeline configured to prefetch was SPM, as the input data was otherwise
+    read through a memmap (i.e. only loading necessary portions of the file to memory) 
+    with Lustre. Flushing would normally be necessary with
     only pipeline configured to prefetch was SPM, as the input data was otherwise
@@
 
+    All experiments were executed using 1, 8 and 16 processes.
+    Each process consisted of a single application call processing a single fMRI image.
     All experiments were executed using 1, 8 and 16 processes.
@@
     The Human Connectome Project (HCP) dataset is available at \url{https://www.humanconnectome.org} under registered access.
+    \section{Code Availability}
+
+    Scripts used to
+    execute each pipeline can be found
+    at~\url{https://github.com/valhayot/sea-slurm-scripts}. 
+    Dockerfiles containing AFNI, SPM, and FSL with Sea pre-installed are available
+    at~\url{https://github.com/ValHayot/Sea/tree/master/containers/applications}.
+
+    \section{Data Availability}
+
+    The ds001545 dataset is openly available at \url{https://openneuro.org/datasets/ds001545/versions/1.1.1} in BIDS format.
+    The PREVENT-AD dataset is available at \url{https://openpreventad.loris.ca} under registered access.
+    The Human Connectome Project (HCP) dataset is available at \url{https://www.humanconnectome.org} under registered access.
 

### 2023-08-08 paper/sea-neuro/paper-neuro.tex
Edits until the end

--- 2023-08-08 paper/sea-neuro/paper-neuro.tex
@@
 	Neuroimaging open-data initiatives have led to increased availability of large
-  scientific datasets. While these datasets are shifting the processing
   scientific datasets. While these datasets are shifting the processing
@@
     their resulting workflow. To give
-    Although engines such as Nipype~\cite{nipype}
-    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
-    do not prohibit the use of data-management
-    strategies, they do not facilitate the integration of these strategies into
-    their resulting workflow. To give
     neuroimaging applications data management capabilities, the applications
@@
     HPC systems rely on scalable network-based parallel file systems (e.g.,
-    HPC systems rely on scalable network-based parallel file systems (e.g.,
     Lustre) for storage. While such file systems offer excellent performance,
@@
     throttling~\cite{huang2020ooops} the data-intensive workloads or
-    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
     recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
@@
     system must also be compliant to the format. One method to ensure that these
-    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
     conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
@@
     storage location it can write to based on the details provided in an
-    storage location it can write to based on the details provided in an
     initialization file called \texttt{sea.ini}. This file informs Sea of which
@@
     duration, Sea provides functionality to flush and evict data to persistent
-    shared storage. To avoid interrupting ongoing
-    processing with data management operations, this is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
     shared storage. To avoid interrupting ongoing
@@
 \end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
-denotes the total time between execution launch and completion of the last computing task.}
+\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
 \label{fig:seaneuro:slashbin}
@@
 \subsection{Speedups observed in controlled environment}
-\TG{I think the methods need to be summarized in 2-3 sentences here}
-\TG{I would also say "The SPM pipeline" rather than "SPM" (same for all pipelines)}
 
-Speedups are obtained by using Sea when Lustre
-performance has been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}), although performance is variable. On average, a single SPM
+We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
+(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB) with 3 pipelines 
+from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
+dedicated to our experiments. We deployed ``busy writers'' in the cluster
+that simulated the load from other user applications by writing to the shared file system at a controlled rate.
+
+Speedups were obtained by using Sea when Lustre
+performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
+although performance was variable. On average, a single SPM pipeline
 process preprocessing a single HCP image produced the greatest speedup
@@
 excellent speedups, which is likely due to a mix of prefetching the initial
-input files and the I/O patterns of the application. While it was expected that
-AFNI would benefit the most from Sea due to its
+input files and the I/O patterns of the application. While it was expected that the
+AFNI pipeline would benefit the most from Sea due to its
 minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro-pipelines},
-AFNI performs a very high number of glibc calls. While the overhead of an
+from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
+AFNI pipeline performs a very high number of glibc calls. While the overhead of an
 individual call is likely very minimal, it is possible that the culmination of
-all these calls results in a large enough bottleneck that affects both Sea and
+all these calls results in a large enough bottleneck that affected both Sea and
 Baseline executions to an extensive amount. Moreover, the number of glibc calls
-aimed at interacting with data on the Lustre storage are very few. AFNI still
+aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
 obtained an average of 2$\times$ speedup across all experimental conditions.
@@
 line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup. We do occasionally observe slowdowns from using Sea. These slowdowns
-may arise from the initial read of the data, as they appear to occur less
-frequently with SPM, or due to increased CPU contention caused by Sea's rapid I/O.
+greater the speedup. We did occasionally observe slowdowns from using Sea. These slowdowns
+may arise from the initial read of the data, as they appeared to occur less
+frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.
 
@@
   Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
-  shows us that although performance in production clusters is highly variable,
-  we can still obtain important speedups by using Sea. It is very likely that when we
+  show that although performance in production clusters was highly variable,
+  we could still obtain important speedups by using Sea. It is very likely that when we
   executed our experiments without flushing
@@
     Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
-    Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
     in computation and data-intensiveness of the different pipelines on a single
@@
     \footnotesize{$*$ Measurements taken on the dedicated cluster}
-    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
-    \label{table:seaneuro-pipelines}
-  \end{table*}
     \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
@@
      Sea's benefits are not only limited to performance as
-     Sea's benefits are not only limited to performance as
     users can also benefit from limiting the number of files created on Lustre

### 2024-01-28 paper/sea-neuro/paper-neuro.tex
added comment on makespan
Minor update to sea neuro paper text

@@
   unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
-  where its benefits are limited. Overall, Sea is beneficial, even when performance gain
-  is minimal, as it can be used to limit the number of files created on parallel file systems.
+  where its benefits are limited. Even in scenarios with minimal performance gain, Sea
+  can improve overall filesystem performance by limiting the number of files on shared filesystems.
 \end{abstract}
@@
     management challenges, from storage infrastructure to application processing times.
-    Many tools have been developed by the community to address various needs in 
-    neuroimaging data management.
+    To address the community's data management needs, many tools have been developped.
+    % Many tools have been developed by the community to address various needs in 
+    % neuroimaging data management.
     %To meet the storage, accessibility and security demands of neuroimaging data,
@@
     population of individuals, such as those found in MRI datasets, that
-    collectively amount to very large datasets. Differences in the number and
+    amount to very large datasets. Differences in the number and
     size of files result in different data transfer overheads. Whereas large
@@
     access to a wide variety of infrastructures, HPC clusters are a
-    cost-effective solution to accessing a wide array of resources for
+    cost-effective solution to accessing a wide array of hardware resources for
     researchers. %In this paper, we will focus on HPC processing.
@@
     with data-intensive applications can effectively deteriorate the performance
-    of the shared file system for all users on the cluster. Solutions for
+    of the shared file system for all active users. Solutions for
     improving shared file system performance include
@@
 
-    In this paper, we present Sea~\cite{hayot2022sea}, a data-management library designed to reduce
-    transfer-related overheads in user space. Sea leverages the \texttt{LD\_PRELOAD} trick to redirect applications read and write calls to 
-    local or remote storage transparently. We view Sea as a complement
+    In this paper, we evaluate Sea~\cite{hayot2022sea}, a data-management library designed to reduce
+    transfer-related overheads in user space, on real neuroimaging applications and quatify its
+    benefits to neuroimaging research. We view Sea as a complement
     to the existing stack of neuroimaging tools and standards meant to facilitate 
@@
     glibc) on Linux systems. This enables Sea to redirect write calls
-    aimed at slower storage devices to a faster devices whenever possible.
+    aimed at slower storage devices to faster devices whenever possible.
     Similarly, when intercepting read calls, Sea can choose to read from a
@@
     shared storage. To avoid interrupting ongoing
-    processing with data management operations, this is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
+    processing with data management operations, this is accomplished via a 
+    separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage.
+    Users must inform Sea of files
     that need to be persisted to storage within a file called
@@
 
-    To interact with Sea, a mountpoint is used. The mountpoint
+  Interfacing with Sea is achieved using a user-specified mountpoint. The mountpoint
     is an empty directory that behaves as a view to all the files and

### 2024-01-29 paper/sea-neuro/paper-neuro.tex
Some comments on t-test and speedup values

--- 2023-08-08 paper/sea-neuro/paper-neuro.tex
@@
     Furthermore, recent developments in neuroimaging pipelines have addressed computation time 
-    limitations by adopting machine-learning approaches~\cite{henschel2020fastsurfer,hoffmann2021synthmorph}.
+    limitations by adopting machine-learning approaches that are an order of magnitude faster than traditional image processing solutions~\cite{henschel2020fastsurfer,hoffmann2021synthmorph}.
     While all these solutions to Big Data-related data management exist, certain aspects, 
@@
 We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
-(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB) with 3 pipelines 
+(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB, see
+Table~\ref{table:sea-neuro:data}) with 3 pipelines 
 from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
@@
 
+\TG{Revise the numbers in this paragraph according to current notebook}
 Speedups were obtained by using Sea when Lustre
@@
 process preprocessing a single HCP image produced the greatest speedup
-(13$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
+(13$\times$) \TG{I think this number should be 32x, according to the current notebook}. The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
 using a single process was the next fastest pipeline, with an average speedup of
@@
 The HCP dataset, on average, obtained the greatest speedups with
-Sea. Out of the three datasets, HCP has the largest images (see
+Sea \TG{Mention value}. Out of the three datasets, HCP has the largest images (see
 Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
@@
 
-Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
-busy writers, Baseline performance deviated dramatically from the identity
-line. We observe that the more Baseline deviates from this identity line, the
-greater the speedup. We did occasionally observe slowdowns from using Sea. These slowdowns
+Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} \TG{(p=x, two-sample T-test)}. However, with
+busy writers, the makespan obtained with Sea was smaller than baseline \TG{(p=x, two-sample T-test)}. We observe that the more Baseline deviates from this identity line, the
+greater the speedup \TG{I would remove this sentence, it's obvious}. We did occasionally observe slowdowns from using Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less
@@
 (Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
-behave mostly identically. Given these results, we conclude that the
+did not lead to significant differences in average makespan (\TG{p=x, two-sampled T-test}). Given these results, we conclude that the
 overhead of Sea is minimal and that the differences in makespan may be
@@
 
-    Our results show that the speedups obtained from using Sea can be 
+    Our results show that the speedups obtained from using Sea when pre-processing functional MRI data can be 
     substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
     file system. The highest speed-ups observed on a production cluster were
-    found to be 11$\times$.
+    found to be 11$\times$ \TG{Double check this number}.
 
@@
     32$\times$ in the controlled environment and 11$\times$ in a production
-    cluster. It was also found that speedups were greatest with a larger
+    cluster \TG{Double check this number}. It was also found that speedups were greatest with a larger
     filesize, as long as the compute time and non-Lustre I/O-related overheads

### 2024-01-29 paper/sea-neuro/paper-neuro.tex
Some comments on t-test and speedup values

@@
 	Neuroimaging open-data initiatives have led to increased availability of large
+  scientific datasets. While these datasets are shifting the processing
   scientific datasets. While these datasets are shifting the processing
@@
     duration, Sea provides functionality to flush and evict data to persistent
+    shared storage. To avoid interrupting ongoing
+    processing with data management operations, this is accomplished via a separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
     shared storage. To avoid interrupting ongoing
@@
 1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
+(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
+1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
 excellent speedups, which is likely due to a mix of prefetching the initial
@@
 The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
+The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
 applications. Not only did it spend an extensive amount of time computing, the
 amount of output data generated was least of the three. Due to the
+compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
 compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
@@
 
+  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
+  shared with other users,
+  using the same datasets and pipelines as for the controlled cluster. 
   We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
@@
   \end{table*}
+    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
+    \label{table:seaneuro-pipelines}
+  \end{table*}
 
     \subsection{Controls}
+    Our objective was to measure the speedup brought by Sea on real neuroimaging
     Our objective was to measure the speedup brought by Sea on real neuroimaging

### 2024-03-13 paper/sea-neuro/paper-neuro.tex
Update paper

@@
     the \texttt{LD\_PRELOAD}
-    trick to intercept POSIX file system calls (more specifically, file access calls to the GNU C library,
-    glibc) on Linux systems. This enables Sea to redirect write calls
-    aimed at slower storage devices to a faster devices whenever possible.
+    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
+    glibc, on Linux systems. This enables Sea to redirect write calls
+    aimed at slower storage devices to a faster device whenever possible.
     Similarly, when intercepting read calls, Sea can choose to read from a
@@
 process preprocessing a single HCP image produced the greatest speedup
-(13$\times$) \TG{I think this number should be 32x, according to the current notebook}. The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
+(12.6$\times$),  \TG{I think this number should be 32x, according to the current notebook}. The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
 using a single process was the next fastest pipeline, with an average speedup of
-(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
-1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
-(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
+(4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
 1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
@@
 amount of output data generated was least of the three. Due to the
-compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
-compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
+compute-intensive nature of the FSL Feat pipeline, it is expected to have resulted in the
 least significant speedups.
@@
 Sea \TG{Mention value}. Out of the three datasets, HCP has the largest images (see
-Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
+Table~\ref{table:sea-neuro:data}). Larger individual images
 occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
@@
 Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} \TG{(p=x, two-sample T-test)}. However, with
-busy writers, the makespan obtained with Sea was smaller than baseline \TG{(p=x, two-sample T-test)}. We observe that the more Baseline deviates from this identity line, the
-greater the speedup \TG{I would remove this sentence, it's obvious}. We did occasionally observe slowdowns from using Sea. These slowdowns
+busy writers, the makespan obtained with Sea was smaller than baseline \TG{(p=x, two-sample T-test)}. We did occasionally observe slowdowns from using Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less

### 2024-03-13 paper/sea-neuro/paper-neuro.tex
Some fixes made during meeting with @valhayot

--- 2024-01-29 paper/sea-neuro/paper-neuro.tex
@@
 
-\TG{Revise the numbers in this paragraph according to current notebook}
 Speedups were obtained by using Sea when Lustre
 performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
-although performance was variable. On average, a single SPM pipeline
+although performance was variable. The largest speedup across all conditions
+was found to be 32$\times$ and occurred when SPM was processing a single
+HCP image with 6 active busy writers.
+On average, however, a single SPM pipeline
 process preprocessing a single HCP image produced the greatest speedup
-(13$\times$) \TG{I think this number should be 32x, according to the current notebook}. The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
+(12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
 using a single process was the next fastest pipeline, with an average speedup of
-(5$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
+(4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
 1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
@@
 
-Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} \TG{(p=x, two-sample T-test)}. However, with
-busy writers, the makespan obtained with Sea was smaller than baseline \TG{(p=x, two-sample T-test)}. We observe that the more Baseline deviates from this identity line, the
-greater the speedup \TG{I would remove this sentence, it's obvious}. We did occasionally observe slowdowns from using Sea. These slowdowns
+Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} (p=0.7, two-sample unpaired T-test). However, with
+busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample T-test). We did occasionally observe slowdowns from using Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less
@@
     Our results show that the speedups obtained from using Sea when pre-processing functional MRI data can be 
-    substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
+    substantial (up to 32$\times$) depending on the degradation of the Lustre
     file system. The highest speed-ups observed on a production cluster were
-    found to be 11$\times$ \TG{Double check this number}.
+    found to be 11$\times$.
 
@@
     32$\times$ in the controlled environment and 11$\times$ in a production
-    cluster \TG{Double check this number}. It was also found that speedups were greatest with a larger
+    cluster. It was also found that speedups were greatest with a larger
     filesize, as long as the compute time and non-Lustre I/O-related overheads

### 2024-03-13 paper/sea-neuro/paper-neuro.tex

@@
 
-The HCP dataset, on average, obtained the greatest speedups with
-Sea \TG{Mention value}. Out of the three datasets, HCP has the largest images (see
+The HCP dataset obtained the greatest speedups with
+Sea, with $2.3\times$ on average vs $1.3\times$ for PREVENT-AD and $1.9\times$ for ds001545.
+Out of the three datasets, HCP has the largest images (see
 Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
@@
 (Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
-did not lead to significant differences in average makespan (\TG{p=x, two-sampled T-test}). Given these results, we conclude that the
+did not lead to significant differences in average makespan (p=0.9, two-sampled unpaired T-test). Given these results, we conclude that the
 overhead of Sea is minimal and that the differences in makespan may be

### 2024-03-13 paper/sea-neuro/paper-neuro.tex

--- 2024-03-13 paper/sea-neuro/paper-neuro.tex
@@
 performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
-although performance was variable. On average, a single SPM pipeline
+although performance was variable. The largest speedup across all conditions
+was found to be 32$\times$ and occurred when SPM was processing a single
+HCP image with 6 active busy writers.
+On average, however, a single SPM pipeline
 process preprocessing a single HCP image produced the greatest speedup
-(12.6$\times$),  \TG{I think this number should be 32x, according to the current notebook}. The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
+(12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
 using a single process was the next fastest pipeline, with an average speedup of
@@
 
-The HCP dataset, on average, obtained the greatest speedups with
-Sea \TG{Mention value}. Out of the three datasets, HCP has the largest images (see
-Table~\ref{table:sea-neuro:data}). Larger individual images
+The HCP dataset obtained the greatest speedups with
+Sea, with $2.3\times$ on average vs $1.3\times$ for PREVENT-AD and $1.9\times$ for ds001545.
+Out of the three datasets, HCP has the largest images (see
+Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
 occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
@@
 
-Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} \TG{(p=x, two-sample T-test)}. However, with
-busy writers, the makespan obtained with Sea was smaller than baseline \TG{(p=x, two-sample T-test)}. We did occasionally observe slowdowns from using Sea. These slowdowns
+Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} (p=0.7, two-sample unpaired T-test). However, with
+busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample T-test). We did occasionally observe slowdowns from using Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less
@@
     Our results show that the speedups obtained from using Sea when pre-processing functional MRI data can be 
-    substantial (up to 32$\times$ \TG{this value should be mentioned in the results}) depending on the degradation of the Lustre
+    substantial (up to 32$\times$) depending on the degradation of the Lustre
     file system. The highest speed-ups observed on a production cluster were

### 2024-03-13 paper/sea-neuro/paper-neuro.tex
Closed comments

@@
     neuroimaging data management.
-    %To meet the storage, accessibility and security demands of neuroimaging data,
-    %large datasets have been stored in object stores provided by cloud storage providers
-    %\TG{mentioning clouds so early in the paper may give the impression that the paper will 
-    % focus on them.}.
     For instance, standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{gorgolewski2016brain},
@@
     Although engines such as Nipype~\cite{nipype}
-    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
-    do not prohibit the use of data-management
-    strategies, they do not facilitate the integration of these strategies into
-    their resulting workflow. To give
-    Although engines such as Nipype~\cite{nipype}
-    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
     do not prohibit the use of data-management
@@
 
-\TG{Revise the numbers in this paragraph according to current notebook}
 Speedups were obtained by using Sea when Lustre
@@
 \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
-\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
 \label{fig:seaneuro:beluga-wflush}
@@
     file system. The highest speed-ups observed on a production cluster were
-    found to be 11$\times$ \TG{Double check this number}.
+    found to be 11$\times$.
 
@@
   \end{table*}
-    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
-    \label{table:seaneuro-pipelines}
-  \end{table*}
 

### 2024-03-13 paper/sea-neuro/paper-neuro.tex
Removed duplicates

--- 2024-03-13 paper/sea-neuro/paper-neuro.tex
@@
   (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
-  computing clusters. Our results demonstrate that given the features of the
-  applications and datasets, Sea provides large speedups (up to 32$\times$) when
+  computing clusters. Our results demonstrate that Sea provides large speedups (up to 32$\times$) when
   the shared file system's (e.g. Lustre) performance is deteriorated. When the
@@
 
-    Big Data in neuroimaging max exist in two formats: 1) very large files
+    Big Data in neuroimaging may exist in two formats: 1) very large files
     (e.g., BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
@@
     Although engines such as Nipype~\cite{nipype}
-    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
     do not prohibit the use of data-management
@@
 amount of output data generated was least of the three. Due to the
-compute-intensive nature of the FSL Feat pipeline, it is expected that it resulted in the
+compute-intensive nature of the FSL Feat pipeline, it is expected to have resulted in the
 least significant speedups.
@@
 level of the applications in our experiments. As a result, each application
+process was likely attempting to utilize all cores, resulting in longer wait times
+with the increase in parallelism. With this unintentional additional wait time that
 process was likely attempting to utilize all cores, resulting in longer wait times
@@
   using the same datasets and pipelines as for the controlled cluster. 
+  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
+  shared with other users,
+  using the same datasets and pipelines as for the controlled cluster. 
   Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
@@
 
+  Our results with flushing enabled confirmed this theory. When running with
+  flushing enabled, which had the additional overhead of having to
+  ensure that all the data was copied to Lustre, we noticed that occasionally, we
+  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
   Our results with flushing enabled confirmed this theory. When running with
@@
 
+    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
+     when the data is written at a rate which exceeds the rate at which the
     Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
@@
 
+    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
+    where we had
+    control over the global cluster usage and were able to demonstrate that
+    significant speedups could be obtained when all the Lustre storage devices were
+    busy writing other data concurrently. However, even these experiments were
+    limited in demonstrating potential speedups with Sea as there were many
     Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster

### 2024-03-13 paper/sea-neuro/paper-neuro.tex
Removed more duplicates

@@
     ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
-    shared storage. To avoid interrupting ongoing
-    processing with data management operations, this is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
     that need to be persisted to storage within a file called
@@
 with the increase in parallelism. With this unintentional additional wait time that
-process was likely attempting to utilize all cores, resulting in longer wait times
-with the increase in parallelism. With this unintentional additional wait time that
 is added to the applications due to compute contention, the benefits of using
@@
   using the same datasets and pipelines as for the controlled cluster. 
-  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
-  shared with other users,
-  using the same datasets and pipelines as for the controlled cluster. 
   Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
@@
 
-  Our results with flushing enabled confirmed this theory. When running with
-  flushing enabled, which had the additional overhead of having to
-  ensure that all the data was copied to Lustre, we noticed that occasionally, we
-  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
   Our results with flushing enabled confirmed this theory. When running with
@@
     obtained. Our controlled experiments demonstrated that using Sea with a
-    of the degradation was less than that of possible speedups
-    obtained. Our controlled experiments demonstrated that using Sea with a
     degraded Lustre file system can result in average speedups of up to
@@
 
-    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
-     when the data is written at a rate which exceeds the rate at which the
     Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
@@
     the only way to augment data intensiveness of the applications is
-    our controlled experiments did not showcase the benefits of Sea in a highly
-    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
-    the only way to augment data intensiveness of the applications is
     through increased parallelism which is limited by the number of available
@@
     Lustre performance had been degraded.
-    cores. As a result, we could only demonstrate the speedups brought by Sea when
-    Lustre performance had been degraded.
 
-    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
-    In such cases, the performance bottleneck is the
     Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
@@
 
-    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
-    where we had
-    control over the global cluster usage and were able to demonstrate that
-    significant speedups could be obtained when all the Lustre storage devices were
-    busy writing other data concurrently. However, even these experiments were
-    limited in demonstrating potential speedups with Sea as there were many
     Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
@@
     alongside an Apache Spark application that continuously read and wrote approximately
-    writers, the fMRI preprocessing pipelines had exclusive access to the
-    cluster. With busy writers, the fMRI preprocessing pipelines were executed
-    alongside an Apache Spark application that continuously read and wrote approximately
     1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
@@
     Our experiments generally did not use prefetching, flushing or eviction. The
-    only pipeline configured to prefetch was SPM, as the input data was otherwise
-    read through a memmap (i.e. only loading necessary portions of the file to memory) 
-    with Lustre. Flushing would normally be necessary with
     only pipeline configured to prefetch was SPM, as the input data was otherwise
@@
 
-    All experiments were executed using 1, 8 and 16 processes.
-    Each process consisted of a single application call processing a single fMRI image.
     All experiments were executed using 1, 8 and 16 processes.

### 2024-03-13 paper/sea-neuro/paper-neuro.tex

@@
 The HCP dataset obtained the greatest speedups with
-Sea, with $2.3\times$ on average vs $1.3\times$ for PREVENT-AD and $1.9\times$ for ds001545.
+Sea, with $2.3\times$ on average vs. $1.9\times$ for ds001545  and $1.3\times$ for PREVENT-AD.
 Out of the three datasets, HCP has the largest images (see
@@
 
-Baseline performance without busy writers was comparable to that of Sea's as can be seen in Figure~\ref{fig:seaneuro:slashbin} (p=0.7, two-sample unpaired T-test). However, with
-busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample T-test). We did occasionally observe slowdowns from using Sea. These slowdowns
+Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
+busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). We did occasionally observe slowdowns from using Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less
@@
 (Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
-did not lead to significant differences in average makespan (p=0.9, two-sampled unpaired T-test). Given these results, we conclude that the
+did not lead to significant differences in average makespan (p=0.9, two-sampled unpaired t-test). Given these results, we conclude that the
 overhead of Sea is minimal and that the differences in makespan may be
@@
     via a 20Gbps ethernet connection. The cluster's hardware was similar to that
-    of Compute Canada's Beluga cluster.
+    of the Beluga cluster at the Digital Alliance of Canada.
 
-    Compute Canada's Beluga cluster was used as the production cluster.
+    The Beluga cluster was used as the production cluster.
     It consisted of a total of 977 compute nodes, of which we

### 2024-03-14 paper/sea-neuro/paper-neuro.tex

@@
 process was likely attempting to utilize all cores, resulting in longer wait times
-with the increase in parallelism. With this unintentional additional wait time that
-is added to the applications due to compute contention, the benefits of using
+with the increase in parallelism. With the added compute contention overhead, the benefits of using
 Sea becomes more limited.
@@
 Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
-busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). We did occasionally observe slowdowns from using Sea. These slowdowns
+busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). Occassional slowdowns did occur with the use of Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less
@@
 
-    While on the production cluster, Sea and Baseline generally performed similarly,
-    it is impossible for cluster users to know if Lustre's
+    It is unrealistic for cluster users to know whether Lustre's
     performance will be deteriorated during the lifetime of their application
@@
     degraded Lustre file system can result in average speedups of up to
-    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
-    less important that the potential speedups that can arise.
+    2.5$\times$, which is likely to benefit scientific application deployments in production.
 
@@
     factors that can affect reported results, others can include when the
-    experiment were executed, what kind of applications were being executed
+    experiment was executed, what kind of applications were being executed
     alongside the experiment, scratch file system usage during experiments, and

### 2024-08-02 paper/sea-neuro/paper-neuro.tex
Add missing subsections to neuro paper

@@
   Neuroimaging open-data initiatives have led to increased availability of large
-  scientific datasets. While these datasets are shifting the processing
-  bottleneck from compute-intensive to data-intensive, current standardized
-  analysis tools have yet to adopt strategies that mitigate the costs
-  associated with large data transfers. A major challenge in adapting neuroimaging
-  applications for data-intensive processing is that they must be entirely
-  rewritten. To facilitate data management for standardized neuroimaging tools, we
-  developed Sea, a library that intercepts and redirects application read and
-  write calls to minimize data transfer time. In this paper, we investigate the
-  performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
-  SPM and AFNI), using three neuroimaging datasets of different sizes
-  (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
-  computing clusters. Our results demonstrate that Sea provides large speedups (up to 32$\times$) when
-  the shared file system's (e.g. Lustre) performance is deteriorated. When the
-  shared file system is not overburdened by other users, performance is
-  unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
-  where its benefits are limited. Overall, Sea is beneficial, even when performance gain
-  is minimal, as it can be used to limit the number of files created on parallel file systems.
+  scientific datasets. While these datasets are shifting the processing bottleneck
+  from compute-intensive to data-intensive, current standardized analysis tools
+  have yet to adopt strategies that mitigate the costs associated with large data
+  transfers. A major challenge in adapting neuroimaging applications for data-intensive
+  processing is that they must be entirely rewritten. To facilitate data management for
+  standardized neuroimaging tools, we developed Sea, a library that intercepts and
+  redirects application read and write calls to minimize data transfer time. In this
+  paper, we investigate the performance of Sea on three preprocessing pipelines applied
+  to three different neuroimaging datasets on two high-performance computing clusters.
+  Our results demonstrate that Sea provides large speedups (up to 32$\times$) when the
+  shared file system’s performance is deteriorated. When the shared file system is not
+  overburdened by other users, performance is unaffected by Sea, suggesting that Sea’s
+  overhead is minimal even in cases where its benefits are limited.
+
+	% Neuroimaging open-data initiatives have led to increased availability of large
+  % scientific datasets. While these datasets are shifting the processing
+  % bottleneck from compute-intensive to data-intensive, current standardized
+  % analysis tools have yet to adopt strategies that mitigate the costs
+  % associated with large data transfers. A major challenge in adapting neuroimaging
+  % applications for data-intensive processing is that they must be entirely
+  % rewritten. To facilitate data management for standardized neuroimaging tools, we
+  % developed Sea, a library that intercepts and redirects application read and
+  % write calls to minimize data transfer time. In this paper, we investigate the
+  % performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
+  % SPM and AFNI), using three neuroimaging datasets of different sizes
+  % (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
+  % computing clusters. Our results demonstrate that Sea provides large speedups (up to 32$\times$) when
+  % the shared file system's (e.g. Lustre) performance is deteriorated. When the
+  % shared file system is not overburdened by other users, performance is
+  % unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
+  % where its benefits are limited. Overall, Sea is beneficial, even when performance gain
+  % is minimal, as it can be used to limit the number of files created on parallel file systems.
 \end{abstract}
@@
 
+    \subsection{Conclusion and Future work}
+
+    We developed a lightweight data-management library, Sea, to seamlessly
+    provide data management capabilities to existing neuroimaging workflows. We
+    have tested our library on three of the most commonly used fMRI
+    preprocessing libraries and found that Sea was able to intercept and
+    redirect all application I/O calls without issue.
+
+    When benchmarking the performance of Sea compared to Baseline, it was found that Sea provided speedups
+    when Lustre performance was degraded. We were able to obtain up to a
+    32$\times$ in the controlled environment and 11$\times$ in a production
+    cluster. It was also found that speedups were greatest with a larger
+    filesize, as long as the compute time and non-Lustre I/O-related overheads
+    were also minimal.
+
+    Sea's benefits are not only limited to performance as
+    users can also benefit from limiting the number of files created on Lustre
+    with Sea. Archiving of the output directory on Lustre with Sea to further
+    reduce number of files may be an interesting addition to Sea.
+
     %user
@@
     %Busy writers
-    \section{Conclusion and Future work}
 
-    We developed a lightweight data-management library, Sea, to seamlessly
-    provide data management capabilities to existing neuroimaging workflows. We
-    have tested our library on three of the most commonly used fMRI
-    preprocessing libraries and found that Sea was able to intercept and
-    redirect all application I/O calls without issue.
+    \section{Data Availability}
 
-    When benchmarking the performance of Sea compared to Baseline, it was found that Sea provided speedups
-    when Lustre performance was degraded. We were able to obtain up to a
-    32$\times$ in the controlled environment and 11$\times$ in a production
-    cluster. It was also found that speedups were greatest with a larger
-    filesize, as long as the compute time and non-Lustre I/O-related overheads
-    were also minimal.
-
-    Sea's benefits are not only limited to performance as
-    users can also benefit from limiting the number of files created on Lustre
-    with Sea. Archiving of the output directory on Lustre with Sea to further
-    reduce number of files may be an interesting addition to Sea.
+    The ds001545 dataset is openly available at \url{https://openneuro.org/datasets/ds001545/versions/1.1.1} in BIDS format.
+    The PREVENT-AD dataset is available at \url{https://openpreventad.loris.ca} under registered access.
+    The Human Connectome Project (HCP) dataset is available at \url{https://www.humanconnectome.org} under registered access.
 
@@
 
-    \section{Data Availability}
+    \section{Acknowledgements}
+    Val\'erie Hayot-Sasson was funded by the Canadian Open
+    Neuroscience Platform (CONP) Scholar Award. This work
+    was also supported by the Canada Foundation for Innova-
+    tion as well as the Canada Research Chairs program.
 
-    The ds001545 dataset is openly available at \url{https://openneuro.org/datasets/ds001545/versions/1.1.1} in BIDS format.
-    The PREVENT-AD dataset is available at \url{https://openpreventad.loris.ca} under registered access.
-    The Human Connectome Project (HCP) dataset is available at \url{https://www.humanconnectome.org} under registered access.
+    \section{Author Contributions}
+    VHS contributed to idea development, experimental and model design, implementation of experiments and
+    evaluation. TG contributed to idea development, experimental design and implementation.
+
+    \section{Competing Interests}
+    The authors declare no competing interests.
 

### 2024-11-06 paper/sea-neuro/paper-neuro.tex
Change latex template

@@
-\documentclass[fleqn,10pt]{wlscirep}
+\documentclass[utf8]{FrontiersinHarvard}
+\usepackage{url,hyperref,lineno,microtype,subcaption}
+\usepackage[onehalfspacing]{setspace}
 \usepackage[utf8]{inputenc}
@@
 
-\title{Hierarchical storage management in user space for neuroimaging applications}
+%\title{Hierarchical storage management in user space for neuroimaging applications}
 
-\author{Val\'erie Hayot-Sasson}
-\author{Tristan Glatard}
-\affil{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+\def\firstAuthorLast{V. Hayot-Sasson {et~al.}} %use et al only if is more than 1 author
+\def\Authors{Val\'erie Hayot-Sasson\,$^{1,*}$ and Tristan Glatard\,$^{1}$}
+% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
+% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
+\def\Address{$^{1}$Department of Computer Science and Software Engineering, Concordia University, Montreal, Qu\'ebec, Canada}
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
+\def\corrAuthor{Corresponding Author}
 
+\def\corrEmail{vhayot@uchicago.edu}
+% \author{Val\'erie Hayot-Sasson}
+% \author{Tristan Glatard}
+% \affil{}
+
+\begin{document}
+    \onecolumn
+    \firstpage{1}
+
+  \title[Sea]{Hierarchical storage management in user space for neuroimaging applications}
+
+  \author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+  \address{} %This field will be automatically populated
+  \correspondance{} 
+  \extraAuth{}
+
+  \maketitle
   \begin{abstract}
@@
 \end{abstract}
-\begin{document}
 
-\flushbottom
-\maketitle
-%  Click the title above to edit the author information and abstract
+% \flushbottom
+% \maketitle
+% %  Click the title above to edit the author information and abstract
 
-\thispagestyle{empty}
+% \thispagestyle{empty}
 
@@
 
-    \section{Results}
-
-    \subsection{Sea}
-
-\begin{figure*}
-
-    \centering
-    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
-\caption{The Sea data-management library overview}
-\label{fig:seaneuro:diagram}
-\end{figure*}
-    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
-    the \texttt{LD\_PRELOAD}
-    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
-    glibc, on Linux systems. This enables Sea to redirect write calls
-    aimed at slower storage devices to a faster device whenever possible.
-    Similarly, when intercepting read calls, Sea can choose to read from a
-    faster device if a copy is available on that device. Sea decides which
-    storage location it can write to based on the details provided in an
-    initialization file called \texttt{sea.ini}. This file informs Sea of which
-    locations it can use to read and write to, and well as their order of
-    priority. Importantly, Sea is not a filesystem, but rather a
-    lightweight tool that redirects files to more optimal storage devices.
-
-    We define all locations in which Sea can redirect writes to as a cache. Such
-    caches include any available storage located on the compute nodes, such as
-    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs).
-
-    As neuroimaging pipeline results typically require post-processing and
-    HPC compute-local resources are only accessible during the reserved
-    duration, Sea provides functionality to flush and evict data to persistent
-    shared storage. To avoid interrupting ongoing
-    processing with data management operations, this is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
-    that need to be persisted to storage within a file called
-    \texttt{.sea\_flushlist}, and temporary files which can be removed from
-    cache within a file called the \texttt{.sea\_evictlist}. Both these files
-    can be populated using regular expressions to denote the paths to flush and
-    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
-    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
-    operation and simply move the file from local to persistent storage rather
-    than copying it. Files that will be reused by the application should only be
-    flushed rather than flushed and evicted, as files can benefit from speedups
-    related to reading from fast rather than slow storage. Sea currently also
-    provides a rudimentary prefetch thread that can move files located within
-    Sea to the fastest available cache. To use Sea's prefetch
-    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
-    populated using regular expressions like the flushing and eviction files.
-
-    To interact with Sea, a mountpoint is used. The mountpoint
-    is an empty directory that behaves as a view to all the files and
-    directories stored within Sea. In order to keep track of the locations of
-    files within the mountpoint, Sea mirrors the directory structure of each
-    storage location across all caches. It is generally advisable to provide
-    empty storage locations for Sea to use as the mirroring of large directories
-    can take some time.
-
-    Sea can easily be executed directly using the available containers on the
-    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
-    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.
-
-% \tristan{I find this section lacks a clear focus. What are the key points from
-% these results? You should focus on them and possibly name your paragraphs
-% accordingly, for improved clarity. My takeaways would be the following:
-% \begin{enumerate} \item In controlled conditions, important speed-ups are
-% observed. Mention min, average and max speed-up for all conditions. \item
-% Speed-up is correlated with baseline makespan \item Other determinants of
-% speed-up are unclear \item Overhead is minor \item Speed-ups are also observed
-% in production cluster \end{enumerate}
-% }
-\begin{figure*}
-
-\begin{subfigure}{0.5\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
-    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
-    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
-\end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
-denotes the total time between execution launch and completion of the last computing task.}
-\label{fig:seaneuro:slashbin}
-\end{figure*}
-
-\subsection{Speedups observed in controlled environment}
-
-We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
-(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB, see
-Table~\ref{table:sea-neuro:data}) with 3 pipelines 
-from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
-dedicated to our experiments. We deployed ``busy writers'' in the cluster
-that simulated the load from other user applications by writing to the shared file system at a controlled rate.
-
-Speedups were obtained by using Sea when Lustre
-performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
-although performance was variable. The largest speedup across all conditions
-was found to be 32$\times$ and occurred when SPM was processing a single
-HCP image with 6 active busy writers.
-On average, however, a single SPM pipeline
-process preprocessing a single HCP image produced the greatest speedup
-(12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
-using a single process was the next fastest pipeline, with an average speedup of
-(4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
-1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
-excellent speedups, which is likely due to a mix of prefetching the initial
-input files and the I/O patterns of the application. While it was expected that the
-AFNI pipeline would benefit the most from Sea due to its
-minimal compute time and the amount of data it generates, it did not benefit
-from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
-AFNI pipeline performs a very high number of glibc calls. While the overhead of an
-individual call is likely very minimal, it is possible that the culmination of
-all these calls results in a large enough bottleneck that affected both Sea and
-Baseline executions to an extensive amount. Moreover, the number of glibc calls
-aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
-obtained an average of 2$\times$ speedup across all experimental conditions.
-
-The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
-applications. Not only did it spend an extensive amount of time computing, the
-amount of output data generated was least of the three. Due to the
-compute-intensive nature of the FSL Feat pipeline, it is expected to have resulted in the
-least significant speedups.
-
-The HCP dataset obtained the greatest speedups with
-Sea, with $2.3\times$ on average vs. $1.9\times$ for ds001545  and $1.3\times$ for PREVENT-AD.
-Out of the three datasets, HCP has the largest images (see
-Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
-occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
-the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
-single image) results in the next largest speedups, followed by PREVENT-AD, the
-dataset with the smallest functional images.
-
-For both the pipelines and the datasets, we noticed that obtained speedups
-decreased with increased parallelism. We did not modify the default parallelism
-level of the applications in our experiments. As a result, each application
-process was likely attempting to utilize all cores, resulting in longer wait times
-with the increase in parallelism. With the added compute contention overhead, the benefits of using
-Sea becomes more limited.
-
-\subsection{Speedup correlated with Lustre degradation}
-
-Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
-busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). Occassional slowdowns did occur with the use of Sea. These slowdowns
-may arise from the initial read of the data, as they appeared to occur less
-frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.
-
-\subsection{Overhead of Sea was minimal}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.7\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
-\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
-\label{fig:seaneuro:tmpfs}
-\end{figure*}
-
-To determine the overhead of Sea, we compared the average performance of Sea with
-the performance of pipeline executing entirely within memory
-(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
-did not lead to significant differences in average makespan (p=0.9, two-sampled unpaired t-test). Given these results, we conclude that the
-overhead of Sea is minimal and that the differences in makespan may be
-due to Sea's initial reading of the data from Lustre.
-
-\subsection{Speedups observed in production environments}
-
-\begin{figure*}
-
-\begin{subfigure}{0.5\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
-    \caption{Complete results}\label{fig:seaneuro:belugafull}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
-    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
-\end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
-\label{fig:seaneuro:beluga-noflush}
-\end{figure*}
-
-\begin{figure*}
-\begin{subfigure}{0.5\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
-    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-    \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
-    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
-\end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
-\label{fig:seaneuro:beluga-wflush}
-\end{figure*}
-
-  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
-  shared with other users,
-  using the same datasets and pipelines as for the controlled cluster. 
-  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
-  show that although performance in production clusters was highly variable,
-  we could still obtain important speedups by using Sea. It is very likely that when we
-  executed our experiments without flushing
-  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
-  resulting in Sea and Baseline performing
-  quite similarly to each other.
-
-  Our results with flushing enabled confirmed this theory. When running with
-  flushing enabled, which had the additional overhead of having to
-  ensure that all the data was copied to Lustre, we noticed that occasionally, we
-  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
-  performance was degraded. The
-  maximum speedup observed when running on the production cluster with flushing enabled was
-  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
-  3.2$\times$, processing 8 ds001545 images.
-
-    \section{Discussion}
-    \subsection{Significance of speedups with Sea}
-
-    Our results show that the speedups obtained from using Sea when pre-processing functional MRI data can be 
-    substantial (up to 32$\times$) depending on the degradation of the Lustre
-    file system. The highest speed-ups observed on a production cluster were
-    found to be 11$\times$.
-
-    It is unrealistic for cluster users to know whether Lustre's
-    performance will be deteriorated during the lifetime of their application
-    execution, nor to predict how such a deterioration would impact the
-    application. The fact that Sea and Baseline perform similarly in the average
-    case is an indication that runtime is generally not deteriorated by Sea, as
-    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
-    is safe to use in the event where Lustre performance is sufficiently
-    degraded.
-
-    Although Sea usage did result in performance degradation at times, the magnitude
-    of the degradation was less than that of possible speedups
-    obtained. Our controlled experiments demonstrated that using Sea with a
-    degraded Lustre file system can result in average speedups of up to
-    2.5$\times$, which is likely to benefit scientific application deployments in production.
-
-    \subsection{Limitations of using Sea}
-
-    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
-     when the data is written at a rate which exceeds the rate at which the
-    compute node's page cache can flush results to Lustre. In our experiments,
-    the pipeline which consisted of both the shortest duration and largest
-    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
-    approximately 100~GB of data, the controlled cluster would have had
-    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
-    100~GB of page cache. Since all writes would not have occurred at the same
-    time and computation time was extended as a result of CPU contention, even
-    our controlled experiments did not showcase the benefits of Sea in a highly
-    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
-    the only way to augment data intensiveness of the applications is
-    through increased parallelism which is limited by the number of available
-    cores. As a result, we could only demonstrate the speedups brought by Sea when
-    Lustre performance had been degraded.
-
-    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
-    In such cases, the performance bottleneck is the
-    CPU time, which Sea does not address. Moreover, depending on the
-    application's I/O patterns, it is possible that there will be sufficient
-    page cache space and time to flush all the results asynchronously
-    to Lustre. However, even with minimal speedups, Sea can still be useful in
-    these cases to limit the number of new files created on Lustre, which can be
-    particularly useful in instances where users have strict file limits.
-
-    %\subsection{Sea uses more memory} \subsection{Other use cases} writes a lot
-    %but it's actually more complicated. When dataset increases
-    \subsection{Predicting speedups}
-
-    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
-    where we had
-    control over the global cluster usage and were able to demonstrate that
-    significant speedups could be obtained when all the Lustre storage devices were
-    busy writing other data concurrently. However, even these experiments were
-    limited in demonstrating potential speedups with Sea as there were many
-    controlled variables, some of which may vary from cluster to cluster. For
-    example, the production cluster had a total of 977 nodes and only 38 disks associated with
-    their Lustre scratch file system. We can imagine that if 900 of the nodes
-    are busy writing to the scratch file system, we would observe a speedup
-    larger than what has been reported on our dedicated cluster with only 8 compute
-    nodes and 44 disks associated to Lustre. However, these are not the only
-    factors that can affect reported results, others can include when the
-    experiment was executed, what kind of applications were being executed
-    alongside the experiment, scratch file system usage during experiments, and
-    what were the application read and write patterns.
-
-    Users cannot control or predict the cluster status at the time of their
-    experiments.
-    Our experiments show that when parallelism is optimal (no competing threads
-    and minimized compute time), we obtain speedups with Sea if Lustre
-    performance is degraded. Otherwise, we should experience the same
-    performance without Sea. The more data the application writes, the bigger
-    the speedups, since what cannot be written to page cache can still benefit from
-    being written to local storage.
-
-    The frequency and the size of the user application I/O also play an
-    important role in determining speedups. An application which writes a lot of
-    data, but spreads out the writes evenly, may result in no speedups from
-    using Sea, whereas an application which writes large amounts of data in
-    bursts may benefit from significant speedups. Furthermore, applications
-    which create many small files may also experience larger speedups with Sea
-    as the these small files may overburden the Lustre metadata server.
-
-    %  difficulties with summarizing speedups into guidelines. variability is
-    %  important and it's coming from the interaction between infrastructure and
-    %  application. Benefits of sea depend on parameters such as load of the
-    %  filesystem which is multifold and the application compute and file system
-    %  configuration. page cache examples:
-    %   - large dataset - > more computation
-    %   - data access patterns and bursty-ness
-    %   - small datasets + metadata accesses
-
-    \subsection{Prefetching, flushing and eviction with Sea}
-
-    Sea provides prefetching, flushing and eviction mechanisms to further
-    improve the speedups that can be obtained. Prefetching allows input data to
-    be moved to and manipulated from memory. All of our experiments with the SPM pipeline
-    used prefetching as SPM updates the input data via the use of a memory map.
-    Without prefetching, updates to the input files would have been performed
-    directly on Lustre, thus exhibiting a less important speedup. It was
-    favorable to proceed with prefetching, in this case, as the input files
-    themselves were not large compared to available memory space. However, in
-    cases with very large input files, it may be preferable to maintain the
-    memory map to Lustre.
-
-    Whereas flushing and eviction were not used in the controlled experiments,
-    they were enabled for the production cluster experiments. We made the assumption, that
-    users would want access to all available intermediate data, although this
-    would likely not be the case with the processing of full datasets. Thus,
-    eviction was mostly used to ensure that files eventually deleted by the
-    pipelines would not be copied to Lustre. Slowdowns experienced with Sea may
-    be a result of flushing occurring outside of the computation. We assume that
-    in large scale analyses, users will not need access to all intermediate
-    results, leading to better usage of tmpfs and also minimized writes to
-    Lustre. Such a scenario should lead to performance results more akin to
-    those obtained using Big Data frameworks.
-
-    \subsection{Glibc interception of neuroimaging pipelines}
-
-    As demonstrated by our Sea experiments, we have found that glibc interception
-    is effective at capturing all file system calls executed by the
-    applications. This means that Sea is compatible with three of the most
-    prominently used neuroimaging preprocessing pipelines. Although we have not
-    tested it extensively against all possible parameters and toolbox
-    applications, we assume that use of glibc within the toolboxes themselves are
-    likely to be consistent. Despite not testing against all parameters and
-    toolbox applications, Sea has been extensively tested using many integration
-    tests on 3 different Linux operating systems (Ubuntu, CentOS and Fedora),
-    and a total of 8 different OS versions.
-
-    The ability to use glibc interception over more common alternative methods
-    such as kernel-based or FUSE-based file systems was essential in ensuring
-    ease-of-use and minimal performance overheads. As both our controlled and
-    production environment experiments demonstrated, Sea's overhead is minimal
-    and the magnitude of the speedups to be gained by the use of Sea is
-    greater than the magnitude of the slowdowns.
-
-    %opportunities for other optimizations Sea may be beneficial
-    \subsection{Should Sea always be used ?}
-
-    Our results show that performance with Sea is generally hard to predict as
-    users do not have control over the variables outside the application they
-    are executing. We have previously discussed the conditions in which Sea may
-    not be favorable to use, however, differences in the execution environment
-    can still shift performance results to be favorable to Sea. As we cannot
-    guarantee performance with Sea, we break down this discussion from the
-    perspective of a cluster administrator and that of a user.
-
-    From the perspective of a cluster administrator, Sea is considered useful to
-    maintaining the global health of the system. By using available local
-    resources, Sea is able to alleviate strain on globally shared cluster
-    resources, thus improving the experience for all users. Sea is able to do
-    this through the use of eviction which ensures that data that is not
-    required never makes it to Lustre.
-
-    From the perspective of a user, Sea provides speedups when
-    Lustre performance is degraded and ensures limited overhead. It is most efficient
-    when the compute time and non-Lustre I/O overhead is minimal. Even in
-    instances where Sea does not improve performance, it helps reduce and ensure
-    that the number of files that make it to Lustre are the exact ones required,
-    therefore not exceeding any predetermined quota a user might have on the
-    shared file system.
-
-    \subsection{Conclusion and Future work}
-
-    We developed a lightweight data-management library, Sea, to seamlessly
-    provide data management capabilities to existing neuroimaging workflows. We
-    have tested our library on three of the most commonly used fMRI
-    preprocessing libraries and found that Sea was able to intercept and
-    redirect all application I/O calls without issue.
-
-    When benchmarking the performance of Sea compared to Baseline, it was found that Sea provided speedups
-    when Lustre performance was degraded. We were able to obtain up to a
-    32$\times$ in the controlled environment and 11$\times$ in a production
-    cluster. It was also found that speedups were greatest with a larger
-    filesize, as long as the compute time and non-Lustre I/O-related overheads
-    were also minimal.
-
-    Sea's benefits are not only limited to performance as
-    users can also benefit from limiting the number of files created on Lustre
-    with Sea. Archiving of the output directory on Lustre with Sea to further
-    reduce number of files may be an interesting addition to Sea.
-
-    %user
-
-    %sysadmin
-
-    %write intensive
-
-    %\subsection{Effective use of available cluster resources}
-    %\subsection{Comparisons to best case scenario}
-
-    %talk about testing
-
-    %Flushing - rely on modification time
     \section{Materials and Methods}
@@
 
+    \section{Results}
+
+    \subsection{Sea}
+
+\begin{figure*}
+
+    \centering
+    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
+    \captionsetup{justification=centering}
+    \captionof{figure}{The Sea data-management library overview}
+    \label{fig:seaneuro:diagram}
+\end{figure*}
+
+    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
+    the \texttt{LD\_PRELOAD}
+    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
+    glibc, on Linux systems. This enables Sea to redirect write calls
+    aimed at slower storage devices to a faster device whenever possible.
+    Similarly, when intercepting read calls, Sea can choose to read from a
+    faster device if a copy is available on that device. Sea decides which
+    storage location it can write to based on the details provided in an
+    initialization file called \texttt{sea.ini}. This file informs Sea of which
+    locations it can use to read and write to, and well as their order of
+    priority. Importantly, Sea is not a filesystem, but rather a
+    lightweight tool that redirects files to more optimal storage devices.
+
+    We define all locations in which Sea can redirect writes to as a cache. Such
+    caches include any available storage located on the compute nodes, such as
+    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs).
+
+    As neuroimaging pipeline results typically require post-processing and
+    HPC compute-local resources are only accessible during the reserved
+    duration, Sea provides functionality to flush and evict data to persistent
+    shared storage. To avoid interrupting ongoing
+    processing with data management operations, this is accomplished via a separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
+    that need to be persisted to storage within a file called
+    \texttt{.sea\_flushlist}, and temporary files which can be removed from
+    cache within a file called the \texttt{.sea\_evictlist}. Both these files
+    can be populated using regular expressions to denote the paths to flush and
+    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
+    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
+    operation and simply move the file from local to persistent storage rather
+    than copying it. Files that will be reused by the application should only be
+    flushed rather than flushed and evicted, as files can benefit from speedups
+    related to reading from fast rather than slow storage. Sea currently also
+    provides a rudimentary prefetch thread that can move files located within
+    Sea to the fastest available cache. To use Sea's prefetch
+    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
+    populated using regular expressions like the flushing and eviction files.
+
+    To interact with Sea, a mountpoint is used. The mountpoint
+    is an empty directory that behaves as a view to all the files and
+    directories stored within Sea. In order to keep track of the locations of
+    files within the mountpoint, Sea mirrors the directory structure of each
+    storage location across all caches. It is generally advisable to provide
+    empty storage locations for Sea to use as the mirroring of large directories
+    can take some time.
+
+    Sea can easily be executed directly using the available containers on the
+    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
+    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.
+
+% \tristan{I find this section lacks a clear focus. What are the key points from
+% these results? You should focus on them and possibly name your paragraphs
+% accordingly, for improved clarity. My takeaways would be the following:
+% \begin{enumerate} \item In controlled conditions, important speed-ups are
+% observed. Mention min, average and max speed-up for all conditions. \item
+% Speed-up is correlated with baseline makespan \item Other determinants of
+% speed-up are unclear \item Overhead is minor \item Speed-ups are also observed
+% in production cluster \end{enumerate}
+% }
+
+%\setcounter{figure}{2}
+%\setcounter{subfigure}{0}
+\begin{subfigure}
+%\setcounter{figure}{2}
+%\setcounter{subfigure}{0}
+    \centering
+    \begin{minipage}[b]{0.5\textwidth}
+      \centering
+        \includegraphics[width=\linewidth]{figures/slashbin_nozoom.pdf}
+        \captionsetup{justification=centering}
+        \captionof{subfigure}{Complete results}\label{fig:seaneuro:slashbinfull}
+    \end{minipage}%
+    \begin{minipage}[b]{0.5\textwidth}
+      \centering
+        \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
+        \captionsetup{justification=centering}
+        \captionof{subfigure}{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
+    \end{minipage}
+
+%\setcounter{figure}{2}
+\setcounter{subfigure}{-1}
+    \caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
+    denotes the total time between execution launch and completion of the last computing task.}
+    \label{fig:seaneuro:slashbin}
+\end{subfigure}
+
+% \begin{figure*}
+
+% % \begin{subfigure}{0.5\textwidth}
+%     \centering
+%     \captionsetup{width=.85\linewidth}
+%     \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
+%     \caption{Complete results}\label{fig:seaneuro:slashbinfull}
+% % \end{subfigure}
+% % \begin{subfigure}{0.5\textwidth}
+% %     \centering
+% %     \captionsetup{width=.85\linewidth}
+% %     \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
+% %     \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
+% % \end{subfigure}
+% \caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
+% denotes the total time between execution launch and completion of the last computing task.}
+% %\label{fig:seaneuro:slashbin}
+% \end{figure*}
+
+\subsection{Speedups observed in controlled environment}
+
+We benchmarked Sea by processing 3 functional MRI datasets of increasing sizes
+(ds001545: 27~GB, PREVENT-AD: 283~GB, Human Connectome Project: 79~TB, see
+Table~\ref{table:sea-neuro:data}) with 3 pipelines 
+from the most commonly-used neuroimaging toolboxes (AFNI, SPM, FSL), using a controlled HPC cluster 
+dedicated to our experiments. We deployed ``busy writers'' in the cluster
+that simulated the load from other user applications by writing to the shared file system at a controlled rate.
+
+Speedups were obtained by using Sea when Lustre
+performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
+although performance was variable. The largest speedup across all conditions
+was found to be 32$\times$ and occurred when SPM was processing a single
+HCP image with 6 active busy writers.
+On average, however, a single SPM pipeline
+process preprocessing a single HCP image produced the greatest speedup
+(12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
+using a single process was the next fastest pipeline, with an average speedup of
+(4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
+1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
+excellent speedups, which is likely due to a mix of prefetching the initial
+input files and the I/O patterns of the application. While it was expected that the
+AFNI pipeline would benefit the most from Sea due to its
+minimal compute time and the amount of data it generates, it did not benefit
+from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
+AFNI pipeline performs a very high number of glibc calls. While the overhead of an
+individual call is likely very minimal, it is possible that the culmination of
+all these calls results in a large enough bottleneck that affected both Sea and
+Baseline executions to an extensive amount. Moreover, the number of glibc calls
+aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
+obtained an average of 2$\times$ speedup across all experimental conditions.
+
+The FSL Feat pipeline, in contrast, appeared to be the most compute-bound of the three
+applications. Not only did it spend an extensive amount of time computing, the
+amount of output data generated was least of the three. Due to the
+compute-intensive nature of the FSL Feat pipeline, it is expected to have resulted in the
+least significant speedups.
+
+The HCP dataset obtained the greatest speedups with
+Sea, with $2.3\times$ on average vs. $1.9\times$ for ds001545  and $1.3\times$ for PREVENT-AD.
+Out of the three datasets, HCP has the largest images (see
+Table~\ref{table:sea-neuro:data}). Larger individual images mean that they
+occupy more page cache space and take longer to flush to Lustre. Unsurprisingly,
+the next dataset with the largest images (\SI{282}{\mebi\byte} compressed for a
+single image) results in the next largest speedups, followed by PREVENT-AD, the
+dataset with the smallest functional images.
+
+For both the pipelines and the datasets, we noticed that obtained speedups
+decreased with increased parallelism. We did not modify the default parallelism
+level of the applications in our experiments. As a result, each application
+process was likely attempting to utilize all cores, resulting in longer wait times
+with the increase in parallelism. With the added compute contention overhead, the benefits of using
+Sea becomes more limited.
+
+\subsection{Speedup correlated with Lustre degradation}
+
+Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
+busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). Occassional slowdowns did occur with the use of Sea. These slowdowns
+may arise from the initial read of the data, as they appeared to occur less
+frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.
+
+\subsection{Overhead of Sea was minimal}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\linewidth]{figures/beluga_sea_tmpfs.pdf}%
+\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
+\label{fig:seaneuro:tmpfs}
+\end{figure*}
+
+To determine the overhead of Sea, we compared the average performance of Sea with
+the performance of pipeline executing entirely within memory
+(Figure~\ref{fig:seaneuro:tmpfs}). The results demonstrated that Sea and tmpfs
+did not lead to significant differences in average makespan (p=0.9, two-sampled unpaired t-test). Given these results, we conclude that the
+overhead of Sea is minimal and that the differences in makespan may be
+due to Sea's initial reading of the data from Lustre.
+
+\subsection{Speedups observed in production environments}
+
+\begin{subfigure}
+  %\setcounter{figure}{2}
+  %\setcounter{subfigure}{0}
+      \centering
+      \begin{minipage}[b]{0.5\textwidth}
+        \centering
+          \includegraphics[width=\linewidth]{figures/beluga_withoutflush_nozoom.pdf}
+          \captionsetup{justification=centering}
+          \captionof{subfigure}{Complete results}\label{fig:seaneuro:belugafull}
+      \end{minipage}%
+      \begin{minipage}[b]{0.5\textwidth}
+        \centering
+          \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
+          \captionsetup{justification=centering}
+          \captionof{subfigure}{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
+      \end{minipage}
+
+  %\setcounter{figure}{2}
+  \setcounter{subfigure}{-1}
+      \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
+      \label{fig:seaneuro:beluga-noflush}
+  \end{subfigure}
+
+% \begin{figure*}
+
+% % \begin{subfigure}{0.5\textwidth}
+%     \centering
+%     \captionsetup{width=.85\linewidth}
+%     \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
+%     \caption{Complete results}\label{fig:seaneuro:belugafull}
+% % \end{subfigure}
+% % \begin{subfigure}{0.5\textwidth}
+% %     \centering
+% %     \captionsetup{width=.85\linewidth}
+% %     \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
+% %     \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
+% % \end{subfigure}
+% \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
+% \label{fig:seaneuro:beluga-noflush}
+% \end{figure*}
+
+\begin{subfigure}
+  %\setcounter{figure}{2}
+  %\setcounter{subfigure}{0}
+      \centering
+      \begin{minipage}[b]{0.5\textwidth}
+        \centering
+          \includegraphics[width=\linewidth]{figures/beluga_withflush_nozoom.pdf}
+          \captionsetup{justification=centering}
+          \captionof{subfigure}{Complete results}\label{fig:seaneuro:belugafullwf}
+      \end{minipage}%
+      \begin{minipage}[b]{0.5\textwidth}
+        \centering
+          \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
+          \captionsetup{justification=centering}
+          \captionof{subfigure}{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
+      \end{minipage}
+
+  %\setcounter{figure}{2}
+  \setcounter{subfigure}{-1}
+      \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
+      \label{fig:seaneuro:beluga-wflush}
+  \end{subfigure}
+
+% \begin{figure*}
+% %\begin{subfigure}{0.5\textwidth}
+%     \centering
+%     \captionsetup{width=.85\linewidth}
+%     \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
+%     \caption{Complete results}\label{fig:seaneuro:belugafullwf}
+% % \end{subfigure}
+% % \begin{subfigure}{0.5\textwidth}
+% %     \centering
+% %     \captionsetup{width=.85\linewidth}
+% %     \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
+% %     \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
+% % \end{subfigure}
+% \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
+% \label{fig:seaneuro:beluga-wflush}
+% \end{figure*}
+
+  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
+  shared with other users,
+  using the same datasets and pipelines as for the controlled cluster. 
+  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
+  show that although performance in production clusters was highly variable,
+  we could still obtain important speedups by using Sea. It is very likely that when we
+  executed our experiments without flushing
+  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
+  resulting in Sea and Baseline performing
+  quite similarly to each other.
+
+  Our results with flushing enabled confirmed this theory. When running with
+  flushing enabled, which had the additional overhead of having to
+  ensure that all the data was copied to Lustre, we noticed that occasionally, we
+  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
+  performance was degraded. The
+  maximum speedup observed when running on the production cluster with flushing enabled was
+  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
+  3.2$\times$, processing 8 ds001545 images.
+
+    \section{Discussion}
+    \subsection{Significance of speedups with Sea}
+
+    Our results show that the speedups obtained from using Sea when pre-processing functional MRI data can be 
+    substantial (up to 32$\times$) depending on the degradation of the Lustre
+    file system. The highest speed-ups observed on a production cluster were
+    found to be 11$\times$.
+
+    It is unrealistic for cluster users to know whether Lustre's
+    performance will be deteriorated during the lifetime of their application
+    execution, nor to predict how such a deterioration would impact the
+    application. The fact that Sea and Baseline perform similarly in the average
+    case is an indication that runtime is generally not deteriorated by Sea, as
+    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
+    is safe to use in the event where Lustre performance is sufficiently
+    degraded.
+
+    Although Sea usage did result in performance degradation at times, the magnitude
+    of the degradation was less than that of possible speedups
+    obtained. Our controlled experiments demonstrated that using Sea with a
+    degraded Lustre file system can result in average speedups of up to
+    2.5$\times$, which is likely to benefit scientific application deployments in production.
+
+    \subsection{Limitations of using Sea}
+
+    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
+     when the data is written at a rate which exceeds the rate at which the
+    compute node's page cache can flush results to Lustre. In our experiments,
+    the pipeline which consisted of both the shortest duration and largest
+    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
+    approximately 100~GB of data, the controlled cluster would have had
+    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
+    100~GB of page cache. Since all writes would not have occurred at the same
+    time and computation time was extended as a result of CPU contention, even
+    our controlled experiments did not showcase the benefits of Sea in a highly
+    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
+    the only way to augment data intensiveness of the applications is
+    through increased parallelism which is limited by the number of available
+    cores. As a result, we could only demonstrate the speedups brought by Sea when
+    Lustre performance had been degraded.
+
+    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
+    In such cases, the performance bottleneck is the
+    CPU time, which Sea does not address. Moreover, depending on the
+    application's I/O patterns, it is possible that there will be sufficient
+    page cache space and time to flush all the results asynchronously
+    to Lustre. However, even with minimal speedups, Sea can still be useful in
+    these cases to limit the number of new files created on Lustre, which can be
+    particularly useful in instances where users have strict file limits.
+
+    %\subsection{Sea uses more memory} \subsection{Other use cases} writes a lot
+    %but it's actually more complicated. When dataset increases
+    \subsection{Predicting speedups}
+
+    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
+    where we had
+    control over the global cluster usage and were able to demonstrate that
+    significant speedups could be obtained when all the Lustre storage devices were
+    busy writing other data concurrently. However, even these experiments were
+    limited in demonstrating potential speedups with Sea as there were many
+    controlled variables, some of which may vary from cluster to cluster. For
+    example, the production cluster had a total of 977 nodes and only 38 disks associated with
+    their Lustre scratch file system. We can imagine that if 900 of the nodes
+    are busy writing to the scratch file system, we would observe a speedup
+    larger than what has been reported on our dedicated cluster with only 8 compute
+    nodes and 44 disks associated to Lustre. However, these are not the only
+    factors that can affect reported results, others can include when the
+    experiment was executed, what kind of applications were being executed
+    alongside the experiment, scratch file system usage during experiments, and
+    what were the application read and write patterns.
+
+    Users cannot control or predict the cluster status at the time of their
+    experiments.
+    Our experiments show that when parallelism is optimal (no competing threads
+    and minimized compute time), we obtain speedups with Sea if Lustre
+    performance is degraded. Otherwise, we should experience the same
+    performance without Sea. The more data the application writes, the bigger
+    the speedups, since what cannot be written to page cache can still benefit from
+    being written to local storage.
+
+    The frequency and the size of the user application I/O also play an
+    important role in determining speedups. An application which writes a lot of
+    data, but spreads out the writes evenly, may result in no speedups from
+    using Sea, whereas an application which writes large amounts of data in
+    bursts may benefit from significant speedups. Furthermore, applications
+    which create many small files may also experience larger speedups with Sea
+    as the these small files may overburden the Lustre metadata server.
+
+    %  difficulties with summarizing speedups into guidelines. variability is
+    %  important and it's coming from the interaction between infrastructure and
+    %  application. Benefits of sea depend on parameters such as load of the
+    %  filesystem which is multifold and the application compute and file system
+    %  configuration. page cache examples:
+    %   - large dataset - > more computation
+    %   - data access patterns and bursty-ness
+    %   - small datasets + metadata accesses
+
+    \subsection{Prefetching, flushing and eviction with Sea}
+
+    Sea provides prefetching, flushing and eviction mechanisms to further
+    improve the speedups that can be obtained. Prefetching allows input data to
+    be moved to and manipulated from memory. All of our experiments with the SPM pipeline
+    used prefetching as SPM updates the input data via the use of a memory map.
+    Without prefetching, updates to the input files would have been performed
+    directly on Lustre, thus exhibiting a less important speedup. It was
+    favorable to proceed with prefetching, in this case, as the input files
+    themselves were not large compared to available memory space. However, in
+    cases with very large input files, it may be preferable to maintain the
+    memory map to Lustre.
+
+    Whereas flushing and eviction were not used in the controlled experiments,
+    they were enabled for the production cluster experiments. We made the assumption, that
+    users would want access to all available intermediate data, although this
+    would likely not be the case with the processing of full datasets. Thus,
+    eviction was mostly used to ensure that files eventually deleted by the
+    pipelines would not be copied to Lustre. Slowdowns experienced with Sea may
+    be a result of flushing occurring outside of the computation. We assume that
+    in large scale analyses, users will not need access to all intermediate
+    results, leading to better usage of tmpfs and also minimized writes to
+    Lustre. Such a scenario should lead to performance results more akin to
+    those obtained using Big Data frameworks.
+
+    \subsection{Glibc interception of neuroimaging pipelines}
+
+    As demonstrated by our Sea experiments, we have found that glibc interception
+    is effective at capturing all file system calls executed by the
+    applications. This means that Sea is compatible with three of the most
+    prominently used neuroimaging preprocessing pipelines. Although we have not
+    tested it extensively against all possible parameters and toolbox
+    applications, we assume that use of glibc within the toolboxes themselves are
+    likely to be consistent. Despite not testing against all parameters and
+    toolbox applications, Sea has been extensively tested using many integration
+    tests on 3 different Linux operating systems (Ubuntu, CentOS and Fedora),
+    and a total of 8 different OS versions.
+
+    The ability to use glibc interception over more common alternative methods
+    such as kernel-based or FUSE-based file systems was essential in ensuring
+    ease-of-use and minimal performance overheads. As both our controlled and
+    production environment experiments demonstrated, Sea's overhead is minimal
+    and the magnitude of the speedups to be gained by the use of Sea is
+    greater than the magnitude of the slowdowns.
+
+    %opportunities for other optimizations Sea may be beneficial
+    \subsection{Should Sea always be used ?}
+
+    Our results show that performance with Sea is generally hard to predict as
+    users do not have control over the variables outside the application they
+    are executing. We have previously discussed the conditions in which Sea may
+    not be favorable to use, however, differences in the execution environment
+    can still shift performance results to be favorable to Sea. As we cannot
+    guarantee performance with Sea, we break down this discussion from the
+    perspective of a cluster administrator and that of a user.
+
+    From the perspective of a cluster administrator, Sea is considered useful to
+    maintaining the global health of the system. By using available local
+    resources, Sea is able to alleviate strain on globally shared cluster
+    resources, thus improving the experience for all users. Sea is able to do
+    this through the use of eviction which ensures that data that is not
+    required never makes it to Lustre.
+
+    From the perspective of a user, Sea provides speedups when
+    Lustre performance is degraded and ensures limited overhead. It is most efficient
+    when the compute time and non-Lustre I/O overhead is minimal. Even in
+    instances where Sea does not improve performance, it helps reduce and ensure
+    that the number of files that make it to Lustre are the exact ones required,
+    therefore not exceeding any predetermined quota a user might have on the
+    shared file system.
+
+    \section{Conclusion and Future work}
+
+    We developed a lightweight data-management library, Sea, to seamlessly
+    provide data management capabilities to existing neuroimaging workflows. We
+    have tested our library on three of the most commonly used fMRI
+    preprocessing libraries and found that Sea was able to intercept and
+    redirect all application I/O calls without issue.
+
+    When benchmarking the performance of Sea compared to Baseline, it was found that Sea provided speedups
+    when Lustre performance was degraded. We were able to obtain up to a
+    32$\times$ in the controlled environment and 11$\times$ in a production
+    cluster. It was also found that speedups were greatest with a larger
+    filesize, as long as the compute time and non-Lustre I/O-related overheads
+    were also minimal.
+
+    Sea's benefits are not only limited to performance as
+    users can also benefit from limiting the number of files created on Lustre
+    with Sea. Archiving of the output directory on Lustre with Sea to further
+    reduce number of files may be an interesting addition to Sea.
+
+    %user
+
+    %sysadmin
+
+    %write intensive
+
+    %\subsection{Effective use of available cluster resources}
+    %\subsection{Comparisons to best case scenario}
+
+    %talk about testing
+
+    %Flushing - rely on modification time
+
     %Busy writers
@@
 
+    \bibliographystyle{Frontiers-Harvard}
     \bibliography{biblio}

### 2025-07-19 paper/sea-neuro/paper-neuro.tex
Add paper updates and remove unused classes

--- 2024-08-02 paper/sea-neuro/paper-neuro.tex
@@
-\documentclass[fleqn,10pt]{wlscirep}
+\documentclass[pdflatex,sn-mathphys-num]{sn-jnl}
+
+\usepackage[nopatch=footnote]{microtype}
+\usepackage{url}
+\usepackage{lineno}
+% \usepackage[onehalfspacing]{setspace}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
-\usepackage{caption}
+% \usepackage{caption}
 \usepackage{subcaption}
 \usepackage{xspace}
-\usepackage{todonotes}
+% \usepackage{graphicx}
+% \usepackage{siunitx}
 
-\newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
 
-\linenumbers
+% \usepackage{todonotes}
+% \usepackage{float}
+% \newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}
 
-\title{Hierarchical storage management in user space for neuroimaging applications}
+% \linenumbers
 
-\author{Val\'erie Hayot-Sasson}
-\author{Tristan Glatard}
-\affil{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
 
-\begin{abstract}
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
 
-  Neuroimaging open-data initiatives have led to increased availability of large
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+
+\begin{document}
+
+\title[Hierarchical storage management in user space for neuroimaging applications]{Hierarchical storage management in user space for neuroimaging applications}
+
+%%% Use the \authfn to add symbols for additional footnotes, if any. 1 is reserved for correspondence emails; then continuing with 2 etc for contributions.
+
+\author[1]{\fnm{Val\'erie} \sur{Hayot-Sasson}}%\email{valerie.hayot-sasson@concordia.ca}
+
+\author*[1]{\fnm{Tristan} \sur{Glatard}}\email{tristan.glatard@concordia.ca}
+
+\affil[1]{\orgdiv{Department of Computer Science and Software Engineering}, \orgname{Concordia University}, \orgaddress{\city{Montr\'eal}, \state{Qu\'ebec}, \country{Canada}}}
+
+% \author[1]{Val\'erie Hayot-Sasson}
+% \author[1,\authfn{1}]{Tristan Glatard}
+
+% \affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Qu\'ebec, Canada}
+
+%%% Author Notes
+% \authnote{\authfn{1}tristan.glatard@concordia.ca}
+
+% %%% Paper category
+% \papercat{Paper}
+
+% %%% "Short" author for running page header
+% \runningauthor{Hayot-Sasson and Glatard}
+
+% \begin{frontmatter}
+    % \onecolumn
+    % \firstpage{1}
+
+  %\title[Sea]{Hierarchical storage management in user space for neuroimaging applications}
+
+  % \author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+  % \address{} %This field will be automatically populated
+  % \correspondance{} 
+  % \extraAuth{}
+
+  \abstract{Neuroimaging open-data initiatives have led to increased availability of large
   scientific datasets. While these datasets are shifting the processing bottleneck
@@
   overburdened by other users, performance is unaffected by Sea, suggesting that Sea’s
-  overhead is minimal even in cases where its benefits are limited.
+  overhead is minimal even in cases where its benefits are limited.}
 
@@
   % is minimal, as it can be used to limit the number of files created on parallel file systems.
-\end{abstract}
-\begin{document}
+\keywords{data management, neuroimaging, high-performance computing}
 
-\flushbottom
 \maketitle
-%  Click the title above to edit the author information and abstract
+% \end{frontmatter}
 
-\thispagestyle{empty}
+% \flushbottom
+% \maketitle
+% %  Click the title above to edit the author information and abstract
+
+% \thispagestyle{empty}
 
@@
 
+    \section{Materials and Methods}
+
+    \subsection{Datasets and Pipelines}
+
+    To determine the performance gain Sea brings to neuroimaging analyses, we
+    must evaluate the value of Sea on a variety of neuroimaging applications.
+    For our analysis, we selected different fMRI preprocessing applications as
+    fMRI processing has some of the most well-established tools for neuroimaging
+    and some of the largest datasets. Of course, different modalities and tools
+    may result in vastly different data access patterns and compute times.
+
+    \subsubsection{Datasets}
+    fMRI datasets may vary greatly in total number of images and number of
+    volumes within each image. To adequately capture the diversity of datasets
+    and the applicability of Sea, we have selected three datasets of varying
+    time and space resolutions (Table~\ref{table:sea-neuro:data}): 1)
+    OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD
+    dataset~\cite{preventad}, and 3) the HCP dataset~\cite{HCP}. The ds001545
+    dataset is a total of \SI{46}{\giga\byte} (1,778 files) and consists of data
+    collected from 30 participants in a single session watching 3 different
+    clips (i.e. three runs/participants) of The Grand Budapest Hotel.
+
+    The PREVENT-AD dataset is an open dataset consisting of data from 330
+    participants with a familial history of Alzheimer's Disease. At the time of
+    the experiments, the dataset contained \SI{255}{\giga\byte} (53,061 files)
+    originating from the 308 subjects available on
+    DataLad~\cite{halchenko2021datalad}.
+
+    The HCP project, whose aim is to characterize brain connectivity using data
+    collected from 1,200 subjects, is the largest of the three datasets. At the
+    time of our experiments, the dataset obtained consisted \SI{85}{\tera\byte}
+    (15,716,060 files) of data collected from 1,113 subjects.
+
+    By experimenting with datasets of different spatial and temporal
+    resolutions, the amount of compute may be affected. Naturally, with more
+    data there may inevitably be more computations performed, but also, the code
+    itself may perform entirely different computations as a result of
+    differences in resolution. Moreover, parallelism and program I/O behaviour
+    may be impacted depending on the particular characteristics of the data.
+    Capturing this potential variety in processing will provide us with a better
+    understanding of when and where Sea can be useful.
+
+    \begin{table*}[t]
+      %\small
+      \centering
+    \resizebox{\textwidth}{!}{\begin{tabular}{|c c c c c|}
+      \hline
+      Dataset & Total Size (MB) & Total Number of images & Number of images per experiment & Total compressed size processed (MB) \\
+      \hline
+      \multirow{3}{*}{PREVENT-AD} & \multirow{3}{*}{289,532} &
+      \multirow{3}{*}{53,061} & 1 & 52\\
+      & & & 8 & 402\\
+      & & & 16 & 732\\
+      \hline
+      \multirow{3}{*}{ds001545} & \multirow{3}{*}{27,377} &
+      \multirow{3}{*}{1,778} & 1 & 282\\
+      & & & 8 & 2115\\
+      & & & 16 & 4167\\
+      \hline
+      \multirow{3}{*}{HCP} & \multirow{3}{*}{83,140,079} &
+      \multirow{3}{*}{15,716,060*} & 1 &  1301\\
+      & & & 8 & 5998\\
+      & & & 16 & 8328\\
+
+      \hline
+
+      \hline
+    \end{tabular}} \footnotesize{*Number estimated based on data we had access
+    to}\\
+    \caption{Dataset characteristics. Compressed size is listed, since all
+    applications, except for SPM, processed compressed
+    data}\label{table:sea-neuro:data}
+    \end{table*}
+
+    %just for example run dataset size, resolution, resolution of voxels matrix
+    %size pixdim1X2 number of slices pixdim number of volumes pixdim4 inplane
+    %resolution pixsize1x2 slice thickness pizsize3 in mm repetition time
+    %pixsize4 in s
+
+    %HCP resting state + x stats
+
+    %# files written and output size, duration and memory consumption.
+    \subsubsection{fMRI Preprocessing Pipelines}
+
+    Similarly to datasets, preprocessing pipelines can also vary in duration as
+    a result of methodological differences. Thus, we preprocessed each dataset
+    using three standard preprocessing pipelines available within the FSL~\cite{fsl},
+    AFNI~\cite{cox1996afni}, and SPM~\cite{spm} toolboxes.
+    Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
+    in computation and data-intensiveness of the different pipelines on a single
+    subject for all three of the datasets.
+
+    Each tool was set to only run the functional preprocessing pipeline. For the FSL preprocessing pipeline, 
+    we used FEAT configured with many of
+    the default parameters. We set the following parameters: slice-timing
+    correction (interleaved), intensity normalization and non-linear
+    registration to standard space. FSL version 6.0.4 was used.
+
+    %% Add SPM here
+    To preprocess the datasets with SPM, we reused the template described in
+    \cite{haitas2021age}, with fieldmap correction removed. As the template was
+    designed for a specific dataset, we updated the paths to the fMRI and
+    anatomical images, as well as the dimensions, such as the number of volumes
+    and slices and the TR and TA, for each dataset. Furthermore, we set the
+    slice order to interleaved in all cases. SPM version 12 was used.
+
+    For AFNI preprocessing, the pipeline was configured to perform slice timing,
+    alignment to Talaraich space, registration, smoothing and brain masking.
+    While AFNI does have some parallelized components that can be controlled via
+    environment variables, we chose to let the pipeline use as much parallelism
+    as required for its execution. AFNI version 21.1.02 was used.
+
+    % We set fMRIPrep to run with \texttt{--fs-no-reconall} flag set and set the
+    % \texttt{--bids\-database-dir} to a tmpfs location. Furthermore, the seeds
+    % were fixed to ensure reproducible results. Unlike the other tools where
+    % only a single fMRI image was preprocessed per process, it was not possible
+    % to do so with fMRIPrep. Thus, fMRIPrep preprocessed entire subject
+    % directories at a time
+
+    %%{provide subject #/session#} of the ds001545 dataset.
+
+    \begin{table*}[t]
+      %\small
+      \centering
+      \resizebox{\textwidth}{!}{\begin{tabular}{|c c c c c c|}
+      \hline
+      Tool & Dataset &  Output Size (MB) & Total glibc calls & Glibc Lustre calls
+      & Compute time (s)* \\
+      \hline
+      \multirow{3}{*}{AFNI} & PREVENT-AD & 540 & 272,342 & 4,118 & 103.25 \\
+      & ds001545 & 3,063 & 281,660 & 4,340 & 280.30 \\
+      & HCP & 18,720 & 305,555 & 5,137 & 816.16 \\
+      \hline
+      \multirow{3}{*}{FSL Feat} & PREVENT-AD & 254 & 191,148 & 28,099 &
+      1,338.29 \\
+      & ds001545 & 551 & 192,404 & 28,371 & 2,145.96 \\
+      & HCP & 1,608 & 192,445 & 28,997 & 6,596.46 \\
+      \hline
+      \multirow{3}{*}{SPM} & PREVENT-AD & 331.0 & 42,329 & 18,257 &  483.67 \\
+      & ds001545 & 744 & 54,481 & 27,770 & 446.53 \\
+      & HCP & 2,083 & 62,234 & 33,477 & 715.43 \\
+
+      \hline
+    \end{tabular}}
+    \footnotesize{$*$ Measurements taken on the dedicated cluster}
+    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
+    \label{table:seaneuro-pipelines}
+  \end{table*}
+
+    \subsection{Controls}
+    Our objective was to measure the speedup brought by Sea on real neuroimaging
+    pipelines. Although Sea is agnostic to the internals of the pipelines and
+    their output results, incorrect parameter selection and resulting outputs
+    may result in I/O patterns that would not be observed had the pipeline been
+    executed properly.
+
+    To validate that the outputs produced were sufficiently correct, we
+    performed visual quality control of the preprocessed images to ensure that
+    they were correctly skull-stripped and registered to the template brain (MNI
+    template). We repeated this step for all datasets and all pipelines
+    processing a single image. In addition, we compared output number of files
+    and output size to ensure that the pipeline was not producing different
+    outputs depending on the file system used.
+
+    % compare md5sums of uncompressed files.
+
+    \subsection{Infrastructure}
+
+    Our experiments were executed on two different clusters: 1) a controlled, dedicated cluster
+    and 2) and a production cluster. We chose this experimental setup as the dedicated
+    cluster allowed us to understand Sea's performance without activity of other users.
+    The production cluster gave us insight into what performance typical users might
+    be able to expect on a shared cluster.
+
+    The dedicated cluster was composed of 8
+    compute nodes with \SI{256}{\gibi\byte} memory and \SI{125}{\gibi\byte}
+    tmpfs, running Centos 8.1.1911.
+    Storage on the dedicated cluster consisted of 4 Lustre ZFS storage nodes with 44
+    HDD object storage targets (OSTs) and 1 metadata server (MDS) with a single
+    metadata target (MDT). The storage nodes were connected to the compute nodes
+    via a 20Gbps ethernet connection. The cluster's hardware was similar to that
+    of the Beluga cluster at the Digital Alliance of Canada.
+
+    The Beluga cluster was used as the production cluster.
+    It consisted of a total of 977 compute nodes, of which we
+    used up to 16 general compute nodes at a time with \SI{186}{\gibi\byte} of
+    available memory, \SI{480}{\gibi\byte} of local scratch SSD space, 2 Intel
+    Gold 6148 Skylake @ 2.4 GHz CPUs and Centos 7 installed. Each compute node
+    was connected to the Lustre scratch file system via an Mellanox Infiniband
+    EDR (100 Gbps) network interconnect. The Lustre scratch file system consists
+    of a total \SI{2.6}{\pebi\byte} of space, with 38 OSTs of
+    \SI{69.8}{\tebi\byte} each and two MDTs of \SI{3.3}{\tebi\byte}.
+
+    On the dedicated cluster, we ran the experiments with and without
+    Lustre degradation, which was induced by ``busy writers''. Without busy
+    writers, the fMRI preprocessing pipelines had exclusive access to the
+    cluster. With busy writers, the fMRI preprocessing pipelines were executed
+    alongside an Apache Spark application that continuously read and wrote approximately
+    1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
+    and writes. In our experiments, we either had 6 nodes each executing the
+    Spark application or no busy writers.
+
+    Our experiments generally did not use prefetching, flushing or eviction. The
+    only pipeline configured to prefetch was SPM, as the input data was otherwise
+    read through a memmap (i.e. only loading necessary portions of the file to memory) 
+    with Lustre. Flushing would normally be necessary with
+    preprocessing as the user would always require the output data. To
+    investigate the impacts of flushing on these pipelines, we performed a
+    separate set of experiments where AFNI and SPM would flush all data
+    produced, on the production cluster.
+
+    For all experiments, we compared the performance of Sea with that of Lustre
+    alone (Baseline). To ensure that system performance was equivalent between
+    Sea and our Baseline, we executed Sea and Baseline pipelines together.
+
+    All experiments were executed using 1, 8 and 16 processes.
+    Each process consisted of a single application call processing a single fMRI image.
+
     \section{Results}
@@
 
-\begin{figure*}
+\begin{figure*}[t]
 
     \centering
-    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
-\caption{The Sea data-management library overview}
+    \includegraphics[width=\linewidth]{figures/sea-diagram.pdf}%
+    \captionsetup{justification=centering}
+    \captionof{figure}{The Sea data-management library overview}
     \label{fig:seaneuro:diagram}
 \end{figure*}
+
     Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
@@
 % }
-\begin{figure*}
 
-\begin{subfigure}{0.5\textwidth}
+% %\setcounter{figure}{2}
+% %\setcounter{subfigure}{0}
+% \begin{subfigure}
+% %\setcounter{figure}{2}
+% %\setcounter{subfigure}{0}
+%     \centering
+%     \begin{minipage}[b]{0.5\textwidth}
+%       \centering
+%         \includegraphics[width=\linewidth]{figures/slashbin_nozoom.pdf}
+%         \captionsetup{justification=centering}
+%         \captionof{subfigure}{Complete results}\label{fig:seaneuro:slashbinfull}
+%     \end{minipage}%
+%     \begin{minipage}[b]{0.5\textwidth}
+%       \centering
+%         \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
+%         \captionsetup{justification=centering}
+%         \captionof{subfigure}{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
+%     \end{minipage}
+
+% %\setcounter{figure}{2}
+% \setcounter{subfigure}{-1}
+%     \caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
+%     denotes the total time between execution launch and completion of the last computing task.}
+%     \label{fig:seaneuro:slashbin}
+% \end{subfigure}
+
+\begin{figure*}[ht]
 \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
+
+\begin{subfigure}{0.8\textwidth}
+    \centering
+    \captionsetup{width=\textwidth}
+    \includegraphics[width=\textwidth]{figures/slashbin_nozoom.pdf}%
     \caption{Complete results}\label{fig:seaneuro:slashbinfull}
 \end{subfigure}
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.8\textwidth}
     \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
+    \captionsetup{width=\textwidth}
+    \includegraphics[width=\textwidth]{figures/slashbin_withzoom.pdf}
     \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
@@
 
-\begin{figure*}
+\begin{figure*}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
+\includegraphics[width=0.8\textwidth]{figures/beluga_sea_tmpfs.pdf}%
 \caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
@@
 
-\begin{figure*}
+% \begin{subfigure}
+%   %\setcounter{figure}{2}
+%   %\setcounter{subfigure}{0}
+%       \centering
+%       \begin{minipage}[b]{0.5\textwidth}
+%         \centering
+%           \includegraphics[width=\linewidth]{figures/beluga_withoutflush_nozoom.pdf}
+%           \captionsetup{justification=centering}
+%           \captionof{subfigure}{Complete results}\label{fig:seaneuro:belugafull}
+%       \end{minipage}%
+%       \begin{minipage}[b]{0.5\textwidth}
+%         \centering
+%           \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
+%           \captionsetup{justification=centering}
+%           \captionof{subfigure}{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
+%       \end{minipage}
 
-\begin{subfigure}{0.5\textwidth}
+%   %\setcounter{figure}{2}
+%   \setcounter{subfigure}{-1}
+%       \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
+%       \label{fig:seaneuro:beluga-noflush}
+%   \end{subfigure}
+
+\begin{figure*}[h!]
   \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
+
+\begin{subfigure}{0.8\textwidth}
+    \centering
+    \captionsetup{width=\linewidth}
+    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom.pdf}%
     \caption{Complete results}\label{fig:seaneuro:belugafull}
 \end{subfigure}
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.8\textwidth}
     \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
+    \captionsetup{width=\linewidth}
+    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_zoom.pdf}
     \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
@@
 
-\begin{figure*}
-\begin{subfigure}{0.5\textwidth}
+% \begin{subfigure}
+%   %\setcounter{figure}{2}
+%   %\setcounter{subfigure}{0}
+%       \centering
+%       \begin{minipage}[b]{0.5\textwidth}
+%         \centering
+%           \includegraphics[width=\linewidth]{figures/beluga_withflush_nozoom.pdf}
+%           \captionsetup{justification=centering}
+%           \captionof{subfigure}{Complete results}\label{fig:seaneuro:belugafullwf}
+%       \end{minipage}%
+%       \begin{minipage}[b]{0.5\textwidth}
+%         \centering
+%           \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
+%           \captionsetup{justification=centering}
+%           \captionof{subfigure}{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
+%       \end{minipage}
+
+%   %\setcounter{figure}{2}
+%   \setcounter{subfigure}{-1}
+%       \caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
+%       \label{fig:seaneuro:beluga-wflush}
+%   \end{subfigure}
+
+\begin{figure*}[ht]
   \centering
-    \captionsetup{width=.85\linewidth}
+\begin{subfigure}{0.8\textwidth}
+    \centering
+    \captionsetup{width=\linewidth}
     \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
@@
 \end{subfigure}
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.8\textwidth}
     \centering
-    \captionsetup{width=.85\linewidth}
-    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
+    \captionsetup{width=\textwidth}
+    \includegraphics[width=\textwidth]{figures/beluga_withflush_zoom.pdf}
     \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
@@
 
-    \subsection{Conclusion and Future work}
+    \section{Conclusion and Future work}
 
@@
     %Flushing - rely on modification time
-    \section{Materials and Methods}
-
-    \subsection{Datasets and Pipelines}
-
-    To determine the performance gain Sea brings to neuroimaging analyses, we
-    must evaluate the value of Sea on a variety of neuroimaging applications.
-    For our analysis, we selected different fMRI preprocessing applications as
-    fMRI processing has some of the most well-established tools for neuroimaging
-    and some of the largest datasets. Of course, different modalities and tools
-    may result in vastly different data access patterns and compute times.
-
-    \subsubsection{Datasets}
-    fMRI datasets may vary greatly in total number of images and number of
-    volumes within each image. To adequately capture the diversity of datasets
-    and the applicability of Sea, we have selected three datasets of varying
-    time and space resolutions (Table~\ref{table:sea-neuro:data}): 1)
-    OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD
-    dataset~\cite{preventad}, and 3) the HCP dataset~\cite{HCP}. The ds001545
-    dataset is a total of \SI{46}{\giga\byte} (1,778 files) and consists of data
-    collected from 30 participants in a single session watching 3 different
-    clips (i.e. three runs/participants) of The Grand Budapest Hotel.
-
-    The PREVENT-AD dataset is an open dataset consisting of data from 330
-    participants with a familial history of Alzheimer's Disease. At the time of
-    the experiments, the dataset contained \SI{255}{\giga\byte} (53,061 files)
-    originating from the 308 subjects available on
-    DataLad~\cite{halchenko2021datalad}.
-
-    The HCP project, whose aim is to characterize brain connectivity using data
-    collected from 1,200 subjects, is the largest of the three datasets. At the
-    time of our experiments, the dataset obtained consisted \SI{85}{\tera\byte}
-    (15,716,060 files) of data collected from 1,113 subjects.
-
-    By experimenting with datasets of different spatial and temporal
-    resolutions, the amount of compute may be affected. Naturally, with more
-    data there may inevitably be more computations performed, but also, the code
-    itself may perform entirely different computations as a result of
-    differences in resolution. Moreover, parallelism and program I/O behaviour
-    may be impacted depending on the particular characteristics of the data.
-    Capturing this potential variety in processing will provide us with a better
-    understanding of when and where Sea can be useful.
-
-    \begin{table*}[t]
-      \small\centering
-    \resizebox{\textwidth}{!}{\begin{tabular}{|c c c c c|}
-      \hline
-      Dataset & Total Size (MB) & Total Number of images & Number of images per experiment & Total compressed size processed (MB) \\
-      \hline
-      \multirow{3}{*}{PREVENT-AD} & \multirow{3}{*}{289,532} &
-      \multirow{3}{*}{53,061} & 1 & 52\\
-      & & & 8 & 402\\
-      & & & 16 & 732\\
-      \hline
-      \multirow{3}{*}{ds001545} & \multirow{3}{*}{27,377} &
-      \multirow{3}{*}{1,778} & 1 & 282\\
-      & & & 8 & 2115\\
-      & & & 16 & 4167\\
-      \hline
-      \multirow{3}{*}{HCP} & \multirow{3}{*}{83,140,079} &
-      \multirow{3}{*}{15,716,060*} & 1 &  1301\\
-      & & & 8 & 5998\\
-      & & & 16 & 8328\\
-
-      \hline
-
-      \hline
-    \end{tabular}} \footnotesize{*Number estimated based on data we had access
-    to}\\
-    \caption{Dataset characteristics. Compressed size is listed, since all
-    applications, except for SPM, processed compressed
-    data}\label{table:sea-neuro:data}
-    \end{table*}
-
-    %just for example run dataset size, resolution, resolution of voxels matrix
-    %size pixdim1X2 number of slices pixdim number of volumes pixdim4 inplane
-    %resolution pixsize1x2 slice thickness pizsize3 in mm repetition time
-    %pixsize4 in s
-
-    %HCP resting state + x stats
-
-    %# files written and output size, duration and memory consumption.
-    \subsubsection{fMRI Preprocessing Pipelines}
-
-    Similarly to datasets, preprocessing pipelines can also vary in duration as
-    a result of methodological differences. Thus, we preprocessed each dataset
-    using three standard preprocessing pipelines available within the FSL~\cite{fsl},
-    AFNI~\cite{cox1996afni}, and SPM~\cite{spm} toolboxes.
-    Table~\ref{table:seaneuro-pipelines} provides a reference of the differences
-    in computation and data-intensiveness of the different pipelines on a single
-    subject for all three of the datasets.
-
-    Each tool was set to only run the functional preprocessing pipeline. For the FSL preprocessing pipeline, 
-    we used FEAT configured with many of
-    the default parameters. We set the following parameters: slice-timing
-    correction (interleaved), intensity normalization and non-linear
-    registration to standard space. FSL version 6.0.4 was used.
-
-    %% Add SPM here
-    To preprocess the datasets with SPM, we reused the template described in
-    \cite{haitas2021age}, with fieldmap correction removed. As the template was
-    designed for a specific dataset, we updated the paths to the fMRI and
-    anatomical images, as well as the dimensions, such as the number of volumes
-    and slices and the TR and TA, for each dataset. Furthermore, we set the
-    slice order to interleaved in all cases. SPM version 12 was used.
-
-    For AFNI preprocessing, the pipeline was configured to perform slice timing,
-    alignment to Talaraich space, registration, smoothing and brain masking.
-    While AFNI does have some parallelized components that can be controlled via
-    environment variables, we chose to let the pipeline use as much parallelism
-    as required for its execution. AFNI version 21.1.02 was used.
-
-    % We set fMRIPrep to run with \texttt{--fs-no-reconall} flag set and set the
-    % \texttt{--bids\-database-dir} to a tmpfs location. Furthermore, the seeds
-    % were fixed to ensure reproducible results. Unlike the other tools where
-    % only a single fMRI image was preprocessed per process, it was not possible
-    % to do so with fMRIPrep. Thus, fMRIPrep preprocessed entire subject
-    % directories at a time
-
-    %%{provide subject #/session#} of the ds001545 dataset.
-
-    \begin{table*}[t]
-      \small\centering
-      \resizebox{\textwidth}{!}{\begin{tabular}{|c c c c c c|}
-      \hline
-      Tool & Dataset &  Output Size (MB) & Total glibc calls & Glibc Lustre calls
-      & Compute time (s)* \\
-      \hline
-      \multirow{3}{1em}{AFNI} & PREVENT-AD & 540 & 272,342 & 4,118 & 103.25 \\
-      & ds001545 & 3,063 & 281,660 & 4,340 & 280.30 \\
-      & HCP & 18,720 & 305,555 & 5,137 & 816.16 \\
-      \hline
-      \multirow{3}{1em}{FSL Feat} & PREVENT-AD & 254 & 191,148 & 28,099 &
-      1,338.29 \\
-      & ds001545 & 551 & 192,404 & 28,371 & 2,145.96 \\
-      & HCP & 1,608 & 192,445 & 28,997 & 6,596.46 \\
-      \hline
-      \multirow{3}{1em}{SPM} & PREVENT-AD & 331.0 & 42,329 & 18,257 &  483.67 \\
-      & ds001545 & 744 & 54,481 & 27,770 & 446.53 \\
-      & HCP & 2,083 & 62,234 & 33,477 & 715.43 \\
-
-      \hline
-    \end{tabular}}
-    \footnotesize{$*$ Measurements taken on the dedicated cluster}
-    \caption{Pipeline execution characteristics based on the processing of a single fMRI image using a single application process}
-    \label{table:seaneuro-pipelines}
-  \end{table*}
-
-    \subsection{Controls}
-    Our objective was to measure the speedup brought by Sea on real neuroimaging
-    pipelines. Although Sea is agnostic to the internals of the pipelines and
-    their output results, incorrect parameter selection and resulting outputs
-    may result in I/O patterns that would not be observed had the pipeline been
-    executed properly.
-
-    To validate that the outputs produced were sufficiently correct, we
-    performed visual quality control of the preprocessed images to ensure that
-    they were correctly skull-stripped and registered to the template brain (MNI
-    template). We repeated this step for all datasets and all pipelines
-    processing a single image. In addition, we compared output number of files
-    and output size to ensure that the pipeline was not producing different
-    outputs depending on the file system used.
-
-    % compare md5sums of uncompressed files.
-
-    \subsection{Infrastructure}
-
-    Our experiments were executed on two different clusters: 1) a controlled, dedicated cluster
-    and 2) and a production cluster. We chose this experimental setup as the dedicated
-    cluster allowed us to understand Sea's performance without activity of other users.
-    The production cluster gave us insight into what performance typical users might
-    be able to expect on a shared cluster.
-
-    The dedicated cluster was composed of 8
-    compute nodes with \SI{256}{\gibi\byte} memory and \SI{125}{\gibi\byte}
-    tmpfs, running Centos 8.1.1911.
-    Storage on the dedicated cluster consisted of 4 Lustre ZFS storage nodes with 44
-    HDD object storage targets (OSTs) and 1 metadata server (MDS) with a single
-    metadata target (MDT). The storage nodes were connected to the compute nodes
-    via a 20Gbps ethernet connection. The cluster's hardware was similar to that
-    of the Beluga cluster at the Digital Alliance of Canada.
-
-    The Beluga cluster was used as the production cluster.
-    It consisted of a total of 977 compute nodes, of which we
-    used up to 16 general compute nodes at a time with \SI{186}{\gibi\byte} of
-    available memory, \SI{480}{\gibi\byte} of local scratch SSD space, 2 Intel
-    Gold 6148 Skylake @ 2.4 GHz CPUs and Centos 7 installed. Each compute node
-    was connected to the Lustre scratch file system via an Mellanox Infiniband
-    EDR (100 Gbps) network interconnect. The Lustre scratch file system consists
-    of a total \SI{2.6}{\pebi\byte} of space, with 38 OSTs of
-    \SI{69.8}{\tebi\byte} each and two MDTs of \SI{3.3}{\tebi\byte}.
-
-    On the dedicated cluster, we ran the experiments with and without
-    Lustre degradation, which was induced by ``busy writers''. Without busy
-    writers, the fMRI preprocessing pipelines had exclusive access to the
-    cluster. With busy writers, the fMRI preprocessing pipelines were executed
-    alongside an Apache Spark application that continuously read and wrote approximately
-    1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
-    and writes. In our experiments, we either had 6 nodes each executing the
-    Spark application or no busy writers.
-
-    Our experiments generally did not use prefetching, flushing or eviction. The
-    only pipeline configured to prefetch was SPM, as the input data was otherwise
-    read through a memmap (i.e. only loading necessary portions of the file to memory) 
-    with Lustre. Flushing would normally be necessary with
-    preprocessing as the user would always require the output data. To
-    investigate the impacts of flushing on these pipelines, we performed a
-    separate set of experiments where AFNI and SPM would flush all data
-    produced, on the production cluster.
-
-    For all experiments, we compared the performance of Sea with that of Lustre
-    alone (Baseline). To ensure that system performance was equivalent between
-    Sea and our Baseline, we executed Sea and Baseline pipelines together.
-
-    All experiments were executed using 1, 8 and 16 processes.
-    Each process consisted of a single application call processing a single fMRI image.
 
@@
 
-    \section{Data Availability}
+    \section*{Data Availability}
 
@@
 
-    \section{Code Availability}
+    \section*{Code Availability}
 
@@
 
-    \section{Acknowledgements}
+    \section*{Declarations}
     Val\'erie Hayot-Sasson was funded by the Canadian Open
     Neuroscience Platform (CONP) Scholar Award. This work
-    was also supported by the Canada Foundation for Innova-
-    tion as well as the Canada Research Chairs program.
+    was also supported by the Canada Foundation for Innovation as well as the Canada Research Chairs program.
 
-    \section{Author Contributions}
+    \section*{Author Contributions}
     VHS contributed to idea development, experimental and model design, implementation of experiments and
@@
 
-    \section{Competing Interests}
+    \section*{Competing Interests}
     The authors declare no competing interests.
 
+    % \bibliographystyle{Frontiers-Harvard}
     \bibliography{biblio}

### 2025-07-20 bin/sea_flusher.sh
Fix Fedora 41 dockerfile to include python2 source install and fix bugs in test cases

@@ -184,8 +184,6 @@ flush () {
         fi
         flush_files+=$(echo ${all_files} | tr " " "\n" | grep -Eo "${rgx}" || true)
     done
-    #echo "re_flush $re_flush"
-    #echo "flush_files ${flush_files}"
 
     # if .sea_evictlist file contains regex
     for rgx in ${re_evict[@]+"${re_evict[@]}"}
@@ -195,10 +193,6 @@ flush () {
             evict_files+=" "
         fi
 
-        if [[ $(ls ${rgx} 2> /dev/null) == "" ]]
-        then
-            continue
-        fi
         evict_files+=$(echo ${all_files} | tr " " "\n" | grep -Eo "${rgx}" || true)
     done
 
@@ -260,6 +254,10 @@ flush () {
     flush_files=${tmp_flush}
     evict_files=${tmp_evict}
 
+    log "EVICT : ${evict_files}"
+    log "FLUSH : ${flush_files}"
+    log "FLEVICT: ${fe_files}"
+
     #echo "FLUSH ${flush_files}"
     #echo "EVICT ${evict_files}"
 
@@ -310,18 +308,18 @@ get_sources () {
     do
         source_lvl=$(cat ${conf_file} | grep "^\s*cache_$i" | cut -d "=" -f 2 | tr -d ' ;')
         IFS=',' read -ra curr_sources <<< "$source_lvl"
+        log "Fluser ${i}"
 
         resolved_sources=()
         for src in "${curr_sources[@]}"
         do
-            if [ ! -d ${src} ]
+            envvar=$(printenv | grep "^${src}=" | sed "s/^${src}=//g" || echo "")
+            if [[ ${envvar} != "" ]]
             then
-                envvar=$(printenv | grep "^${src}=" | sed "s/^${src}=//g" || echo "")
-                if [[ ${envvar} != "" ]]
-                then
-                    resolved_sources+=( ${envvar} )
-                fi
+                log "Flusher1 ${envvar}"
+                resolved_sources+=( ${envvar} )
             else
+                log "Flusher 2 ${src}"
                 resolved_sources+=( ${src} )
             fi
         done
@@ -333,13 +331,10 @@ get_sources () {
     done
 
     base_source=$(cat ${conf_file} | grep "^\s*cache_$i" | cut -d "=" -f 2 | tr -d ' ;')
-    if [ ! -d ${base_source} ]
[…]

### 2025-07-20 bin/sea_prefetch.sh

@@ -115,13 +115,10 @@ get_sources () {
         resolved_sources=()
         for src in "${curr_sources[@]}"
         do
-            if [ ! -d ${src} ]
+            envvar=$(printenv | grep "^${src}=" | sed "s/^${src}=//g" || echo "")
+            if [[ ${envvar} != "" ]]
             then
-                envvar=$(printenv | grep "^${src}=" | sed "s/^${src}=//g" || echo "")
-                if [[ ${envvar} != "" ]]
-                then
-                    resolved_sources+=( ${envvar} )
-                fi
+                resolved_sources+=( ${envvar} )
             else
                 resolved_sources+=( ${src} )
             fi
@@ -133,13 +130,10 @@ get_sources () {
 
     base_source=$(cat ${conf_file} | grep "^\s*cache_$i" | cut -d "=" -f 2 | tr -d ' ;')
 
-    if [ ! -d ${base_source} ]
+    envvar=$(printenv | grep "^${base_source}=" | sed "s/^${base_source}=//g" || echo "")
+    if [[ ${envvar} != "" ]]
     then
-        envvar=$(printenv | grep "^${base_source}=" | sed "s/^${base_source}=//g" || echo "")
-        if [[ ${envvar} != "" ]]
-        then
-            base_source=${envvar}
-        fi
+        base_source=${envvar}
     fi
 
     log "Prefetch: Base dir ${base_source}"

### 2025-07-20 src/config.cpp

@@ -8,6 +8,7 @@ extern "C"
 #include <sstream>
 #include <algorithm>
 #include <ctime>
+#include <cstring>
 #include <cstdlib>
 #include "config.h"
 #include "logger.h"
@@ -84,7 +85,7 @@ void parse_config()
     }
     if ((sea_config.n_levels = iniparser_getint(config_dict, "sea:n_levels", 0)) == 0)
     {
-        printf("Missing n_caches in config file %s\n", config_file);
+        printf("Missing n_levels in config file %s\n", config_file);
         exit(1);
     }
     std::vector<std::string> all_sources;

### 2025-07-20 src/functions.cpp

@@ -961,7 +961,7 @@ extern "C"
         init_path("unlinkat", abspath, passpath, 0, 1);
 
         struct stat buffer;
-        if (((funcptr___xstat)libc___xstat)(_STAT_VER_LINUX, passpath, &buffer) == 0)
+        if (((funcptr_stat)libc_stat)(passpath, &buffer) == 0)
         {
 
             if (S_ISDIR(buffer.st_mode))

### 2025-10-14 paper/sea-neuro/paper-neuro.tex
Update paper and generated figures

@@
 \usepackage{listings}%
-
 % \usepackage{todonotes}
@@
 % \newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}
-
-% \linenumbers
+\newcommand{\vhs}[1]{\textcolor{red}{#1}}
+\linenumbers
 
@@
 \newtheorem{definition}{Definition}%
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{myyamlstyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
 
@@
 
+    \subsection{Sea}
+
+\begin{figure*}[t]
+
+    \centering
+    \includegraphics[width=\linewidth]{figures/sea-diagram.pdf}%
+    \captionsetup{justification=centering}
+    \captionof{figure}{The Sea data-management library overview}
+    \label{fig:seaneuro:diagram}
+\end{figure*}
+
+    Sea (\autoref{fig:seaneuro:diagram}) is a data-management library that leverages
+    the \texttt{LD\_PRELOAD}
+    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
+    glibc, on Linux systems. Its goal is to easily enable users to intergrate efficient data management within their
+    applications within their existing workflows. Using glibc interception, Sea can redirect write calls
+    aimed at slower storage devices to a faster device whenever possible.
+    Similarly, when intercepting read calls, Sea can choose to read from a
+    faster device if a copy is available on that device. Sea decides which
+    storage location it can write to based on the details provided in an
+    initialization file called \texttt{sea.ini} (\autoref{lst:seaini}). This file informs Sea of which
+    locations it can use to read and write to, and well as their order of
+    priority. Importantly, Sea is not a filesystem, but rather a
+    lightweight tool that redirects files to more optimal storage devices.
+
+    \begin{figure}
+      \centering
+       \begin{lstlisting}[style=myyamlstyle, caption={Example Sea configuration file (sea.ini)}, label=lst:seaini]
+[Sea]                                                                  
+mount_dir = /lustre/seamount ;                                    
+n_levels = 3 ;                                                        
+cache_0 = /dev/shm/seasource ;                                           
+cache_1 = /localscratch/seasource ;                                         
+cache_2 = /lustre/seasource ;                                    
+log_level = 0 ;                                 
+log_file = /lustre/sea.log ;                             
+max_fs = 646971392 ;                                                   
+n_threads = 16 ;    \end{lstlisting}
+    \end{figure}
+
+    \vhs{To configure Sea, a few parameters need to be configured within a \texttt{sea.ini} file located in the directory from where the application will be launched.
+To interact with Sea, a mountpoint is used.
+The mountpoint
+    is an empty directory that behaves as a view to all the files and
+    directories stored within Sea. In order to keep track of the locations of
+    files within the mountpoint, Sea mirrors the directory structure of each
+    storage location across all caches. It is generally advisable to provide
+    empty storage locations for Sea to use as the mirroring of large directories
+    can take some time.}
+
+    \vhs{We define all locations in which Sea can redirect reads and  writes to as a cache. Such
+    caches include any available storage located on the compute nodes, such as
+    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs). When the application attempts to write
+    data to the Sea mountpoint, Sea will write the data to one of these cache locations, by order of preference (cache\_0 being the most optimal).
+    Since many cache locations may be available, the user must specify the number of cache levels available (n\_levels) followed by the locations for each of these levels.}
+
+    \vhs{Optionally, logging may be enabled, giving insight the glibc commands that were executed and how the filepaths were treated by Sea. There are five log levels
+    available (0:None, 1:Error, 2: Warning, 3: Info, 4: Debug).\autoref{fig:seaneuro:logs} shows how logs can be interpreted to gain insight into the decisions made by Sea.}
+
+    \vhs{It is important to note that Sea does not chunk files, and therefore, there is no risk of partitioned files across storage devices. Rather, it relies on configuration parameters
+    \texttt{max\_fs} and \texttt{n\_threads} to determine whether there is sufficient space on a storage device to write the file to. \texttt{max\_fs} represents an upper bound estimate of the maximum
+    filesize generated by the application in bytes. }
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=\textwidth]{figures/ex_log.png}
+       \captionsetup{justification=centering}
+    \captionof{figure}{Example logs outputted at log level 3 (INFO)}
+    \label{fig:seaneuro:logs}
+    \end{figure}
+
+    As neuroimaging pipeline results typically require post-processing and
+    HPC compute-local resources are only accessible during the reserved
+    duration, Sea provides functionality to flush and evict data to persistent
+    shared storage. To avoid interrupting ongoing
+    processing with data management operations, this is accomplished via a separate thread (known as the
+    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
+    that need to be persisted to storage within a file called
+    \texttt{.sea\_flushlist}, and temporary files which can be removed from
+    cache within a file called the \texttt{.sea\_evictlist}. Both these files
+    can be populated using regular expressions to denote the paths to flush and
+    evict (e.g., ~\autoref{fig:seaneuro:flushlist}). If a file occurs in both the \texttt{.sea\_flushlist} and
+    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
+    operation and simply move the file from local to persistent storage rather
+    than copying it. Files that will be reused by the application should only be
+    flushed rather than flushed and evicted, as files can benefit from speedups
+    related to reading from fast rather than slow storage. Sea currently also
+    provides a rudimentary prefetch thread that can move files located within
+    Sea to the fastest available cache. To use Sea's prefetch
+    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
+    populated using regular expressions like the flushing and eviction files.
+    When used, the prefetch, flush and eviction configuration files must all be located within the directory from where the application will be launched.
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.5\textwidth]{figures/ex_flushlist.png}
+       \captionsetup{justification=centering}
+    \captionof{figure}{Example \texttt{.sea\_flishlist} used for AFNI processing}
+    \label{fig:seaneuro:flushlist}
+    \end{figure}
+
+    Sea can easily be executed directly using the available containers on the
+    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
+    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.
+
     \subsection{Datasets and Pipelines}
@@
 
-    \subsection{Sea}
-
-\begin{figure*}[t]
-
-    \centering
-    \includegraphics[width=\linewidth]{figures/sea-diagram.pdf}%
-    \captionsetup{justification=centering}
-    \captionof{figure}{The Sea data-management library overview}
-    \label{fig:seaneuro:diagram}
-\end{figure*}
-
-    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
-    the \texttt{LD\_PRELOAD}
-    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
-    glibc, on Linux systems. This enables Sea to redirect write calls
-    aimed at slower storage devices to a faster device whenever possible.
-    Similarly, when intercepting read calls, Sea can choose to read from a
-    faster device if a copy is available on that device. Sea decides which
-    storage location it can write to based on the details provided in an
-    initialization file called \texttt{sea.ini}. This file informs Sea of which
-    locations it can use to read and write to, and well as their order of
-    priority. Importantly, Sea is not a filesystem, but rather a
-    lightweight tool that redirects files to more optimal storage devices.
-
-    We define all locations in which Sea can redirect writes to as a cache. Such
-    caches include any available storage located on the compute nodes, such as
-    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs).
-
-    As neuroimaging pipeline results typically require post-processing and
-    HPC compute-local resources are only accessible during the reserved
-    duration, Sea provides functionality to flush and evict data to persistent
-    shared storage. To avoid interrupting ongoing
-    processing with data management operations, this is accomplished via a separate thread (known as the
-    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
-    that need to be persisted to storage within a file called
-    \texttt{.sea\_flushlist}, and temporary files which can be removed from
-    cache within a file called the \texttt{.sea\_evictlist}. Both these files
-    can be populated using regular expressions to denote the paths to flush and
-    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
-    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
-    operation and simply move the file from local to persistent storage rather
-    than copying it. Files that will be reused by the application should only be
-    flushed rather than flushed and evicted, as files can benefit from speedups
-    related to reading from fast rather than slow storage. Sea currently also
-    provides a rudimentary prefetch thread that can move files located within
-    Sea to the fastest available cache. To use Sea's prefetch
-    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
-    populated using regular expressions like the flushing and eviction files.
-
-    To interact with Sea, a mountpoint is used. The mountpoint
-    is an empty directory that behaves as a view to all the files and
-    directories stored within Sea. In order to keep track of the locations of
-    files within the mountpoint, Sea mirrors the directory structure of each
-    storage location across all caches. It is generally advisable to provide
-    empty storage locations for Sea to use as the mirroring of large directories
-    can take some time.
-
-    Sea can easily be executed directly using the available containers on the
-    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
-    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.
-
 % \tristan{I find this section lacks a clear focus. What are the key points from
@@
 
-\begin{figure*}[ht]
-\centering
-
-\begin{subfigure}{0.8\textwidth}
+\begin{figure*}
+\centering
+
+\begin{subfigure}{\textwidth}
     \centering
     \captionsetup{width=\textwidth}
-    \includegraphics[width=\textwidth]{figures/slashbin_nozoom.pdf}%
+    \includegraphics[width=\textwidth]{figures/slashbin_nozoom_updated.pdf}%
     \caption{Complete results}\label{fig:seaneuro:slashbinfull}
 \end{subfigure}
-\begin{subfigure}{0.8\textwidth}
+\begin{subfigure}{\textwidth}
     \centering
     \captionsetup{width=\textwidth}
-    \includegraphics[width=\textwidth]{figures/slashbin_withzoom.pdf}
-    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
+    \includegraphics[width=\textwidth]{figures/slashbin_withzoom_updated.pdf}
+    \caption{Zoomed between 0 and 3000 seconds}\label{fig:seaneuro:slashbinzoom}
 \end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
-denotes the total time between execution launch and completion of the last computing task.}
+\caption{Makespan comparison between Sea and baseline (I/O to and from Lustre) on controlled dedicated cluster. Makespan
+denotes the total time between execution launch and completion of the last computing task. Five repetitions were recorded for each experimental condition (including with and without busy writers)
+using Sea and baseline. The black dots (without busy writers) represent pipeline makespan when pipeline has exclusive access to the cluster.
+a) shows that using Sea on SPM processing of large 
+datasets produces important speedups (marker below the y=x line) when compared to baseline, due to its data intensive nature.
+However, as seen in (b), for short application makespans, Sea performance is comparable to baseline. }
 \label{fig:seaneuro:slashbin}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/beluga_sea_tmpfs.pdf}%
-\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
+\includegraphics[width=\textwidth]{figures/beluga_sea_tmpfs_updated.pdf}%
+\caption{Makespan comparison between Sea and tmpfs (baseline) on the production cluster with flushing disabled. Mean pipeline execution time with Sea was recorded across five repetitions and compared
+to time it took to execute the same pipeline exclusively using data stored on tmpfs.
+As shown by the figure, Sea's mean runtime is comparable to that of writing directly to RAM, across pipelines, datasets and parallelism.}
 \label{fig:seaneuro:tmpfs}
@@
 
-\begin{subfigure}{0.8\textwidth}
+\begin{subfigure}{\textwidth}
     \centering
     \captionsetup{width=\linewidth}
-    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom.pdf}%
+    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom_updated.pdf}%
     \caption{Complete results}\label{fig:seaneuro:belugafull}
 \end{subfigure}
-\begin{subfigure}{0.8\textwidth}
+\begin{subfigure}{\textwidth}
     \centering
     \captionsetup{width=\linewidth}
-    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_zoom.pdf}
-    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
+    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom_updated.pdf}
+    \caption{Zoomed between 0 and 3000 seconds}\label{fig:seaneuro:belugazoom}
 \end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
+\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled. }
 \label{fig:seaneuro:beluga-noflush}
@@
 
-\begin{figure*}[ht]
-  \centering
-\begin{subfigure}{0.8\textwidth}
+\begin{figure*}[t]
+  \centering
+\begin{subfigure}{\textwidth}
     \centering
     \captionsetup{width=\linewidth}
-    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
+    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom_updated.pdf}%
     \caption{Complete results}\label{fig:seaneuro:belugafullwf}
 \end{subfigure}
-\begin{subfigure}{0.8\textwidth}
+\begin{subfigure}{\textwidth}
     \centering
     \captionsetup{width=\textwidth}
-    \includegraphics[width=\textwidth]{figures/beluga_withflush_zoom.pdf}
+    \includegraphics[width=\textwidth]{figures/beluga_withflush_zoom_updated.pdf}
     \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
 \end{subfigure}
-\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
+\caption{Makespan comparison between Sea and baseline (Lustre) on the production cluster with flushing enabled for all files.}
 \label{fig:seaneuro:beluga-wflush}

### 2025-10-15 paper/sea-neuro/paper-neuro.tex
Add more paper edits and updated figures

@@
 
+\definecolor{verylightgrey}{rgb}{0.92,0.92,0.92}
 \definecolor{codegreen}{rgb}{0,0.6,0}
@@
 
-    \subsection{Sea}
+    \subsection{The Sea data management library}
 
@@
 
-    \vhs{It is important to note that Sea does not chunk files, and therefore, there is no risk of partitioned files across storage devices. Rather, it relies on configuration parameters
+    \vhs{It is important to note that Sea does not chunk files, and therefore, there is no risk of corruption due to partitioned files across storage devices.
+Rather, it relies on configuration parameters
     \texttt{max\_fs} and \texttt{n\_threads} to determine whether there is sufficient space on a storage device to write the file to. \texttt{max\_fs} represents an upper bound estimate of the maximum
-    filesize generated by the application in bytes. }
+    filesize generated by the application in bytes, whereas \texttt{n\_threads} is the maximum number of parallel threads that could be writing a file of size \texttt{max\_fs} concurrently. }
+
+    \vhs{While Sea can function with applications that memory-map (memmap) files, it does not enhance the performance in instances where the file is memmapped.
+Memmap is an operation by which a small fragment of a large file is loaded (mapped) into memory.
+As the path of a memmaped file remains the same and data is updated in place, Sea will not be able to enhance performance of the I/O operations.
+}
 
@@
 
-    Sea can easily be executed directly using the available containers on the
-    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
+    \vhs{Sea can run alongside any application that dynamically loads and calls the libc library for read/write operations.
+Such applications commonly include interpreted applications, such as Python and Matlab/Octave.
+Sea can additionally work seamlessly from within a containerized environment, as long as the necessary cache directories required by Sea are mounted to the container.
+We provide various Docker containers preinstalled with Sea within the Github Container Registry (\url{https://github.com/orgs/big-data-lab-team/packages?repo_name=Sea}).
+Alternatively, Sea can be compiled from source using Make.} Once
     compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.
+
+        \begin{figure}
+      \centering
+       \begin{lstlisting}[language=bash,frame=single,
+    backgroundcolor=\color{verylightgrey}, caption={Example Sea deployment command}, basicstyle=\ttfamily, label=lst:launch]
+        $ SEA_HOME=$PWD ./sea_launch.sh <myprogram>
+      \end{lstlisting}
+    \end{figure}
 
@@
     To preprocess the datasets with SPM, we reused the template described in
-    \cite{haitas2021age}, with fieldmap correction removed. As the template was
+    \cite{haitas2021age}, with fieldmap correction removed. \vhs{As with the other applications, MNI registration was performed.} Since the template was
     designed for a specific dataset, we updated the paths to the fMRI and
@@
 
-    For AFNI preprocessing, the pipeline was configured to perform slice timing,
-    alignment to Talaraich space, registration, smoothing and brain masking.
+    For AFNI preprocessing, the pipeline was configured to perform slice timing, \vhs{EPI alignment to anatomical, 
+    linear alignment to MNI space,} registration, smoothing \vhs{(blur size of 4)} and brain masking.
     While AFNI does have some parallelized components that can be controlled via
@@
 
-    To validate that the outputs produced were sufficiently correct, we
+    To validate that the outputs produced were \vhs{unaltered by Sea, we fixed the application seeds and}
     performed visual quality control of the preprocessed images to ensure that
     they were correctly skull-stripped and registered to the template brain (MNI
-    template). We repeated this step for all datasets and all pipelines
-    processing a single image. In addition, we compared output number of files
+    template). We repeated this step for all datasets and all pipelines \vhs{(with and without using Sea)}
+    processing a single image. In addition, we compared output number of files\vhs{, file hashes (MD5),} 
     and output size to ensure that the pipeline was not producing different
@@
 (12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
-using a single process was the next fastest pipeline, with an average speedup of
+using a single process had the \vhs{next greatest speedup}, with an average speedup of
 (4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of

### 2025-10-16 paper/sea-neuro/paper-neuro.tex
Add few more paper edits

@@
     glibc, on Linux systems. Its goal is to easily enable users to intergrate efficient data management within their
-    applications within their existing workflows. Using glibc interception, Sea can redirect write calls
+    applications within their existing workflows.
+\vhs{Sea benefits most applications that process large quantities of data.
+These can include applications that process large files, as they would benefit from increased bandwidth provided by faster storage devices, as well as large quantities of small files, as such applications would benefit from decreased latency brought on by higher performance storage devices.
+Applications relying on shared filesystems whose performance may be affected by other user behaviour may also benefit from Sea usage, as node-local temporary storage is most commonly dedicated to the user's in HPC environments}.
+
+    With glibc interception, Sea can redirect write calls
     aimed at slower storage devices to a faster device whenever possible.
@@
 Sea can additionally work seamlessly from within a containerized environment, as long as the necessary cache directories required by Sea are mounted to the container.
-We provide various Docker containers preinstalled with Sea within the Github Container Registry (\url{https://github.com/orgs/big-data-lab-team/packages?repo_name=Sea}).
+We provide various Docker containers preinstalled with Sea within the Github Container Registry\footnote{\url{https://github.com/orgs/big-data-lab-team/packages?repo_name=Sea}}.
 Alternatively, Sea can be compiled from source using Make.} Once
@@
     correction (interleaved), intensity normalization and non-linear
-    registration to standard space. FSL version 6.0.4 was used.
+    registration to standard space. \vhs{Outputted files were compressed NIfTI (.nii.gz) format.} FSL version 6.0.4 was used.
 
@@
     To preprocess the datasets with SPM, we reused the template described in
-    \cite{haitas2021age}, with fieldmap correction removed. \vhs{As with the other applications, MNI registration was performed.} Since the template was
+    \cite{haitas2021age}, with fieldmap correction removed.
+\vhs{SPM was configured to perform slice timing correction, motion correction for within-subject registration, structural alignment to standard space MNI template, and an 8mm gaussian blur.} Since the template was
     designed for a specific dataset, we updated the paths to the fMRI and
@@
     and slices and the TR and TA, for each dataset. Furthermore, we set the
-    slice order to interleaved in all cases. SPM version 12 was used.
-
-    For AFNI preprocessing, the pipeline was configured to perform slice timing, \vhs{EPI alignment to anatomical, 
-    linear alignment to MNI space,} registration, smoothing \vhs{(blur size of 4)} and brain masking.
+    slice order to interleaved in all cases. \vhs{Output files were in uncompressed NIfTI (.nii) format.} SPM version 12 was used.
+
+    For AFNI preprocessing, the pipeline was configured to perform slice timing correction, \vhs{EPI alignment to anatomical with robust motion correction, 
+    warping of both anatomical and EPI to MNI space,} smoothing \vhs{(blur size of 4)} and brain masking.
     While AFNI does have some parallelized components that can be controlled via
     environment variables, we chose to let the pipeline use as much parallelism
-    as required for its execution. AFNI version 21.1.02 was used.
+    as required for its execution. \vhs{Output files generated were in uncompressed .BRIK/.HEAD format.} AFNI version 21.1.02 was used.
 
@@
 Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
-busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). Occassional slowdowns did occur with the use of Sea. These slowdowns
+busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). Occasional slowdowns did occur with the use of Sea. These slowdowns
 may arise from the initial read of the data, as they appeared to occur less
@@
     Dockerfiles containing AFNI, SPM, and FSL with Sea pre-installed are available
-    at~\url{https://github.com/ValHayot/Sea/tree/master/containers/applications}.
+    at~\url{https://github.com/big-data-lab-team/Sea/tree/main/containers/applications}. Sea implementation is available at \url{https://github.com/big-data-lab-team/Sea}.
 

### 2025-10-16 paper/sea-neuro/paper-neuro.tex
Update paper section on actions

@@
     toolbox applications, Sea has been extensively tested using many integration
-    tests on 3 different Linux operating systems (Ubuntu, CentOS and Fedora),
-    and a total of 8 different OS versions.
+    tests on four different Linux operating systems (Ubuntu, Rocky Linux, AlmaLinux,and Fedora),
+    and a total of nine different OS versions.
 
